### 2022-02-17 moduleoneanalysis.Rmd
coding module one

---
title: "moduleoneanalysis"
output:
  html_document:
    fig_width: 7
    fig_height: 6
    fig_caption: true
    keep_md: true
    theme: paper
    df_print: paged
---

```{r setup, include=FALSE}

```

```{r}

```

```{r}

```