### 2022-02-27 moduletwoanalysis.Rmd
improved perceived risk graph

---
title: "module2_eco_pol_factors"
output:
  html_document:
    fig_width: 7
    fig_height: 6
    fig_caption: true
    keep_md: true
    theme: paper
    df_print: paged
---

```{r load packages, warning = FALSE, message = FALSE}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(readxl))
suppressPackageStartupMessages(library(hablar))
suppressPackageStartupMessages(library(tsibble))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(DT))
suppressPackageStartupMessages(library(ggplot2))
```

```{r}
moduletwo <- read_excel( "~/Desktop/thesis_surveydata/political_factors_complete.xlsx")
save(moduletwo, file = "moduletwo.rda")
module2 <- data.frame(moduletwo)
module2 %>%
  filter(!row_number() %in% c(1,2))

# loaded data set for module 2 eco -pol variables and removes the rows with question number and question text
```

```{r, eval =TRUE}
#This chunk contains the coding schemes of various scales used in survey one: eco-political factors, kahan scale and acceptance scale
codedmodule2 <- module2 %>%

#remove row 1
  filter(!row_number() %in% c(1,2)) %>%

# replace risky likert scale with numbers
  mutate_at(vars(starts_with("Risky")), funs(case_when(. =="Not at all risky" ~ 1, 
                                                       . =="Slightly risky" ~ 2, 
                                                       . =="Moderately risky" ~ 3, 
                                                       . =="Very risky" ~ 4, 
                                                       . =="Extremely risky" ~ 5))) %>%

# replace beneficial likert scale with numbers  
  mutate_at(vars(starts_with("Ben")), funs(case_when(. =="Not at all beneficial" ~ 1,
                                                     . =="Slightly beneficial" ~ 2,
                                                     . =="Moderately beneficial" ~ 3,
                                                     . =="Very beneficial" ~ 4,
                                                     . =="Extremely beneficial" ~ 5 ))) %>%

# replace nuclear acceptance likert scale with numbers
  mutate_at(vars(N_accept,N_reluctantlyaccept,N_reject), funs(case_when(. == "Strongly disagree" ~ 1, 
                                                                        . == "Somewhat disagree" ~ 2,
                                                                        . == "Neither agree nor disagree" ~3,
                                                                        . == "Somewhat agree" ~ 4,
                                                                        . == "Strongly agree" ~ 5))) %>%

# code likert scale for variables for Kahan scale into numbers
  mutate_at(vars(starts_with (c("K_I","K_H","DISPLACE", "POLLUTE", "HEALTH", "JOBS", "BEAUTY", "PRIDE", "NPRIDE", "DEV", "PROSPER", "RELY"))), funs(case_when(. == "Strongly disagree" ~ 1, 
               . == "Somewhat disagree" ~ 2,
               . == "Neither agree nor disagree" ~3,
               . == "Somewhat agree" ~ 4,
               . == "Strongly agree" ~ 5))) %>%

# reverse code for likert scale for variables for Kahan scale into numbers
  mutate_at(vars(starts_with (c("K_S","K_E"))), funs(case_when(. == "Strongly disagree" ~ 5, 
                                                               . == "Somewhat disagree" ~ 4,
                                                               . == "Neither agree nor disagree" ~3,
                                                               . == "Somewhat agree" ~ 2,
                                                               . == "Strongly agree" ~ 1))) %>%

# code eco-pol scale variables into numbers
  mutate_at(vars(SYSTEMDEMO,SYSTEMRELIGION,SYSTEMTECHNO,SYSTEMTOTAL,WEALTHLIM,MECHANISATION,DECISIONDECEN,INDUSTRYLARGE,ECONOMYGLOBAL,OWNERPVT,OWNERNOREG), funs(case_when(. == "Strongly disagree" ~ 1, 
                        . == "Somewhat disagree" ~ 2,
                        . == "Neither agree nor disagree" ~3,
                        . == "Somewhat agree" ~ 4,
                        . == "Strongly agree" ~ 5))) %>%

# reverse code eco-pol scale variables into numbers 
 mutate_at(vars(DECISIONCEN, INDUSTRYSMALL, ECONOMYLOCAL, ENVOVERDEV,OWNERPUB, OWNERREG), funs(case_when(. == "Strongly disagree" ~ 5, 
                                                                                                         . == "Somewhat disagree" ~ 4,
                                                                                                         . == "Neither agree nor disagree" ~3,
                                                                                                         . == "Somewhat agree" ~ 2,
                                                                                                         . == "Strongly agree" ~ 1)))

codedmodule2

```
```{r, eval=TRUE}

# summary statistics for risks from different energy sources

risksummary <- codedmodule2 %>%
  summarize_at(vars(starts_with(c("RISKY"))), list(~mean(., na.rm = TRUE), ~median(., na.rm = TRUE), ~sd(.,na.rm = TRUE),~var(.,na.rm = TRUE), ~n())) %>%
  #add row index so later spreading indexed correctly
  add_rownames()%>%
  #melt to long format
  gather(technology, value, -rowname) %>%
  # separate risky from variable suffix
  separate(technology, c("Risky", "var"), extra = "merge", fill = "left") %>%
  #separate mean from variable prefix
  separate(var, c("technology", "summary")) %>%
  # spread summary values back to wide form
  spread(summary,value) %>%
  #clean up
  select(-rowname, -Risky) %>%
  arrange(mean)%>%
  setNames(paste0('perceivedrisk.', names(.)))%>%
  rename(technology = perceivedrisk.technology)

risksummary

```

```{r}
#plotting perceived risk from different technologies

module2%>%
  filter(!row_number() %in% c(1,2)) %>% 
  select(starts_with("RISKY"))%>%
  gather(Technology, Perceived_Risk, Risky_Hydro:Risky_INDLPG, factor_key = TRUE)%>%
  separate(Technology, c("Risky", "Technology"), extra = "merge", fill = "left")%>%
  select(-Risky) %>%
  group_by(Technology, Perceived_Risk) %>%
  summarise(n=n())

```

```{r}

colorsequence<- c("#8C4646","#D9695F","#F2A679","#F2D091","#5D8C7B","#808080")
df_barplot<-module2%>%
  filter(!row_number() %in% c(1,2)) %>% 
  select(starts_with("RISKY"))%>%
  gather(Technology, Perceived_Risk, Risky_Hydro:Risky_INDLPG, factor_key = TRUE)%>%
  separate(Technology, c("Risky", "Technology"), extra = "merge", fill = "left")%>%
  select(-Risky) %>%
  group_by(Technology, Perceived_Risk) %>%
  summarise(n=n())

df_barplot %>% ggplot(aes(fill= fct_rev(factor(Perceived_Risk, levels=c("Extremely risky", "Very risky","Moderately risky","Slightly risky","Not at all risky","Rather not say/Don't know"))), y=n , x=Technology)) + 
  geom_bar(position="fill",stat="identity")+
  scale_fill_manual("legend", values = c("Extremely risky" = "#8C4646", "Very risky" = "#D9695F", "Moderately risky" = "#F2A679", "Slightly risky" = "#F2D091","Not at all risky" = "#5D8C7B", "Rather not say/Don't know" = "#808080"))+
  coord_flip()+
  theme_classic()

# fix ugly visual. Rearrange from extremely risky to not at all risky. Red to green, rather not say/Don't know should be grey. Also facet wrap by INDividually owned and centrally owned technology.

```

```{r}

# summary statistics for benefits from different energy sources

benefitsummary <- codedmodule2 %>%
  summarize_at(vars(starts_with(c("Ben"))), list(~mean(., na.rm = TRUE), ~median(., na.rm = TRUE), ~sd(.,na.rm = TRUE),~var(.,na.rm = TRUE), ~n())) %>%
  #add row index so later spreading indexed correctly
  add_rownames()%>%
  #melt to long format
  gather(technology, value, -rowname) %>%
  # separate Ben from variable suffix
  separate(technology, c("Ben", "var"), extra = "merge", fill = "left") %>%
  #separate mean from variable prefix
  separate(var, c("technology", "summary")) %>%
  # spread summary values back to wide form
  spread(summary,value) %>%
  #clean up
  select(-rowname, -Ben) %>%
  arrange(mean)%>%
  setNames(paste0('perceivedbenefit.', names(.))) %>%
  rename(technology = perceivedbenefit.technology)
benefitsummary

#round off to two decimal places
```

```{r}

module2%>%
  filter(!row_number() %in% c(1,2)) %>% 
  select(starts_with("Ben"))%>%
  gather(Technology, Perceived_Benefit, Ben_Hydro:Ben_INDLPG, factor_key = TRUE)%>%
  separate(Technology, c("Ben", "Technology"), extra = "merge", fill = "left")%>%
  select(-Ben) %>%
  group_by(Technology, Perceived_Benefit) %>%
  summarise(n=n())%>%
ggplot(aes(fill= Perceived_Benefit, y=n, x=Technology)) + 
  geom_bar(position="fill", stat="identity")+
  coord_flip()

# fix ugly visual. Rearrange from extremely beneficial to not at all beneficial. Green to red, rather not say/Don't know should be grey. Also facet wrap by INDividually owned and centrally owned technology. Also arrange descending from most beneficial to least beneficial.

```

```{r}
#summary statistics for nuclear energy accept, reject and reluctantly accept scale
codedmodule2 %>%
  summarize_at(vars(N_accept,N_reject,N_reluctantlyaccept), list(~mean(., na.rm = TRUE),~median(., na.rm = TRUE), ~sd(.,na.rm = TRUE),~var(.,na.rm = TRUE),~n())) %>%
  add_rownames() %>%
  gather(acceptance, value, -rowname) %>%
  separate(acceptance, c("N_accept", "var") ,  extra = "merge", fill = "left") %>%
  separate(var, c("Acceptance", "summary")) %>%
  spread(summary, value) %>%
  select(-rowname, -N_accept)%>%
  arrange(mean)

```

```{r}

# chi square test for risk perception from nuclear energy and gender, race

# first combine gender and caste variables

chitest <- codedmodule2 %>%
  select(Risky_Nuclear,Risky_Hydro, Risky_Hydro, Risky_Solar,Urban_Rural, Gender, Caste) %>%
  na.omit() %>%
  filter(!Caste=="Rather not say/Don't know")%>%
  unite(gendercaste, Gender, Caste, sep = "_", remove = FALSE)

chitest

```

```{r}
full_join(risksummary,benefitsummary, by = NULL, copy = FALSE, keep = FALSE) %>%
  select(technology, perceivedrisk.mean, perceivedbenefit.mean) %>%
   gather(key = "level", value = "mean",
       perceivedrisk.mean, perceivedbenefit.mean) %>%
  ggplot(aes(x=technology, y= mean, fill = level))+
  geom_bar(stat="identity", position=position_dodge())+
  coord_flip()

# arrange descending and facet wrap by IND prefix

```

```{r}
full_join(risksummary,benefitsummary, by = NULL, copy = FALSE, keep = FALSE) %>%
  select(technology, perceivedrisk.median, perceivedbenefit.median) %>%
   gather(key = "level", value = "median",
       perceivedrisk.median, perceivedbenefit.median) %>%
  ggplot(aes(x=technology, y= median, fill = level))+
  geom_bar(stat="identity", position=position_dodge())+
  coord_flip()

# arrange descending and facet wrap by IND prefix

```

```{r}

chisq.test(chitest$Risky_Nuclear, chitest$Urban_Rural)
```