### 2022-03-14 module2graphsandPCA.Rmd

---
title: "module2graphsandPCA"
output: 
output:
  html_document:
    fig_width: 7
    fig_height: 6
    fig_caption: yes
    keep_md: yes
    theme: paper
    df_print: paged

---

```{r load packages, warning = FALSE, message = FALSE, echo=FALSE}
  knitr::opts_chunk$set(warning = FALSE, message = FALSE)

suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(readxl))
suppressPackageStartupMessages(library(hablar))
suppressPackageStartupMessages(library(tsibble))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(DT))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(ggbiplot))

```

```{r, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
moduletwo <- read_excel( "~/Desktop/thesis_surveydata/political_factors_complete.xlsx")
save(moduletwo, file = "moduletwo.rda")
module2 <- data.frame(moduletwo)
module2[module2 == "Rather not say/ Don't know"] <- "Rather not say/Don't know"
module2 %>%
  filter(!row_number() %in% c(1,2))

# loaded data set for module 2 eco -pol variables and removes the rows with question number and question text
```

```{r, eval =TRUE, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
#This chunk contains the coding schemes of various scales used in survey one: eco-political factors, kahan scale and acceptance scale
codedmodule2 <- module2 %>%

#remove row 1
  filter(!row_number() %in% c(1,2)) %>%

# replace risky likert scale with numbers
  mutate_at(vars(starts_with("Risky")), funs(case_when(. =="Not at all risky" ~ 1, 
                                                       . =="Slightly risky" ~ 2, 
                                                       . =="Moderately risky" ~ 3, 
                                                       . =="Very risky" ~ 4, 
                                                       . =="Extremely risky" ~ 5))) %>%

# replace beneficial likert scale with numbers  
  mutate_at(vars(starts_with("Ben")), funs(case_when(. =="Not at all beneficial" ~ 1,
                                                     . =="Slightly beneficial" ~ 2,
                                                     . =="Moderately beneficial" ~ 3,
                                                     . =="Very beneficial" ~ 4,
                                                     . =="Extremely beneficial" ~ 5 ))) %>%

# replace nuclear acceptance likert scale with numbers
  mutate_at(vars(N_accept,N_reluctantlyaccept,N_reject), funs(case_when(. == "Strongly disagree" ~ 1, 
                                                                        . == "Somewhat disagree" ~ 2,
                                                                        . == "Neither agree nor disagree" ~3,
                                                                        . == "Somewhat agree" ~ 4,
                                                                        . == "Strongly agree" ~ 5))) %>%

# code likert scale for variables for Kahan scale into numbers
  mutate_at(vars(starts_with (c("K_I","K_H","DISPLACE", "POLLUTE", "HEALTH", "JOBS", "BEAUTY", "PRIDE", "NPRIDE", "DEV", "PROSPER", "RELY"))), funs(case_when(. == "Strongly disagree" ~ 1, 
               . == "Somewhat disagree" ~ 2,
               . == "Neither agree nor disagree" ~3,
               . == "Somewhat agree" ~ 4,
               . == "Strongly agree" ~ 5))) %>%

# reverse code for likert scale for variables for Kahan scale into numbers
  mutate_at(vars(starts_with (c("K_S","K_E"))), funs(case_when(. == "Strongly disagree" ~ 5, 
                                                               . == "Somewhat disagree" ~ 4,
                                                               . == "Neither agree nor disagree" ~3,
                                                               . == "Somewhat agree" ~ 2,
                                                               . == "Strongly agree" ~ 1))) %>%

# code eco-pol scale variables into numbers
  mutate_at(vars(SYSTEMDEMO,SYSTEMRELIGION,SYSTEMTECHNO,SYSTEMTOTAL,WEALTHLIM,MECHANISATION,DECISIONDECEN,INDUSTRYLARGE,ECONOMYGLOBAL,OWNERPVT,OWNERNOREG), funs(case_when(. == "Strongly disagree" ~ 1, 
                        . == "Somewhat disagree" ~ 2,
                        . == "Neither agree nor disagree" ~3,
                        . == "Somewhat agree" ~ 4,
                        . == "Strongly agree" ~ 5))) %>%

# reverse code eco-pol scale variables into numbers 
 mutate_at(vars(DECISIONCEN, INDUSTRYSMALL, ECONOMYLOCAL, ENVOVERDEV,OWNERPUB, OWNERREG), funs(case_when(. == "Strongly disagree" ~ 5, 
                                                                                                         . == "Somewhat disagree" ~ 4,
                                                                                                         . == "Neither agree nor disagree" ~3,
                                                                                                         . == "Somewhat agree" ~ 2,
                                                                                                         . == "Strongly agree" ~ 1)))

codedmodule2
```

1. Plot the  mean risk scores (Q1)as histogram ordered from high to low risk – so basically the red bars only in reverse order (aka nuclear as high risk).

```{r, eval=TRUE, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

# summary statistics for risks from different energy sources

risksumcentral <- codedmodule2 %>%
  select(!starts_with(c("RISKY_IND")))%>%
  summarize_at(vars(starts_with(c("RISKY"))), list(~mean(., na.rm = TRUE), ~median(., na.rm = TRUE), ~sd(.,na.rm = TRUE),~var(.,na.rm = TRUE), ~n())) %>%
  #add row index so later spreading indexed correctly
  add_rownames()%>%
  #melt to long format
  gather(technology, value, -rowname) %>%
  # separate risky from variable suffix
  separate(technology, c("Risky", "var"), extra = "merge", fill = "left") %>%
  #separate mean from variable prefix
  separate(var, c("technology", "summary")) %>%
  # spread summary values back to wide form
  spread(summary,value) %>%
  #clean up
  select(-rowname, -Risky) %>%
  arrange(mean) %>%
  setNames(paste0('perceivedrisk.', names(.))) %>%
  dplyr::rename(technology= perceivedrisk.technology)

risksumcentral

```

```{r, echo=FALSE}

knitr::opts_chunk$set(warning = FALSE, message = FALSE)

risksumcentral %>%
  select(technology, perceivedrisk.mean) %>%
  mutate(technology = fct_reorder(technology, perceivedrisk.mean, .desc = FALSE))%>%
  ggplot(aes(x=technology, y= perceivedrisk.mean))+
  geom_bar(stat="identity", position=position_dodge(), fill = "firebrick2")+
  ylim(0,4)+
  labs(y = "Mean Values of Perceived Risk", x = "Energy Technologies")+
  ggtitle("Your Title Here")+
  coord_flip()

```

```{r, eval=TRUE, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

# summary statistics for risks from different energy sources

risksumIND <- codedmodule2 %>%
  summarize_at(vars(starts_with(c("RISKY_IND"))), list(~mean(., na.rm = TRUE), ~median(., na.rm = TRUE), ~sd(.,na.rm = TRUE),~var(.,na.rm = TRUE), ~n())) %>%
  #add row index so later spreading indexed correctly
  add_rownames()%>%
  #melt to long format
  gather(technology, value, -rowname) %>%
  # separate risky from variable suffix
  separate(technology, c("Risky", "var"), extra = "merge", fill = "left") %>%
  #separate mean from variable prefix
  separate(var, c("technology", "summary")) %>%
  # spread summary values back to wide form
  spread(summary,value) %>%
  #clean up
  select(-rowname, -Risky) %>%
  arrange(mean) %>%
  setNames(paste0('perceivedrisk.', names(.))) %>%
  dplyr::rename(technology= perceivedrisk.technology)

risksumIND

```

```{r, echo=FALSE}

knitr::opts_chunk$set(warning = FALSE, message = FALSE)

risksumIND %>%
  select(technology, perceivedrisk.mean) %>%
  mutate(technology = fct_reorder(technology, perceivedrisk.mean, .desc = FALSE))%>%
  ggplot(aes(x=technology, y= perceivedrisk.mean))+
  geom_bar(stat="identity", position=position_dodge(), fill = "firebrick2")+
  ylim(0,4)+
  labs(y = "Mean Values of Perceived Risk", x = "Energy Technologies")+
  ggtitle("Your Title Here")+
  coord_flip()

```

```{r, eval=TRUE, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

# summary statistics for risks from different energy sources

bensumcentral <- codedmodule2 %>%
  select(!starts_with(c("Ben_IND")))%>%
  summarize_at(vars(starts_with(c("Ben"))), list(~mean(., na.rm = TRUE), ~median(., na.rm = TRUE), ~sd(.,na.rm = TRUE),~var(.,na.rm = TRUE), ~n())) %>%
  #add row index so later spreading indexed correctly
  add_rownames()%>%
  #melt to long format
  gather(technology, value, -rowname) %>%
  # separate risky from variable suffix
  separate(technology, c("Ben", "var"), extra = "merge", fill = "left") %>%
  #separate mean from variable prefix
  separate(var, c("technology", "summary")) %>%
  # spread summary values back to wide form
  spread(summary,value) %>%
  #clean up
  select(-rowname, -Ben) %>%
  arrange(mean) %>%
  setNames(paste0('perceivedbenefit.', names(.))) %>%
  dplyr::rename(technology= perceivedbenefit.technology)

bensumcentral

```

```{r, echo=FALSE}

knitr::opts_chunk$set(warning = FALSE, message = FALSE)

bensumcentral %>%
  select(technology, perceivedbenefit.mean) %>%
  mutate(technology = fct_reorder(technology, perceivedbenefit.mean, .desc = FALSE))%>%
  ggplot(aes(x=technology, y= perceivedbenefit.mean))+
  geom_bar(stat="identity", position=position_dodge(), fill = "palegreen3")+
  ylim(0,4)+
  labs(y = "Mean Values of Perceived Benefit", x = "Energy Technologies")+
  ggtitle("Your Title Here")+
  coord_flip()

```

```{r, eval=TRUE, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

# summary statistics for risks from different energy sources

bensumIND <- codedmodule2 %>%
  summarize_at(vars(starts_with(c("Ben_IND"))), list(~mean(., na.rm = TRUE), ~median(., na.rm = TRUE), ~sd(.,na.rm = TRUE),~var(.,na.rm = TRUE), ~n())) %>%
  #add row index so later spreading indexed correctly
  add_rownames()%>%
  #melt to long format
  gather(technology, value, -rowname) %>%
  # separate risky from variable suffix
  separate(technology, c("Ben", "var"), extra = "merge", fill = "left") %>%
  #separate mean from variable prefix
  separate(var, c("technology", "summary")) %>%
  # spread summary values back to wide form
  spread(summary,value) %>%
  #clean up
  select(-rowname, -Ben) %>%
  arrange(mean) %>%
  setNames(paste0('perceivedbenefit.', names(.))) %>%
  dplyr::rename(technology= perceivedbenefit.technology)

bensumIND

```

```{r, echo=FALSE}

knitr::opts_chunk$set(warning = FALSE, message = FALSE)

bensumIND %>%
  select(technology, perceivedbenefit.mean) %>%
  mutate(technology = fct_reorder(technology, perceivedbenefit.mean, .desc = FALSE))%>%
  ggplot(aes(x=technology, y= perceivedbenefit.mean))+
  geom_bar(stat="identity", position=position_dodge(), fill = "palegreen3")+
  ylim(0,4)+
  labs(y = "Mean Values of Perceived Benefit", x = "Energy Technologies")+
  ggtitle("Your Title Here")+
  coord_flip()

```

```{r, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

center_barplot <- full_join( bensumcentral,risksumcentral, by = NULL, copy = FALSE, keep = FALSE) %>%
  select(technology, perceivedbenefit.mean, perceivedrisk.mean) %>%
  mutate(technology = fct_reorder(technology, perceivedbenefit.mean, .desc = FALSE))%>%
   gather(key = "level", value = "mean", perceivedbenefit.mean, perceivedrisk.mean)

center_barplot$level <- as.factor(center_barplot$level)                                          
class(center_barplot$level)
levels(center_barplot$level)

 center_barplot%>% mutate(level = fct_relevel(level, 
            "perceivedbenefit.mean", "perceivedrisk.mean"))+
  ggplot(aes(y=technology, x= mean, fill =  level))+
  geom_bar(stat="identity", position=position_dodge())+
  xlim(0,4)+
  scale_fill_manual("legend", values = c("perceivedbenefit.mean" = "palegreen3","perceivedrisk.mean" = "firebrick2"))

```

```{r, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

full_join(bensumIND,risksumIND, by = NULL, copy = FALSE, keep = FALSE) %>%
  select(technology, perceivedbenefit.mean, perceivedrisk.mean) %>%
  mutate(technology = fct_reorder(technology, perceivedbenefit.mean, .desc = FALSE))%>%
  gather(key = "level", value = "mean", perceivedbenefit.mean, perceivedrisk.mean) %>%
  ggplot(aes(x=technology, y= mean, fill = level ))+
  geom_bar(stat="identity", position=position_dodge())+
  ylim(0,4)+
  scale_fill_manual("legend", values = c("perceivedbenefit.mean" = "palegreen3","perceivedrisk.mean" = "firebrick2"))+
  coord_flip()

```