### 2022-03-14 module2graphsandPCA.Rmd
did PCA on all scales

---
title: "module2graphsandPCA"
output: 
output:
  html_document:
    fig_width: 7
    fig_height: 6
    fig_caption: yes
    keep_md: yes
    theme: paper
    df_print: paged

---

```{r load packages, warning = FALSE, message = FALSE, echo=FALSE}
  knitr::opts_chunk$set(warning = FALSE, message = FALSE)

suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(readxl))
suppressPackageStartupMessages(library(hablar))
suppressPackageStartupMessages(library(tsibble))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(DT))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(ggbiplot))

```

```{r, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
moduletwo <- read_excel( "~/Desktop/thesis_surveydata/political_factors_complete.xlsx")
save(moduletwo, file = "moduletwo.rda")
module2 <- data.frame(moduletwo)
module2[module2 == "Rather not say/ Don't know"] <- "Rather not say/Don't know"
module2 %>%
  filter(!row_number() %in% c(1,2))

# loaded data set for module 2 eco -pol variables and removes the rows with question number and question text
```

```{r, eval =TRUE, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
#This chunk contains the coding schemes of various scales used in survey one: eco-political factors, kahan scale and acceptance scale
codedmodule2 <- module2 %>%

#remove row 1
  filter(!row_number() %in% c(1,2)) %>%

# replace risky likert scale with numbers
  mutate_at(vars(starts_with("Risky")), funs(case_when(. =="Not at all risky" ~ 1, 
                                                       . =="Slightly risky" ~ 2, 
                                                       . =="Moderately risky" ~ 3, 
                                                       . =="Very risky" ~ 4, 
                                                       . =="Extremely risky" ~ 5))) %>%

# replace beneficial likert scale with numbers  
  mutate_at(vars(starts_with("Ben")), funs(case_when(. =="Not at all beneficial" ~ 1,
                                                     . =="Slightly beneficial" ~ 2,
                                                     . =="Moderately beneficial" ~ 3,
                                                     . =="Very beneficial" ~ 4,
                                                     . =="Extremely beneficial" ~ 5 ))) %>%

# replace nuclear acceptance likert scale with numbers
  mutate_at(vars(N_accept,N_reluctantlyaccept,N_reject), funs(case_when(. == "Strongly disagree" ~ 1, 
                                                                        . == "Somewhat disagree" ~ 2,
                                                                        . == "Neither agree nor disagree" ~3,
                                                                        . == "Somewhat agree" ~ 4,
                                                                        . == "Strongly agree" ~ 5))) %>%

# code likert scale for variables for Kahan scale into numbers
  mutate_at(vars(starts_with (c("K_I","K_H","DISPLACE", "POLLUTE", "HEALTH", "JOBS", "BEAUTY", "PRIDE", "NPRIDE", "DEV", "PROSPER", "RELY"))), funs(case_when(. == "Strongly disagree" ~ 1, 
               . == "Somewhat disagree" ~ 2,
               . == "Neither agree nor disagree" ~3,
               . == "Somewhat agree" ~ 4,
               . == "Strongly agree" ~ 5))) %>%

# reverse code for likert scale for variables for Kahan scale into numbers
  mutate_at(vars(starts_with (c("K_S","K_E"))), funs(case_when(. == "Strongly disagree" ~ 5, 
                                                               . == "Somewhat disagree" ~ 4,
                                                               . == "Neither agree nor disagree" ~3,
                                                               . == "Somewhat agree" ~ 2,
                                                               . == "Strongly agree" ~ 1))) %>%

# code eco-pol scale variables into numbers
  mutate_at(vars(SYSTEMDEMO,SYSTEMRELIGION,SYSTEMTECHNO,SYSTEMTOTAL,WEALTHLIM,MECHANISATION,DECISIONDECEN,INDUSTRYLARGE,ECONOMYGLOBAL,OWNERPVT,OWNERNOREG), funs(case_when(. == "Strongly disagree" ~ 1, 
                        . == "Somewhat disagree" ~ 2,
                        . == "Neither agree nor disagree" ~3,
                        . == "Somewhat agree" ~ 4,
                        . == "Strongly agree" ~ 5))) %>%

# reverse code eco-pol scale variables into numbers 
 mutate_at(vars(DECISIONCEN, INDUSTRYSMALL, ECONOMYLOCAL, ENVOVERDEV,OWNERPUB, OWNERREG), funs(case_when(. == "Strongly disagree" ~ 5, 
                                                                                                         . == "Somewhat disagree" ~ 4,
                                                                                                         . == "Neither agree nor disagree" ~3,
                                                                                                         . == "Somewhat agree" ~ 2,
                                                                                                         . == "Strongly agree" ~ 1)))

codedmodule2
```

1. Plot the  mean risk scores (Q1)as histogram ordered from high to low risk – so basically the red bars only in reverse order (aka nuclear as high risk).

```{r, eval=TRUE, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

# summary statistics for risks from different energy sources

risksumcentral <- codedmodule2 %>%
  select(!starts_with(c("RISKY_IND")))%>%
  summarize_at(vars(starts_with(c("RISKY"))), list(~mean(., na.rm = TRUE), ~median(., na.rm = TRUE), ~sd(.,na.rm = TRUE),~var(.,na.rm = TRUE), ~n())) %>%
  #add row index so later spreading indexed correctly
  add_rownames()%>%
  #melt to long format
  gather(technology, value, -rowname) %>%
  # separate risky from variable suffix
  separate(technology, c("Risky", "var"), extra = "merge", fill = "left") %>%
  #separate mean from variable prefix
  separate(var, c("technology", "summary")) %>%
  # spread summary values back to wide form
  spread(summary,value) %>%
  #clean up
  select(-rowname, -Risky) %>%
  arrange(mean) %>%
  setNames(paste0('perceivedrisk.', names(.))) %>%
  dplyr::rename(technology= perceivedrisk.technology)

risksumcentral

```

```{r, echo=FALSE}

knitr::opts_chunk$set(warning = FALSE, message = FALSE)

risksumcentral %>%
  select(technology, perceivedrisk.mean) %>%
  mutate(technology = fct_reorder(technology, perceivedrisk.mean, .desc = FALSE))%>%
  ggplot(aes(x=technology, y= perceivedrisk.mean))+
  geom_bar(stat="identity", position=position_dodge(), fill = "firebrick2")+
  ylim(0,4)+
  labs(y = "Mean Values of Perceived Risk", x = "Energy Technologies")+
  ggtitle("Your Title Here")+
  coord_flip()

```

```{r, eval=TRUE, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

# summary statistics for risks from different energy sources

risksumIND <- codedmodule2 %>%
  summarize_at(vars(starts_with(c("RISKY_IND"))), list(~mean(., na.rm = TRUE), ~median(., na.rm = TRUE), ~sd(.,na.rm = TRUE),~var(.,na.rm = TRUE), ~n())) %>%
  #add row index so later spreading indexed correctly
  add_rownames()%>%
  #melt to long format
  gather(technology, value, -rowname) %>%
  # separate risky from variable suffix
  separate(technology, c("Risky", "var"), extra = "merge", fill = "left") %>%
  #separate mean from variable prefix
  separate(var, c("technology", "summary")) %>%
  # spread summary values back to wide form
  spread(summary,value) %>%
  #clean up
  select(-rowname, -Risky) %>%
  arrange(mean) %>%
  setNames(paste0('perceivedrisk.', names(.))) %>%
  dplyr::rename(technology= perceivedrisk.technology)

risksumIND

```

```{r, echo=FALSE}

knitr::opts_chunk$set(warning = FALSE, message = FALSE)

risksumIND %>%
  select(technology, perceivedrisk.mean) %>%
  mutate(technology = fct_reorder(technology, perceivedrisk.mean, .desc = FALSE))%>%
  ggplot(aes(x=technology, y= perceivedrisk.mean))+
  geom_bar(stat="identity", position=position_dodge(), fill = "firebrick2")+
  ylim(0,4)+
  labs(y = "Mean Values of Perceived Risk", x = "Energy Technologies")+
  ggtitle("Your Title Here")+
  coord_flip()

```

```{r, eval=TRUE, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

# summary statistics for risks from different energy sources

bensumcentral <- codedmodule2 %>%
  select(!starts_with(c("Ben_IND")))%>%
  summarize_at(vars(starts_with(c("Ben"))), list(~mean(., na.rm = TRUE), ~median(., na.rm = TRUE), ~sd(.,na.rm = TRUE),~var(.,na.rm = TRUE), ~n())) %>%
  #add row index so later spreading indexed correctly
  add_rownames()%>%
  #melt to long format
  gather(technology, value, -rowname) %>%
  # separate risky from variable suffix
  separate(technology, c("Ben", "var"), extra = "merge", fill = "left") %>%
  #separate mean from variable prefix
  separate(var, c("technology", "summary")) %>%
  # spread summary values back to wide form
  spread(summary,value) %>%
  #clean up
  select(-rowname, -Ben) %>%
  arrange(mean) %>%
  setNames(paste0('perceivedbenefit.', names(.))) %>%
  dplyr::rename(technology= perceivedbenefit.technology)

bensumcentral

```

```{r, echo=FALSE}

knitr::opts_chunk$set(warning = FALSE, message = FALSE)

bensumcentral %>%
  select(technology, perceivedbenefit.mean) %>%
  mutate(technology = fct_reorder(technology, perceivedbenefit.mean, .desc = FALSE))%>%
  ggplot(aes(x=technology, y= perceivedbenefit.mean))+
  geom_bar(stat="identity", position=position_dodge(), fill = "palegreen3")+
  ylim(0,4)+
  labs(y = "Mean Values of Perceived Benefit", x = "Energy Technologies")+
  ggtitle("Your Title Here")+
  coord_flip()

```

```{r, eval=TRUE, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

# summary statistics for risks from different energy sources

bensumIND <- codedmodule2 %>%
  summarize_at(vars(starts_with(c("Ben_IND"))), list(~mean(., na.rm = TRUE), ~median(., na.rm = TRUE), ~sd(.,na.rm = TRUE),~var(.,na.rm = TRUE), ~n())) %>%
  #add row index so later spreading indexed correctly
  add_rownames()%>%
  #melt to long format
  gather(technology, value, -rowname) %>%
  # separate risky from variable suffix
  separate(technology, c("Ben", "var"), extra = "merge", fill = "left") %>%
  #separate mean from variable prefix
  separate(var, c("technology", "summary")) %>%
  # spread summary values back to wide form
  spread(summary,value) %>%
  #clean up
  select(-rowname, -Ben) %>%
  arrange(mean) %>%
  setNames(paste0('perceivedbenefit.', names(.))) %>%
  dplyr::rename(technology= perceivedbenefit.technology)

bensumIND

```

```{r, echo=FALSE}

knitr::opts_chunk$set(warning = FALSE, message = FALSE)

bensumIND %>%
  select(technology, perceivedbenefit.mean) %>%
  mutate(technology = fct_reorder(technology, perceivedbenefit.mean, .desc = FALSE))%>%
  ggplot(aes(x=technology, y= perceivedbenefit.mean))+
  geom_bar(stat="identity", position=position_dodge(), fill = "palegreen3")+
  ylim(0,4)+
  labs(y = "Mean Values of Perceived Benefit", x = "Energy Technologies")+
  ggtitle("Your Title Here")+
  coord_flip()

```

```{r, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

center_barplot <- full_join( bensumcentral,risksumcentral, by = NULL, copy = FALSE, keep = FALSE) %>%
  select(technology, perceivedbenefit.mean, perceivedrisk.mean) %>%
  mutate(technology = fct_reorder(technology, perceivedbenefit.mean, .desc = FALSE))%>%
   gather(key = "level", value = "mean", perceivedbenefit.mean, perceivedrisk.mean)

center_barplot$level <- as.factor(center_barplot$level)                                          
class(center_barplot$level)
levels(center_barplot$level)

 center_barplot%>% 
  ggplot(aes(x=technology, y= mean, fill =  center_barplot$level))+
  geom_bar(stat="identity", position=position_dodge())+
  ylim(0,4)+
  scale_fill_manual("legend", values = c("perceivedbenefit.mean" = "palegreen3","perceivedrisk.mean" = "firebrick2"))+
  coord_flip()

```

```{r, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

full_join(bensumIND,risksumIND, by = NULL, copy = FALSE, keep = FALSE) %>%
  select(technology, perceivedbenefit.mean, perceivedrisk.mean) %>%
  mutate(technology = fct_reorder(technology, perceivedbenefit.mean, .desc = FALSE))%>%
  gather(key = "level", value = "mean", perceivedbenefit.mean, perceivedrisk.mean) %>%
  ggplot(aes(x=technology, y= mean, fill = level ))+
  geom_bar(stat="identity", position=position_dodge())+
  ylim(0,4)+
  scale_fill_manual("legend", values = c("perceivedbenefit.mean" = "palegreen3","perceivedrisk.mean" = "firebrick2"))+
  coord_flip()

```

```{r, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

#PCA on Kahan scale

pcaresultsK <- codedmodule2 %>%
  select(starts_with(c("K_")))%>%
  na.omit()%>%
  prcomp(center = TRUE, scale. = TRUE)

rotations = pcaresultsK["rotation"]

#calculate total variance explained by each principal component
var_explainedK <- pcaresultsK$sdev^2 / sum(pcaresultsK$sdev^2)

rotationsK <- do.call(rbind.data.frame, rotations)

rotationsK

```

```{r, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

KPC_var <-data.frame(var_explainedK, row.names=names(rotationsK)) %>%
          tibble::rownames_to_column("PC")

KPC_var

```

```{r, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

qplot(c(1:12), var_explainedK) + 
  geom_line() + 
  xlab("Principal Component") + 
  ylab("Variance Explained") +
  ggtitle("Scree Plot") +
  ylim(0, 0.30)+
  xlim(1,7)

```

```{r, echo= FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
ecopolscalescores <- codedmodule2 %>%
  select(starts_with(c("Risky","SYSTEMDEMO","SYSTEMRELIGION","SYSTEMTECHNO","SYSTEMTOTAL","WEALTHLIM","MECHANISATION","DECISIONDECEN","DECISIONCEN","INDUSTRYLARGE","INDUSTRYSMALL","ECONOMYGLOBAL","ECONOMYLOCAL", "ENVOVERDEV","DEVOVERENV","OWNERPVT","OWNERNOREG","OWNERPUB", "OWNERREG", "Gender","Caste","Religion", "State")))%>%
  rowwise()%>%
  na.omit()%>%
  dplyr::mutate(Pro_decentralisation = mean(c_across("DECISIONDECEN":"DECISIONCEN")))%>%
   dplyr::mutate(Pro_largescale = mean(c_across("INDUSTRYLARGE":"INDUSTRYSMALL")))%>%
   dplyr::mutate(Pro_global = mean(c_across("ECONOMYGLOBAL":"ECONOMYLOCAL")))%>%
   dplyr::mutate(Pro_ecogrowth = mean(c_across("ENVOVERDEV":"DEVOVERENV")))%>%
   dplyr::mutate(Pro_private = mean(c_across("OWNERPVT":"OWNERREG")))%>%
  dplyr::select(!c("DECISIONDECEN","DECISIONCEN", "INDUSTRYLARGE","INDUSTRYSMALL","ECONOMYGLOBAL", "ECONOMYLOCAL","ENVOVERDEV",     "DEVOVERENV","OWNERPVT","OWNERNOREG","OWNERPUB", "OWNERREG"))

ecopolscalescores
```

```{r, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
# PCA test on qualities of the perceiver

ecopolscalepca <- ecopolscalescores %>%
  select(!starts_with(c("Risky","Gender","Caste","Religion","State")))%>%
  prcomp(center = TRUE,scale. = TRUE)

var_explained_ecopol = ecopolscalepca$sdev^2 / sum(ecopolscalepca$sdev^2)

summary(ecopolscalepca)

```
```{r, echo=FALSE}

rotationsecopol = ecopolscalepca["rotation"]
rotationsDFecopol <- do.call(rbind.data.frame, rotationsecopol)
rotationsDFecopol
```

```{r, echo=FALSE}

qplot(c(1:11), var_explained_ecopol) + 
  geom_line() + 
  xlab("Principal Component") + 
  ylab("Variance Explained") +
  ggtitle("Scree Plot") +
  ylim(0.05, 0.20)+
  xlim(1,7)

```

```{r}

knitr::opts_chunk$set(warning = FALSE, message = FALSE)

#PCA on Kahan scale

Solar <- codedmodule2 %>%
  select("DISPLACESOLAR":"RELYSOLAR")

#%>%
  separate(Var, c("prefix", "Var", "technology"), sep = c(0,-5), remove = T)%>%
  select(!"prefix")

ROOFS<- codedmodule2 %>%
  select("DISPLACEROOFS":"RELYROOFS")

# %>%
  pivot_longer(cols ="DISPLACEROOFS":"RELYROOFS", names_to = "Var", values_to = "score") %>%
  separate(Var, c("prefix", "Var", "technology"), sep = c(0,-5), remove = T)%>%
  select(!"prefix")

DIESEL<- codedmodule2 %>%
  select("DISPLACEDIESEL":"RELYDIESEL")%>%
  pivot_longer(cols ="DISPLACEDIESEL":"RELYDIESEL", names_to = "Var", values_to = "score") %>%
  separate(Var, c("prefix", "Var", "technology"), sep = c(0,-6), remove = T)%>%
  select(!"prefix")

FIREWOOD<- codedmodule2 %>%
  select("DISPLACEFIRE":"RELYFIRE")%>%
  pivot_longer(cols ="DISPLACEFIRE":"RELYFIRE", names_to = "Var", values_to = "score") %>%
  separate(Var, c("prefix", "Var", "technology"), sep = c(0,-8), remove = T)%>%
  select(!"prefix")

COAL<- codedmodule2 %>%
  select("DISPLACECOAL":"RELYCOAL")%>%
  pivot_longer(cols ="DISPLACECOAL":"RELYCOAL", names_to = "Var", values_to = "score") %>%
  separate(Var, c("prefix", "Var", "technology"), sep = c(0,-4), remove = T)%>%
  select(!"prefix")

NUCLEAR<- codedmodule2 %>%
  select("DISPLACENUCLEAR":"RELYNUCLEAR")%>%
  pivot_longer(cols ="DISPLACENUCLEAR":"RELYNUCLEAR", names_to = "Var", values_to = "score") %>%
  separate(Var, c("prefix", "Var", "technology"), sep = c(0,-7), remove = T)%>%
  select(!"prefix")

Solar.pca <- Solar %>%
  na.omit()%>%
  prcomp(center = TRUE, scale. = TRUE)

rotations = Solar.pca["rotation"]

#calculate total variance explained by each principal component
var_explainedSolar <- Solar.pca$sdev^2 / sum(Solar.pca$sdev^2)

rotationsSolar <- do.call(rbind.data.frame, rotations)

rotationsSolar

```

```{r, echo=FALSE}

qplot(c(1:10), var_explainedSolar) + 
  geom_line() + 
  xlab("Principal Component") + 
  ylab("Variance Explained") +
  ggtitle("Scree Plot") +
  ylim(0.01, 0.40)+
  xlim(1,7)

```