### 2022-03-21 module2/PCAs/module2PCAs.Rmd
moved files into folders

---
title: "PCAs"
output: 
  output:
  html_document:
    fig_width: 9
    fig_height: 6
    fig_caption: yes
    keep_md: yes
    theme: paper
    df_print: paged

---

```{r load packages, warning = FALSE, message = FALSE, echo=FALSE}
  knitr::opts_chunk$set(warning = FALSE, message = FALSE)

suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(readxl))

suppressPackageStartupMessages(library(tsibble))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(DT))

suppressPackageStartupMessages(library(psych)) # psychometrics

```

```{r, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
moduletwo <- read_excel( "~/Desktop/thesis_surveydata/political_factors_complete.xlsx")
save(moduletwo, file = "moduletwo.rda")
module2 <- data.frame(moduletwo)
module2[module2 == "Rather not say/ Don't know"] <- "Rather not say/Don't know"
#module2 %>% filter(!row_number() %in% c(1,2))

# loaded data set for module 2 eco -pol variables and removes the rows with question number and question text
```

## Coded Dataset

```{r, eval =TRUE, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
#This chunk contains the coding schemes of various scales used in survey one: eco-political factors, kahan scale and acceptance scale
codedmodule2 <- module2 %>%

#remove row 1
  filter(!row_number() %in% c(1,2)) %>%

# replace risky likert scale with numbers
  mutate_at(vars(starts_with("Risky")), funs(case_when(. =="Not at all risky" ~ 1, 
                                                       . =="Slightly risky" ~ 2, 
                                                       . =="Moderately risky" ~ 3, 
                                                       . =="Very risky" ~ 4, 
                                                       . =="Extremely risky" ~ 5))) %>%

# replace beneficial likert scale with numbers  
  mutate_at(vars(starts_with("Ben")), funs(case_when(. =="Not at all beneficial" ~ 1,
                                                     . =="Slightly beneficial" ~ 2,
                                                     . =="Moderately beneficial" ~ 3,
                                                     . =="Very beneficial" ~ 4,
                                                     . =="Extremely beneficial" ~ 5 ))) %>%

# replace nuclear acceptance likert scale with numbers
  mutate_at(vars(N_accept,N_reluctantlyaccept,N_reject), funs(case_when(. == "Strongly disagree" ~ 1, 
                                                                        . == "Somewhat disagree" ~ 2,
                                                                        . == "Neither agree nor disagree" ~3,
                                                                        . == "Somewhat agree" ~ 4,
                                                                        . == "Strongly agree" ~ 5))) %>%

# code likert scale for variables for Kahan scale into numbers
  mutate_at(vars(starts_with (c("K_I","K_H","DISPLACE", "POLLUTE", "HEALTH", "JOBS", "BEAUTY", "PRIDE", "NPRIDE", "DEV", "PROSPER", "RELY"))), funs(case_when(. == "Strongly disagree" ~ 1, 
               . == "Somewhat disagree" ~ 2,
               . == "Neither agree nor disagree" ~3,
               . == "Somewhat agree" ~ 4,
               . == "Strongly agree" ~ 5))) %>%

# reverse code for likert scale for variables for Kahan scale into numbers
mutate_at(vars(starts_with (c("K_S","K_E"))), funs(case_when(. == "Strongly disagree" ~ 1, 
                                                               . == "Somewhat disagree" ~ 2,
                                                               . == "Neither agree nor disagree" ~3,
                                                               . == "Somewhat agree" ~ 4,
                                                               . == "Strongly agree" ~ 5))) %>%

# code eco-pol scale variables into numbers
  mutate_at(vars(SYSTEMDEMO,SYSTEMRELIGION,SYSTEMTECHNO,SYSTEMTOTAL,WEALTHLIM,MECHANISATION,DECISIONDECEN,INDUSTRYLARGE,ECONOMYGLOBAL,OWNERPVT,OWNERNOREG), funs(case_when(. == "Strongly disagree" ~ 1, 
                        . == "Somewhat disagree" ~ 2,
                        . == "Neither agree nor disagree" ~3,
                        . == "Somewhat agree" ~ 4,
                        . == "Strongly agree" ~ 5))) %>%

# reverse code eco-pol scale variables into numbers 
 mutate_at(vars(DECISIONCEN, INDUSTRYSMALL, ECONOMYLOCAL, ENVOVERDEV,OWNERPUB, OWNERREG), funs(case_when(. == "Strongly disagree" ~ 1, 
                                                                                                         . == "Somewhat disagree" ~ 2,
                                                                                                         . == "Neither agree nor disagree" ~3,
                                                                                                         . == "Somewhat agree" ~ 4,
                                                                                                         . == "Strongly agree" ~ 5)))

codedmodule2
```

# **Kahan et al (2007) scale**

- **K_IINTRFER**	The government interferes far too much in our everyday lives.
- **K_IPRIVACY**	The government should stop telling people how to live their lives.
- **K_SHARM**	    Sometimes the government needs to make laws that keep people from hurting themselves.
- **K_IPROTECT**	It’s not the government’s business to try to protect people from themselves.
- **K_SLIMCHOI**	The government should put limits on the choices individuals can make so they don’t get in the way of what’s good for society.
- **K_SPROTECT**	The government should do more to advance society’s goals, even if that means limiting the freedom and choices of individuals.

- **K_HEQUAL**	  We have gone too far in pushing equal rights in this country.
- **K_HREVDIS1**	Nowadays it seems like there is just as much discrimination against upper castes as there is against Dalits.
- **K_EDISCRIM**	Discrimination against minorities is still a very serious problem in our society.
- **K_ERADEQ1**	  We need to dramatically reduce inequalities between the rich and the poor.
- **K_EWEALTH**	  Our society would be better off if the distribution of wealth was more equal.
- **K_ERADEQ2**	  We need to dramatically reduce inequalities between men and women.

## Cronbach's alpha test on the Kahan et al (2007) scale

```{r, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

library(ltm)

#cronbach alpha

KahanI <- codedmodule2 %>%
  dplyr::select(starts_with(c("K_S")))%>%
  na.omit()

#K_S items produce a strong alpha but not K_I

psych::alpha(KahanI)

KahanE <- codedmodule2 %>%
  dplyr::select(starts_with(c("K_H", "K_E")))%>%
  na.omit()
psych::alpha(KahanE)

```

## Factor Analysis on Kahan et al (2007) scale

```{r, echo=FALSE}

#factor analysis on Kahan scale
Kahan <- codedmodule2 %>%
  dplyr::select(starts_with(c("K_")))%>%
  na.omit()

pcaresultsK <- fa(Kahan, nfactors=4, scores=TRUE, rotate="varimax", fm="pa")
print.psych(pcaresultsK, cut = 0.3, sort = TRUE)
fa.diagram(pcaresultsK)

```

# **The economic -political values of the perceiver scale**

- **DECISIONDECEN**	 Local politicians shouldn't have to ask permission from the central government to implement policies
- **DECISIONCEN**	   Laws and policies would be implemented more smoothly if more power lay with the central government.
- **SYSTEMTOTAL**	   It is good to have a strong leader who does not have to bother with elections.
- **SYSTEMTECHNO**	 Experts, not the government, should make decisions according to what they think is best for the country.
- **SYSTEMDEMO**	   It is very important to have a democratic political system because it ensures that no individual leader has too much power.
- **SYSTEMRELIGION** There should be a system governed by religious law.
- **INDUSTRYSMALL**	 Large corporations are destroying the local industries in India and benefiting only a handful of people.
- **INDUSTRYLARGE**	 Large scale industries are required for the development of the country that will benefit everyone
- **ECONOMYLOCAL**	 India would be better off if foreign companies didn't come to here
- **ECONOMYGLOBAL**	 Foreign companies have led to a range of benefits for the Indian people and society
- **DEVOVERENV**	   Economic growth and creating jobs should be prioritized over environmental protection
- **ENVOVERDEV**	   Polluting industries that spoil the environment should be shut down even if it costs people their jobs
- **OWNERPVT**	     All businesses and industries should be owned privately
- **OWNERPUB**	     The government should own most large businesses and industries
- **OWNERREG**	     Regardless of ownership, the government should pass strong regulations and implement them
- **OWNERNOREG**	   There is too much red-tape and the government should not interfere with businesses and industries
- **WEALTHLIM**	     A limit should be put to how much wealth a person can amass
- **MECHANISATION**	 Rapid mechanization of work is taking away jobs from workers in this country

## Cronbach's alpha test on the economic -political values of the perceiver
```{r, echo= FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
ecopolscalescores <- codedmodule2 %>%
 dplyr::select(starts_with(c("SYSTEMDEMO","SYSTEMRELIGION","SYSTEMTECHNO","SYSTEMTOTAL","WEALTHLIM","MECHANISATION","DECISIONDECEN","DECISIONCEN","INDUSTRYLARGE","INDUSTRYSMALL","ECONOMYGLOBAL","ECONOMYLOCAL", "ENVOVERDEV","DEVOVERENV","OWNERPVT","OWNERNOREG","OWNERPUB", "OWNERREG")))%>%
  rowwise()%>%
  na.omit()%>%
  dplyr::mutate(Pro_decentralisation = mean(c_across("DECISIONDECEN":"DECISIONCEN")))%>%
   dplyr::mutate(Pro_largescale = mean(c_across("INDUSTRYLARGE":"INDUSTRYSMALL")))%>%
   dplyr::mutate(Pro_global = mean(c_across("ECONOMYGLOBAL":"ECONOMYLOCAL")))%>%
   dplyr::mutate(Pro_ecogrowth = mean(c_across("ENVOVERDEV":"DEVOVERENV")))%>%
   dplyr::mutate(Pro_private = mean(c_across("OWNERPVT":"OWNERREG")))%>%
  dplyr::select(!c("DECISIONDECEN","DECISIONCEN", "INDUSTRYLARGE","INDUSTRYSMALL","ECONOMYGLOBAL", "ECONOMYLOCAL","ENVOVERDEV",     "DEVOVERENV","OWNERPVT","OWNERNOREG","OWNERPUB", "OWNERREG"))

#cronbach alpha

psych::alpha(ecopolscalescores, check.keys = TRUE)

```

```{r, echo=FALSE}

#factor analysis with 2, 4 and 6 factor spaces

ecopolscale.fa <- fa(ecopolscalescores, nfactors=2, scores=TRUE, rotate="varimax", fm="pa")
print.psych(ecopolscale.fa, cut = 0.3, sort = TRUE)
fa.diagram(ecopolscale.fa)

```

\newpage

# **Economic and Political characteristics of the energy technology scale -  Solar Energy and Nuclear Energy**

- **DISPLACESOLAR**	 Solar energy is leading to displacement of people from their land
- **POLLUTESOLAR**	 Solar energy increases pollution of air/water/land
- **HEALTHSOLAR**	   Solar energy poses a great risk to the health of people living around it
- **JOBSSOLAR**	     Solar energy will bring jobs to the local community
- **BEAUTYSOLAR**	   Solar energy spoils the natural beauty of the landscape
- **PRIDESOLAR**	   I would be proud if my community used solar energy
- **NPRIDESOLAR**	   Solar energy is a mark of pride for our nation
- **DEVSOLAR**	     Solar energy pushes forward the country's development
- **PROSPERSOLAR**	 Solar energy brings economic prosperity to the surrounding regions
- **RELYSOLAR**	     I don't like the idea that I have to rely on the government for electricity from (large scale) solar energy

## Cronbach's alpha test on eco-pol characteristics of the technology -  Solar Energy

```{r, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

#alpha test on Eco-pol characteristics of the technology

Solar <- codedmodule2 %>%
dplyr::select("DISPLACESOLAR":"RELYSOLAR") %>%
dplyr::rename("Solar energy is leading to displacement of people from their land" = "DISPLACESOLAR") %>%
dplyr::rename("Solar energy increases pollution of air/water/land" = "POLLUTESOLAR")%>%
dplyr::rename("Solar energy poses a great risk to the health of people living around it" = "HEALTHSOLAR")%>%
dplyr::rename("Solar energy will bring jobs to the local community" = "JOBSSOLAR")%>%
dplyr::rename("Solar energy spoils the natural beauty of the landscape" = "BEAUTYSOLAR")%>%
dplyr::rename("I would be proud if my community used solar energy" = "PRIDESOLAR")%>%
dplyr::rename("Solar energy is a mark of pride for our nation" = "NPRIDESOLAR") %>%
dplyr::rename("Solar energy pushes forward the country's development" = "DEVSOLAR")%>%
dplyr::rename("Solar energy brings economic prosperity to the surrounding regions" = "PROSPERSOLAR")%>%
dplyr::rename("I don't like the idea that I have to rely on the government for electricity from (large scale solar energy)" = "RELYSOLAR")

psych::alpha(Solar, check.keys = TRUE)

```

```{r, echo=FALSE}

#factor analysis with 2, 4 and 6 factor spaces

solar.fa <- fa(Solar, nfactors=2, scores=TRUE, rotate="varimax", fm="pa")
print.psych(solar.fa, cut = 0.4, sort = TRUE)
fa.diagram(solar.fa)

```
```{r}
ROOFS<- codedmodule2 %>%
 dplyr::select("DISPLACEROOFS":"RELYROOFS")

psych::alpha(ROOFS, check.keys = TRUE)

```

```{r}
#factor analysis with 2, 4 and 6 factor spaces

roofs.fa <- fa(ROOFS, nfactors=4, scores=TRUE, rotate="varimax", fm="pa")
print.psych(roofs.fa, cut = 0.4, sort = TRUE)
fa.diagram(roofs.fa)

unclass(roofs.fa)
```