### 2022-03-21 module2/module2extra.Rmd

---
title: "module2extra"
output:
  html_document:
    fig_width: 7
    fig_height: 6
    fig_caption: yes
    keep_md: yes
    theme: paper
    df_print: paged
---

```{r}

likertmodule2 <- module2%>%
  filter(!row_number() %in% c(1,2)) %>% 
  select(starts_with("Ben"))%>%
  gather(Technology, Perceived_Benefit, Ben_Hydro:Ben_INDKerosene)%>%
  separate(Technology, c("Ben", "Technology"), extra = "merge", fill = "left")%>%
  select(-Ben) %>% 
  group_by(Technology, Perceived_Benefit) %>%
  dplyr::summarise(n=n())%>%
  mutate(Perceived_Benefit = factor(Perceived_Benefit,  levels = c("Rather not say/Don't know","Not at all beneficial","Slightly beneficial","Moderately beneficial","Very beneficial", "Extremely beneficial")))%>%
  pivot_wider(names_from = Perceived_Benefit, values_from = n)%>%
  rename("NA1"="Rather not say/Don't know")%>%
  select("Technology","Not at all beneficial","Slightly beneficial","Moderately beneficial","Very beneficial", "Extremely beneficial", "NA1","NA")%>%
  unite("Rather not say/Don't know", "NA1":"NA", na.rm = TRUE, remove = FALSE)%>%
  select(!c("NA1","NA"))

```

```{r}
#summary statistics for nuclear energy accept, reject and reluctantly accept scale
codedmodule2 %>%
  summarize_at(vars(N_accept,N_reject,N_reluctantlyaccept), list(~mean(., na.rm = TRUE),~median(., na.rm = TRUE), ~sd(.,na.rm = TRUE),~var(.,na.rm = TRUE),~n())) %>%
  add_rownames() %>%
  gather(acceptance, value, -rowname) %>%
  separate(acceptance, c("N_accept", "var") ,  extra = "merge", fill = "left") %>%
  separate(var, c("Acceptance", "summary")) %>%
  spread(summary, value) %>%
  select(-rowname, -N_accept)%>%
  arrange(mean)

```

```{r}

# chi square test for risk perception from nuclear energy and gender, race

# first combine gender and caste variables

chitest <- codedmodule2 %>%
  select(Risky_Nuclear,Risky_Hydro, Risky_Hydro, Risky_Solar,Urban_Rural, Gender, Caste) %>%
  na.omit() %>%
  filter(!Caste=="Rather not say/Don't know")%>%
  unite(gendercaste, Gender, Caste, sep = "_", remove = FALSE)

chitest

```

```{r, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

KPC_var <-data.frame(var_explainedK, row.names=names(rotationsK)) %>%
          tibble::rownames_to_column("PC")

KPC_var

```

```{r}
#%>%
  #separate(Var, c("prefix", "Var", "technology"), sep = c(0,-5), remove = T)%>%
 # select(!"prefix")
# %>%
  #pivot_longer(cols ="DISPLACEROOFS":"RELYROOFS", names_to = "Var", values_to = "score") %>%
  #separate(Var, c("prefix", "Var", "technology"), sep = c(0,-5), remove = T)%>%
  #select(!"prefix")

#DIESEL<- codedmodule2 %>%
  #select("DISPLACEDIESEL":"RELYDIESEL")%>%
  #pivot_longer(cols ="DISPLACEDIESEL":"RELYDIESEL", names_to = "Var", values_to = "score") %>%
  #separate(Var, c("prefix", "Var", "technology"), sep = c(0,-6), remove = T)%>%
  #select(!"prefix")
```

```{r}
rotations = pcaresultsK["rotation"]

#calculate total variance explained by each principal component
var_explainedK <- pcaresultsK$sdev^2 / sum(pcaresultsK$sdev^2)

rotationsK <- do.call(rbind.data.frame, rotations)%>%
              tibble::rownames_to_column("Variables")

rotationsK

pcatableK  <- summary(pcaresultsK)

UnclassedK <- unclass(pcatableK)

ImportanceK <- UnclassedK["importance"]

varexplK <- do.call(rbind.data.frame, ImportanceK)

KPCAtable <- data.frame(t(varexplK), row.names=names(varexplK))

as_tibble(KPCAtable) %>%
 add_column("Principal Components" = c("PC1", "PC2", "PC3", "PC4", "PC5", "PC6", "PC7", "PC8", "PC9", "PC10", "PC11", "PC12"))%>%
 dplyr::select("Principal Components",  importance.Proportion.of.Variance, importance.Cumulative.Proportion ) %>%  
 dplyr::rename(Proportion_of_Variance = importance.Proportion.of.Variance, Cumulative_Proportion = importance.Cumulative.Proportion)
```

\newpage

```{r, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

qplot(c(1:12), var_explainedK) + 
  geom_line() + 
  xlab("Principal Component") + 
  ylab("Variance Explained") +
  ggtitle("Scree Plot") +
  ylim(0, 0.30)+
  xlim(1,7)

```

\newpage

# **The economic -political values of the perceiver scale**

- **DECISIONDECEN**	 Local politicians shouldn't have to ask permission from the central government to implement policies
- **DECISIONCEN**	   Laws and policies would be implemented more smoothly if more power lay with the central government.
- **SYSTEMTOTAL**	   It is good to have a strong leader who does not have to bother with elections.
- **SYSTEMTECHNO**	 Experts, not the government, should make decisions according to what they think is best for the country.
- **SYSTEMDEMO**	   It is very important to have a democratic political system because it ensures that no individual leader has too much power.
- **SYSTEMRELIGION** There should be a system governed by religious law.
- **INDUSTRYSMALL**	 Large corporations are destroying the local industries in India and benefiting only a handful of people.
- **INDUSTRYLARGE**	 Large scale industries are required for the development of the country that will benefit everyone
- **ECONOMYLOCAL**	 India would be better off if foreign companies didn't come to here
- **ECONOMYGLOBAL**	 Foreign companies have led to a range of benefits for the Indian people and society
- **DEVOVERENV**	   Economic growth and creating jobs should be prioritized over environmental protection
- **ENVOVERDEV**	   Polluting industries that spoil the environment should be shut down even if it costs people their jobs
- **OWNERPVT**	     All businesses and industries should be owned privately
- **OWNERPUB**	     The government should own most large businesses and industries
- **OWNERREG**	     Regardless of ownership, the government should pass strong regulations and implement them
- **OWNERNOREG**	   There is too much red-tape and the government should not interfere with businesses and industries
- **WEALTHLIM**	     A limit should be put to how much wealth a person can amass
- **MECHANISATION**	 Rapid mechanization of work is taking away jobs from workers in this country

## Cronbach's alpha test on the economic -political values of the perceiver
```{r, echo= FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
ecopolscalescores <- codedmodule2 %>%
 dplyr::select(starts_with(c("Risky","SYSTEMDEMO","SYSTEMRELIGION","SYSTEMTECHNO","SYSTEMTOTAL","WEALTHLIM","MECHANISATION","DECISIONDECEN","DECISIONCEN","INDUSTRYLARGE","INDUSTRYSMALL","ECONOMYGLOBAL","ECONOMYLOCAL", "ENVOVERDEV","DEVOVERENV","OWNERPVT","OWNERNOREG","OWNERPUB", "OWNERREG", "Gender","Caste","Religion", "State")))%>%
  rowwise()%>%
  na.omit()%>%
  dplyr::mutate(Pro_decentralisation = mean(c_across("DECISIONDECEN":"DECISIONCEN")))%>%
   dplyr::mutate(Pro_largescale = mean(c_across("INDUSTRYLARGE":"INDUSTRYSMALL")))%>%
   dplyr::mutate(Pro_global = mean(c_across("ECONOMYGLOBAL":"ECONOMYLOCAL")))%>%
   dplyr::mutate(Pro_ecogrowth = mean(c_across("ENVOVERDEV":"DEVOVERENV")))%>%
   dplyr::mutate(Pro_private = mean(c_across("OWNERPVT":"OWNERREG")))%>%
  dplyr::select(!c("DECISIONDECEN","DECISIONCEN", "INDUSTRYLARGE","INDUSTRYSMALL","ECONOMYGLOBAL", "ECONOMYLOCAL","ENVOVERDEV",     "DEVOVERENV","OWNERPVT","OWNERNOREG","OWNERPUB", "OWNERREG"))

#cronbach alpha

cronbach.alpha(ecopolscalescores, na.rm = TRUE, CI=TRUE, standardized=TRUE)

```

## PCA on the economic -political values of the perceiver

```{r, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
# PCA test on qualities of the perceiver

ecopolscalepca <- ecopolscalescores %>%
  dplyr::select(!starts_with(c("Risky","Gender","Caste","Religion","State")))%>%
  prcomp(center = TRUE,scale. = TRUE)

var_explained_ecopol = ecopolscalepca$sdev^2 / sum(ecopolscalepca$sdev^2)

pcatableecopol  <- summary(ecopolscalepca)

Unclassedecopol <- unclass(pcatableecopol)

Importanceecopol <- Unclassedecopol["importance"]

varexplecopol <- do.call(rbind.data.frame, Importanceecopol)

ecopolPCAtable <- data.frame(t(varexplecopol), row.names=names(varexplecopol))

as_tibble(ecopolPCAtable) %>%
 add_column("Principal Components" = c("PC1", "PC2", "PC3", "PC4", "PC5", "PC6", "PC7", "PC8", "PC9", "PC10", "PC11"))%>%
 dplyr::select("Principal Components",  importance.Proportion.of.Variance, importance.Cumulative.Proportion ) %>%  
 dplyr::rename(Proportion_of_Variance = importance.Proportion.of.Variance, Cumulative_Proportion = importance.Cumulative.Proportion)

```

## Variables in the PCs

```{r, echo=FALSE}

rotationsecopol = ecopolscalepca["rotation"]
rotationsDFecopol <- do.call(rbind.data.frame, rotationsecopol)
rotationsDFecopol
```

```{r, echo=FALSE}

qplot(c(1:11), var_explained_ecopol) + 
  geom_line() + 
  xlab("Principal Component") + 
  ylab("Variance Explained") +
  ggtitle("Scree Plot") +
  ylim(0.05, 0.20)+
  xlim(1,7)

```

\newpage

# **eco - pol characteristics of the technology scale-  Solar Energy and Nuclear Energy**

- **DISPLACESOLAR**	 Solar energy is leading to displacement of people from their land
- **POLLUTESOLAR**	 Solar energy increases pollution of air/water/land
- **HEALTHSOLAR**	   Solar energy poses a great risk to the health of people living around it
- **JOBSSOLAR**	     Solar energy will bring jobs to the local community
- **BEAUTYSOLAR**	   Solar energy spoils the natural beauty of the landscape
- **PRIDESOLAR**	   I would be proud if my community used solar energy
- **NPRIDESOLAR**	   Solar energy is a mark of pride for our nation
- **DEVSOLAR**	     Solar energy pushes forward the country's development
- **PROSPERSOLAR**	 Solar energy brings economic prosperity to the surrounding regions
- **RELYSOLAR**	     I don't like the idea that I have to rely on the government for electricity from (large scale) solar energy

## Cronbach's alpha test on eco-pol characteristics of the technology -  Solar Energy

```{r, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

#PCA on eco-pol characterisitics of the technology

Solar <- codedmodule2 %>%
 dplyr::select("DISPLACESOLAR":"RELYSOLAR")

cronbach.alpha(Solar, na.rm = TRUE, CI=TRUE, standardized=TRUE)

ROOFS<- codedmodule2 %>%
 dplyr::select("DISPLACEROOFS":"RELYROOFS")

```

## PCA on eco-pol characteristics of the technology -  Solar Energy

```{r, echo = FALSE}

Solar.pca <- Solar %>%
  na.omit()%>%
  prcomp(center = TRUE, scale. = TRUE)

pcatablesolar  <- summary(Solar.pca)

Unclassedsolar <- unclass(pcatablesolar)

Importancesolar <- Unclassedsolar["importance"]

varexplsolar <- do.call(rbind.data.frame, Importancesolar)

solarPCAtable <- data.frame(t(varexplsolar), row.names=names(varexplsolar))

as_tibble(solarPCAtable) %>%
 add_column("Principal Components" = c("PC1", "PC2", "PC3", "PC4", "PC5", "PC6", "PC7", "PC8", "PC9", "PC10"))%>%
 dplyr::select("Principal Components",  importance.Proportion.of.Variance, importance.Cumulative.Proportion ) %>%  
 dplyr::rename(Proportion_of_Variance = importance.Proportion.of.Variance, Cumulative_Proportion = importance.Cumulative.Proportion)
```

```{r, echo=FALSE}

rotations = Solar.pca["rotation"]

#calculate total variance explained by each principal component
var_explainedSolar <- Solar.pca$sdev^2 / sum(Solar.pca$sdev^2)

rotationsSolar <- do.call(rbind.data.frame, rotations)

rotationsSolar

```

```{r, echo=FALSE}

qplot(c(1:10), var_explainedSolar) + 
  geom_line() + 
  xlab("Principal Component") + 
  ylab("Variance Explained") +
  ggtitle("Scree Plot") +
  ylim(0.01, 0.40)+
  xlim(1,7)

```

## PCA on eco-pol characteristics of the technology -  Nuclear Energy

```{r, echo = FALSE}

Nuclear<- codedmodule2 %>%
 dplyr::select("DISPLACENUCLEAR":"RELYNUCLEAR")

cronbach.alpha(Nuclear, na.rm = TRUE, CI=TRUE, standardized=TRUE)

```

```{r, echo = FALSE}

Nuclear.pca <- Nuclear %>%
  na.omit()%>%
  prcomp(center = TRUE, scale. = TRUE)

pcatablenuclear  <- summary(Nuclear.pca)

Unclassednuclear <- unclass(pcatablenuclear)

Importancenuclear <- Unclassednuclear["importance"]

varexplnuclear <- do.call(rbind.data.frame, Importancenuclear)

nuclearPCAtable <- data.frame(t(varexplnuclear), row.names=names(varexplnuclear))

as_tibble(nuclearPCAtable) %>%
 add_column("Principal Components" = c("PC1", "PC2", "PC3", "PC4", "PC5", "PC6", "PC7", "PC8", "PC9", "PC10"))%>%
 dplyr::select("Principal Components",  importance.Proportion.of.Variance, importance.Cumulative.Proportion ) %>%  
 dplyr::rename(Proportion_of_Variance = importance.Proportion.of.Variance, Cumulative_Proportion = importance.Cumulative.Proportion)

```

```{r, echo=FALSE}

rotations = Nuclear.pca["rotation"]

#calculate total variance explained by each principal component
var_explainedNuclear <- Nuclear.pca$sdev^2 / sum(Nuclear.pca$sdev^2)

rotationsNuclear <- do.call(rbind.data.frame, rotations)

rotationsNuclear

```

```{r, echo=FALSE}

qplot(c(1:10), var_explainedNuclear) + 
  geom_line() + 
  xlab("Principal Component") + 
  ylab("Variance Explained") +
  ggtitle("Scree Plot") +
  ylim(0.01, 0.30)+
  xlim(1,7)

```

```{r}
#trying to work with likert packages here

likertstack <- module2%>%
  filter(!row_number() %in% c(1,2)) %>% 
  select(starts_with("RISKY"))%>%
  gather(Technology, Perceived_Risk, Risky_Hydro:Risky_INDLPG, factor_key = TRUE)%>%
  separate(Technology, c("Risky", "Technology"), extra = "merge", fill = "left")%>%
  select(-Risky) %>% 
  group_by(Technology, Perceived_Risk) %>%
  summarise(n=n()) %>%
  pivot_wider(names_from = "Perceived_Risk", values_from = "n")

likert(Measure ~ ., data=likertstack, ylab=NULL,
       ReferenceZero=3, as.percent=TRUE,
       positive.order=TRUE, 
       main = list("Pereived risk from energy technologies in India",x=unit(.55, "npc")), 
       sub= list("Perceived risk",x=unit(.57, "npc")), 
       xlim=c(-40,-20,0,20,40,60,80,100),
       strip=FALSE, 
       par.strip.text=list(cex=.7))
```