### 2022-04-05 module1/moduleoneanalysis.Rmd
coded some LSI and ISI variables

---
title: "moduleoneanalysis"
output:
  html_document:
    fig_width: 7
    fig_height: 6
    fig_caption: true
    keep_md: true
    theme: paper
    df_print: paged
---

```{r load packages, warning = FALSE, message = FALSE}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(readxl))
suppressPackageStartupMessages(library(dplyr))

```

```{r, eval=TRUE}

moduleone <- read_excel(here::here("thesis_surveydata/livelihood_insecurity_complete.xlsx"))
save(moduleone, file = "moduleone.rda")
module1 <- data.frame(moduleone)
module1 %>%
  filter(!row_number() %in% c(1,2))

# loaded data set for module 2 eco -pol variables and removes the rows with question number and question text
```

```{r, eval =TRUE}
#This chunk contains the coding schemes of various scales used in survey one: eco-political factors, kahan scale and acceptance scale
codedmodule1 <- module1 %>%

#remove row 1
  filter(!row_number() %in% c(1,2)) %>%

# replace risky likert scale with numbers
  mutate_at(vars(starts_with("Risky")), funs(case_when(. =="Not at all risky" ~ 1, 
                                                       . =="Slightly risky" ~ 2, 
                                                       . =="Moderately risky" ~ 3, 
                                                       . =="Very risky" ~ 4, 
                                                       . =="Extremely risky" ~ 5))) %>%

# replace beneficial likert scale with numbers  
  mutate_at(vars(starts_with("Ben")), funs(case_when(. =="Not at all beneficial" ~ 1,
                                                     . =="Slightly beneficial" ~ 2,
                                                     . =="Moderately beneficial" ~ 3,
                                                     . =="Very beneficial" ~ 4,
                                                     . =="Extremely beneficial" ~ 5 ))) %>%

# replace nuclear acceptance likert scale with numbers
  mutate_at(vars(N_accept,N_reluctantlyaccept,N_reject), funs(case_when(. == "Strongly disagree" ~ 1, 
                                                                        . == "Somewhat disagree" ~ 2,
                                                                        . == "Neither agree nor disagree" ~3,
                                                                        . == "Somewhat agree" ~ 4,
                                                                        . == "Strongly agree" ~ 5))) %>%

  mutate_at(vars(Livelihood_Income), funs(case_when(. == "(a) own farming/fishing/animal rearing/business" ~ "LSI", 
                                                    . == "(b) employed for wages/salary" ~ "ISI",
                                                    . == "(c) student" ~"student",
                                                    . == "(d) currently unemployed" ~ "unemployed",
                                                    . == "(e) homemaker" ~ "homemaker",
                                                    . == "(f) retired" ~ "retired",
                                                    . == "(g) unable to work" ~ "unable to work" ))) %>%

# coding education
   mutate_at(vars(education), funs(case_when(. == "No schooling completed" ~ 0, 
                                                                        . == "Primary school" ~ 1,
                                                                        . == "Grade 10th" ~ 2,
                                                                        . == "Grade 12th" ~ 3,
                                                                        . == "Some college, no degree" ~ 4,
                                                                        . == "Trade/technical/vocational training" ~ 5,
                                                                        . == "Bachelor's degree" ~ 6,
                                                                        . == "Master's degree" ~ 7,
                                                                        . == "Professional degree" ~ 8,
                                                                        . == "Doctorate degree" ~ 9 ))) %>%
# coding yes and no questions
  mutate_at(vars(Iassociation, govtscheme, wateravail, polposition, polparty), funs(case_when(.== "Yes" ~ 1,
                                            .== "No" ~ 0))) %>%
# reverse coding yes and no questions
  mutate_at(vars(illness,severeillness), funs(case_when(.== "Yes" ~ 0,
                                            .== "No" ~ 1))) %>%

# coding How confident are you that you can continue your current livelihood if you want to? 
 mutate_at(vars(Lconfidence), funs(case_when(. == "Not at all confident" ~ 1, 
                                                                        . == "Not much confident" ~ 2,
                                                                        . == "Neutral" ~3,
                                                                        . == "Somewhat confident" ~ 4,
                                                                        . == "Very confident" ~ 5))) %>%

 mutate_at(vars(Lboat, houseown), funs(case_when(.== "Own it" ~ 1,
                                                 .== "Rent it" ~ 0)))%>%

# J security index coding

 mutate_at(vars(JISRkeep), funs(case_when(. == "Strongly disagree" ~ 1, 
                                                                        . == "Somewhat disagree" ~ 2,
                                                                        . == "Neither agree nor disagree" ~3,
                                                                        . == "Somewhat agree" ~ 4,
                                                                        . == "Strongly agree" ~ 5))) %>%

# reverse code JIS 
  mutate_at(vars(JISlosefuture, JISlose, JISfuture), funs(case_when(. == "Strongly disagree" ~ 5, 
                                                                        . == "Somewhat disagree" ~ 4,
                                                                        . == "Neither agree nor disagree" ~3,
                                                                        . == "Somewhat agree" ~ 2,
                                                                        . == "Strongly agree" ~ 1))) %>%

  mutate_at(vars(totalincome, otherincome), funs(case_when(. == "Less than Rs 1,00,000" ~ 1,
                                                                        . == "Rs 1,00,000 to Rs 2,50,000"  ~ 2,
                                                                        . == "Rs 2,50,001 to Rs 5,00,000" ~ 3,
                                                                        . == "Rs 5,00,001 to Rs 10,00,000" ~ 4,
                                                                        . == "Rs 10,00,001 and above"  ~ 5))) %>%

  mutate_at(vars(workknowledge, subsidyknow), funs(case_when(. == "Very poor" ~ 1, 
                                                                        . == "Poor" ~ 2,
                                                                        . == "Fair" ~3,
                                                                        . == "Good" ~ 4,
                                                                        . == "Very good" ~ 5))) %>%

mutate_at(vars(drinkingwater, cookingwater), funs(case_when(. == "Bottled water/ Filtered Water" ~ 1,                                       
                                                          .  == "Piped into dwelling" ~ 2,                                                  
                                                          .  == "Piped to neighbour" ~ 3,                                                   
                                                          .  == "Piped to yard/plot" ~ 4, 
                                                          .  == "Protected well" ~ 5, 
                                                          .  == "Public tap/standpipe" ~6,
                                                          .  == "Surface Water  (River/Dam/Lake/Pond/Stream/Canal/Irrigation Channel)" ~7, 
                                                          .  == "Tanker Truck" ~ 8,
                                                          .  == "Tubewell/borehole" ~ 9,
                                                          .  == "Unprotected Spring" ~ 10))) %>%

  mutate_at(vars(toilet), funs(case_when(. == "No facility/bush/field" ~ 1,                                       
                                                          .  == "Hanging toilet/hanging latrine" ~ 2,                                             
                                                          .  == "Bucket toilet"  ~ 3,                                                   
                                                          .  == "Composting toilet"  ~ 4, 
                                                          .  == "Pit latrine without slab/open pit" ~ 5, 
                                                          .  == "Pit latrine with slab" ~6,
                                                          .  == "Ventilated improved pit laterine"  ~7, 
                                                          .  == "Flush to somewhere else"  ~ 8,
                                                          .  == "Flush to pit latrine"  ~ 9,
                                                          .  == "Flush to septic tank" ~ 10,
                                                          .  == "Flush to piped sewer system" ~ 11))) %>%

  mutate_at(vars(house), funs(case_when(. == "Hut made of mud, bamboo, straw etc. (kutcha house)" ~ 1, 
                                        . == "Permanent single story house (pukka house)" ~ 2,
                                        . == "Apartment/Flat in a multi-story building"  ~3,
                                        . == "Multi-story house" ~ 4)))

#coded dependent variables for module 1 livelihood and income security survey

codedmodule1
```

```{r}
# combining quantitative livelihood-income security variables

# Livelihood security

x <- as.factor(codedmodule1$cookstove)

levels(x)

```