### 2022-04-09 module2/module2IVgraphs.Rmd
governance preference graphs

---
title: "module2IVgraphs"
output: 
html_document:
    fig_width: 9
    fig_height: 6
    fig_caption: yes
    keep_md: yes
    theme: paper
    df_print: paged

pdf_document:

    toc: true
    number_sections: true
    toc_depth: 2
    fig_width: 9
    fig_height: 6
    fig_caption: yes
---

```{r load packages, warning = FALSE, message = FALSE, echo=FALSE}
  knitr::opts_chunk$set(warning = FALSE, message = FALSE)

suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(readxl))
suppressPackageStartupMessages(library(hablar))
suppressPackageStartupMessages(library(tsibble))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(DT))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(ggbiplot))
suppressPackageStartupMessages(library(psych)) # psychometrics

```

```{r, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
moduletwo <- read_excel(here::here("thesis_surveydata/political_factors_complete.xlsx"))
save(moduletwo, file = "moduletwo.rda")
module2 <- data.frame(moduletwo)
module2[module2 == "Rather not say/ Don't know"] <- "Rather not say/Don't know"
module2 %>%
  filter(!row_number() %in% c(1,2))

# loaded data set for module 2 eco -pol variables and removes the rows with question number and question text
```

```{r, echo = FALSE}

#This chunk contains the coding schemes of various scales used in survey one: eco-political factors, kahan scale and acceptance scale
codedmodule2 <- module2 %>%

#remove row 1
  filter(!row_number() %in% c(1,2)) %>%

# replace risky likert scale with numbers
  mutate_at(vars(starts_with("Risky")), funs(case_when(. =="Not at all risky" ~ 1, 
                                                       . =="Slightly risky" ~ 2, 
                                                       . =="Moderately risky" ~ 3, 
                                                       . =="Very risky" ~ 4, 
                                                       . =="Extremely risky" ~ 5))) %>%

# replace beneficial likert scale with numbers  
  mutate_at(vars(starts_with("Ben")), funs(case_when(. =="Not at all beneficial" ~ 1,
                                                     . =="Slightly beneficial" ~ 2,
                                                     . =="Moderately beneficial" ~ 3,
                                                     . =="Very beneficial" ~ 4,
                                                     . =="Extremely beneficial" ~ 5 ))) %>%

# replace nuclear acceptance likert scale with numbers
  mutate_at(vars(N_accept,N_reluctantlyaccept,N_reject), funs(case_when(. == "Strongly disagree" ~ 1, 
                                                                        . == "Somewhat disagree" ~ 2,
                                                                        . == "Neither agree nor disagree" ~3,
                                                                        . == "Somewhat agree" ~ 4,
                                                                        . == "Strongly agree" ~ 5))) %>%

# code likert scale for variables for Kahan scale into numbers
  mutate_at(vars(starts_with (c("K_I","K_H","DISPLACE", "POLLUTE", "HEALTH", "JOBS", "BEAUTY", "PRIDE", "NPRIDE", "DEV", "PROSPER", "RELY"))), funs(case_when(. == "Strongly disagree" ~ 1, 
               . == "Somewhat disagree" ~ 2,
               . == "Neither agree nor disagree" ~3,
               . == "Somewhat agree" ~ 4,
               . == "Strongly agree" ~ 5))) %>%

# reverse code for likert scale for variables for Kahan scale into numbers
  mutate_at(vars(starts_with (c("K_S","K_E"))), funs(case_when(. == "Strongly disagree" ~ 5, 
                                                               . == "Somewhat disagree" ~ 4,
                                                               . == "Neither agree nor disagree" ~3,
                                                               . == "Somewhat agree" ~ 2,
                                                               . == "Strongly agree" ~ 1))) %>%

# code eco-pol scale variables into numbers
  mutate_at(vars(SYSTEMDEMO,SYSTEMRELIGION,SYSTEMTECHNO,SYSTEMTOTAL,WEALTHLIM,MECHANISATION,DECISIONDECEN,INDUSTRYLARGE,ECONOMYGLOBAL,OWNERPVT,OWNERNOREG), funs(case_when(. == "Strongly disagree" ~ 1, 
                        . == "Somewhat disagree" ~ 2,
                        . == "Neither agree nor disagree" ~3,
                        . == "Somewhat agree" ~ 4,
                        . == "Strongly agree" ~ 5))) %>%

# reverse code eco-pol scale variables into numbers 
 mutate_at(vars(DECISIONCEN, INDUSTRYSMALL, ECONOMYLOCAL, ENVOVERDEV,OWNERPUB, OWNERREG), funs(case_when(. == "Strongly disagree" ~ 5, 
                                                                                                         . == "Somewhat disagree" ~ 4,
                                                                                                         . == "Neither agree nor disagree" ~3,
                                                                                                         . == "Somewhat agree" ~ 2,
                                                                                                         . == "Strongly agree" ~ 1)))

codedmodule2

```

```{r}
demorisk <- codedmodule2%>%
  group_by(SYSTEMDEMO)%>%
  filter(SYSTEMDEMO == 5)%>%
  summarize_at(vars(starts_with(c("RISKY"))), list(~mean(., na.rm = TRUE), ~median(., na.rm = TRUE), ~sd(.,na.rm = TRUE),~var(.,na.rm = TRUE), ~n())) %>%
  #melt to long format
  gather(key = "technology", value = "summary", Risky_Hydro_mean, Risky_Solar_mean, Risky_Wind_mean, Risky_Nuclear_mean, Risky_Coal_mean, Risky_Gas_mean, Risky_Oil_mean, Risky_Hydro_sd, Risky_Solar_sd, Risky_Wind_sd, Risky_Nuclear_sd, Risky_Coal_sd, Risky_Gas_sd, Risky_Oil_sd) %>%
  select(SYSTEMDEMO, technology, summary) %>%
  # separate risky from variable suffix
  separate(technology, c("Risky", "var"), extra = "merge", fill = "left") %>%
  #separate mean from variable prefix
  separate(var, c("technology", "meansd")) %>%
  # spread summary values back to wide form
  spread(meansd,summary) %>%
  #clean up
  select(-Risky)%>%
  arrange(mean)
```

```{r, echo=FALSE}
techrisk <- codedmodule2%>%
  group_by(SYSTEMTECHNO)%>%
  filter(SYSTEMTECHNO == 5)%>%
  summarize_at(vars(starts_with(c("RISKY"))), list(~mean(., na.rm = TRUE), ~median(., na.rm = TRUE), ~sd(.,na.rm = TRUE),~var(.,na.rm = TRUE), ~n())) %>%
  #melt to long format
  gather(key = "technology", value = "summary", Risky_Hydro_mean, Risky_Solar_mean, Risky_Wind_mean, Risky_Nuclear_mean, Risky_Coal_mean, Risky_Gas_mean, Risky_Oil_mean, Risky_Hydro_sd, Risky_Solar_sd, Risky_Wind_sd, Risky_Nuclear_sd, Risky_Coal_sd, Risky_Gas_sd, Risky_Oil_sd) %>%
  select(SYSTEMTECHNO, technology, summary) %>%
  # separate risky from variable suffix
  separate(technology, c("Risky", "var"), extra = "merge", fill = "left") %>%
  #separate mean from variable prefix
  separate(var, c("technology", "meansd")) %>%
  # spread summary values back to wide form
  spread(meansd,summary) %>%
  #clean up
  select(-Risky)%>%
  arrange(mean)
```

```{r, echo=FALSE}
religionrisk <- codedmodule2%>%
  group_by(SYSTEMRELIGION)%>%
  filter(SYSTEMRELIGION == 5)%>%
  summarize_at(vars(starts_with(c("RISKY"))), list(~mean(., na.rm = TRUE), ~median(., na.rm = TRUE), ~sd(.,na.rm = TRUE),~var(.,na.rm = TRUE), ~n())) %>%
  #melt to long format
  gather(key = "technology", value = "summary", Risky_Hydro_mean, Risky_Solar_mean, Risky_Wind_mean, Risky_Nuclear_mean, Risky_Coal_mean, Risky_Gas_mean, Risky_Oil_mean, Risky_Hydro_sd, Risky_Solar_sd, Risky_Wind_sd, Risky_Nuclear_sd, Risky_Coal_sd, Risky_Gas_sd, Risky_Oil_sd) %>%
  select(SYSTEMRELIGION, technology, summary) %>%
  # separate risky from variable suffix
  separate(technology, c("Risky", "var"), extra = "merge", fill = "left") %>%
  #separate mean from variable prefix
  separate(var, c("technology", "meansd")) %>%
  # spread summary values back to wide form
  spread(meansd,summary) %>%
  #clean up
  select(-Risky)%>%
  arrange(mean)
```

```{r, echo=FALSE}
totalrisk <- codedmodule2%>%
  group_by(SYSTEMTOTAL)%>%
  filter(SYSTEMTOTAL == 5)%>%
  summarize_at(vars(starts_with(c("RISKY"))), list(~mean(., na.rm = TRUE), ~median(., na.rm = TRUE), ~sd(.,na.rm = TRUE),~var(.,na.rm = TRUE), ~n())) %>%
  #melt to long format
  gather(key = "technology", value = "summary", Risky_Hydro_mean, Risky_Solar_mean, Risky_Wind_mean, Risky_Nuclear_mean, Risky_Coal_mean, Risky_Gas_mean, Risky_Oil_mean, Risky_Hydro_sd, Risky_Solar_sd, Risky_Wind_sd, Risky_Nuclear_sd, Risky_Coal_sd, Risky_Gas_sd, Risky_Oil_sd) %>%
  select(SYSTEMTOTAL, technology, summary) %>%
  # separate risky from variable suffix
  separate(technology, c("Risky", "var"), extra = "merge", fill = "left") %>%
  #separate mean from variable prefix
  separate(var, c("technology", "meansd")) %>%
  # spread summary values back to wide form
  spread(meansd,summary) %>%
  #clean up
  select(-Risky)%>%
  arrange(mean)
```

```{r, echo=FALSE}

df1 <- left_join(religionrisk, totalrisk, by = "technology")
df2 <- left_join(demorisk, techrisk, by = "technology" )
```

```{r, echo=FALSE}

codedmodule2 %>%
   select(starts_with(c("Risky", "SYSTEM"))) %>%
  select(!starts_with(c("Risky_IND")))%>%
  filter(SYSTEMTOTAL == 5) +
  ggplot(aes(x= ))
  +
  #filter(SYSTEMTOTAL ==  "5"), "SYSTEMTEHNO" == 5, "SYSTEMDEMO" == 5, "SYSTEMRELIGION" == 5)
  ggplot(subset("SYSTEMTOTAL" == 5))

```