### 2022-04-12 module2/module2IVgraphs.Rmd

---
title: "module2IVgraphs"
output: 
  html_document:
    fig_width: 9
    fig_height: 6
    fig_caption: yes
    keep_md: yes
    theme: paper
    df_print: paged

  pdf_document:

    toc: true
    number_sections: true
    toc_depth: 2
    fig_width: 9
    fig_height: 6
    fig_caption: yes
---

```{r load packages, warning = FALSE, message = FALSE, echo=FALSE}
  knitr::opts_chunk$set(warning = FALSE, message = FALSE)

suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(readxl))
suppressPackageStartupMessages(library(tsibble))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(DT))
suppressPackageStartupMessages(library(ggplot2))

```

```{r, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
moduletwo <- read_excel(here::here("thesis_surveydata/political_factors_complete.xlsx"))
save(moduletwo, file = "moduletwo.rda")
module2 <- data.frame(moduletwo)
module2[module2 == "Rather not say/ Don't know"] <- "Rather not say/Don't know"

# loaded data set for module 2 eco -pol variables and removes the rows with question number and question text
```

# **Coded Dataset for survey module 2**
```{r, echo = FALSE}

#This chunk contains the coding schemes of various scales used in survey one: eco-political factors, kahan scale and acceptance scale
codedmodule2 <- module2 %>%

#remove row 1
  filter(!row_number() %in% c(1,2)) %>%

# replace risky likert scale with numbers
  mutate_at(vars(starts_with("Risky")), funs(case_when(. =="Not at all risky" ~ 1, 
                                                       . =="Slightly risky" ~ 2, 
                                                       . =="Moderately risky" ~ 3, 
                                                       . =="Very risky" ~ 4, 
                                                       . =="Extremely risky" ~ 5))) %>%

# replace beneficial likert scale with numbers  
  mutate_at(vars(starts_with("Ben")), funs(case_when(. =="Not at all beneficial" ~ 1,
                                                     . =="Slightly beneficial" ~ 2,
                                                     . =="Moderately beneficial" ~ 3,
                                                     . =="Very beneficial" ~ 4,
                                                     . =="Extremely beneficial" ~ 5 ))) %>%

# replace nuclear acceptance likert scale with numbers
  mutate_at(vars(N_accept,N_reluctantlyaccept,N_reject), funs(case_when(. == "Strongly disagree" ~ 1, 
                                                                        . == "Somewhat disagree" ~ 2,
                                                                        . == "Neither agree nor disagree" ~3,
                                                                        . == "Somewhat agree" ~ 4,
                                                                        . == "Strongly agree" ~ 5))) %>%

# code likert scale for variables for Kahan scale into numbers
  mutate_at(vars(starts_with (c("K_I","K_H","DISPLACE", "POLLUTE", "HEALTH", "JOBS", "BEAUTY", "PRIDE", "NPRIDE", "DEV", "PROSPER", "RELY"))), funs(case_when(. == "Strongly disagree" ~ 1, 
               . == "Somewhat disagree" ~ 2,
               . == "Neither agree nor disagree" ~3,
               . == "Somewhat agree" ~ 4,
               . == "Strongly agree" ~ 5))) %>%

# reverse code for likert scale for variables for Kahan scale into numbers
  mutate_at(vars(starts_with (c("K_S","K_E"))), funs(case_when(. == "Strongly disagree" ~ 5, 
                                                               . == "Somewhat disagree" ~ 4,
                                                               . == "Neither agree nor disagree" ~3,
                                                               . == "Somewhat agree" ~ 2,
                                                               . == "Strongly agree" ~ 1))) %>%

# code eco-pol scale variables into numbers
  mutate_at(vars(SYSTEMDEMO,SYSTEMRELIGION,SYSTEMTECHNO,SYSTEMTOTAL,WEALTHLIM,MECHANISATION,DECISIONDECEN,INDUSTRYLARGE,ECONOMYGLOBAL,OWNERPVT,OWNERNOREG), funs(case_when(. == "Strongly disagree" ~ 1, 
                        . == "Somewhat disagree" ~ 2,
                        . == "Neither agree nor disagree" ~3,
                        . == "Somewhat agree" ~ 4,
                        . == "Strongly agree" ~ 5))) %>%

# reverse code eco-pol scale variables into numbers 
 mutate_at(vars(DECISIONCEN, INDUSTRYSMALL, ECONOMYLOCAL, ENVOVERDEV,OWNERPUB, OWNERREG), funs(case_when(. == "Strongly disagree" ~ 5, 
                                                                                                         . == "Somewhat disagree" ~ 4,
                                                                                                         . == "Neither agree nor disagree" ~3,
                                                                                                         . == "Somewhat agree" ~ 2,
                                                                                                         . == "Strongly agree" ~ 1)))

data.frame(codedmodule2)

```

```{r, echo = FALSE}
demorisk <- codedmodule2%>%
  group_by(SYSTEMDEMO)%>%
  filter(SYSTEMDEMO == 5)%>%
  summarize_at(vars(starts_with(c("RISKY"))), list(~mean(., na.rm = TRUE), ~median(., na.rm = TRUE), ~sd(.,na.rm = TRUE),~var(.,na.rm = TRUE), ~n())) %>%
  #melt to long format
  gather(key = "technology", value = "summary", Risky_Hydro_mean, Risky_Solar_mean, Risky_Wind_mean, Risky_Nuclear_mean, Risky_Coal_mean, Risky_Gas_mean, Risky_Oil_mean, Risky_Hydro_sd, Risky_Solar_sd, Risky_Wind_sd, Risky_Nuclear_sd, Risky_Coal_sd, Risky_Gas_sd, Risky_Oil_sd,Risky_Hydro_n, Risky_Solar_n, Risky_Wind_n, Risky_Nuclear_n, Risky_Coal_n, Risky_Gas_n, Risky_Oil_n) %>%
  select(SYSTEMDEMO, technology, summary) %>%
  # separate risky from variable suffix
  separate(technology, c("Risky", "var"), extra = "merge", fill = "left") %>%
  #separate mean from variable prefix
  separate(var, c("technology", "meansd")) %>%
  # spread summary values back to wide form
  spread(meansd,summary) %>%
  #clean up
  select(-Risky)%>%
  arrange(mean) %>%
  mutate_at(vars(SYSTEMDEMO), funs(case_when(. == 5 ~ "democractic"))) %>%
  dplyr::rename(Govt = SYSTEMDEMO)

techrisk <- codedmodule2%>%
  group_by(SYSTEMTECHNO)%>%
  filter(SYSTEMTECHNO == 5)%>%
  summarize_at(vars(starts_with(c("RISKY"))), list(~mean(., na.rm = TRUE), ~median(., na.rm = TRUE), ~sd(.,na.rm = TRUE),~var(.,na.rm = TRUE), ~n())) %>%
  #melt to long format
  gather(key = "technology", value = "summary", Risky_Hydro_mean, Risky_Solar_mean, Risky_Wind_mean, Risky_Nuclear_mean, Risky_Coal_mean, Risky_Gas_mean, Risky_Oil_mean, Risky_Hydro_sd, Risky_Solar_sd, Risky_Wind_sd, Risky_Nuclear_sd, Risky_Coal_sd, Risky_Gas_sd, Risky_Oil_sd, Risky_Hydro_n, Risky_Solar_n, Risky_Wind_n, Risky_Nuclear_n, Risky_Coal_n, Risky_Gas_n, Risky_Oil_n) %>%
  select(SYSTEMTECHNO, technology, summary) %>%
  # separate risky from variable suffix
  separate(technology, c("Risky", "var"), extra = "merge", fill = "left") %>%
  #separate mean from variable prefix
  separate(var, c("technology", "meansd")) %>%
  # spread summary values back to wide form
  spread(meansd,summary) %>%
  #clean up
  select(-Risky)%>%
  arrange(mean) %>%
  mutate_at(vars(SYSTEMTECHNO), funs(case_when(. == 5 ~ "technocratic"))) %>%
  dplyr::rename(Govt = SYSTEMTECHNO)

religionrisk <- codedmodule2%>%
  group_by(SYSTEMRELIGION)%>%
  filter(SYSTEMRELIGION == 5)%>%
  summarize_at(vars(starts_with(c("RISKY"))), list(~mean(., na.rm = TRUE), ~median(., na.rm = TRUE), ~sd(.,na.rm = TRUE),~var(.,na.rm = TRUE), ~n())) %>%
  #melt to long format
  gather(key = "technology", value = "summary", Risky_Hydro_mean, Risky_Solar_mean, Risky_Wind_mean, Risky_Nuclear_mean, Risky_Coal_mean, Risky_Gas_mean, Risky_Oil_mean, Risky_Hydro_sd, Risky_Solar_sd, Risky_Wind_sd, Risky_Nuclear_sd, Risky_Coal_sd, Risky_Gas_sd, Risky_Oil_sd, Risky_Hydro_n, Risky_Solar_n, Risky_Wind_n, Risky_Nuclear_n, Risky_Coal_n, Risky_Gas_n, Risky_Oil_n) %>%
  select(SYSTEMRELIGION, technology, summary) %>%
  # separate risky from variable suffix
  separate(technology, c("Risky", "var"), extra = "merge", fill = "left") %>%
  #separate mean from variable prefix
  separate(var, c("technology", "meansd")) %>%
  # spread summary values back to wide form
  spread(meansd,summary) %>%
  #clean up
  select(-Risky)%>%
  arrange(mean)%>%
  mutate_at(vars(SYSTEMRELIGION), funs(case_when(. == 5 ~ "religious"))) %>%
  dplyr::rename(Govt = SYSTEMRELIGION)

totalrisk <- codedmodule2%>%
  group_by(SYSTEMTOTAL)%>%
  filter(SYSTEMTOTAL == 5)%>%
  summarize_at(vars(starts_with(c("RISKY"))), list(~mean(., na.rm = TRUE), ~median(., na.rm = TRUE), ~sd(.,na.rm = TRUE),~var(.,na.rm = TRUE), ~n())) %>%
  #melt to long format
  gather(key = "technology", value = "summary", Risky_Hydro_mean, Risky_Solar_mean, Risky_Wind_mean, Risky_Nuclear_mean, Risky_Coal_mean, Risky_Gas_mean, Risky_Oil_mean, Risky_Hydro_sd, Risky_Solar_sd, Risky_Wind_sd, Risky_Nuclear_sd, Risky_Coal_sd, Risky_Gas_sd, Risky_Oil_sd, Risky_Hydro_n, Risky_Solar_n, Risky_Wind_n, Risky_Nuclear_n, Risky_Coal_n, Risky_Gas_n, Risky_Oil_n) %>%
  select(SYSTEMTOTAL, technology, summary) %>%
  # separate risky from variable suffix
  separate(technology, c("Risky", "var"), extra = "merge", fill = "left") %>%
  #separate mean from variable prefix
  separate(var, c("technology", "meansd")) %>%
  # spread summary values back to wide form
  spread(meansd,summary) %>%
  #clean up
  select(-Risky)%>%
  arrange(mean)%>%
  mutate_at(vars(SYSTEMTOTAL), funs(case_when(. == 5 ~ "totalitarian"))) %>%
  dplyr::rename(Govt = SYSTEMTOTAL)
```

# **Perceived risk by Preferred Government Type**

The following is a list of statements about governance. How strongly do you agree or disagree with each of these statements? Do you strongly disagree(1), somewhat disagree(2), neither agree nor disagree(3), somewhat agree(4) or strongly agree(5)?

Plotted only strongly agree(5) for four government types.

```{r, echo=FALSE}

df1 <- full_join(religionrisk, totalrisk)
df2 <- full_join(demorisk, techrisk)
govtgraph <- full_join(df1, df2) %>%
  group_by(technology)

data.frame(govtgraph)

p <- ggplot(govtgraph, aes(x = technology, y= mean, group = Govt))+
  geom_line(aes(colour = Govt, line = Govt)) +
  labs(y = "mean perceived risk", x = "energy technology")

update_labels(p, list(colour = "Preferred Government Type"))
  #ylim(0,4)
  #scale_color_manual(values = c("darkred", "steelblue"))

```

# **Perceived risk by Support for Democratic government**

The following is a list of statements about governance. How strongly do you agree or disagree with each of these statements? Do you strongly disagree(1), somewhat disagree(2), neither agree nor disagree(3), somewhat agree(4) or strongly agree(5)? - It is very important to have a democratic political system because it ensures that no individual leader has too much power.

Plotted only strongly disagree(1) and strongly agree(5)

```{r, echo=FALSE}

demorisk2 <- codedmodule2%>%
  group_by(SYSTEMDEMO)%>%
  filter(SYSTEMDEMO %in% c(1, 5))%>%
  summarize_at(vars(starts_with(c("RISKY"))), list(~mean(., na.rm = TRUE), ~median(., na.rm = TRUE), ~sd(.,na.rm = TRUE),~var(.,na.rm = TRUE), ~n())) %>%
  #melt to long format
  gather(key = "technology", value = "summary", Risky_Hydro_mean, Risky_Solar_mean, Risky_Wind_mean, Risky_Nuclear_mean, Risky_Coal_mean, Risky_Gas_mean, Risky_Oil_mean, Risky_Hydro_sd, Risky_Solar_sd, Risky_Wind_sd, Risky_Nuclear_sd, Risky_Coal_sd, Risky_Gas_sd, Risky_Oil_sd, Risky_Hydro_n, Risky_Solar_n, Risky_Wind_n, Risky_Nuclear_n, Risky_Coal_n, Risky_Gas_n, Risky_Oil_n) %>%
  select(SYSTEMDEMO, technology, summary) %>%
  # separate risky from variable suffix
  separate(technology, c("Risky", "var"), extra = "merge", fill = "left") %>%
  #separate mean from variable prefix
  separate(var, c("technology", "meansd")) %>%
  # spread summary values back to wide form
  spread(meansd,summary) %>%
  #clean up
  select(-Risky)%>%
  arrange(SYSTEMDEMO)

data.frame(demorisk2)

d <- demorisk2 %>%
  group_by(technology) %>%
  ggplot(aes(x = technology, y= mean, group = SYSTEMDEMO))+
  geom_line(aes(colour = SYSTEMDEMO, line = SYSTEMDEMO)) +
   labs(y = "mean perceived risk", x = "energy technology")

update_labels(d, list(colour = "Support for Democratic Government"))

```

# **Perceived risk by Support for decentralised governance**

```{r, echo=FALSE}

decenrisk2 <- codedmodule2%>%
    select(starts_with(c("Risky","Ben","SYSTEMDEMO","SYSTEMRELIGION","SYSTEMTECHNO","SYSTEMTOTAL","WEALTHLIM","MECHANISATION","DECISIONDECEN","DECISIONCEN","INDUSTRYLARGE","INDUSTRYSMALL","ECONOMYGLOBAL","ECONOMYLOCAL", "ENVOVERDEV","DEVOVERENV","OWNERPVT","OWNERNOREG","OWNERPUB", "OWNERREG")))%>%
  rowwise()%>%
  dplyr::mutate(Pro_decentralisation = mean(c_across("DECISIONDECEN":"DECISIONCEN"))) %>%
  group_by(Pro_decentralisation)%>%
  filter(Pro_decentralisation %in% c(1, 5))%>%
  summarize_at(vars(starts_with(c("RISKY"))), list(~mean(., na.rm = TRUE), ~median(., na.rm = TRUE), ~sd(.,na.rm = TRUE),~var(.,na.rm = TRUE), ~n())) %>%
  #melt to long format
  gather(key = "technology", value = "summary", Risky_Hydro_mean, Risky_Solar_mean, Risky_Wind_mean, Risky_Nuclear_mean, Risky_Coal_mean, Risky_Gas_mean, Risky_Oil_mean, Risky_Hydro_sd, Risky_Solar_sd, Risky_Wind_sd, Risky_Nuclear_sd, Risky_Coal_sd, Risky_Gas_sd, Risky_Oil_sd,Risky_Hydro_n, Risky_Solar_n, Risky_Wind_n, Risky_Nuclear_n, Risky_Coal_n, Risky_Gas_n, Risky_Oil_n ) %>%
  select(Pro_decentralisation, technology, summary) %>%
  # separate risky from variable suffix
  separate(technology, c("Risky", "var"), extra = "merge", fill = "left") %>%
  #separate mean from variable prefix
  separate(var, c("technology", "meansd")) %>%
  # spread summary values back to wide form
  spread(meansd,summary) %>%
  #clean up
  select(-Risky)%>%
  arrange(Pro_decentralisation)

data.frame(decenrisk2)

decen <- decenrisk2 %>%
  group_by(technology) %>%
  ggplot(aes(x = technology, y= mean, group = Pro_decentralisation))+
  geom_line(aes(colour = Pro_decentralisation, line = Pro_decentralisation))+
  ylim(1,4)+
  labs(y = "mean perceived risk", x = "energy technology")

update_labels(decen, list(colour = "Pro decentralised governance"))
```

# **Perceived risk by Support for large scale industry**

```{r, echo=FALSE}

largerisk2 <- codedmodule2%>%
    select(starts_with(c("Risky","Ben","SYSTEMDEMO","SYSTEMRELIGION","SYSTEMTECHNO","SYSTEMTOTAL","WEALTHLIM","MECHANISATION","DECISIONDECEN","DECISIONCEN","INDUSTRYLARGE","INDUSTRYSMALL","ECONOMYGLOBAL","ECONOMYLOCAL", "ENVOVERDEV","DEVOVERENV","OWNERPVT","OWNERNOREG","OWNERPUB", "OWNERREG")))%>%
  rowwise()%>%
  dplyr::mutate(Pro_largescale = sum(c_across("INDUSTRYLARGE":"INDUSTRYSMALL"))) %>%
  group_by(Pro_largescale)%>%
  filter(Pro_largescale%in% c(2, 10))%>%
  summarize_at(vars(starts_with(c("RISKY"))), list(~mean(., na.rm = TRUE), ~median(., na.rm = TRUE), ~sd(.,na.rm = TRUE),~var(.,na.rm = TRUE), ~n())) %>%
  #melt to long format
  gather(key = "technology", value = "summary", Risky_Hydro_mean, Risky_Solar_mean, Risky_Wind_mean, Risky_Nuclear_mean, Risky_Coal_mean, Risky_Gas_mean, Risky_Oil_mean, Risky_Hydro_sd, Risky_Solar_sd, Risky_Wind_sd, Risky_Nuclear_sd, Risky_Coal_sd, Risky_Gas_sd, Risky_Oil_sd, Risky_Hydro_n, Risky_Solar_n, Risky_Wind_n, Risky_Nuclear_n, Risky_Coal_n, Risky_Gas_n, Risky_Oil_n) %>%
  select(Pro_largescale, technology, summary) %>%
  # separate risky from variable suffix
  separate(technology, c("Risky", "var"), extra = "merge", fill = "left") %>%
  #separate mean from variable prefix
  separate(var, c("technology", "meansd")) %>%
  # spread summary values back to wide form
  spread(meansd,summary) %>%
  #clean up
  select(-Risky)%>%
  arrange(Pro_largescale)

data.frame(largerisk2)

l<- largerisk2 %>%
  group_by(technology) %>%
  ggplot(aes(x = technology, y= mean, group = Pro_largescale))+
  geom_line(aes(colour = Pro_largescale, line = Pro_largescale))+
  ylim(1,4)+
  labs(y = "mean perceived risk", x = "energy technology")

update_labels(l, list(colour = "Pro large scale industry"))
```

```{r, echo=FALSE}

#globalrisk2 <- codedmodule2%>%
#    select(starts_with(c("Risky","Ben","SYSTEMDEMO","SYSTEMRELIGION","SYSTEMTECHNO","SYSTEMTOTAL","WEALTHLIM","MECHANISATION","DECISIONDECEN","DECISIONCEN","INDUSTRYLARGE","INDUSTRYSMALL","ECONOMYGLOBAL","ECONOMYLOCAL", "ENVOVERDEV","DEVOVERENV","OWNERPVT","OWNERNOREG","OWNERPUB", "OWNERREG")))%>%
#  rowwise()%>%
#  dplyr::mutate(Pro_global = sum(c_across("ECONOMYGLOBAL":"ECONOMYLOCAL"))) %>%
#  group_by(Pro_global)%>%
#  filter(Pro_global%in% c(2, 10))%>%
#  summarize_at(vars(starts_with(c("RISKY"))), list(~mean(., na.rm = TRUE), ~median(., na.rm = TRUE), ~sd(.,na.rm = TRUE),~var(.,na.rm = TRUE), ~n())) %>%
  #melt to long format
#  gather(key = "technology", value = "summary", Risky_Hydro_mean, Risky_Solar_mean, Risky_Wind_mean, Risky_Nuclear_mean, Risky_Coal_mean, Risky_Gas_mean, Risky_Oil_mean, Risky_Hydro_sd, Risky_Solar_sd, Risky_Wind_sd, Risky_Nuclear_sd, Risky_Coal_sd, Risky_Gas_sd, Risky_Oil_sd) %>%
#  select(Pro_global, technology, summary) %>%
  # separate risky from variable suffix
#  separate(technology, c("Risky", "var"), extra = "merge", fill = "left") %>%
  #separate mean from variable prefix
#  separate(var, c("technology", "meansd")) %>%
  # spread summary values back to wide form
#  spread(meansd,summary) %>%
  #clean up
#  select(-Risky)%>%
#  arrange(mean)

#globalrisk2 %>%
#  group_by(technology) %>%
#  ggplot(aes(x = technology, y= mean, group = Pro_global))+
#  geom_line(aes(colour = Pro_global, line = Pro_global))+
#  ylim(1,4)
```

```{r, echo=FALSE}

#envrisk2 <- codedmodule2%>%
#    select(starts_with(c("Risky","Ben","SYSTEMDEMO","SYSTEMRELIGION","SYSTEMTECHNO","SYSTEMTOTAL","WEALTHLIM","MECHANISATION","DECISIONDECEN","DECISIONCEN","INDUSTRYLARGE","INDUSTRYSMALL","ECONOMYGLOBAL","ECONOMYLOCAL", "ENVOVERDEV","DEVOVERENV","OWNERPVT","OWNERNOREG","OWNERPUB", "OWNERREG")))%>%
#  rowwise()%>%
#  group_by(ENVOVERDEV)%>%
#  filter(ENVOVERDEV%in% c(1, 5))%>%
#  summarize_at(vars(starts_with(c("RISKY"))), list(~mean(., na.rm = TRUE), ~median(., na.rm = TRUE), ~sd(.,na.rm = TRUE),~var(.,na.rm = TRUE), ~n())) %>%
  #melt to long format
#  gather(key = "technology", value = "summary", Risky_Hydro_mean, Risky_Solar_mean, Risky_Wind_mean, Risky_Nuclear_mean, Risky_Coal_mean, Risky_Gas_mean, Risky_Oil_mean, Risky_Hydro_sd, Risky_Solar_sd, Risky_Wind_sd, Risky_Nuclear_sd, Risky_Coal_sd, Risky_Gas_sd, Risky_Oil_sd) %>%
#  select(ENVOVERDEV, technology, summary) %>%
  # separate risky from variable suffix
#  separate(technology, c("Risky", "var"), extra = "merge", fill = "left") %>%
  #separate mean from variable prefix
#  separate(var, c("technology", "meansd")) %>%
  # spread summary values back to wide form
#  spread(meansd,summary) %>%
  #clean up
#  select(-Risky)%>%
#  arrange(mean)

#envrisk2 %>%
#  group_by(technology) %>%
#  ggplot(aes(x = technology, y= mean, group = ENVOVERDEV))+
#  geom_line(aes(colour = ENVOVERDEV, line = ENVOVERDEV))+
#  ylim(1,4)
```

```{r, echo=FALSE}

#devrisk2 <- codedmodule2%>%
#    select(starts_with(c("Risky","Ben","SYSTEMDEMO","SYSTEMRELIGION","SYSTEMTECHNO","SYSTEMTOTAL","WEALTHLIM","MECHANISATION","DECISIONDECEN","DECISIONCEN","INDUSTRYLARGE","INDUSTRYSMALL","ECONOMYGLOBAL","ECONOMYLOCAL", "ENVOVERDEV","DEVOVERENV","OWNERPVT","OWNERNOREG","OWNERPUB", "OWNERREG")))%>%
#  rowwise()%>%
#  group_by(DEVOVERENV)%>%
#  filter(DEVOVERENV%in% c(1, 5))%>%
#  summarize_at(vars(starts_with(c("RISKY"))), list(~mean(., na.rm = TRUE), ~median(., na.rm = TRUE), ~sd(.,na.rm = TRUE),~var(.,na.rm = TRUE), ~n())) %>%
  #melt to long format
#  gather(key = "technology", value = "summary", Risky_Hydro_mean, Risky_Solar_mean, Risky_Wind_mean, Risky_Nuclear_mean, Risky_Coal_mean, Risky_Gas_mean, Risky_Oil_mean, Risky_Hydro_sd, Risky_Solar_sd, Risky_Wind_sd, Risky_Nuclear_sd, Risky_Coal_sd, Risky_Gas_sd, Risky_Oil_sd) %>%
#  select(DEVOVERENV, technology, summary) %>%
  # separate risky from variable suffix
#  separate(technology, c("Risky", "var"), extra = "merge", fill = "left") %>%
  #separate mean from variable prefix
#  separate(var, c("technology", "meansd")) %>%
  # spread summary values back to wide form
#  spread(meansd,summary) %>%
  #clean up
#  select(-Risky)%>%
#  arrange(mean)

#devrisk2 %>%
#  group_by(technology) %>%
#  ggplot(aes(x = technology, y= mean, group = DEVOVERENV))+
#  geom_line(aes(colour = DEVOVERENV, line = DEVOVERENV))+
#  ylim(1,4)
```

```{r, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
#calculating individualism score and hierarchy score for each respondent

Kscalescores <- codedmodule2 %>%
  select(starts_with(c("Risky","K_I","K_S", "K_H", "K_E")))%>%
  rowwise()%>%
  na.omit()%>%
  dplyr::mutate(Individualism_score = mean(c_across(K_IINTRFER:K_SPROTECT)))%>%
  dplyr::mutate(Hierarchy_score = mean(c_across(K_HEQUAL:K_ERADEQ2)))%>%
  dplyr::select(!starts_with(c("K_I","K_S", "K_H", "K_E")))

```

# **Scatter plot of Kahan scale scores around the median scores on Individualism and Hierarchy scales**

```{r, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

#scatter plot of Kahan scale scores around the median scores on Individualism and Hierarchy scales.

Kscalescores %>%
  ggplot(aes(Individualism_score, Hierarchy_score))+
  geom_count(aes(color = ..n.., size = ..n..))+
  guides(color = 'legend')+
  geom_hline(yintercept=2, colour="black", lwd=1)+
  geom_vline(xintercept=3, colour="black", lwd=1)+
  labs(x = " Communitarianism to Individualism", y = "Egalitarianism to Hierarchy")

```

# **Scatter plot for pro decentralised governance against perceived risks from rooftop solar**

```{r, echo=FALSE}

ecopolscalescores <- codedmodule2 %>%
  select(starts_with(c("Risky","Ben","SYSTEMDEMO","SYSTEMRELIGION","SYSTEMTECHNO","SYSTEMTOTAL","WEALTHLIM","MECHANISATION","DECISIONDECEN","DECISIONCEN","INDUSTRYLARGE","INDUSTRYSMALL","ECONOMYGLOBAL","ECONOMYLOCAL", "ENVOVERDEV","DEVOVERENV","OWNERPVT","OWNERNOREG","OWNERPUB", "OWNERREG")))%>%
  rowwise()%>%
  dplyr::mutate(Pro_decentralisation = sum(c_across("DECISIONDECEN":"DECISIONCEN")))%>%
   dplyr::mutate(Pro_largescale = mean(c_across("INDUSTRYLARGE":"INDUSTRYSMALL")))%>%
   dplyr::mutate(Pro_global = mean(c_across("ECONOMYGLOBAL":"ECONOMYLOCAL")))%>%
   dplyr::mutate(Pro_ecogrowth = mean(c_across("ENVOVERDEV":"DEVOVERENV")))%>%
   dplyr::mutate(Pro_private = mean(c_across("OWNERPVT":"OWNERREG")))%>%
  #dplyr::select(!c("DECISIONDECEN","DECISIONCEN", "INDUSTRYLARGE","INDUSTRYSMALL","ECONOMYGLOBAL", "ECONOMYLOCAL","ENVOVERDEV",     "DEVOVERENV","OWNERPVT","OWNERNOREG","OWNERPUB", "OWNERREG")) %>%
    na.omit()

ggplot(ecopolscalescores, aes(Pro_decentralisation, Risky_INDSolar))+
  geom_count()+
  labs(x = "Pro decentralised governance", y = " Perceived risks from rooftop solar")

```