### 2023-02-04 combined data set/DVgraphs.Rmd
conducting the Kruskal wallis test on DVs

---
title: "Dependent Variable Graphs"
output:
  pdf_document:
    toc: yes
    number_sections: yes
    toc_depth: 2
    fig_width: 9
    fig_height: 6
    fig_caption: yes
  html_document:
    fig_width: 9
    fig_height: 6
    fig_caption: yes
    keep_md: yes
    theme: paper
    df_print: paged
---

```{r load packages, warning = FALSE, message = FALSE, echo=FALSE}
  knitr::opts_chunk$set(warning = FALSE, message = FALSE)

suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(readxl))
suppressPackageStartupMessages(library(hablar))
suppressPackageStartupMessages(library(tsibble))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(DT))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(psych)) # psychometrics

```

```{r, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
moduletwo <- read_excel(here::here("thesis_surveydata/political_factors_complete.xlsx"))

module2 <- data.frame(moduletwo)
module2[module2 == "Rather not say/ Don't know"] <- "Rather not say/Don't know"

# loaded data set for module 2 eco -pol variables and removes the rows with question number and question text
```

```{r, eval =TRUE, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
#This chunk contains the coding schemes of various scales used in survey one: eco-political factors, kahan scale and acceptance scale
codedmodule2 <- module2 %>%

#remove row 1
  filter(!row_number() %in% c(1,2)) %>%

# replace risky likert scale with numbers
  mutate_at(vars(starts_with("Risky")), funs(case_when(. =="Not at all risky" ~ 1, 
                                                       . =="Slightly risky" ~ 2, 
                                                       . =="Moderately risky" ~ 3, 
                                                       . =="Very risky" ~ 4, 
                                                       . =="Extremely risky" ~ 5))) %>%

# replace beneficial likert scale with numbers  
  mutate_at(vars(starts_with("Ben")), funs(case_when(. =="Not at all beneficial" ~ 1,
                                                     . =="Slightly beneficial" ~ 2,
                                                     . =="Moderately beneficial" ~ 3,
                                                     . =="Very beneficial" ~ 4,
                                                     . =="Extremely beneficial" ~ 5 ))) %>%

# replace nuclear acceptance likert scale with numbers
  mutate_at(vars(N_accept,N_reluctantlyaccept,N_reject), funs(case_when(. == "Strongly disagree" ~ 1, 
                                                                        . == "Somewhat disagree" ~ 2,
                                                                        . == "Neither agree nor disagree" ~3,
                                                                        . == "Somewhat agree" ~ 4,
                                                                        . == "Strongly agree" ~ 5))) %>%

# code likert scale for variables for Kahan scale into numbers
  mutate_at(vars(starts_with (c("K_I","K_H","DISPLACE", "POLLUTE", "HEALTH", "JOBS", "BEAUTY", "PRIDE", "NPRIDE", "DEV", "PROSPER", "RELY"))), funs(case_when(. == "Strongly disagree" ~ 1, 
               . == "Somewhat disagree" ~ 2,
               . == "Neither agree nor disagree" ~3,
               . == "Somewhat agree" ~ 4,
               . == "Strongly agree" ~ 5))) %>%

# reverse code for likert scale for variables for Kahan scale into numbers
  mutate_at(vars(starts_with (c("K_S","K_E"))), funs(case_when(. == "Strongly disagree" ~ 5, 
                                                               . == "Somewhat disagree" ~ 4,
                                                               . == "Neither agree nor disagree" ~3,
                                                               . == "Somewhat agree" ~ 2,
                                                               . == "Strongly agree" ~ 1))) %>%

# code eco-pol scale variables into numbers
  mutate_at(vars(SYSTEMDEMO,SYSTEMRELIGION,SYSTEMTECHNO,SYSTEMTOTAL,WEALTHLIM,MECHANISATION,DECISIONDECEN,INDUSTRYLARGE,ECONOMYGLOBAL,OWNERPVT,OWNERNOREG), funs(case_when(. == "Strongly disagree" ~ 1, 
                        . == "Somewhat disagree" ~ 2,
                        . == "Neither agree nor disagree" ~3,
                        . == "Somewhat agree" ~ 4,
                        . == "Strongly agree" ~ 5))) %>%

# reverse code eco-pol scale variables into numbers 
 mutate_at(vars(DECISIONCEN, INDUSTRYSMALL, ECONOMYLOCAL, ENVOVERDEV,OWNERPUB, OWNERREG), funs(case_when(. == "Strongly disagree" ~ 5, 
                                                                                                         . == "Somewhat disagree" ~ 4,
                                                                                                         . == "Neither agree nor disagree" ~3,
                                                                                                         . == "Somewhat agree" ~ 2,
                                                                                                         . == "Strongly agree" ~ 1)))
```

```{r, eval=TRUE, echo=FALSE}

moduleone <- read_excel(here::here("thesis_surveydata/livelihood_insecurity_complete.xlsx"))

module1 <- data.frame(moduleone)

# loaded data set for module 2 eco -pol variables and removes the rows with question number and question text
```

```{r, eval =TRUE, echo=FALSE}
#This chunk contains the coding schemes of various scales used in survey one: eco-political factors, kahan scale and acceptance scale
codedmodule1 <- module1 %>%

#remove row 1
  filter(!row_number() %in% c(1,2)) %>%

# replace risky likert scale with numbers
  mutate_at(vars(starts_with("Risky")), funs(case_when(. =="Not at all risky" ~ 1, 
                                                       . =="Slightly risky" ~ 2, 
                                                       . =="Moderately risky" ~ 3, 
                                                       . =="Very risky" ~ 4, 
                                                       . =="Extremely risky" ~ 5))) %>%

# replace beneficial likert scale with numbers  
  mutate_at(vars(starts_with("Ben")), funs(case_when(. =="Not at all beneficial" ~ 1,
                                                     . =="Slightly beneficial" ~ 2,
                                                     . =="Moderately beneficial" ~ 3,
                                                     . =="Very beneficial" ~ 4,
                                                     . =="Extremely beneficial" ~ 5 ))) %>%

# replace nuclear acceptance likert scale with numbers
  mutate_at(vars(N_accept,N_reluctantlyaccept,N_reject), funs(case_when(. == "Strongly disagree" ~ 1, 
                                                                        . == "Somewhat disagree" ~ 2,
                                                                        . == "Neither agree nor disagree" ~3,
                                                                        . == "Somewhat agree" ~ 4,
                                                                        . == "Strongly agree" ~ 5)))

#coded dependent variables for module 1 livelihood and income security survey

```

```{r, echo=FALSE}

suppressPackageStartupMessages(library(tidyselect))
# combine codedmodule1 and codedmodule2. Therefore these graphs are made on the combined data set from both the surveys.

numcol2 <- codedmodule2 %>% add_column(Respondent = 1:1099, .before = "Language") 
numcol1 <- codedmodule1 %>% add_column(Respondent = 1099:2160, .before = "Language")

joineddf <-  full_join(numcol2, numcol1, by = c("Respondent", "Language", "Survey_Date", "State", "Risky_Hydro", "Risky_Solar", "Risky_Wind", "Risky_Nuclear", "Risky_Coal", "Risky_Gas", "Risky_Oil", "Risky_INDHydro", "Risky_INDSolar", "Risky_INDWind", "Risky_INDBiogas", "Risky_INDDiesel", "Risky_INDKerosene", "Risky_INDFirewoodetc", "Risky_INDLPG", "Ben_Hydro", "Ben_Solar", "Ben_Wind", "Ben_Nuclear", "Ben_Coal", "Ben_Gas", "Ben_Oil", "Ben_INDHydro", "Ben_INDSolar", "Ben_INDWind", "Ben_INDDiesel", "Ben_INDBiogas", "Ben_INDFirewoodetc", "Ben_INDLPG", "Ben_INDKerosene", "N_accept", "N_reluctantlyaccept", "N_reject", "age", "urban_rural", "gender", "caste", "religion", "religiontext"))

```

\newpage

## Q1. Below is a list of various centrally managed sources of energy. How risky are each of these energy technologies according to you?  Please tell me whether you see each technology as: not at all risky(1), slightly risky(2), moderately risky(3), very risky(4), or extremely risky(5)

```{r, eval=TRUE, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

# summary statistics for risks from different energy sources

risksumcentral <- joineddf %>%
  dplyr::select(!starts_with(c("RISKY_IND")))%>%
  summarize_at(vars(starts_with(c("RISKY"))), list(~mean(., na.rm = TRUE), ~median(., na.rm = TRUE), ~sd(.,na.rm = TRUE),~var(.,na.rm = TRUE), ~n())) %>%
  #add row index so later spreading indexed correctly
  add_rownames()%>%
  #melt to long format
  gather(technology, value, -rowname) %>%
  # separate risky from variable suffix
  separate(technology, c("Risky", "var"), extra = "merge", fill = "left") %>%
  #separate mean from variable prefix
  separate(var, c("technology", "summary")) %>%
  # spread summary values back to wide form
  spread(summary,value) %>%
  #clean up
  dplyr::select(-rowname, -Risky) %>%
  arrange(mean) %>%
  setNames(paste0('perceivedrisk.', names(.))) %>%
  dplyr::rename(technology= perceivedrisk.technology)

data.frame(risksumcentral)

```

```{r, echo=FALSE}

knitr::opts_chunk$set(warning = FALSE, message = FALSE)

risksumcentral %>%
  dplyr::select(technology, perceivedrisk.mean, perceivedrisk.sd) %>%
  mutate(technology = fct_reorder(technology, perceivedrisk.mean, .desc = FALSE))%>%
  ggplot(aes(x=technology, y= perceivedrisk.mean))+
  geom_bar(stat="identity", position=position_dodge(), fill = "firebrick2")+
  ylim(0,5)+
  geom_errorbar(aes(ymin=perceivedrisk.mean-perceivedrisk.sd, ymax=perceivedrisk.mean + perceivedrisk.sd), width=.2,
                 position=position_dodge(.9))+
  labs(y = "Mean Values of Perceived Risk", x = "Energy Technologies")+
  ggtitle("Perceived risk from centrally managed energy technologies")+
  coord_flip()

```

\newpage

## Q2. Following is a list of various individually or community-owned and managed sources of energy. How risky are each of these energy technologies according to you? Please tell me whether you see each technology as: not at all risky(1), slightly risky(2), moderately risky(3), very risky(4), or extremely risky(5).

prefix IND stands for individually or community-owned and managed.

```{r, eval=TRUE, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

# summary statistics for risks from different energy sources

risksumIND <- joineddf %>%
  summarize_at(vars(starts_with(c("RISKY_IND"))), list(~mean(., na.rm = TRUE), ~median(., na.rm = TRUE), ~sd(.,na.rm = TRUE),~var(.,na.rm = TRUE), ~n())) %>%
  #add row index so later spreading indexed correctly
  add_rownames()%>%
  #melt to long format
  gather(technology, value, -rowname) %>%
  # separate risky from variable suffix
  separate(technology, c("Risky", "var"), extra = "merge", fill = "left") %>%
  #separate mean from variable prefix
  separate(var, c("technology", "summary")) %>%
  # spread summary values back to wide form
  spread(summary,value) %>%
  #clean up
  dplyr::select(-rowname, -Risky) %>%
  arrange(mean) %>%
  setNames(paste0('perceivedrisk.', names(.))) %>%
  dplyr::rename(technology= perceivedrisk.technology)

data.frame(risksumIND)

```

```{r, echo=FALSE}

knitr::opts_chunk$set(warning = FALSE, message = FALSE)

risksumIND %>%
  dplyr::select(technology, perceivedrisk.mean,perceivedrisk.sd) %>%
  mutate(technology = fct_reorder(technology, perceivedrisk.mean, .desc = FALSE))%>%
  ggplot(aes(x=technology, y= perceivedrisk.mean))+
  geom_bar(stat="identity", position=position_dodge(), fill = "firebrick2")+
  geom_errorbar(aes(ymin=perceivedrisk.mean-perceivedrisk.sd, ymax=perceivedrisk.mean + perceivedrisk.sd), width=.2,
                 position=position_dodge(.9))+
  ylim(0,5)+
  labs(y = "Mean Values of Perceived Risk", x = "Energy Technologies")+
  ggtitle("Perceived risk from individually or community owned energy technologies")+
  coord_flip()

```
\newpage

## Q3. Following is a list of centrally managed sources of energy. How beneficial are each of these according to you? Please tell me whether you see each technology as: not at all beneficial(1), slightly beneficial(2), moderately beneficial(3), very beneficial(4), or extremely beneficial(5)?

```{r, eval=TRUE, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

# summary statistics for risks from different energy sources

bensumcentral <- joineddf %>%
  dplyr::select(!starts_with(c("Ben_IND")))%>%
  summarize_at(vars(starts_with(c("Ben"))), list(~mean(., na.rm = TRUE), ~median(., na.rm = TRUE), ~sd(.,na.rm = TRUE),~var(.,na.rm = TRUE), ~n())) %>%
  #add row index so later spreading indexed correctly
  add_rownames()%>%
  #melt to long format
  gather(technology, value, -rowname) %>%
  # separate risky from variable suffix
  separate(technology, c("Ben", "var"), extra = "merge", fill = "left") %>%
  #separate mean from variable prefix
  separate(var, c("technology", "summary")) %>%
  # spread summary values back to wide form
  spread(summary,value) %>%
  #clean up
  dplyr::select(-rowname, -Ben) %>%
  arrange(mean) %>%
  setNames(paste0('perceivedbenefit.', names(.))) %>%
  dplyr::rename(technology= perceivedbenefit.technology)

data.frame(bensumcentral)

```

```{r, echo=FALSE}

knitr::opts_chunk$set(warning = FALSE, message = FALSE)

bensumcentral %>%
  dplyr::select(technology, perceivedbenefit.mean,perceivedbenefit.sd) %>%
  mutate(technology = fct_reorder(technology, perceivedbenefit.mean, .desc = FALSE))%>%
  ggplot(aes(x=technology, y= perceivedbenefit.mean))+
  geom_bar(stat="identity", position=position_dodge(), fill = "palegreen3")+
   geom_errorbar(aes(ymin=perceivedbenefit.mean-perceivedbenefit.sd, ymax=perceivedbenefit.mean + perceivedbenefit.sd), width=.2,
                 position=position_dodge(.9))+
  ylim(0,5)+
  labs(y = "Mean Values of Perceived Benefit", x = "Energy Technologies")+
  ggtitle("Perceived benefit from centrally managed energy technologies")+
  coord_flip()

```

\newpage

## Q4. Following is a list of various individually or community owned and managed sources of energy. How beneficial are each of these according to you? Please tell me whether you see each technology as: not at all beneficial(1), slightly beneficial(2), moderately beneficial(3), very beneficial(4) or extremely beneficial(5) ?

prefix IND stands for individually or community-owned and managed.

```{r, eval=TRUE, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

# summary statistics for risks from different energy sources

bensumIND <- joineddf %>%
  summarize_at(vars(starts_with(c("Ben_IND"))), list(~mean(., na.rm = TRUE), ~median(., na.rm = TRUE), ~sd(.,na.rm = TRUE),~var(.,na.rm = TRUE), ~n())) %>%
  #add row index so later spreading indexed correctly
  add_rownames()%>%
  #melt to long format
  gather(technology, value, -rowname) %>%
  # separate risky from variable suffix
  separate(technology, c("Ben", "var"), extra = "merge", fill = "left") %>%
  #separate mean from variable prefix
  separate(var, c("technology", "summary")) %>%
  # spread summary values back to wide form
  spread(summary,value) %>%
  #clean up
  dplyr::select(-rowname, -Ben) %>%
  arrange(mean) %>%
  setNames(paste0('perceivedbenefit.', names(.))) %>%
  dplyr::rename(technology= perceivedbenefit.technology)

data.frame(bensumIND)

```

```{r, echo=FALSE}

knitr::opts_chunk$set(warning = FALSE, message = FALSE)

bensumIND %>%
  dplyr::select(technology, perceivedbenefit.mean, perceivedbenefit.sd) %>%
  mutate(technology = fct_reorder(technology, perceivedbenefit.mean, .desc = FALSE))%>%
  ggplot(aes(x=technology, y= perceivedbenefit.mean))+
  geom_bar(stat="identity", position=position_dodge(), fill = "palegreen3")+
  geom_errorbar(aes(ymin=perceivedbenefit.mean-perceivedbenefit.sd, ymax=perceivedbenefit.mean + perceivedbenefit.sd), width=.2,
                 position=position_dodge(.9))+
  ylim(0,5)+
  labs(y = "Mean Values of Perceived Benefit", x = "Energy Technologies")+
  ggtitle("Perceived benefit from individually or community owned energy technologies")+
  coord_flip()

```

\newpage

```{r, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

center_barplot <- full_join( bensumcentral,risksumcentral, by = NULL, copy = FALSE, keep = FALSE) %>%
 dplyr::select(technology, perceivedbenefit.mean, perceivedrisk.mean, perceivedbenefit.sd, perceivedrisk.sd) %>%
  mutate(technology = fct_reorder(technology, perceivedbenefit.mean, .desc = FALSE))%>%
   gather(key = "level", value = "mean", perceivedbenefit.mean, perceivedrisk.mean)

center_barplot$level <- as.factor(center_barplot$level)

 center_barplot%>% 
  ggplot(aes(x=technology, y= mean, fill =  (factor(level, levels = c("perceivedrisk.mean", "perceivedbenefit.mean")))))+
  geom_bar(stat="identity", position=position_dodge())+

#geom_errorbar(aes(ymin= mean - sd, ymax= mean + sd), width =.2, position=position_dodge(.9))+
  ylim(0,4)+
  scale_fill_manual("legend", values = c("perceivedbenefit.mean" = "palegreen3","perceivedrisk.mean" = "firebrick2"))+
  labs(y = "Mean Values", x = "Energy Technologies")+
  ggtitle("Perceived risk and benefit from centrally owned energy technologies")+
  coord_flip()

```

Prefix IND stands for individually or community-owned and managed.

```{r, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

IND_barplot <- full_join(bensumIND,risksumIND, by = NULL, copy = FALSE, keep = FALSE) %>%
 dplyr::select(technology, perceivedbenefit.mean, perceivedrisk.mean) %>%
  mutate(technology = fct_reorder(technology, perceivedbenefit.mean, .desc = FALSE))%>%
  gather(key = "level", value = "mean", perceivedbenefit.mean, perceivedrisk.mean)

IND_barplot$level <- as.factor(IND_barplot$level)

IND_barplot%>%  ggplot(aes(x=technology, y= mean, fill = (factor(IND_barplot$level, levels = c("perceivedrisk.mean", "perceivedbenefit.mean")))))+
  geom_bar(stat="identity", position=position_dodge())+
  ylim(0,4)+
  scale_fill_manual("legend", values = c("perceivedbenefit.mean" = "palegreen3","perceivedrisk.mean" = "firebrick2"))+
  labs(y = "Mean Values", x = "Energy Technologies")+
  ggtitle("Perceived risk and benefit from individually owned energy technologies")+
  coord_flip()

```

```{r, echo=FALSE}

library(dplyr)
library(tidyr)

perceivedrisk <-joineddf %>% select(starts_with("Risky")) %>%
  gather(key = "technology", value = "perceived_risk") %>% 
  mutate(technology = gsub("Risky_", "", technology)) %>%
  na.omit()

perceivedrisk

```

```{r}
# checking assumptions for kruskal wallis test

duplicated(joineddf)
sum(duplicated(joineddf))

# there are no duplicate responses in our dataframe :).

ggplot(perceivedrisk, aes(x = perceived_risk, y = technology)) + 
  geom_boxplot()

levels(perceivedrisk$technology)

bartlett.test(perceived_risk ~ technology, data = perceivedrisk)

str(perceivedrisk)

```

```{r, echo=FALSE}
# create a list of technology levels
tech_levels <- unique(perceivedrisk$technology)

# perform Shapiro test for each technology
shapiro_results <- lapply(tech_levels, function(x) {
  shapiro.test(perceivedrisk$perceived_risk[perceivedrisk$technology == x])
})

#extract the W and p-value from the htest object 
shapiro_results_df <- data.frame(W = sapply(shapiro_results, function(x) x$statistic),
                                  p.value = sapply(shapiro_results, function(x) x$p.value), 
                                  technology = tech_levels)

```

```{r}
# load the stats package
library(stats)

perceivedrisk$technology <- factor(perceivedrisk$technology)

# perform the Kruskal-Wallis H-test, the technologies are treated as independent groups and the mean of perceived risk is of interest. 
kruskal_result <- kruskal.test(perceived_risk ~ technology, data = perceivedrisk)

# print the test statistic and p-value
print(kruskal_result)

library(emmeans)
emm_results <- emmeans(kruskal_results, ~ technology)
summary(emm_results, infer = c(T, T))

```

```{r}

library(dplyr)
ggplot(perceivedrisk %>%
  mutate(technology = reorder(technology, perceived_risk, FUN = median)), 
         aes(x = technology, y = perceived_risk)) +
  geom_boxplot(width = 0.5) +
  geom_hline(aes(yintercept = median(perceived_risk)), linetype = "dashed", color = "red") +
  coord_flip()

```

```{r}
library(dplyr)
perceivedrisk %>%
  group_by(technology) %>%
  summarise(
    count = n(),
    mean = mean(perceived_risk, na.rm = TRUE),
    sd = sd(perceived_risk, na.rm = TRUE),
    median = median(perceived_risk, na.rm = TRUE),
    IQR = IQR(perceived_risk, na.rm = TRUE)
  )

```
```{r}

pairwise.wilcox.test(perceivedrisk$perceived_risk, perceivedrisk$technology, p.adjust.method = "BH")

```

```{r}

# visualising the pairwaise wilcox test results. the non significant pairs are grey

library(broom)
result <- pairwise.wilcox.test(perceivedrisk$perceived_risk, perceivedrisk$technology, p.adjust.method = "BH")
tidy_result <- tidy(result)

library(ggplot2)
ggplot(tidy_result, aes(x = group1, y = group2, fill = p.value)) + 
  geom_tile(color = "white") + 
  scale_fill_gradient2(low = "red", high = "white", midpoint = 0.025, 
                       limits = c(0, 0.05), space = "Lab", 
                       name = "P-value") + 
  theme_classic() + 
  labs(x = "Technology 1", y = "Technology 2") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

```{r}
#now do the same for perceived benefit
#null hypothesis - no sognificant difference between perceived benefit from technologies. 
#push to github - figure out what happened! why not pushing anymore yaar 
#

```

```{r, echo=FALSE}
library(dplyr)
library(tidyr)

perceivedbenefit <-joineddf %>% select(starts_with("Ben")) %>%
  gather(key = "technology", value = "perceived_benefit") %>% 
  mutate(technology = gsub("Ben_", "", technology)) %>%
  na.omit()

perceivedbenefit

```

```{r}
# checking assumptions for kruskal wallis test

duplicated(joineddf)
sum(duplicated(joineddf))

# there are no duplicate responses in our dataframe :).

#what was I doing here - check again ? assumptions for running it on perceived benefit

ggplot(perceivedbenefit, aes(x = perceived_benefit, y = technology)) + 
  geom_boxplot()

levels(perceivedbenefit$technology)

bartlett.test(perceived_benefit ~ technology, data = perceivedbenefit)

str(perceivedbenefit)

```

```{r}
# load the stats package
library(stats)

perceivedbenefit$technology <- factor(perceivedbenefit$technology)

# perform the Kruskal-Wallis H-test, the technologies are treated as independent groups and the mean of perceived risk is of interest. 
kruskal_resultben <- kruskal.test(perceived_benefit ~ technology, data = perceivedbenefit)

# print the test statistic and p-value
print(kruskal_resultben)

```

```{r}

library(dplyr)
ggplot(perceivedbenefit %>%
  mutate(technology = reorder(technology, perceived_benefit, FUN = median)), 
         aes(x = technology, y = perceived_benefit)) +
  geom_boxplot(width = 0.5) +
  geom_hline(aes(yintercept = median(perceived_benefit)), linetype = "dashed", color = "red") +
  coord_flip()

```
```{r}
library(dplyr)
perceivedbenefit %>%
  group_by(technology) %>%
  summarise(
    count = n(),
    mean = mean(perceived_benefit, na.rm = TRUE),
    sd = sd(perceived_benefit, na.rm = TRUE),
    median = median(perceived_benefit, na.rm = TRUE),
    IQR = IQR(perceived_benefit, na.rm = TRUE)
  )

```

```{r}

pairwise.wilcox.test(perceivedbenefit$perceived_benefit, perceivedbenefit$technology, p.adjust.method = "BH")

```

```{r}

# visualising the pairwaise wilcox test results. the non significant pairs are grey

library(broom)
result <- pairwise.wilcox.test(perceivedbenefit$perceived_benefit, perceivedbenefit$technology, p.adjust.method = "BH")
tidy_result <- tidy(result)

library(ggplot2)
ggplot(tidy_result, aes(x = group1, y = group2, fill = p.value)) + 
  geom_tile(color = "white") + 
  scale_fill_gradient2(low = "green", high = "white", midpoint = 0.025, 
                       limits = c(0, 0.05), space = "Lab", 
                       name = "P-value") + 
  theme_classic() + 
  labs(x = "Technology 1", y = "Technology 2") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```