### 2023-02-10 combined data set/demovariablesregression.Rmd
mean to median in graphs

---
title: "demographicvariablesregression"

output:
  html_document:
    fig_width: 9
    fig_height: 6
    fig_caption: yes
    keep_md: yes
    theme: paper
    df_print: paged
  pdf_document:

    toc: yes
    number_sections: yes
    toc_depth: 2
    fig_width: 9
    fig_height: 6
    fig_caption: yes
---

```{r load packages, warning = FALSE, message = FALSE, echo=FALSE}
  knitr::opts_chunk$set(warning = FALSE, message = FALSE)

suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(readxl))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(tidyr))
suppressPackageStartupMessages(library(ggplot2))

```

```{r, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
moduletwo <- read_excel(here::here("thesis_surveydata/political_factors_complete.xlsx"))
module2 <- data.frame(moduletwo)
module2[module2 == "Rather not say/ Don't know"] <- "Rather not say/Don't know"

# loaded data set for module 2 eco -pol variables and removes the rows with question number and question text
```

```{r, eval =TRUE, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
#This chunk contains the coding schemes of various scales used in survey one: eco-political factors, kahan scale and acceptance scale
codedmodule2 <- module2 %>%

#remove row 1
  filter(!row_number() %in% c(1,2)) %>%

# replace risky likert scale with numbers
  mutate_at(vars(starts_with("Risky")), funs(case_when(. =="Not at all risky" ~ 1, 
                                                       . =="Slightly risky" ~ 2, 
                                                       . =="Moderately risky" ~ 3, 
                                                       . =="Very risky" ~ 4, 
                                                       . =="Extremely risky" ~ 5))) %>%

# replace beneficial likert scale with numbers  
  mutate_at(vars(starts_with("Ben")), funs(case_when(. =="Not at all beneficial" ~ 1,
                                                     . =="Slightly beneficial" ~ 2,
                                                     . =="Moderately beneficial" ~ 3,
                                                     . =="Very beneficial" ~ 4,
                                                     . =="Extremely beneficial" ~ 5 ))) %>%

# replace nuclear acceptance likert scale with numbers
  mutate_at(vars(N_accept,N_reluctantlyaccept,N_reject), funs(case_when(. == "Strongly disagree" ~ 1, 
                                                                        . == "Somewhat disagree" ~ 2,
                                                                        . == "Neither agree nor disagree" ~3,
                                                                        . == "Somewhat agree" ~ 4,
                                                                        . == "Strongly agree" ~ 5))) %>%

# code likert scale for variables for Kahan scale into numbers
  mutate_at(vars(starts_with (c("K_I","K_H","DISPLACE", "POLLUTE", "HEALTH", "JOBS", "BEAUTY", "PRIDE", "NPRIDE", "DEV", "PROSPER", "RELY"))), funs(case_when(. == "Strongly disagree" ~ 1, 
               . == "Somewhat disagree" ~ 2,
               . == "Neither agree nor disagree" ~3,
               . == "Somewhat agree" ~ 4,
               . == "Strongly agree" ~ 5))) %>%

# reverse code for likert scale for variables for Kahan scale into numbers
  mutate_at(vars(starts_with (c("K_S","K_E"))), funs(case_when(. == "Strongly disagree" ~ 5, 
                                                               . == "Somewhat disagree" ~ 4,
                                                               . == "Neither agree nor disagree" ~3,
                                                               . == "Somewhat agree" ~ 2,
                                                               . == "Strongly agree" ~ 1))) %>%

# code eco-pol scale variables into numbers
  mutate_at(vars(SYSTEMDEMO,SYSTEMRELIGION,SYSTEMTECHNO,SYSTEMTOTAL,WEALTHLIM,MECHANISATION,DECISIONDECEN,INDUSTRYLARGE,ECONOMYGLOBAL,OWNERPVT,OWNERNOREG), funs(case_when(. == "Strongly disagree" ~ 1, 
                        . == "Somewhat disagree" ~ 2,
                        . == "Neither agree nor disagree" ~3,
                        . == "Somewhat agree" ~ 4,
                        . == "Strongly agree" ~ 5))) %>%

# reverse code eco-pol scale variables into numbers 
 mutate_at(vars(DECISIONCEN, INDUSTRYSMALL, ECONOMYLOCAL, ENVOVERDEV,OWNERPUB, OWNERREG), funs(case_when(. == "Strongly disagree" ~ 5, 
                                                                                                         . == "Somewhat disagree" ~ 4,
                                                                                                         . == "Neither agree nor disagree" ~3,
                                                                                                         . == "Somewhat agree" ~ 2,
                                                                                                         . == "Strongly agree" ~ 1)))

```

```{r, eval=TRUE, echo=FALSE}

moduleone <- read_excel(here::here("thesis_surveydata/livelihood_insecurity_complete.xlsx"))
module1 <- data.frame(moduleone)

# loaded data set for module 2 eco -pol variables and removes the rows with question number and question text
```

```{r, eval =TRUE, echo=FALSE}
#This chunk contains the coding schemes of various scales used in survey one: eco-political factors, kahan scale and acceptance scale
codedmodule1 <- module1 %>%

#remove row 1
  filter(!row_number() %in% c(1,2)) %>%

# replace risky likert scale with numbers
  mutate_at(vars(starts_with("Risky")), funs(case_when(. =="Not at all risky" ~ 1, 
                                                       . =="Slightly risky" ~ 2, 
                                                       . =="Moderately risky" ~ 3, 
                                                       . =="Very risky" ~ 4, 
                                                       . =="Extremely risky" ~ 5))) %>%

# replace beneficial likert scale with numbers  
  mutate_at(vars(starts_with("Ben")), funs(case_when(. =="Not at all beneficial" ~ 1,
                                                     . =="Slightly beneficial" ~ 2,
                                                     . =="Moderately beneficial" ~ 3,
                                                     . =="Very beneficial" ~ 4,
                                                     . =="Extremely beneficial" ~ 5 ))) %>%

# replace nuclear acceptance likert scale with numbers
  mutate_at(vars(N_accept,N_reluctantlyaccept,N_reject), funs(case_when(. == "Strongly disagree" ~ 1, 
                                                                        . == "Somewhat disagree" ~ 2,
                                                                        . == "Neither agree nor disagree" ~3,
                                                                        . == "Somewhat agree" ~ 4,
                                                                        . == "Strongly agree" ~ 5)))

```

```{r, echo = FALSE}

suppressPackageStartupMessages(library(tidyselect))
# combine codedmodule1 and codedmodule2

numcol2 <- codedmodule2 %>% add_column(Respondent = 1:1099, .before = "Language") 
numcol1 <- codedmodule1 %>% add_column(Respondent = 1099:2160, .before = "Language")

joineddf <-  full_join(numcol2, numcol1, by = c("Respondent", "Language", "Survey_Date", "State", "Risky_Hydro", "Risky_Solar", "Risky_Wind", "Risky_Nuclear", "Risky_Coal", "Risky_Gas", "Risky_Oil", "Risky_INDHydro", "Risky_INDSolar", "Risky_INDWind", "Risky_INDBiogas", "Risky_INDDiesel", "Risky_INDKerosene", "Risky_INDFirewoodetc", "Risky_INDLPG", "Ben_Hydro", "Ben_Solar", "Ben_Wind", "Ben_Nuclear", "Ben_Coal", "Ben_Gas", "Ben_Oil", "Ben_INDHydro", "Ben_INDSolar", "Ben_INDWind", "Ben_INDDiesel", "Ben_INDBiogas", "Ben_INDFirewoodetc", "Ben_INDLPG", "Ben_INDKerosene", "N_accept", "N_reluctantlyaccept", "N_reject", "age", "urban_rural", "gender", "caste", "religion", "religiontext"))

summary(joineddf)

```

```{r}
library(dplyr)

demovars <- joineddf %>%
  select(starts_with("Risky"), starts_with("Ben"),urban_rural, gender, caste, religion, age, State)%>%
  mutate(religion = ifelse(religion == "Other (Please mention)", "Other", religion))

summary(demovars)
table(demovars$gender)

```

```{r}
library(dplyr)

demovars_frisk <- demovars %>%
  select(starts_with("Risky"), gender) %>%
  filter(gender == "Female")%>%
  na.omit()

demovars_mrisk <- demovars %>%
  select(starts_with("Risky"), gender) %>%
  filter(gender == "Male") %>%
  na.omit()

sum_frisk <- demovars_frisk %>% 
    summarise(across(starts_with("Risky"), list(mean = mean, sd = sd, median = median, n = nrow))) %>%
    pivot_longer(cols = starts_with("Risky"), 
                                          names_to = c("Risky", "Technology", "stat"), 
                                          names_sep =  "_",
                                          values_to = "Value")%>%
    select(-Risky) %>%
    pivot_wider(names_from = stat, values_from = Value)%>%
    mutate(gender = "f")

sum_mrisk <- demovars_mrisk %>% 
    summarise(across(starts_with("Risky"), list(mean = mean, sd = sd, median = median, n = nrow))) %>%
    pivot_longer(cols = starts_with("Risky"), 
                                          names_to = c("Risky", "Technology", "stat"), 
                                          names_sep =  "_",
                                          values_to = "Value")%>%

    select(-Risky) %>%
    pivot_wider(names_from = stat, values_from = Value)%>%
    mutate(gender = "m")

```

```{r}
library(ggplot2)

# Combine the two datasets
sum_gender <- rbind(sum_frisk, sum_mrisk)

# Plot

p1 <- sum_gender%>% 
  filter(Technology %in% c("Wind","Nuclear","Solar","Oil","Hydro","Gas","Coal")) %>%
  ggplot(aes(x=Technology, y= median, fill = gender))+
  geom_bar(stat="identity", position=position_dodge())+
  #geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=0.2, position=position_dodge(0.9)) +
  ylim(0,5)+
  scale_fill_manual(values=c("red", "blue"), name="legend")+
  labs(y = "Median perceivd risk", x = "Energy Technologies")+
  ggtitle("Perceived risk and benefit from centrally owned energy technologies by gender")+
  theme(axis.text.x = element_text(angle=90, hjust=1, vjust=0.5))

p2 <- sum_gender%>% 
  filter(Technology %in% c("INDHydro", "INDSolar", "INDWind", "INDBiogas","INDDiesel", "INDKerosene","INDFirewoodetc","INDLPG")) %>%
  ggplot(aes(x=Technology, y= median, fill = gender))+
  geom_bar(stat="identity", position=position_dodge())+
  #geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=0.2, position=position_dodge(0.9)) +
  ylim(0,5)+
  scale_fill_manual(values=c("red", "blue"), name="legend")+
  labs(y = "Median perceived risk", x = "Energy Technologies")+
  ggtitle("Perceived risk and benefit from individually owned energy technologies by gender")+
  theme(axis.text.x = element_text(angle=90, hjust=1, vjust=0.5))

library(gridExtra)
grid.arrange(p1, p2, ncol = 2)

```

```{r}

# conducting non-parametric procedure : "Wilcoxon rank-sum test" to see how perception of risk may differ according to gender. 
library(stats)
library(dplyr)
library(purrr)
library(broom)

male_risk <- demovars[demovars$gender == "Male",]$Risky_Nuclear
female_risk <- demovars[demovars$gender == "Female",]$Risky_Nuclear

wilcox.test(male_risk, female_risk, alternative = "two.sided")

wilcox_gender <- demovars %>% 
  filter(gender %in% c("Male", "Female"))%>%
  select(starts_with("Risky")) %>% 
  lapply(function(col) wilcox.test(col ~ gender, data = demovars))

wilcox_gender <- do.call(rbind, wilcox_gender)

wilcox_gender

```
```{r}

# for making a data frame from wilcox test
#wilcox_gender <- demovars %>% 
  #filter(gender %in% c("Male", "Female"))%>%
  #select(starts_with("Risky")) %>% 
  #lapply(function(col) broom::tidy(wilcox.test(col ~ gender, data = demovars)))

#wilcox_gender <- do.call(rbind, wilcox_gender)

#wilcox_gender$variable <- rownames(wilcox_gender)

#wilcox_gender <- as.data.frame(wilcox_gender)
```