### 2023-02-15 combined data set/demovariablesregression.Rmd
wilcoxon rank sum test on caste and urban_rural

---
title: "demographicvariablesregression"

output:
  html_document:
    fig_width: 9
    fig_height: 6
    fig_caption: yes
    keep_md: yes
    theme: paper
    df_print: paged
  pdf_document:

    toc: yes
    number_sections: yes
    toc_depth: 2
    fig_width: 9
    fig_height: 6
    fig_caption: yes
---

```{r load packages, warning = FALSE, message = FALSE, echo=FALSE}
  knitr::opts_chunk$set(warning = FALSE, message = FALSE)

suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(readxl))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(tidyr))
suppressPackageStartupMessages(library(ggplot2))

```

```{r, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
moduletwo <- read_excel(here::here("thesis_surveydata/political_factors_complete.xlsx"))
module2 <- data.frame(moduletwo)
module2[module2 == "Rather not say/ Don't know"] <- "Rather not say/Don't know"

# loaded data set for module 2 eco -pol variables and removes the rows with question number and question text
```

```{r, eval =TRUE, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
#This chunk contains the coding schemes of various scales used in survey one: eco-political factors, kahan scale and acceptance scale
codedmodule2 <- module2 %>%

#remove row 1
  filter(!row_number() %in% c(1,2)) %>%

# replace risky likert scale with numbers
  mutate_at(vars(starts_with("Risky")), funs(case_when(. =="Not at all risky" ~ 1, 
                                                       . =="Slightly risky" ~ 2, 
                                                       . =="Moderately risky" ~ 3, 
                                                       . =="Very risky" ~ 4, 
                                                       . =="Extremely risky" ~ 5))) %>%

# replace beneficial likert scale with numbers  
  mutate_at(vars(starts_with("Ben")), funs(case_when(. =="Not at all beneficial" ~ 1,
                                                     . =="Slightly beneficial" ~ 2,
                                                     . =="Moderately beneficial" ~ 3,
                                                     . =="Very beneficial" ~ 4,
                                                     . =="Extremely beneficial" ~ 5 ))) %>%

# replace nuclear acceptance likert scale with numbers
  mutate_at(vars(N_accept,N_reluctantlyaccept,N_reject), funs(case_when(. == "Strongly disagree" ~ 1, 
                                                                        . == "Somewhat disagree" ~ 2,
                                                                        . == "Neither agree nor disagree" ~3,
                                                                        . == "Somewhat agree" ~ 4,
                                                                        . == "Strongly agree" ~ 5))) %>%

# code likert scale for variables for Kahan scale into numbers
  mutate_at(vars(starts_with (c("K_I","K_H","DISPLACE", "POLLUTE", "HEALTH", "JOBS", "BEAUTY", "PRIDE", "NPRIDE", "DEV", "PROSPER", "RELY"))), funs(case_when(. == "Strongly disagree" ~ 1, 
               . == "Somewhat disagree" ~ 2,
               . == "Neither agree nor disagree" ~3,
               . == "Somewhat agree" ~ 4,
               . == "Strongly agree" ~ 5))) %>%

# reverse code for likert scale for variables for Kahan scale into numbers
  mutate_at(vars(starts_with (c("K_S","K_E"))), funs(case_when(. == "Strongly disagree" ~ 5, 
                                                               . == "Somewhat disagree" ~ 4,
                                                               . == "Neither agree nor disagree" ~3,
                                                               . == "Somewhat agree" ~ 2,
                                                               . == "Strongly agree" ~ 1))) %>%

# code eco-pol scale variables into numbers
  mutate_at(vars(SYSTEMDEMO,SYSTEMRELIGION,SYSTEMTECHNO,SYSTEMTOTAL,WEALTHLIM,MECHANISATION,DECISIONDECEN,INDUSTRYLARGE,ECONOMYGLOBAL,OWNERPVT,OWNERNOREG), funs(case_when(. == "Strongly disagree" ~ 1, 
                        . == "Somewhat disagree" ~ 2,
                        . == "Neither agree nor disagree" ~3,
                        . == "Somewhat agree" ~ 4,
                        . == "Strongly agree" ~ 5))) %>%

# reverse code eco-pol scale variables into numbers 
 mutate_at(vars(DECISIONCEN, INDUSTRYSMALL, ECONOMYLOCAL, ENVOVERDEV,OWNERPUB, OWNERREG), funs(case_when(. == "Strongly disagree" ~ 5, 
                                                                                                         . == "Somewhat disagree" ~ 4,
                                                                                                         . == "Neither agree nor disagree" ~3,
                                                                                                         . == "Somewhat agree" ~ 2,
                                                                                                         . == "Strongly agree" ~ 1)))

```

```{r, eval=TRUE, echo=FALSE}

moduleone <- read_excel(here::here("thesis_surveydata/livelihood_insecurity_complete.xlsx"))
module1 <- data.frame(moduleone)

# loaded data set for module 2 eco -pol variables and removes the rows with question number and question text
```

```{r, eval =TRUE, echo=FALSE}
#This chunk contains the coding schemes of various scales used in survey one: eco-political factors, kahan scale and acceptance scale
codedmodule1 <- module1 %>%

#remove row 1
  filter(!row_number() %in% c(1,2)) %>%

# replace risky likert scale with numbers
  mutate_at(vars(starts_with("Risky")), funs(case_when(. =="Not at all risky" ~ 1, 
                                                       . =="Slightly risky" ~ 2, 
                                                       . =="Moderately risky" ~ 3, 
                                                       . =="Very risky" ~ 4, 
                                                       . =="Extremely risky" ~ 5))) %>%

# replace beneficial likert scale with numbers  
  mutate_at(vars(starts_with("Ben")), funs(case_when(. =="Not at all beneficial" ~ 1,
                                                     . =="Slightly beneficial" ~ 2,
                                                     . =="Moderately beneficial" ~ 3,
                                                     . =="Very beneficial" ~ 4,
                                                     . =="Extremely beneficial" ~ 5 ))) %>%

# replace nuclear acceptance likert scale with numbers
  mutate_at(vars(N_accept,N_reluctantlyaccept,N_reject), funs(case_when(. == "Strongly disagree" ~ 1, 
                                                                        . == "Somewhat disagree" ~ 2,
                                                                        . == "Neither agree nor disagree" ~3,
                                                                        . == "Somewhat agree" ~ 4,
                                                                        . == "Strongly agree" ~ 5)))

```

```{r, echo = FALSE}

suppressPackageStartupMessages(library(tidyselect))
# combine codedmodule1 and codedmodule2

numcol2 <- codedmodule2 %>% add_column(Respondent = 1:1099, .before = "Language") 
numcol1 <- codedmodule1 %>% add_column(Respondent = 1099:2160, .before = "Language")

joineddf <-  full_join(numcol2, numcol1, by = c("Respondent", "Language", "Survey_Date", "State", "Risky_Hydro", "Risky_Solar", "Risky_Wind", "Risky_Nuclear", "Risky_Coal", "Risky_Gas", "Risky_Oil", "Risky_INDHydro", "Risky_INDSolar", "Risky_INDWind", "Risky_INDBiogas", "Risky_INDDiesel", "Risky_INDKerosene", "Risky_INDFirewoodetc", "Risky_INDLPG", "Ben_Hydro", "Ben_Solar", "Ben_Wind", "Ben_Nuclear", "Ben_Coal", "Ben_Gas", "Ben_Oil", "Ben_INDHydro", "Ben_INDSolar", "Ben_INDWind", "Ben_INDDiesel", "Ben_INDBiogas", "Ben_INDFirewoodetc", "Ben_INDLPG", "Ben_INDKerosene", "N_accept", "N_reluctantlyaccept", "N_reject", "age", "urban_rural", "gender", "caste", "religion", "religiontext"))

```

```{r}
library(dplyr)

demovars <- joineddf %>%
  select(starts_with("Risky"), starts_with("Ben"),urban_rural, gender, caste, religion, age, State)%>%
  mutate(religion = ifelse(religion == "Other (Please mention)", "Other", religion))

```

```{r}
library(dplyr)

demovars_frisk <- demovars %>%
  select(starts_with("Risky"), gender) %>%
  filter(gender == "Female")%>%
  na.omit()

demovars_mrisk <- demovars %>%
  select(starts_with("Risky"), gender) %>%
  filter(gender == "Male") %>%
  na.omit()

sum_frisk <- demovars_frisk %>% 
    summarise(across(starts_with("Risky"), list(mean = mean, sd = sd, median = median, n = nrow))) %>%
    pivot_longer(cols = starts_with("Risky"), 
                                          names_to = c("Risky", "Technology", "stat"), 
                                          names_sep =  "_",
                                          values_to = "Value")%>%
    select(-Risky) %>%
    pivot_wider(names_from = stat, values_from = Value)%>%
    mutate(gender = "f")

sum_mrisk <- demovars_mrisk %>% 
    summarise(across(starts_with("Risky"), list(mean = mean, sd = sd, median = median, n = nrow))) %>%
    pivot_longer(cols = starts_with("Risky"), 
                                          names_to = c("Risky", "Technology", "stat"), 
                                          names_sep =  "_",
                                          values_to = "Value")%>%

    select(-Risky) %>%
    pivot_wider(names_from = stat, values_from = Value)%>%
    mutate(gender = "m")

```

```{r}
library(ggplot2)

# Combine the two datasets
sum_gender <- rbind(sum_frisk, sum_mrisk)

# Plot

p1 <- sum_gender%>% 
  filter(Technology %in% c("Wind","Nuclear","Solar","Oil","Hydro","Gas","Coal")) %>%
  ggplot(aes(x=Technology, y= median, fill = gender))+
  geom_bar(stat="identity", position=position_dodge())+
  #geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=0.2, position=position_dodge(0.9)) +
  ylim(0,5)+
  scale_fill_manual(values=c("red", "blue"), name="legend")+
  labs(y = "Median perceivd risk", x = "Energy Technologies")+
  ggtitle("Perceived risk and benefit from centrally owned energy technologies by gender")+
  theme(axis.text.x = element_text(angle=90, hjust=1, vjust=0.5))+
  annotate("text", x=c("Solar", "Nuclear", "Coal"), y=4.8, label="*", size=10, fontface="bold") +
  theme(axis.text.x = element_text(angle=90, hjust=1, vjust=0.5, size=12, margin=margin(l=50)))

p2 <- sum_gender%>% 
  filter(Technology %in% c("INDHydro", "INDSolar", "INDWind", "INDBiogas","INDDiesel", "INDKerosene","INDFirewoodetc","INDLPG")) %>%
  ggplot(aes(x=Technology, y= median, fill = gender))+
  geom_bar(stat="identity", position=position_dodge())+
  #geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=0.2, position=position_dodge(0.9)) +
  ylim(0,5)+
  scale_fill_manual(values=c("red", "blue"), name="legend")+
  labs(y = "Median perceived risk", x = "Energy Technologies")+
  ggtitle("Perceived risk and benefit from individually owned energy technologies by gender")+
  theme(axis.text.x = element_text(angle=90, hjust=1, vjust=0.5)) +
  annotate("text", x=c("INDSolar", "INDBiogas", "INDDiesel"), y=4.8, label="*", size=10, fontface="bold") +
  theme(axis.text.x = element_text(angle=90, hjust=1, vjust=0.5, size=12, margin=margin(l=50)))

library(gridExtra)
grid.arrange(p1, p2, ncol = 2, widths = NULL)

```

```{r}

# conducting non-parametric procedure : "Wilcoxon rank-sum test" to see how perception of risk may differ according to gender.

library(broom)
library(stats)
library(dplyr)

male_risk<- demovars[demovars$gender == "Male",]$Risky_Nuclear 
female_risk <- demovars[demovars$gender == "Female",]$Risky_Nuclear

wilcox.test(male_risk, female_risk, alternative = "two.sided")

Cmale_risk<- demovars[demovars$gender == "Male",]$Risky_Coal 
Cfemale_risk <- demovars[demovars$gender == "Female",]$Risky_Coal

wilcox.test(Cmale_risk, Cfemale_risk, alternative = "two.sided")

demovars_mf <- demovars %>% 
  filter(gender %in% c("Male", "Female"))

wilcox_gender <- demovars_mf %>%
  select(starts_with("Risky")) %>%
  lapply(function(col) wilcox.test(col ~ gender, data = demovars_mf))

wilcox_mf <- do.call(rbind, wilcox_gender)

# results from this are used to mark significant differences in the bar graph above.

```

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)

demovars_caste_long <- demovars_caste %>%
  pivot_longer(cols = starts_with("Risky_"), names_to = "technology", values_to = "perceived_risk") %>%
  mutate(technology = gsub("Risky_", "", technology)) %>%
        mutate(caste = ifelse(grepl("SC|Scheduled caste", caste), "SC", caste),
              caste = ifelse(grepl("ST|Scheduled Tribe", caste), "ST", caste),
              caste = ifelse(grepl("OBC|Other Backward Classes", caste), "OBC", caste))

```

```{r}

library(ggplot2)

# Subset dataframe with IND prefix
demovars_caste_long_ind <- demovars_caste_long %>% 
  filter(str_starts(technology, "IND"))%>%
   group_by(caste, technology) %>% 
  summarize(median_perceived_risk = median(perceived_risk))

table(demovars_caste_long_ind)

# Create bar graph with IND prefix 
ggplot(demovars_caste_long_ind, aes(x = technology, y = median_perceived_risk, fill = caste)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Median Perceived Risk by Caste Groups for Technologies with IND Prefix",
       x = "Technology", y = "Median Perceived Risk") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) # Rotate x-axis labels

# Subset dataframe without IND prefix
demovars_caste_long_not_ind <- demovars_caste_long %>% 
  filter(!str_starts(technology, "IND"))%>%
   group_by(caste, technology) %>% 
  summarize(median_perceived_risk = median(perceived_risk))

# Create bar graph without IND prefix

ggplot(demovars_caste_long_not_ind, aes(x = technology, y = median_perceived_risk, fill = caste)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Median Perceived Risk by Caste Groups for Technologies without IND Prefix",
       x = "Technology", y = "Median Perceived Risk") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) # Rotate x-axis labels

```
```{r}

# print the head of the file
head(demovars_caste_long)

# Show the group/caste levels

demovars_caste_long$caste <- factor(demovars_caste_long$caste)
levels(demovars_caste_long$caste)

# compute summary statistic by groups 
library(dplyr)
summary_caste <- group_by(demovars_caste_long, caste) %>%
  summarise(
    count = n(),
    mean = mean(perceived_risk, na.rm = TRUE),
    sd = sd(perceived_risk, na.rm = TRUE),
    median = median(perceived_risk, na.rm = TRUE),
    IQR = IQR(perceived_risk, na.rm = TRUE)
  )

summary(summary_caste)

library(ggplot2)

# Create the box plot with median values

```

```{r}

demovars_caste_long$technology <- factor(demovars_caste_long$technology)
levels(demovars_caste_long$technology)

kruskal.test(perceived_risk ~ caste, data = demovars_caste_long)

pairwise.wilcox.test(demovars_caste_long$perceived_risk, demovars_caste_long$caste,
                 p.adjust.method = "BH")

library(dplyr)
#for Nuclear
Nuclear_caste <- demovars_caste_long %>%
  filter(technology == "Nuclear")

pairwise.wilcox.test(Nuclear_caste$perceived_risk, Nuclear_caste$caste,
                 p.adjust.method = "BH")

# For Coal
Coal_caste <- demovars_caste_long %>%
  filter(technology == "Coal")
pairwise.wilcox.test(Coal_caste$perceived_risk, Coal_caste$caste, p.adjust.method = "BH")

# For Gas
Gas_caste <- demovars_caste_long %>%
  filter(technology == "Gas")
pairwise.wilcox.test(Gas_caste$perceived_risk, Gas_caste$caste, p.adjust.method = "BH")

# For Hydro
Hydro_caste <- demovars_caste_long %>%
  filter(technology == "Hydro")
pairwise.wilcox.test(Hydro_caste$perceived_risk, Hydro_caste$caste, p.adjust.method = "BH")

# For INDBiogas
INDBiogas_caste <- demovars_caste_long %>%
  filter(technology == "INDBiogas")
pairwise.wilcox.test(INDBiogas_caste$perceived_risk, INDBiogas_caste$caste, p.adjust.method = "BH")

# For INDDiesel
INDDiesel_caste <- demovars_caste_long %>%
  filter(technology == "INDDiesel")
pairwise.wilcox.test(INDDiesel_caste$perceived_risk, INDDiesel_caste$caste, p.adjust.method = "BH")

# For INDFirewoodetc
INDFirewoodetc_caste <- demovars_caste_long %>%
  filter(technology == "INDFirewoodetc")
pairwise.wilcox.test(INDFirewoodetc_caste$perceived_risk, INDFirewoodetc_caste$caste, p.adjust.method = "BH")

# For INDHydro
INDHydro_caste <- demovars_caste_long %>%
  filter(technology == "INDHydro")
pairwise.wilcox.test(INDHydro_caste$perceived_risk, INDHydro_caste$caste, p.adjust.method = "BH")

# For INDKerosene
INDKerosene_caste <- demovars_caste_long %>%
  filter(technology == "INDKerosene")
pairwise.wilcox.test(INDKerosene_caste$perceived_risk, INDKerosene_caste$caste, p.adjust.method = "BH")

# For INDLPG
INDLPG_caste <- demovars_caste_long %>%
  filter(technology == "INDLPG")
pairwise.wilcox.test(INDLPG_caste$perceived_risk, INDLPG_caste$caste, p.adjust.method = "BH")

# For INDSolar
INDSolar_caste <- demovars_caste_long %>%
  filter(technology == "INDSolar")
pairwise.wilcox.test(INDSolar_caste$perceived_risk, INDSolar_caste$caste, p.adjust.method = "BH")

# For INDWind
INDWind_caste <- demovars_caste_long %>%
  filter(technology == "INDWind")
pairwise.wilcox.test(INDWind_caste$perceived_risk, INDWind_caste$caste, p.adjust.method = "BH")

# For Oil
Oil_caste <- demovars_caste_long %>%
  filter(technology == "Oil")
pairwise.wilcox.test(Oil_caste$perceived_risk, Oil_caste$caste, p.adjust.method = "BH")

# For Solar
Solar_caste <- demovars_caste_long %>%
  filter(technology == "Solar")
pairwise.wilcox.test(Solar_caste$perceived_risk, Solar_caste$caste, p.adjust.method = "BH")

# For Wind
Wind_caste <- demovars_caste_long %>%
  filter(technology == "Wind")
pairwise.wilcox.test(Wind_caste$perceived_risk, Wind_caste$caste, p.adjust.method = "BH")

#not finding any significant differences in the medians

```
```{r}

demovars_urban_rural <-  demovars %>%
  select(starts_with("Risky"), urban_rural) %>%
  pivot_longer(cols = starts_with("Risky_"), names_to = "technology", values_to = "perceived_risk") %>%
  mutate(technology = gsub("Risky_", "", technology)) %>%
  na.omit()

pairwise.wilcox.test(demovars_urban_rural$perceived_risk, demovars_urban_rural$urban_rural, p.adjust.method = "BH")

```