### 2023-09-20 LPGversusFirewood.Rmd

---
title: "LPG v/s Firewood"
output:
  pdf_document:
    toc: yes
    toc_depth: 2
    keep_tex: yes
  header-includes:
  - \usepackage{dcolumn}
  - \usepackage{array}
  keep_md: yes

---

```{r load packages, warning = FALSE, message = FALSE, echo = FALSE}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(readxl))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(psych))
suppressPackageStartupMessages(library(tidyr))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(lavaan))
suppressPackageStartupMessages(library(tsibble))
suppressPackageStartupMessages(library(DT))
suppressPackageStartupMessages(library(knitr))

```

```{r setup, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

moduletwo <- read_excel(here::here("thesis_surveydata/political_factors_complete.xlsx"))
module2 <- data.frame(moduletwo)
module2[module2 == "Rather not say/ Don't know"] <- "Rather not say/Don't know"

# loaded data set for module 2 eco -pol variables and removes the rows with question number and question text
```

```{r, eval =TRUE, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
#This chunk contains the coding schemes of various scales used in survey one: eco-political factors, kahan scale and acceptance scale
codedmodule2 <- module2 %>%

#remove row 1
  filter(!row_number() %in% c(1,2)) %>%

# replace risky likert scale with numbers
  mutate_at(vars(starts_with("Risky")), funs(case_when(. =="Not at all risky" ~ 1, 
                                                       . =="Slightly risky" ~ 2, 
                                                       . =="Moderately risky" ~ 3, 
                                                       . =="Very risky" ~ 4, 
                                                       . =="Extremely risky" ~ 5))) %>%

# replace beneficial likert scale with numbers  
  mutate_at(vars(starts_with("Ben")), funs(case_when(. =="Not at all beneficial" ~ 1,
                                                     . =="Slightly beneficial" ~ 2,
                                                     . =="Moderately beneficial" ~ 3,
                                                     . =="Very beneficial" ~ 4,
                                                     . =="Extremely beneficial" ~ 5 ))) %>%

# replace nuclear acceptance likert scale with numbers
  mutate_at(vars(N_accept,N_reluctantlyaccept,N_reject), funs(case_when(. == "Strongly disagree" ~ 1, 
                                                                        . == "Somewhat disagree" ~ 2,
                                                                        . == "Neither agree nor disagree" ~3,
                                                                        . == "Somewhat agree" ~ 4,
                                                                        . == "Strongly agree" ~ 5))) %>%

# code likert scale for variables for Kahan scale into numbers
  mutate_at(vars(starts_with (c("K_I","K_H","DISPLACE", "POLLUTE", "HEALTH", "JOBS", "BEAUTY", "PRIDE", "NPRIDE", "DEV", "PROSPER", "RELY"))), funs(case_when(. == "Strongly disagree" ~ 1, 
               . == "Somewhat disagree" ~ 2,
               . == "Neither agree nor disagree" ~3,
               . == "Somewhat agree" ~ 4,
               . == "Strongly agree" ~ 5))) %>%

# reverse code for likert scale for variables for Kahan scale into numbers
  mutate_at(vars(starts_with (c("K_S","K_E"))), funs(case_when(. == "Strongly disagree" ~ 5, 
                                                               . == "Somewhat disagree" ~ 4,
                                                               . == "Neither agree nor disagree" ~3,
                                                               . == "Somewhat agree" ~ 2,
                                                               . == "Strongly agree" ~ 1))) %>%

# code eco-pol scale variables into numbers
  mutate_at(vars(SYSTEMDEMO,SYSTEMRELIGION,SYSTEMTECHNO,SYSTEMTOTAL,WEALTHLIM,MECHANISATION,DECISIONDECEN,INDUSTRYLARGE,ECONOMYGLOBAL,OWNERPVT,OWNERNOREG), funs(case_when(. == "Strongly disagree" ~ 1, 
                        . == "Somewhat disagree" ~ 2,
                        . == "Neither agree nor disagree" ~3,
                        . == "Somewhat agree" ~ 4,
                        . == "Strongly agree" ~ 5))) %>%

# reverse code eco-pol scale variables into numbers 
 mutate_at(vars(DECISIONCEN, INDUSTRYSMALL, ECONOMYLOCAL, ENVOVERDEV,OWNERPUB, OWNERREG), funs(case_when(. == "Strongly disagree" ~ 5, 
                                                                                                         . == "Somewhat disagree" ~ 4,
                                                                                                         . == "Neither agree nor disagree" ~3,
                                                                                                         . == "Somewhat agree" ~ 2,
                                                                                                         . == "Strongly agree" ~ 1))) %>%
  mutate_all(~ifelse(. == "Rather not say/ Don't know", "Rather not say/Don't know", .))%>%
  mutate_all(~ifelse(. == "West bengal", "West Bengal", .))%>%
  mutate_all(~ifelse(. == "WEST BENGAL", "West Bengal", .))%>%
  mutate_all(~ifelse(. == "Up", "Uttar Pradesh", .))%>%
  mutate_all(~ifelse(. == "Tamilnadu", "Tamil Nadu", .))%>%
  mutate_all(~ifelse(. == "UP", "Uttar Pradesh", .))

```

```{r, eval=TRUE, echo = FALSE}

moduleone <- read_excel(here::here("thesis_surveydata/livelihood_insecurity_complete.xlsx"))
save(moduleone, file = "moduleone.rda")
module1 <- data.frame(moduleone)

# loaded data set for module 2 eco -pol variables and removes the rows with question number and question text
```

```{r, echo = FALSE}
#This chunk contains the coding schemes of various scales used in survey one: eco-political factors, kahan scale and acceptance scale
codedmodule1 <- module1 %>%
  #remove row 1 and 2

  filter(!row_number() %in% c(1,2))%>%

  # replace risky likert scale with numbers
  mutate(across(starts_with("Risky"), ~case_when(. =="Not at all risky" ~ 1, 
                                                 . =="Slightly risky" ~ 2, 
                                                 . =="Moderately risky" ~ 3, 
                                                 . =="Very risky" ~ 4, 
                                                 . =="Extremely risky" ~ 5))) %>%

  # replace beneficial likert scale with numbers  
  mutate(across(starts_with("Ben"), ~case_when(. =="Not at all beneficial" ~ 1,
                                               . =="Slightly beneficial" ~ 2,
                                               . =="Moderately beneficial" ~ 3,
                                               . =="Very beneficial" ~ 4,
                                               . =="Extremely beneficial" ~ 5 ))) %>%

  # replace nuclear acceptance likert scale with numbers
  mutate(across(c(N_accept,N_reluctantlyaccept,N_reject), ~case_when(. == "Strongly disagree" ~ 1, 
                                                                     . == "Somewhat disagree" ~ 2,
                                                                     . == "Neither agree nor disagree" ~3,
                                                                     . == "Somewhat agree" ~ 4,
                                                                     . == "Strongly agree" ~ 5))) %>%

  mutate(across(Livelihood_Income, ~case_when(. == "(a) own farming/fishing/animal rearing/business" ~ "LSI", 
                                              . == "(b) employed for wages/salary" ~ "ISI",
                                              . == "(c) student" ~"student",
                                              . == "(d) currently unemployed" ~ "unemployed",
                                              . == "(e) homemaker" ~ "homemaker",
                                              . == "(f) retired" ~ "retired",
                                              . == "(g) unable to work" ~ "unable to work" ))) %>%

  # coding education
  mutate(across(education, ~case_when(. == "No schooling completed" ~ 0, 
                                      . == "Primary school" ~ 1,
                                      . == "Grade 10th" ~ 2,
                                      . == "Grade 12th" ~ 3,
                                      . == "Some college, no degree" ~ 4,
                                      . == "Trade/technical/vocational training" ~ 5,
                                      . == "Bachelor's degree" ~ 6,
                                      . == "Master's degree" ~ 7,
                                      . == "Professional degree" ~ 8,
                                      . == "Doctorate degree" ~ 9 ))) %>%

  # coding yes and no questions
  mutate(across(c(Lassociation,Iassociation, govtscheme, wateravail, polposition, polparty), ~case_when(.== "Yes" ~ 1,
                                                                                                        .== "No" ~ 0))) %>%
  # reverse coding yes and no questions
  mutate(across(c(illness,severeillness), ~case_when(.== "Yes" ~ 0,
                                                     .== "No" ~ 1))) %>%

  # coding How confident are you that you can continue your current livelihood if you want to? 
  mutate(across(Lconfidence, ~case_when(. == "Not at all confident" ~ 1, 
                                        . == "Not much confident" ~ 2,
                                        . == "Neutral" ~3,
                                        . == "Somewhat confident" ~ 4,
                                        . == "Very confident" ~ 5))) %>%

  mutate(across(c(Lboat, houseown), ~case_when(.== "Own it" ~ 1,
                                               .== "Rent it" ~ 0))) %>%

  # J security index coding
  mutate(across(JISRkeep, ~case_when(. == "Strongly disagree" ~ 1, 
                                     . == "Somewhat disagree" ~ 2,
                                     . == "Neither agree nor disagree" ~3,
                                     . == "Somewhat agree" ~ 4,
                                     . == "Strongly agree" ~ 5))) %>%

  # reverse code JIS 
  mutate(across(c(JISlosefuture, JISlose, JISfuture), ~case_when(. == "Strongly disagree" ~ 5, 
                                                                 . == "Somewhat disagree" ~ 4,
                                                                 . == "Neither agree nor disagree" ~3,
                                                                 . == "Somewhat agree" ~ 2,
                                                                 . == "Strongly agree" ~ 1))) %>%

  mutate(across(c(totalincome, otherincome), ~case_when(. == "Less than Rs 1,00,000" ~ 1,
                                                        . == "Rs 1,00,000 to Rs 2,50,000"  ~ 2,
                                                        . == "Rs 2,50,001 to Rs 5,00,000" ~ 3,
                                                        . == "Rs 5,00,001 to Rs 10,00,000" ~ 4,
                                                        . == "Rs 10,00,001 and above"  ~ 5,
                                                        . == "Doesn't apply" ~ 0))) %>%

  mutate(across(c(workknowledge, subsidyknow), ~case_when(. == "Very poor" ~ 1, 
                                                          . == "Poor" ~ 2,
                                                          . == "Fair" ~3,
                                                          . == "Good" ~ 4,
                                                          . == "Very good" ~ 5))) %>%

  mutate(across(c(drinkingwater, cookingwater), ~case_when(. == "Bottled water/ Filtered Water" ~ 1,                                       
                                                           .  == "Piped into dwelling" ~ 2,                                                  
                                                           .  == "Piped to neighbour" ~ 3,                                                   
                                                           .  == "Piped to yard/plot" ~ 4, 
                                                           .  == "Protected well" ~ 5, 
                                                           .  == "Public tap/standpipe" ~6,
                                                           .  == "Surface Water  (River/Dam/Lake/Pond/Stream/Canal/Irrigation Channel)" ~7, 
                                                           .  == "Tanker Truck" ~ 8,
                                                           .  == "Tubewell/borehole" ~ 9,
                                                           .  == "Unprotected Spring" ~ 10))) %>%

  mutate(across(toilet, ~case_when(. == "No facility/bush/field" ~ 1,                                       
                                   .  == "Hanging toilet/hanging latrine" ~ 2,                                             
                                   .  == "Bucket toilet"  ~ 3,                                                   
                                   .  == "Composting toilet"  ~ 4, 
                                   .  == "Pit latrine without slab/open pit" ~ 5, 
                                   .  == "Pit latrine with slab" ~6,
                                   .  == "Ventilated improved pit laterine"  ~7, 
                                   .  == "Flush to somewhere else"  ~ 8,
                                   .  == "Flush to pit latrine"  ~ 9,
                                   .  == "Flush to septic tank" ~ 10,
                                   .  == "Flush to piped sewer system" ~ 11))) %>%

  mutate(across(house, ~case_when(. == "Hut made of mud, bamboo, straw etc. (kutcha house)" ~ 1, 
                                  . == "Permanent single story house (pukka house)" ~ 2,
                                  . == "Apartment/Flat in a multi-story building"  ~3,
                                  . == "Multi-story house" ~ 4))) %>%

  mutate(across(foodtrouble, ~case_when(. == "8 or more months" ~ 1,                                       
                                        .  == "6-8 months"  ~ 2,                                                  
                                        .  == "4-6 months"  ~ 3,                                                   
                                        .  == "2-4 months"  ~ 4, 
                                        .  == "1-2 months" ~ 5, 
                                        .  == "0 months"  ~6))) %>%

  mutate(across(foodadequacy, ~case_when(. == "Extremely inadequate" ~ 1, 
                                         . == "Moderately inadequate" ~ 2,
                                         . == "Slightly inadequate" ~3,
                                         . == "Neither adequate nor inadequate" ~ 4,
                                         . == "Slightly adequate" ~ 5,
                                         . == "Moderately adequate" ~ 6,
                                         . == "Extremely adequate" ~ 7))) %>%

  # separating cooking, vehicle and food cols into separate entries
  separate(col = cookstove, into = c("cook1", "cook2", "cook3", "cook4"), sep = ",") %>%
  separate(col = vehicle, into = c("vehicle1", "vehicle2", "vehicle3", "vehicle4"), sep = ",") %>%
  separate(col = foodsource, into = c("food1", "food2", "food3", "food4"), sep = ",") %>%

  mutate(across(c(cook1, cook2, cook3, cook4), ~case_when(. == "Three Stone Stove/Open Fire" ~ 1, 
                                  . == "Traditional Solid Fuel Stove" ~ 2,
                                  . == "Liquid Fuel Stove" ~3,
                                  . == "Biogas Stove" ~ 4,
                                  . == "Piped Natural Gas Stove" ~ 5,
                                  . == "Liquefied Petroleum Gas (LPG)/Cooking gas stove" ~ 6,
                                  . == "Solar Cooker" ~ 7,
                                  . == "Electric Stove" ~ 7,
                                  is.na(.) ~ 0))) %>%

  mutate(across(c(vehicle1, vehicle2, vehicle3, vehicle4, vehicletext), ~case_when(. == "Bicycle" ~ 1, 
                                    . == "Animal Drawn Cart" ~ 2,
                                    . == "Motorcycle/Scooter" ~3,
                                    . == "Car" ~ 4,
                                    . == "Truck" ~ 4,
                                    . == "Boat with Motor" ~ 4,
                                    . == "Tractor" ~ 4,
                                    . == "E-rickshaw" ~ 1,
                                    . == "Tempo" ~ 3,
                                    . == "Nothing" ~ 0,
                                    . == "Other (Please Specify)" ~ 0,
                                    is.na(.) ~ 0))) %>%

  mutate(across(c(food1, food2, food3,food4), ~case_when(. == "Forest" ~ 1, 
                                  . == "Own Farm" ~ 2,
                                  . == "Market" ~4,
                                  . == "Own livestock" ~ 2,
                                  . == "Subsidised ration shops" ~ 3,
                                  is.na(.) ~ 0)))

```

```{r, echo = FALSE}

suppressPackageStartupMessages(library(tidyselect))
# combine codedmodule1 and codedmodule2

numcol2 <- codedmodule2 %>% add_column(Respondent = 1:1099, .before = "Language") 
numcol1 <- codedmodule1 %>% add_column(Respondent = 1099:2160, .before = "Language")

joineddf <-  full_join(numcol2, numcol1, by = c("Respondent", "Language", "Survey_Date", "State", "Risky_Hydro", "Risky_Solar", "Risky_Wind", "Risky_Nuclear", "Risky_Coal", "Risky_Gas", "Risky_Oil", "Risky_INDHydro", "Risky_INDSolar", "Risky_INDWind", "Risky_INDBiogas", "Risky_INDDiesel", "Risky_INDKerosene", "Risky_INDFirewoodetc", "Risky_INDLPG", "Ben_Hydro", "Ben_Solar", "Ben_Wind", "Ben_Nuclear", "Ben_Coal", "Ben_Gas", "Ben_Oil", "Ben_INDHydro", "Ben_INDSolar", "Ben_INDWind", "Ben_INDDiesel", "Ben_INDBiogas", "Ben_INDFirewoodetc", "Ben_INDLPG", "Ben_INDKerosene", "N_accept", "N_reluctantlyaccept", "N_reject", "age", "urban_rural", "gender", "caste", "religion", "religiontext")) %>%
  mutate_all(~ifelse(. == "Rather not say/ Don't know", "Rather not say/Don't know", .))%>%
  mutate_all(~ifelse(. == "West bengal", "West Bengal", .))%>%
  mutate_all(~ifelse(. == "WEST BENGAL", "West Bengal", .))%>%
  mutate_all(~ifelse(. == "Up", "Uttar Pradesh", .))%>%
  mutate_all(~ifelse(. == "Tamilnadu", "Tamil Nadu", .))%>%
  mutate_all(~ifelse(. == "UP", "Uttar Pradesh", .))

numcolb <- module2 %>%   filter(!row_number() %in% c(1,2))%>% 
                        add_column(Respondent = 1:1099, .before = "Language") 
numcola <- module1 %>%   filter(!row_number() %in% c(1,2)) %>% 
                        add_column(Respondent = 1099:2160, .before = "Language")

uncodedjoineddf <- full_join(numcolb, numcola, by = c("Respondent", "Language", "Survey_Date", "State", "Risky_Hydro", "Risky_Solar", "Risky_Wind", "Risky_Nuclear", "Risky_Coal", "Risky_Gas", "Risky_Oil", "Risky_INDHydro", "Risky_INDSolar", "Risky_INDWind", "Risky_INDBiogas", "Risky_INDDiesel", "Risky_INDKerosene", "Risky_INDFirewoodetc", "Risky_INDLPG", "Ben_Hydro", "Ben_Solar", "Ben_Wind", "Ben_Nuclear", "Ben_Coal", "Ben_Gas", "Ben_Oil", "Ben_INDHydro", "Ben_INDSolar", "Ben_INDWind", "Ben_INDDiesel", "Ben_INDBiogas", "Ben_INDFirewoodetc", "Ben_INDLPG", "Ben_INDKerosene", "N_accept", "N_reluctantlyaccept", "N_reject", "age", "urban_rural", "gender", "caste", "religion", "religiontext")) %>%
  mutate_all(~ifelse(. == "Rather not say/ Don't know", "Rather not say/Don't know", .))%>%
  mutate_all(~ifelse(. == "West bengal", "West Bengal", .))%>%
  mutate_all(~ifelse(. == "WEST BENGAL", "West Bengal", .))%>%
  mutate_all(~ifelse(. == "Up", "Uttar Pradesh", .))%>%
  mutate_all(~ifelse(. == "Tamilnadu", "Tamil Nadu", .))%>%
  mutate_all(~ifelse(. == "UP", "Uttar Pradesh", .))

```
\newpage

# Characteristics of the Sample

The following graph shows that distribution of different demographic variables in our sample of 2,160 from the combined dataset from both surveys.

```{r, echo=FALSE}

library(dplyr)
library(tidyr)

sampledemos <- joineddf %>% dplyr::select("State", "Language", "age", "urban_rural", "gender", "caste" ,"religion") %>%
              mutate_all(~ifelse(. == "Rather not say/ Don't know", "Rather not say/Don't know", .))%>%
              mutate_all(~ifelse(. == "OBC", "Other Backward Classes (OBC)", .))%>%
              mutate_all(~ifelse(. == "SC", "Scheduled caste (SC)", .))%>%
              mutate_all(~ifelse(. == "ST", "Scheduled tribe (ST)", .))%>%
              mutate_all(~ifelse(. == "MAR", "Marathi", .))%>%
              mutate_all(~ifelse(. == "TA", "Tamil", .)) %>%
              mutate_all(~ifelse(. == "EN", "English", .))%>%
              mutate_all(~ifelse(. == "BN", "Bengali", .))%>%
              mutate_all(~ifelse(. == "MAR", "Marathi", .))%>%
              mutate_all(~ifelse(. == "HI", "Hindi", .))%>%
              na.omit()

#changing factor levels for each variable from high to low n

sampledemos$State <- factor(sampledemos$State, levels = c("Maharashtra","Rajasthan", "West Bengal" , "Tamil Nadu", "Uttar Pradesh"))
sampledemos$Language <- factor(sampledemos$Language, levels = c("Hindi","Marathi","Bengali","Tamil"))
sampledemos$age <- factor(sampledemos$age, levels = c( "18-24 years old" ,  "25-34 years old","35-44 years old", "45-54 years old", "55-64 years old","65-74 years old","75 years or older"  ))
sampledemos$urban_rural <- factor(sampledemos$urban_rural, levels = c("Rural","Urban"))
sampledemos$gender <- factor(sampledemos$gender, levels = c("Male","Female"))
sampledemos$caste <- factor(sampledemos$caste, levels = c("Scheduled tribe (ST)", "Scheduled caste (SC)", "Other Backward Classes (OBC)", "General","Brahmin","Rather not say/Don't know"))
sampledemos$religion <- factor(sampledemos$religion, levels = c("Hinduism", "Islam", "Christanity","Buddhism","Agnostism/ Atheism","Other (Please mention)"))

```

```{r, echo=FALSE, fig.width=11, fig.height=14, out.width='80%', out.height='80%'}
# Loading the required libraries
library(tidyverse)
library(gridExtra)
library(scales)

# Function to create bar plot
create_bar_plot <- function(data, variable_name, palette_name) {
  factor_levels <- levels(sampledemos[[variable_name]])
  data %>%
    filter(Variable == variable_name) %>%
    mutate(Category = factor(Category, levels = factor_levels)) %>%
    ggplot(aes(y = "", x = Percentage, fill = Category)) +
    geom_bar(position = "fill", stat = "identity", width = 1) +
    geom_text(aes(label = paste0(round(Percentage, 0), "%")), 
              position = position_fill(vjust = 0.5)) +
    scale_x_continuous(labels = percent) +
    labs(y = NULL, x = "Percentage", fill = variable_name) +
    theme_classic() +
    theme(axis.text.y = element_blank(), strip.text = element_text(size = 10), legend.position = "top") +
    scale_fill_brewer(palette = palette_name)
}

# Gather data to long format and calculate counts and percentages
df_barplot <- sampledemos %>%
  gather(Variable, Category) %>%
  group_by(Variable, Category) %>%
  filter(!is.na(Category)) %>%
  summarise(n = n()) %>%
  mutate(Percentage = n / sum(n) * 100)

# Create separate plots for each variable
plot_state <- create_bar_plot(df_barplot, "State", "Spectral")
plot_language <- create_bar_plot(df_barplot, "Language", "Paired")
plot_age <- create_bar_plot(df_barplot, "age", "Set3")
plot_urban_rural <- create_bar_plot(df_barplot, "urban_rural", "Dark2")
plot_gender <- create_bar_plot(df_barplot, "gender", "Pastel1")
plot_caste <- create_bar_plot(df_barplot, "caste", "Set1")
plot_religion <- create_bar_plot(df_barplot, "religion", "Set2")

# Combine the plots with spacing
combined_plot <- gridExtra::arrangeGrob(plot_state, plot_language, plot_age, plot_urban_rural, plot_gender, plot_caste, plot_religion, ncol = 1, heights = c(1, 1, 1, 1, 1, 1, 1))

# Print the combined plot
grid::grid.newpage()
grid::grid.draw(combined_plot)

```

\newpage
## Individually/Community owned energy technologies (n= 2160)
The percentages are rounded off to whole numbers.

```{r, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

#bar plot for likert responses perceived risk - individually owned technologies 
df_barplot %>% ggplot(aes(fill= fct_rev(factor(Perceived_Risk, levels=c("Extremely risky", "Very risky","Moderately risky","Slightly risky","Not at all risky","Rather not say/Don't know"))), y=n , x=Technology)) + 
  geom_bar(position="fill", stat = "identity")+
  geom_text(data = df_barplot_percent, aes(label=paste0(Percentage, "%"), y=n), position = position_fill(vjust=0.5), size=3) +
  scale_x_discrete(limits = c("INDSolar", "INDFirewoodetc","INDWind", "INDHydro", "INDDiesel" , "INDBiogas","INDKerosene", "INDLPG"))+
  scale_fill_manual("legend", values = c("Extremely risky" = "#8C4646", "Very risky" = "#D9695F", "Moderately risky" = "#F2A679", "Slightly risky" = "#F2D091","Not at all risky" = "#5D8C7B", "Rather not say/Don't know" = "grey94"))+
   labs( y = "Perceived Risk", x ="Individually Owned Energy Technologies")+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

#perceived benefit individually owned technologies

df_barplotben %>% ggplot(aes(fill= factor(Perceived_Benefit, levels=c("Rather not say/Don't know","Not at all beneficial","Slightly beneficial","Moderately beneficial","Very beneficial", "Extremely beneficial")), y=n , x=Technology)) + 
  geom_bar(position="fill",stat="identity")+
   geom_text(data = df_barplotben_percent, aes(label=paste0(Percentage, "%"), y=n), position = position_fill(vjust=0.5), size=3) +
  scale_fill_manual("legend", values = c("Rather not say/Don't know" = "grey94" ,"Not at all beneficial" = "orange","Slightly beneficial"="yellow" ,"Moderately beneficial" = "palegreen","Very beneficial"= "green3", "Extremely beneficial"= "green4"))+
  scale_x_discrete(limits = c( "INDSolar", "INDFirewoodetc" , "INDLPG",  "INDWind", "INDHydro", "INDDiesel" , "INDBiogas","INDKerosene"))+
  labs( y = "Perceived Benefit", x ="Individually Owned Energy Technologies")+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

\newpage

# Perceived risk and Perceived Benefit in comparison

```{r, echo = FALSE}

# exploring LPG and firewood discrepancies 
# first we observe that the  perceived risk from LPG is higher than perceived risk from firewood
joineddf_IND <- joineddf %>%
  dplyr::select(starts_with(c("Ben_IND", "Risky_IND")))

# Melt the data frame to long format
long_formatIND <- joineddf_IND %>% 
  pivot_longer(everything(), names_to = "technology", values_to = "value") %>%
  mutate(risk_or_benefit = if_else(str_detect(technology, "Risky"), "risk", "benefit"),
         technology = str_replace(technology, "Risky_|Ben_", ""))

# Calculate mean and sd for each technology and category (risk or benefit)
summary_dfIND <- long_formatIND %>%
  group_by(technology, risk_or_benefit) %>%
  summarise(mean = mean(value, na.rm = TRUE),
            sd = sd(value, na.rm = TRUE),
            se = sd(value, na.rm = TRUE) / sqrt(n()),
            .groups = "drop")

# Order technologies by mean risk
order_riskIND <- summary_dfIND %>% 
  filter(risk_or_benefit == "risk") %>%
  arrange(mean) %>% 
  pull(technology)

# Change the levels of factor 'technology'
summary_dfIND$technology <- factor(summary_dfIND$technology, levels = order_riskIND)

# Plotting the data
ggplot(summary_dfIND, aes(x = technology, y = mean, fill = risk_or_benefit)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), width = 0.3, position = position_dodge(0.9)) +
  ylim(0, 5) +
  scale_fill_manual("legend", 
                    values = c("benefit" = "palegreen3", "risk" = "firebrick2"), 
                    labels = c("benefit" = "Perceived Benefit", "risk" = "Perceived Risk")) +
  labs(y = "Mean Values", x = "Energy Technologies", fill = "Perceived") +
  ggtitle("Perceived risk and benefit from individually owned energy technologies") +
  coord_flip()
```

\newpage

# T-tests

## Pairwise T-test: Mean perceived risk and  mean perceived benefit (all individually owned energy technologies)

```{r, echo = FALSE, fig.width=14, fig.height=6}
# Now we wanna know if these differences are significant using simple t test

perceivedriskIND <-joineddf %>% select(starts_with(c("Risky_IND"))) %>%
  gather(key = "technology", value = "perceived_risk") %>% 
  mutate(technology = gsub("Risky_", "", technology))

result_ttestIND <- pairwise.t.test(perceivedriskIND$perceived_risk, perceivedriskIND$technology, p.adjust.method = "BH")

# visualising the pairwaise wilcox test results. the non significant pairs are grey

tidy_result_ttestIND <- broom::tidy(result_ttestIND)

heatmap1IND <- ggplot(tidy_result_ttestIND, aes(x = group1, y = group2, fill = p.value)) + 
  geom_tile(color = "white") + 
  scale_fill_gradient2(low = "firebrick2", high = "white", midpoint = 0.025, 
                       limits = c(0, 0.05), space = "Lab", 
                       name = "P-value") + 
  theme_classic() + 
  labs(x = "Technology 1", y = "Technology 2") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
   theme(axis.text.y = element_text(size=12), 
        axis.text.x = element_text(size=12))

## now benIND

perceivedbenIND <-joineddf %>% select(starts_with(c("Ben_IND"))) %>%
  gather(key = "technology", value = "perceived_benefit") %>% 
  mutate(technology = gsub("Ben_", "", technology))

result_ttestINDb <- pairwise.t.test(perceivedbenIND$perceived_benefit, perceivedbenIND$technology, p.adjust.method = "BH")

# visualising the pairwaise t- test results. the non significant pairs are grey

tidy_result_ttestINDb <- broom::tidy(result_ttestINDb)

heatmap2IND <- ggplot(tidy_result_ttestINDb, aes(x = group1, y = group2, fill = p.value)) + 
  geom_tile(color = "white") + 
  scale_fill_gradient2(low = "palegreen3", high = "white", midpoint = 0.025, 
                       limits = c(0, 0.05), space = "Lab", 
                       name = "P-value") + 
  theme_classic() + 
  labs(x = "Technology 1", y = "Technology 2") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
   theme(axis.text.y = element_text(size=12), 
        axis.text.x = element_text(size=12))

library(gridExtra)

grid.arrange(heatmap1IND, heatmap2IND, ncol=2) 
```

## Paired T-test: Comparing mean perceived risk and mean perceived benefit for each technology.

Null hypothesis (H0): The mean difference between perceived risk and perceived benefit from each technology is zero.
All p values are less than 0.05 suggesting that the differences observed in the bargraph are statistically significant.

```{r, echo=FALSE}

# Convert character columns to factors
long_formatIND$technology <- factor(long_formatIND$technology)
long_formatIND$risk_or_benefit <- factor(long_formatIND$risk_or_benefit)

# Perform the t- Test for each technology
p_values <- long_formatIND %>%
  split(.$technology) %>%
  lapply(function(x) t.test(x$value[x$risk_or_benefit == "risk"], x$value[x$risk_or_benefit == "benefit"], paired = TRUE)) %>%
  lapply(function(x) x$p.value)

p_values

```

# Acceptance = Perceived Benefit - Perceived Risk

Graph representing a combined acceptance scale obtained by subtracting perceived risk from perceived benefit for each respondent

```{r, echo = FALSE}

# combining risk and benefit from each energy technology

INDacceptance <- joineddf %>% select(starts_with(c("Risky_IND", "Ben_IND"))) %>%
  na.omit()%>%
  mutate(Biogas = Ben_INDBiogas - Risky_INDBiogas,
         INDHydro = Ben_INDHydro - Risky_INDHydro,
         INDWind = Ben_INDWind - Risky_INDWind,
         INDSolar = Ben_INDSolar - Risky_INDSolar,
         Diesel = Ben_INDDiesel - Risky_INDDiesel,
         Kerosene = Ben_INDKerosene - Risky_INDKerosene,
         Firewood = Ben_INDFirewoodetc - Risky_INDFirewoodetc,
         LPG = Ben_INDLPG - Risky_INDLPG)%>%
  select(-starts_with(c("Risky", "Ben")))

# Compute the mean of each column
mean_INDacceptance <- data.frame(col = names(INDacceptance), mean_val = sapply(INDacceptance, mean))

# Compute standard error (not deviation)(divided by root of n)
sem <- function(x) sqrt(var(x, na.rm = TRUE) / length(na.omit(x)))
sem_INDacceptance <- data.frame(col = names(INDacceptance), sem_val = sapply(INDacceptance, sem))

mean_INDacceptance <- merge(mean_INDacceptance, sem_INDacceptance, by = "col")

# Plot
ggplot(mean_INDacceptance, aes(x = col, y = mean_val, fill = ifelse(mean_val >= 0, 'Above 0', 'Below 0'))) + 
  geom_bar(stat = "identity") +
  geom_errorbar(aes(ymin = mean_val - sem_val, ymax = mean_val + sem_val), width = 0.2) +
  geom_text(aes(label = round(mean_val, 2)), vjust = -0.9) +
  scale_fill_manual(values = c('Above 0' = 'palegreen3', 'Below 0' = 'firebrick2')) +
  labs(fill = "Legend") +
  xlab("Energy Technologies") +
  ylab("Mean Value") +
  ggtitle("Benefit - Risk for Each Energy Technology") +
  theme_minimal()+
  ylim(0, 2)

```
\newpage

# Linear Regression Models

## Perceived Risk from Firewood : demographic variables

```{r, echo=FALSE}

joineddf$Uppercaste <- ifelse(joineddf$caste %in% c('Brahmin', 'General'), 1, 0)

# Convert `gender` into binary variable 'male' (1) vs 'female' (0)
joineddf$Male <- ifelse(joineddf$gender == 'Male', 1, 0)

# Convert `religion` into binary variable 'Hinduism' (1) vs 'Non-Hinduism' (0)
joineddf$Hindu <- ifelse(joineddf$religion == 'Hinduism', 1, 0)

# # Convert `caste` into binary variable 'upper' (1) vs 'non-upper' (0)
codedmodule1$Uppercaste <- ifelse(codedmodule1$caste %in% c('Brahmin', 'General'), 1, 0)

# Convert `gender` into binary variable 'male' (1) vs 'female' (0)
codedmodule1$Male <- ifelse(codedmodule1$gender == 'Male', 1, 0)

# Convert `religion` into binary variable 'Hinduism' (1) vs 'Non-Hinduism' (0)
codedmodule1$Hindu <- ifelse(codedmodule1$religion == 'Hinduism', 1, 0)

falldemos <- joineddf %>%
            select(Risky_INDFirewoodetc,Uppercaste, Male, Hindu, urban_rural, age, State)%>%
            na.omit()

falldemoslm1 <- lm(Risky_INDFirewoodetc ~ Uppercaste + Male + Hindu + urban_rural + age, data = falldemos)
# Print model summary
#summary(falldemoslm1)
#plot(alldemoslm1)

codedmodule1$totalincome <- as.numeric(codedmodule1$totalincome)
codedmodule1$otherincome <- as.numeric(codedmodule1$otherincome)

falldemos2 <-codedmodule1 %>%
            select(Risky_INDFirewoodetc,Uppercaste, Male, Hindu, urban_rural, age, State, totalincome, otherincome)%>%
            mutate(income = totalincome + otherincome)%>%
            na.omit()

#lapply(falldemos2, unique)

falldemos$State <- as.factor(falldemos$State)
#levels(falldemos$State)

falldemos$age <- as.factor(falldemos$age)
#levels(falldemos$age)

# 2) lm with all demographic variables
falldemoslm2<- lm(Risky_INDFirewoodetc ~ Uppercaste + Male + Hindu + urban_rural + age + State , data = falldemos)
# Print model summary
#summary(falldemoslm2)
#plot(alldemoslm1)

```

```{r, echo=FALSE, results='asis'}
library(stargazer)

cat("\\begingroup")
cat("\\setlength{\\tabcolsep}{1pt}")
cat("\\renewcommand{\\arraystretch}{0.7}")
stargazer(falldemoslm1, falldemoslm2,
          title = "Results from 2 linear regression models",
          type = "latex"
)
cat("\\endgroup")

```

## Perceived Risk from LPG : demographic variables 
```{r, echo=FALSE}
#same linear models with LPG

LPGalldemos <- joineddf %>%
            select(Risky_INDLPG,Uppercaste, Male, Hindu, urban_rural, age, State)%>%
            na.omit()

LPGalldemoslm1 <- lm(Risky_INDLPG ~ Uppercaste + Male + Hindu + urban_rural + age , data = LPGalldemos)
# Print model summary
#summary(LPGalldemoslm1)
#plot(alldemoslm1)

LPGalldemos2 <-codedmodule1 %>%
            select(Risky_INDLPG,Uppercaste, Male, Hindu, urban_rural, age, State, totalincome, otherincome)%>%
            mutate(income = totalincome + otherincome)%>%
            na.omit()

#lapply(LPGalldemos2, unique)

# 2) lm with all demographic variables
LPGalldemoslm2<- lm(Risky_INDLPG ~ Uppercaste + Male + Hindu + urban_rural + age + State , data = LPGalldemos)
# Print model summary
#summary(LPGalldemoslm2)
#plot(alldemoslm1)

```

```{r, echo=FALSE, results='asis'}
library(stargazer)

cat("\\begingroup")
cat("\\setlength{\\tabcolsep}{1pt}")
cat("\\renewcommand{\\arraystretch}{0.7}")
stargazer(LPGalldemoslm1,LPGalldemoslm2 , 
          title = "Results from 2 linear regression models",
          type = "latex"
)
cat("\\endgroup")

```

Table 1 and 2 :

1. Uppercaste, Male, Hindu and urban_rural are binary variables. 
2. Reference for age variables is 18-24 years old
3. Reference for State is Maharashtra

# Linear Regression model - stove ownership by four groups - (i)LPG, (ii)Traditonal Stove(Including Firewood), (iii)Both and (iv)Other

```{r, echo = FALSE}

# # Convert `caste` into binary variable 'upper' (1) vs 'non-upper' (0)
module1$Uppercaste <- ifelse(module1$caste %in% c('Brahmin', 'General'), 1, 0)

# Convert `gender` into binary variable 'male' (1) vs 'female' (0)
module1$Male <- ifelse(module1$gender == 'Male', 1, 0)

# Convert `religion` into binary variable 'Hinduism' (1) vs 'Non-Hinduism' (0)
module1$Hindu <- ifelse(module1$religion == 'Hinduism', 1, 0)

stove <-module1 %>%
        filter(!row_number() %in% c(1,2))%>%
        select(Risky_INDFirewoodetc, Risky_INDLPG, urban_rural, Male, Uppercaste, Hindu, age, State, cookstove,)%>%
        mutate(across(starts_with("Risky"), ~case_when(. =="Not at all risky" ~ 1, 
                                                 . =="Slightly risky" ~ 2, 
                                                 . =="Moderately risky" ~ 3, 
                                                 . =="Very risky" ~ 4, 
                                                 . =="Extremely risky" ~ 5))) %>%
  na.omit()

#lapply(stove, unique)
library(dplyr)

grouped_stove <- stove %>%
  mutate(
    group = case_when(
      # Group 3
      cookstove %in% c(
        "Liquefied Petroleum Gas (LPG)/Cooking gas stove,Liquid Fuel Stove,Traditional Solid Fuel Stove",
        "Liquefied Petroleum Gas (LPG)/Cooking gas stove,Three Stone Stove/Open Fire",
        "Liquefied Petroleum Gas (LPG)/Cooking gas stove,Liquid Fuel Stove,Traditional Solid Fuel Stove,Three Stone Stove/Open Fire",
        "Liquefied Petroleum Gas (LPG)/Cooking gas stove,Three Stone Stove/Open Fire,Other(Please Specify)",
        "Liquefied Petroleum Gas (LPG)/Cooking gas stove,Traditional Solid Fuel Stove,Three Stone Stove/Open Fire",
        "Electric Stove,Liquefied Petroleum Gas (LPG)/Cooking gas stove,Traditional Solid Fuel Stove",
        "Liquefied Petroleum Gas (LPG)/Cooking gas stove,Traditional Solid Fuel Stove",
        "Solar Cooker,Liquefied Petroleum Gas (LPG)/Cooking gas stove,Traditional Solid Fuel Stove,Three Stone Stove/Open Fire",
        "Electric Stove,Solar Cooker,Liquefied Petroleum Gas (LPG)/Cooking gas stove,Traditional Solid Fuel Stove,Three Stone Stove/Open Fire",
        "Solar Cooker,Liquefied Petroleum Gas (LPG)/Cooking gas stove,Liquid Fuel Stove,Traditional Solid Fuel Stove,Three Stone Stove/Open Fire",
        "Electric Stove,Liquefied Petroleum Gas (LPG)/Cooking gas stove,Traditional Solid Fuel Stove,Three Stone Stove/Open Fire",
        "Solar Cooker,Liquefied Petroleum Gas (LPG)/Cooking gas stove,Traditional Solid Fuel Stove,No Food Cooked in Household",
        "Electric Stove,Liquefied Petroleum Gas (LPG)/Cooking gas stove,Three Stone Stove/Open Fire",
        "Solar Cooker,Liquefied Petroleum Gas (LPG)/Cooking gas stove,Three Stone Stove/Open Fire"
      ) ~ "Both",

      # Group 1
      cookstove %in% c(
        "Three Stone Stove/Open Fire",
        "Traditional Solid Fuel Stove",
        "Three Stone Stove/Open Fire,Other(Please Specify)",
        "Traditional Solid Fuel Stove,Three Stone Stove/Open Fire",
        "Traditional Solid Fuel Stove,Three Stone Stove/Open Fire,Other(Please Specify)",
        "Solar Cooker,Three Stone Stove/Open Fire",
        "Piped Natural Gas Stove,Traditional Solid Fuel Stove,Three Stone Stove/Open Fire",
        "Biogas Stove,Traditional Solid Fuel Stove",
        "Electric Stove,Biogas Stove,Traditional Solid Fuel Stove",
        "Electric Stove,Traditional Solid Fuel Stove"
      ) ~ "Traditional",

       # Group 2
      cookstove %in% c(
        "Liquefied Petroleum Gas (LPG)/Cooking gas stove,Other(Please Specify)",
        "Liquefied Petroleum Gas (LPG)/Cooking gas stove,Biogas Stove",
        "Electric Stove,Liquefied Petroleum Gas (LPG)/Cooking gas stove",
        "Liquefied Petroleum Gas (LPG)/Cooking gas stove,Biogas Stove,Liquid Fuel Stove",
        "Solar Cooker,Liquefied Petroleum Gas (LPG)/Cooking gas stove",
        "Solar Cooker,Liquefied Petroleum Gas (LPG)/Cooking gas stove,Liquid Fuel Stove",
        "Liquefied Petroleum Gas (LPG)/Cooking gas stove,No Food Cooked in Household",
        "Liquefied Petroleum Gas (LPG)/Cooking gas stove"
      ) ~ "LPG",

      # Default to "Other" for any cookstove values not specified in the conditions above
      TRUE ~ "Other"
    )
  )

count2 <- grouped_stove %>% count(group)

grouped_stove$group <- as.factor(grouped_stove$group)

grouped_stove$group <- factor(grouped_stove$group, levels = c("LPG", "Traditional", "Both", "Other"))

#levels(grouped_stove$group)

# lm with stove ownership 
LPGstove<- lm(Risky_INDLPG ~ urban_rural + Hindu + Male + Uppercaste + age +  group , data = grouped_stove)
# Print model summary
#summary(LPGstove)

# lm with stove ownership 
fstove<- lm(Risky_INDFirewoodetc ~ urban_rural + Hindu + Male + Uppercaste + age +  group , data = grouped_stove)
# Print model summary
#summary(fstove)

count2
```

```{r, echo=FALSE, results='asis'}
library(stargazer)

cat("\\begingroup")
cat("\\setlength{\\tabcolsep}{1pt}")
cat("\\renewcommand{\\arraystretch}{0.7}")
stargazer(LPGstove, fstove, 
          title = "Results from 2 linear regression models",
          type = "latex"
)
cat("\\endgroup")

```