### 2023-09-20 Paper1.Rmd

---
title: "Nuclear Energy: Kahan scale and Economic Political value scale"
output:
  pdf_document:
    toc: yes
    toc_depth: 2
    keep_tex: yes
  header-includes:
  - \usepackage{dcolumn}
  - \usepackage{array}
  - "\\usepackage{placeins}"
  keep_md: yes
---
```{r load packages, warning = FALSE, message = FALSE, echo=FALSE}
  knitr::opts_chunk$set(warning = FALSE, message = FALSE)

suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(readxl))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(tidyr))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(lavaan))
suppressPackageStartupMessages(library(tsibble))
suppressPackageStartupMessages(library(DT))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(psych)) 
```

```{r setup, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

moduletwo <- read_excel(here::here("thesis_surveydata/political_factors_complete.xlsx"))
module2 <- data.frame(moduletwo)
module2[module2 == "Rather not say/ Don't know"] <- "Rather not say/Don't know"

# loaded data set for module 2 eco -pol variables and removes the rows with question number and question text
```

```{r, eval =TRUE, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
#This chunk contains the coding schemes of various scales used in survey one: eco-political factors, kahan scale and acceptance scale
codedmodule2 <- module2 %>%

#remove row 1
  filter(!row_number() %in% c(1,2)) %>%

# replace risky likert scale with numbers
  mutate_at(vars(starts_with("Risky")), funs(case_when(. =="Not at all risky" ~ 1, 
                                                       . =="Slightly risky" ~ 2, 
                                                       . =="Moderately risky" ~ 3, 
                                                       . =="Very risky" ~ 4, 
                                                       . =="Extremely risky" ~ 5))) %>%

# replace beneficial likert scale with numbers  
  mutate_at(vars(starts_with("Ben")), funs(case_when(. =="Not at all beneficial" ~ 1,
                                                     . =="Slightly beneficial" ~ 2,
                                                     . =="Moderately beneficial" ~ 3,
                                                     . =="Very beneficial" ~ 4,
                                                     . =="Extremely beneficial" ~ 5 ))) %>%

# replace nuclear acceptance likert scale with numbers
  mutate_at(vars(N_accept,N_reluctantlyaccept,N_reject), funs(case_when(. == "Strongly disagree" ~ 1, 
                                                                        . == "Somewhat disagree" ~ 2,
                                                                        . == "Neither agree nor disagree" ~3,
                                                                        . == "Somewhat agree" ~ 4,
                                                                        . == "Strongly agree" ~ 5))) %>%

# code likert scale for variables for Kahan scale into numbers
  mutate_at(vars(starts_with (c("K_I","K_H","DISPLACE", "POLLUTE", "HEALTH", "JOBS", "BEAUTY", "PRIDE", "NPRIDE", "DEV", "PROSPER", "RELY"))), funs(case_when(. == "Strongly disagree" ~ 1, 
               . == "Somewhat disagree" ~ 2,
               . == "Neither agree nor disagree" ~3,
               . == "Somewhat agree" ~ 4,
               . == "Strongly agree" ~ 5))) %>%

# reverse code for likert scale for variables for Kahan scale into numbers
  mutate_at(vars(starts_with (c("K_S","K_E"))), funs(case_when(. == "Strongly disagree" ~ 5, 
                                                               . == "Somewhat disagree" ~ 4,
                                                               . == "Neither agree nor disagree" ~3,
                                                               . == "Somewhat agree" ~ 2,
                                                               . == "Strongly agree" ~ 1))) %>%

# code eco-pol scale variables into numbers
mutate_at(vars(SYSTEMDEMO,SYSTEMRELIGION,SYSTEMTECHNO,SYSTEMTOTAL,WEALTHLIM,MECHANISATION,DECISIONDECEN,INDUSTRYLARGE,ECONOMYGLOBAL,OWNERPVT,OWNERNOREG), funs(case_when(. == "Strongly disagree" ~ 1, 
                        . == "Somewhat disagree" ~ 2,
                        . == "Neither agree nor disagree" ~3,
                        . == "Somewhat agree" ~ 4,
                        . == "Strongly agree" ~ 5))) %>%

# reverse code eco-pol scale variables into numbers 
 mutate_at(vars(DECISIONCEN, INDUSTRYSMALL, ECONOMYLOCAL, ENVOVERDEV,OWNERPUB, OWNERREG), funs(case_when(. == "Strongly disagree" ~ 5, 
                                                                                                         . == "Somewhat disagree" ~ 4,
                                                                                                         . == "Neither agree nor disagree" ~3,
                                                                                                         . == "Somewhat agree" ~ 2,
                                                                                                         . == "Strongly agree" ~ 1))) %>%
  mutate_all(~ifelse(. == "Rather not say/ Don't know", "Rather not say/Don't know", .))%>%
  mutate_all(~ifelse(. == "West bengal", "West Bengal", .))%>%
  mutate_all(~ifelse(. == "WEST BENGAL", "West Bengal", .))%>%
  mutate_all(~ifelse(. == "Up", "Uttar Pradesh", .))%>%
  mutate_all(~ifelse(. == "Tamilnadu", "Tamil Nadu", .))%>%
  mutate_all(~ifelse(. == "UP", "Uttar Pradesh", .))

```

```{r, eval=TRUE, echo=FALSE}

moduleone <- read_excel(here::here("thesis_surveydata/livelihood_insecurity_complete.xlsx"))
module1 <- data.frame(moduleone)
# loaded data set for module 2 eco -pol variables and removes the rows with question number and question text
```

```{r, eval =TRUE, echo=FALSE}
#This chunk contains the coding schemes of various scales used in survey one: eco-political factors, kahan scale and acceptance scale
codedmodule1 <- module1 %>%

#remove row 1 and 2 
  filter(!row_number() %in% c(1,2)) %>%

# replace risky likert scale with numbers
  mutate_at(vars(starts_with("Risky")), funs(case_when(. =="Not at all risky" ~ 1, 
                                                       . =="Slightly risky" ~ 2, 
                                                       . =="Moderately risky" ~ 3, 
                                                       . =="Very risky" ~ 4, 
                                                       . =="Extremely risky" ~ 5))) %>%

# replace beneficial likert scale with numbers  
  mutate_at(vars(starts_with("Ben")), funs(case_when(. =="Not at all beneficial" ~ 1,
                                                     . =="Slightly beneficial" ~ 2,
                                                     . =="Moderately beneficial" ~ 3,
                                                     . =="Very beneficial" ~ 4,
                                                     . =="Extremely beneficial" ~ 5 ))) %>%

# replace nuclear acceptance likert scale with numbers
  mutate_at(vars(N_accept,N_reluctantlyaccept,N_reject), funs(case_when(. == "Strongly disagree" ~ 1, 
                                                                        . == "Somewhat disagree" ~ 2,
                                                                        . == "Neither agree nor disagree" ~3,
                                                                        . == "Somewhat agree" ~ 4,
                                                                        . == "Strongly agree" ~ 5)))%>%
  mutate_all(~ifelse(. == "Rather not say/ Don't know", "Rather not say/Don't know", .))%>%
  mutate_all(~ifelse(. == "West bengal", "West Bengal", .))%>%
  mutate_all(~ifelse(. == "WEST BENGAL", "West Bengal", .))%>%
  mutate_all(~ifelse(. == "Up", "Uttar Pradesh", .))%>%
  mutate_all(~ifelse(. == "Tamilnadu", "Tamil Nadu", .))%>%
  mutate_all(~ifelse(. == "UP", "Uttar Pradesh", .))

```

```{r, echo = FALSE}

suppressPackageStartupMessages(library(tidyselect))
# combine codedmodule1 and codedmodule2

numcol2 <- codedmodule2 %>% add_column(Respondent = 1:1099, .before = "Language") 
numcol1 <- codedmodule1 %>% add_column(Respondent = 1099:2160, .before = "Language")

joineddf <-  full_join(numcol2, numcol1, by = c("Respondent", "Language", "Survey_Date", "State", "Risky_Hydro", "Risky_Solar", "Risky_Wind", "Risky_Nuclear", "Risky_Coal", "Risky_Gas", "Risky_Oil", "Risky_INDHydro", "Risky_INDSolar", "Risky_INDWind", "Risky_INDBiogas", "Risky_INDDiesel", "Risky_INDKerosene", "Risky_INDFirewoodetc", "Risky_INDLPG", "Ben_Hydro", "Ben_Solar", "Ben_Wind", "Ben_Nuclear", "Ben_Coal", "Ben_Gas", "Ben_Oil", "Ben_INDHydro", "Ben_INDSolar", "Ben_INDWind", "Ben_INDDiesel", "Ben_INDBiogas", "Ben_INDFirewoodetc", "Ben_INDLPG", "Ben_INDKerosene", "N_accept", "N_reluctantlyaccept", "N_reject", "age", "urban_rural", "gender", "caste", "religion", "religiontext")) %>%
  mutate_at(vars(age), funs(case_when(. == "18-24 years old" ~ 1, 
                                      . == "25-34 years old" ~ 2,
                                      . ==  "35-44 years old" ~3,
                                      . == "45-54 years old" ~ 4,
                                      . == "55-64 years old" ~ 5,
                                      . == "65-74 years old" ~ 6,
                                      . == "75 years or older" ~ 7))) # converting age to continuous variable

joineddf1 <-  full_join(numcol2, numcol1, by = c("Respondent", "Language", "Survey_Date", "State", "Risky_Hydro", "Risky_Solar", "Risky_Wind", "Risky_Nuclear", "Risky_Coal", "Risky_Gas", "Risky_Oil", "Risky_INDHydro", "Risky_INDSolar", "Risky_INDWind", "Risky_INDBiogas", "Risky_INDDiesel", "Risky_INDKerosene", "Risky_INDFirewoodetc", "Risky_INDLPG", "Ben_Hydro", "Ben_Solar", "Ben_Wind", "Ben_Nuclear", "Ben_Coal", "Ben_Gas", "Ben_Oil", "Ben_INDHydro", "Ben_INDSolar", "Ben_INDWind", "Ben_INDDiesel", "Ben_INDBiogas", "Ben_INDFirewoodetc", "Ben_INDLPG", "Ben_INDKerosene", "N_accept", "N_reluctantlyaccept", "N_reject", "age", "urban_rural", "gender", "caste", "religion", "religiontext"))

numcolb <- module2 %>%   filter(!row_number() %in% c(1,2))%>% 
                        add_column(Respondent = 1:1099, .before = "Language") 
numcola <- module1 %>%   filter(!row_number() %in% c(1,2)) %>% 
                        add_column(Respondent = 1099:2160, .before = "Language")

uncodedjoineddf <- full_join(numcolb, numcola, by = c("Respondent", "Language", "Survey_Date", "State", "Risky_Hydro", "Risky_Solar", "Risky_Wind", "Risky_Nuclear", "Risky_Coal", "Risky_Gas", "Risky_Oil", "Risky_INDHydro", "Risky_INDSolar", "Risky_INDWind", "Risky_INDBiogas", "Risky_INDDiesel", "Risky_INDKerosene", "Risky_INDFirewoodetc", "Risky_INDLPG", "Ben_Hydro", "Ben_Solar", "Ben_Wind", "Ben_Nuclear", "Ben_Coal", "Ben_Gas", "Ben_Oil", "Ben_INDHydro", "Ben_INDSolar", "Ben_INDWind", "Ben_INDDiesel", "Ben_INDBiogas", "Ben_INDFirewoodetc", "Ben_INDLPG", "Ben_INDKerosene", "N_accept", "N_reluctantlyaccept", "N_reject", "age", "urban_rural", "gender", "caste", "religion", "religiontext")) %>%
  mutate_all(~ifelse(. == "Rather not say/ Don't know", "Rather not say/Don't know", .))%>%
  mutate_all(~ifelse(. == "West bengal", "West Bengal", .))%>%
  mutate_all(~ifelse(. == "WEST BENGAL", "West Bengal", .))%>%
  mutate_all(~ifelse(. == "Up", "Uttar Pradesh", .))%>%
  mutate_all(~ifelse(. == "Tamilnadu", "Tamil Nadu", .))%>%
  mutate_all(~ifelse(. == "UP", "Uttar Pradesh", .))

```

\newpage

# H1: Nuclear Energy will be seen as riskier than other energy technologies in India.

## Likert Responses (n= 2160)
The percentages are rounded off to whole numbers.

```{r, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
# Q1: how much is the awareness of energy technologies in India ? How does it relate to state specific data ? 
# summary of NA results from the combined sample of 2000 responses (simple percentage bar graphs)
# part 1- how much is the awareness of energy technologies in India -

#plotting perceived risk from different technologies including rather not say don't know
colorsequence<- c("#8C4646","#D9695F","#F2A679","#F2D091","#5D8C7B","#808080")

#following is the data frame for making the barplot 
df_barplot <- uncodedjoineddf %>%
  dplyr::select(starts_with("RISKY"))%>%
  gather(Technology, Perceived_Risk, Risky_Hydro:Risky_INDLPG, factor_key = TRUE)%>%
  separate(Technology, c("Risky", "Technology"), extra = "merge", fill = "left")%>%
  dplyr::select(-Risky) %>% 
  group_by(Technology, Perceived_Risk) %>%
  dplyr::summarise(n=n())

#following data frame gives us percentage for each response 
df_barplot_percent<- df_barplot%>%
  group_by(Technology) %>%
  mutate(Percentage = round(n/sum(n)*100, 0)) %>%
  ungroup()

#plotting perceived benefit from different technologies including rather not say don't know

df_barplotben <- uncodedjoineddf %>%
  dplyr::select(starts_with("Ben"))%>%
  gather(Technology, Perceived_Benefit, Ben_Hydro:Ben_INDKerosene, factor_key = TRUE)%>%
  separate(Technology, c("Ben", "Technology"), extra = "merge", fill = "left")%>%
  dplyr::select(-Ben) %>% 
  group_by(Technology, Perceived_Benefit) %>%
  dplyr::summarise(n=n())

df_barplotben_percent<- df_barplotben%>%
  group_by(Technology) %>%
  mutate(Percentage = round(n/sum(n)*100, 0)) %>%
  ungroup()

#bar plot for likert responses perceived risk - centrally owned technologies 
df_barplot %>% ggplot(aes(fill= fct_rev(factor(Perceived_Risk, levels=c("Extremely risky", "Very risky","Moderately risky","Slightly risky","Not at all risky","Rather not say/Don't know"))), y=n , x=Technology)) + 
  geom_bar(position="fill",stat="identity")+
  geom_text(data = df_barplot_percent, aes(label=paste0(Percentage, "%"), y=n), position = position_fill(vjust=0.5), size=3) +
  scale_x_discrete(limits = c("Solar", "Wind" ,"Hydro", "Oil", "Coal","Gas", "Nuclear"))+
  scale_fill_manual("legend", values = c("Extremely risky" = "#8C4646", "Very risky" = "#D9695F", "Moderately risky" = "#F2A679", "Slightly risky" = "#F2D091","Not at all risky" = "#5D8C7B", "Rather not say/Don't know" = "grey94"))+
  labs( y = "Perceived Risk", x ="Centrally Managed Energy Technologies")+
  theme_classic()

#perceived benefit centrally owned technologies

df_barplotben %>% ggplot(aes(fill= factor(Perceived_Benefit, levels=c("Rather not say/Don't know","Not at all beneficial","Slightly beneficial","Moderately beneficial","Very beneficial", "Extremely beneficial")), y=n , x=Technology)) + 
  geom_bar(position="fill",stat="identity")+
   geom_text(data = df_barplotben_percent, aes(label=paste0(Percentage, "%"), y=n), position = position_fill(vjust=0.5), size=3) +
  scale_fill_manual("legend", values = c("Rather not say/Don't know" = "grey94" ,"Not at all beneficial" = "orange","Slightly beneficial"="yellow" ,"Moderately beneficial" = "palegreen","Very beneficial"= "green3", "Extremely beneficial"= "green4"))+
  scale_x_discrete(limits = c("Solar", "Wind", "Hydro", "Oil", "Coal","Gas","Nuclear"))+
  labs( y = "Perceived Benefit", x ="Centrally Managed Energy Technologies")+
  theme_classic()

```

\newpage

## Mean Perceived Risk and Mean Perceived Benefit for all energy technologies.

```{r, echo=FALSE}
library(tidyverse)

# Filter variables of interest
joineddf_subset <- joineddf %>%
  dplyr::select(!starts_with(c("Ben_IND", "Risky_IND")))%>%
  dplyr::select(starts_with("Risky"), 
         starts_with("Ben"))

# Assuming your data is named joineddf_subset
long_df <- joineddf_subset %>%
  gather(key = "Type", value = "Value")

# Separate out Risky and Ben columns
risky_df <- long_df %>% filter(str_detect(Type, "^Risky"))
ben_df <- long_df %>% filter(str_detect(Type, "^Ben"))

# Extract the technology type (Hydro, Solar, etc.)
risky_df$Tech <- str_replace(risky_df$Type, "Risky_", "")
ben_df$Tech <- str_replace(ben_df$Type, "Ben_", "")

# Calculate the mean for each technology for Risky
risky_mean <- risky_df %>%
  group_by(Tech) %>%
  summarize(Mean_Value = mean(Value, na.rm = TRUE),
          sd = sd(Value, na.rm = TRUE),
          se = sd(Value, na.rm = TRUE) / sqrt(n()),
          .groups = "drop")

# Do the same for Ben technologies
ben_mean <- ben_df %>%
  group_by(Tech) %>%
  summarize(Mean_Value = mean(Value, na.rm = TRUE),
          sd = sd(Value, na.rm = TRUE),
          se = sd(Value, na.rm = TRUE) / sqrt(n()),
          .groups = "drop")

# Combine the two data frames into one for plotting
combined_mean <- bind_rows(
  mutate(risky_mean, Category = "Risky"),
  mutate(ben_mean, Category = "Beneficial")
)
# Reorder the factor levels of the Tech variable by Mean_Value
risky_mean$Tech <- factor(risky_mean$Tech, levels = risky_mean$Tech[order(risky_mean$Mean_Value, decreasing = FALSE)])

```

```{r, echo=FALSE}
library(ggplot2)
library(gridExtra)

mean_risk <- mean(combined_mean$Mean_Value[combined_mean$Category == "Risky"], na.rm = TRUE)
mean_benefit <- mean(combined_mean$Mean_Value[combined_mean$Category == "Beneficial"], na.rm = TRUE)

# Order Technologies
technology_order <- c("Solar","Wind","Hydro","Oil","Coal", "Gas","Nuclear")
ben_mean$Tech <- factor(ben_mean$Tech, levels = technology_order)

# Generate the Ben plot with flipped y-axis
ben_plot <- ggplot(ben_mean, aes(y = Tech, x = Mean_Value)) +
  geom_bar(stat = "identity", fill = "palegreen3") +
  geom_errorbar(aes(xmin = Mean_Value - se, xmax = Mean_Value + se), width = 0.2) +
  coord_cartesian(xlim = c(4, 1)) + 
  labs(title = "Perceived Benefit", x = "Mean Values") +
  #scale_x_reverse() +
  theme(  # remove y-axis labels
        axis.title.y = element_blank()) + # remove the "Tech" axis title
  geom_vline(data = subset(combined_mean, Category == "Beneficial"), 
             aes(xintercept = mean_risk), linetype="dotted", color = "black", size = 0.5)

# Generate the Risky plot
risky_plot <- ggplot(risky_mean, aes(y = Tech, x = Mean_Value)) +
  geom_bar(stat = "identity", fill = "firebrick2") +
  geom_errorbar(aes(xmin = Mean_Value - se, xmax = Mean_Value + se), width = 0.2) +
  coord_cartesian(xlim = c(1, 4)) + 
  labs(title = "Perceived Risk", x = "Mean Values") +
  theme(axis.text.y = element_blank(),
        axis.title.y = element_blank())+
  geom_vline(data = subset(combined_mean, Category == "Risky"), 
             aes(xintercept = mean_risk), linetype="dotted", color = "black", size = 0.5)
 # remove the "Tech" axis title

# Arrange plots side by side
grid.arrange(ben_plot, risky_plot, ncol=2)

```

## Boxplot for perceived benefit and perceived risk by technology

```{r, echo = FALSE, fig.width=16, fig.height=6}

#Box plot to see the distribution

# Loading the required libraries
library(tidyverse)
library(stringr)
library(gridExtra)

# Reshape the data to a long format
joineddf_subset_long <- joineddf_subset %>%
  pivot_longer(cols = everything(), 
               names_to = "variable",
               values_to = "value")

# Separate the data into two: one for "Risky_" variables and one for "Ben_" variables
risky_df <- joineddf_subset_long %>% 
  filter(str_detect(variable, "^Risky_")) %>%
  mutate(variable = str_remove(variable, "Risky_"))  # Remove the "Risky_" prefix

ben_df <- joineddf_subset_long %>% 
  filter(str_detect(variable, "^Ben_")) %>%
  mutate(variable = str_remove(variable, "Ben_"))  # Remove the "Ben_" prefix

# Remove NA values
risky_df <- risky_df %>% drop_na()
ben_df <- ben_df %>% drop_na()

# Create box plots
risky_plot <- ggplot(risky_df, aes(x = variable, y = value, fill = variable)) +
  geom_boxplot() +
  scale_fill_manual(values = rep("firebrick2", length(unique(risky_df$variable)))) +
  labs(x = "Energy Technologies", y = "Likert", fill = "Technologies") +
  ggtitle("Perceived risk") +
  theme_minimal()+
  theme(axis.text.y = element_text(size=20), 
        axis.text.x = element_text(size=20),
        axis.title.y = element_text(size=20),
        axis.title.x = element_text(size=20),
        title = element_text(size=20))+
  guides(fill = FALSE)

ben_plot <- ggplot(ben_df, aes(x = variable, y = value, fill = variable)) +
  geom_boxplot() +
  scale_fill_manual(values = rep("palegreen3", length(unique(ben_df$variable)))) +
  labs(x = "Energy Technologies", y = "Likert Responses", fill = "Technologies") +
  ggtitle("Perceived Benefit") +
  theme_minimal()+
  theme(axis.text.y = element_text(size=20), 
        axis.text.x = element_text(size=20),
        axis.title.y = element_text(size=20),
        axis.title.x = element_text(size=20),
        title = element_text(size=20))+
  guides(fill = FALSE)

# Put the two plots next to each other
grid.arrange(risky_plot, ben_plot, ncol = 2)

```

\newpage

## Pairwise T-test: Mean perceived risk and mean perceived benefit (all energy technologies)

The red and green pairs indicate that there is a statistically significant difference between the means of the two groups. 
White and grey indicate - no differences between the means of the two groups.

```{r, echo = FALSE, fig.width=14, fig.height=6}
#Q2 :How are the perceptions of the risk of producing energy from nuclear fission different from that of other conventional sources of energy like coal or hydro power in India?
#H : Nuclear energy will be perceived as riskier than other sources of energy (coal or hydro) in India like in the west.
perceivedriskc <-joineddf %>% 
  dplyr::select("Risky_Solar", "Risky_Hydro", "Risky_Wind", "Risky_Nuclear", "Risky_Coal", "Risky_Gas", "Risky_Oil") %>%
  gather(key = "technology", value = "perceived_risk") %>% 
  mutate(technology = gsub("Risky_", "", technology)) %>%
  na.omit()

result_ttest <- pairwise.t.test(perceivedriskc$perceived_risk, perceivedriskc$technology, p.adjust.method = "BH")

# visualising the pairwaise t test results. the non significant pairs are grey

tidy_result_ttest <- broom::tidy(result_ttest)

heatmap1 <- ggplot(tidy_result_ttest, aes(x = group1, y = group2, fill = p.value)) + 
  geom_tile(color = "white") + 
  scale_fill_gradient2(low = "firebrick2", high = "white", midpoint = 0.025, 
                       limits = c(0, 0.05), space = "Lab", 
                       name = "P-value") + 
  theme_classic() + 
  labs(x = "", y = "") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
   theme(axis.text.y = element_text(size=20), 
        axis.text.x = element_text(size=20),
        axis.title.y = element_text(size=20),
        axis.title.x = element_text(size=20),
        title = element_text(size=20),
        legend.text = element_text(size=12))

#Question- Why there is significant difference statistically - if the medians are the same in the boxplot.
#Q3: How do perceptions of benefit from nuclear energy compare with other source of energy?

# H : benefits from all energy technologies will be similar 
# include regression, heat maps, bar graphs? - decide which ones are good to include

perceivedbenefitc <-joineddf %>% 
  dplyr::select("Ben_Solar", "Ben_Hydro", "Ben_Wind", "Ben_Nuclear", "Ben_Coal", "Ben_Gas", "Ben_Oil") %>%
  gather(key = "technology", value = "perceived_benefit") %>% 
  mutate(technology = gsub("Ben_", "", technology)) %>%
  na.omit()

result_ttestben <- pairwise.t.test(perceivedbenefitc$perceived_benefit, perceivedbenefitc$technology, p.adjust.method = "BH")

# visualising the pairwaise t test results. the non significant pairs are grey

tidy_resultbc <- broom::tidy(result_ttestben)

library(ggplot2)
heatmap2 <- ggplot(tidy_resultbc, aes(x = group1, y = group2, fill = p.value)) + 
  geom_tile(color = "white") + 
  scale_fill_gradient2(low = "palegreen3", high = "white", midpoint = 0.025, 
                       limits = c(0, 0.05), space = "Lab", 
                       name = "P-value") + 
  theme_classic() + 
  labs(x = "", y = "") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
   theme(axis.text.y = element_text(size=20), 
        axis.text.x = element_text(size=20),
        axis.title.y = element_text(size=20),
        axis.title.x = element_text(size=20),
        title = element_text(size=20),
        legend.text = element_text(size=12))

grid.arrange(heatmap1, heatmap2, ncol=2)

```

```{r, echo=FALSE}

## Paired T-test: Comparing mean perceived risk and mean perceived benefit for each technology.

#Null hypothesis (H0): The mean difference between perceived risk and perceived benefit from each technology is zero.

#All p values are less than 0.05 suggesting that the differences observed in the bar graph are statistically significant.

# Melt the data frame to long format
long_format <- joineddf_subset %>% 
  pivot_longer(everything(), names_to = "technology", values_to = "value") %>%
  mutate(risk_or_benefit = if_else(str_detect(technology, "Risky"), "risk", "benefit"),
         technology = str_replace(technology, "Risky_|Ben_", ""))

# Convert character columns to factors
long_format$technology <- factor(long_format$technology)
long_format$risk_or_benefit <- factor(long_format$risk_or_benefit)

# Perform the t- Test for each technology
p_values <- long_format %>%
  split(.$technology) %>%
  lapply(function(x) t.test(x$value[x$risk_or_benefit == "risk"], x$value[x$risk_or_benefit == "benefit"], paired = TRUE)) %>%
  lapply(function(x) x$p.value)

```

\newpage

# H2: Gender and Caste will have significant impact like Gender and Race in the US studies of risk.

```{r, echo=FALSE}

# # Convert `caste` into binary variable 'upper' (1) vs 'non-upper' (0)
joineddf$Uppercaste <- ifelse(joineddf$caste %in% c('Brahmin', 'General'), 1, 0)
# Convert `gender` into binary variable 'male' (1) vs 'female' (0)
joineddf$Male <- ifelse(joineddf$gender == 'Male', 1, 0)
# Convert `religion` into binary variable 'Hinduism' (1) vs 'Non-Hinduism' (0)
joineddf$Hindu <- ifelse(joineddf$religion == 'Hinduism', 1, 0)
codedmodule2$State <- as.factor(codedmodule2$State)
codedmodule2$gender <- as.factor(codedmodule2$gender)
codedmodule2$Urban <- as.factor(codedmodule2$urban_rural)
joineddf$Urban <- as.factor(joineddf$urban_rural)
codedmodule2$caste <- as.factor(codedmodule2$caste)
codedmodule2$religion <- as.factor(codedmodule2$religion)
# # Convert `caste` into binary variable 'upper' (1) vs 'non-upper' (0)
codedmodule2$Uppercaste <- ifelse(codedmodule2$caste %in% c('Brahmin', 'General'), 1, 0)
# Convert `gender` into binary variable 'male' (1) vs 'female' (0)
codedmodule2$Male <- ifelse(codedmodule2$gender == 'Male', 1, 0)
# Convert `religion` into binary variable 'Hinduism' (1) vs 'Non-Hinduism' (0)
codedmodule2$Hindu <- ifelse(codedmodule2$religion == 'Hinduism', 1, 0)

alldemos <- joineddf %>%
            dplyr::select(Risky_Nuclear, Uppercaste, Male, Hindu, Urban, age, State)

alldemoslm1 <- lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + Urban + age , data = alldemos)
# Print model summary
#summary(alldemoslm1)
#plot(alldemoslm1)

# lm suggests there is no significant white male effect or upper caste male effect for risk perception from nuclear energy in India. 
# the coefficients are very small and so is the r squared - rsquared is minimal.

# 2) lm with all demographic variables
alldemoslm2<- lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + Urban + age + State, data = alldemos)
# Print model summary
#summary(alldemoslm2)
#plot(alldemoslm1)

# most of these results indicate that there is no significant white male effect or upper caste male effect for risk perception from nuclear energy in India. 
# the coefficients are very small and so is the r squared - rsquared is minimal.

```

## two linear regression models
```{r, echo=FALSE, results='asis'}
library(stargazer)

cat("\\begingroup")
cat("\\setlength{\\tabcolsep}{1pt}")
cat("\\renewcommand{\\arraystretch}{0.7}")
stargazer(alldemoslm1, alldemoslm2, 
          title = "Results from 2 linear regression models",
          type = "latex"
)
cat("\\endgroup")

```

\newpage

# H3: Regional differences will have a strong impact

## Linear regression where Mean value is the intercept
Same model with mean value as intercept.

```{r, echo=FALSE}
# Centering the numeric predictor variables
alldemos$Uppercaste_centered <- alldemos$Uppercaste - mean(alldemos$Uppercaste, na.rm = TRUE)
alldemos$Male_centered      <- alldemos$Male - mean(alldemos$Male, na.rm = TRUE)
alldemos$Hindu_centered     <- alldemos$Hindu - mean(alldemos$Hindu, na.rm = TRUE)
alldemos$age_centered       <- alldemos$age - mean(alldemos$age, na.rm = TRUE)

# For factor variable 'Urban', we'll convert it to numeric (0/1) and then center
# Assuming "Rural" is the reference category
alldemos$Urban_numeric      <- ifelse(alldemos$Urban == "Urban", 1, 0)
alldemos$Urban_centered     <- alldemos$Urban_numeric - mean(alldemos$Urban_numeric, na.rm = TRUE)

# Now, run the regression with centered variables
alldemoslm1_centered <- lm(Risky_Nuclear ~ Uppercaste_centered + Male_centered + Hindu_centered + Urban_centered + age_centered, data = alldemos)
#summary(alldemoslm1_centered)

alldemos$State <- as.factor(alldemos$State)
alldemos$State <- relevel(alldemos$State, ref = "Rajasthan")

alldemoslm2_centered <- lm(Risky_Nuclear ~ Uppercaste_centered + Male_centered + Hindu_centered + Urban_centered + age_centered + State, data = alldemos)
#summary(alldemoslm2_centered)

```

```{r, echo=FALSE, results='asis'}
library(stargazer)

cat("\\begingroup")
cat("\\setlength{\\tabcolsep}{1pt}")
cat("\\renewcommand{\\arraystretch}{0.7}")
stargazer(alldemoslm1_centered, alldemoslm2_centered, 
          title = "Results from 2 linear regression models",
          type = "latex"
)
cat("\\endgroup")

```

\newpage

## Regional Differences Graph 
Following is a graph of z scores calculated from mean perceived risk from nuclear energy by state.

```{r, echo = FALSE, fig.width=8, fig.height=4}

staterisky <- joineddf %>%
  dplyr::select(!starts_with(c("Ben_IND", "Risky_IND")))%>%
  dplyr::select(starts_with("Risky"), State)

staterisky_mean <- staterisky %>%
  group_by(State) %>%
  summarise(
    Mean_Risky_Hydro = mean(Risky_Hydro, na.rm = TRUE),
    Mean_Risky_Solar = mean(Risky_Solar, na.rm = TRUE),
    Mean_Risky_Wind = mean(Risky_Wind, na.rm = TRUE),
    Mean_Risky_Nuclear = mean(Risky_Nuclear, na.rm = TRUE),
    Mean_Risky_Coal = mean(Risky_Coal, na.rm = TRUE),
    Mean_Risky_Gas = mean(Risky_Gas, na.rm = TRUE),
    Mean_Risky_Oil = mean(Risky_Oil, na.rm = TRUE)
  )

# Extract only Nuclear
nuclear_data <- staterisky_mean %>%
  dplyr::select(State, Mean_Risky_Nuclear)%>%
  mutate(
    Z_Risky_Nuclear = (Mean_Risky_Nuclear - mean(Mean_Risky_Nuclear, na.rm = TRUE)) / 
                      sd(Mean_Risky_Nuclear, na.rm = TRUE)
  )

mean_nuclear_state <- ggplot(nuclear_data, aes(x = State, y = Mean_Risky_Nuclear, group = 1)) +
  geom_line(color = "#F0BA32", size = 1) +
  geom_point(color = "#F0BA32", size = 3) +
  labs(title = "Means: Perceived Risk of Nuclear Energy",
       x = "State", y = "Mean Perceived Risk") +
  theme_minimal() +
  ylim(1,5)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot the z-scores
zscore_nuclear_state <- ggplot(nuclear_data, aes(x = State, y = Z_Risky_Nuclear, group = 1)) +
  geom_line(color = "#F0BA32", size = 1) +
  geom_point(color = "#F0BA32", size = 3) +
  labs(title = "Z-scores: Perceived Risk of Nuclear Energy",
       x = "State", y = "Z-Score") +
  ylim(-1 ,2)+
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

library(gridExtra)

# Assuming mean_nuclear_state and zscore_nuclear_state are both ggplot objects

# We'll add a spacer grob for spacing between plots
spacer <- grid::rectGrob(gp = grid::gpar(col = "white"))

grid.arrange(mean_nuclear_state, spacer, zscore_nuclear_state, ncol=3, widths=c(0.8, 0.06, 0.8))

```

```{r, echo=FALSE}
library(MASS)

alldemos_factor <-alldemos  %>%
  mutate_at(vars(Risky_Nuclear, Uppercaste, Male, Hindu, Urban, State), as.factor)

model= polr(Risky_Nuclear ~ Uppercaste + Male + Hindu + Urban +State , data = alldemos_factor, Hess = TRUE)
#summary(model)

model3 <- glm(Risky_Nuclear ~ Uppercaste + Male + Hindu + Urban +State , data = alldemos_factor, family = binomial)
#summary(model3)

ci <- confint(model3)
#print(ci)

```

\newpage

```{r, echo=FALSE}
# Q6: Does the Kahan scale explain perception of risk from nuclear energy and other DVs ?

# we conducted fa on the kahan scale 
# CFA for Kahan scale. We shouldn't do FA here since this a well used scale with theoretical basis for factor distinctions and also previous studies using the same scale

library(lavaan)
library(semTools)
library(kableExtra)

fascales <- codedmodule2 %>%
  dplyr::select(starts_with(c("K_","Risky_Nuclear","WEALTHLIM","MECHANISATION","DECISIONDECEN","DECISIONCEN","INDUSTRYLARGE","INDUSTRYSMALL","ECONOMYGLOBAL","ECONOMYLOCAL", "ENVOVERDEV","DEVOVERENV","OWNERPVT","OWNERNOREG","OWNERPUB", "OWNERREG","DISPLACENUCLEAR",  "POLLUTENUCLEAR", "HEALTHNUCLEAR", "JOBSNUCLEAR", "BEAUTYNUCLEAR", "PRIDENUCLEAR", "NPRIDENUCLEAR", "DEVNUCLEAR", "PROSPERNUCLEAR", "RELYNUCLEAR", "Uppercaste", "Male", "Hindu", "Urban", "State","age")))%>%
  mutate_at(vars(age), funs(case_when(. == "18-24 years old" ~ 1, 
                                      . == "25-34 years old" ~ 2,
                                      . ==  "35-44 years old" ~3,
                                      . == "45-54 years old" ~ 4,
                                      . == "55-64 years old" ~ 5,
                                      . == "65-74 years old" ~ 6,
                                      . == "75 years or older" ~ 7))) %>%
  na.omit()

Kahan_fas <- '
  KahanS =~ K_SHARM + K_SLIMCHOI + K_SPROTECT 
  KahanH =~ K_HEQUAL+ K_ERADEQ1 + K_EWEALTH + K_ERADEQ2
'

# Run CFA
Kahan_cfa <- cfa(Kahan_fas, data = fascales, std.lv = TRUE)

# Print the summary of the CFA results, including fit measures
#summary(Kahan_cfa, fit.measure= TRUE, standardized = TRUE)

# RMSEA - between 0.05 and 0.08 (reasonable fit), less than 0.05 close fit and greater than 0.10 is poor fit 
# CFI - greater than 0.90 or 0.95 indicates good fit
#TFI - CFI>TFI >0.90 - indicates good fit
#pvalue should be greater than 0.05 p > 0.0.5

# after having removed a few variables that had very low loading, the model is now a good fit, we have two latent variables - Hierarchy- egalitarian and Communitarian. Again we see that individualistic variables just don't fit here.

# creating table of factor scores

Kahan_scores <- predict(Kahan_cfa, newdata = fascales)

Kahan_Kscores <- cbind(fascales, Kahan_scores)

## RMSEA is still pretty big.

```

```{r fit-measures-display, echo = FALSE}
# Extract fit measures
fit_measures <- fitMeasures(Kahan_cfa, fit.measures = c("CFI", "TLI", "RMSEA", "rmsea.ci.upper", "rmsea.ci.lower"))
fit_measures2 <- fitMeasures(Kahan_cfa)
# Convert to a data frame for presentation
fit_df <- data.frame(Measure = names(fit_measures), Value = fit_measures) %>%
  mutate(Measure = case_when(
    Measure == "cfi" ~ "Comparative Fit Index (CFI) ",
    Measure == "tli" ~ "Tucker-Lewis Index (TLI)",
    Measure == "rmsea" ~ "Root Mean Square Error of Approximation(RMSEA)",
    Measure == "rmsea.ci.upper" ~ "RMSEA 90 Percent confidence interval - lower",
    Measure == "rmsea.ci.lower" ~ "RMSEA 90 Percent confidence interval - upper",
    TRUE ~ as.character(Measure)
  )) %>%
  mutate(Value = round(Value, 3)) %>%
  remove_rownames()

#RMSEA - between 0.05 and 0.08 (reasonable fit), less than 0.05 close fit and greater than 0.10 is poor fit 
#CFI - greater than 0.90 or 0.95 indicates good fit
#TFI - CFI>TFI >0.90 - indicates good fit

```
# Confirmatory Factor Analysis(CFA): Kahan Scale

**Cronbach's Alpha on Kahan et al(2007) Scale: A Note**

The Individualism items (indicated by K_I) were bringing down the Cronbach's alpha values in the Kahan scale. The Alpha for Individualism- Communitarian scale was 0.49. After removing the Individualism items (K_I) the alpha for this factor was 0.71. The reasons for this could be that the individualism items are not well adapted to the Indian population.

```{r, echo =FALSE}
# Display fit measures using kable
knitr::kable(fit_df, caption = "Fit Measures from the CFA", booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped", "hold_position"))

```

\newpage

```{r, echo=FALSE}

# Extract parameter estimates
params <- parameterEstimates(Kahan_cfa, standardized = TRUE)%>%
   slice(1:(n()-10))

params1 <- params %>%
  mutate(rhs = case_when(
    rhs == "K_SHARM" ~ "Sometimes the government needs to make laws that keep people from hurting themselves.",
    rhs == "K_SLIMCHOI" ~ "The government should put limits on the choices individuals can make so they don’t get in the way of what’s good for society.",
    rhs == "K_SPROTECT" ~ "The government should do more to advance society’s goals, even if that means limiting the freedom and choices of individuals.",
    rhs == "K_HEQUAL" ~ "We have gone too far in pushing equal rights in this country.",
    rhs == "K_ERADEQ1" ~ "We need to dramatically reduce inequalities between the rich and the poor.",
    rhs == "K_EWEALTH" ~ "Our society would be better off if the distribution of wealth was more equal.",
    rhs == "K_ERADEQ2" ~ "We need to dramatically reduce inequalities between men and women.",
    TRUE ~ as.character(rhs)
  )) %>%
  dplyr::select(!c(lhs,op, std.nox)) %>%
    rename(
    "Items" = rhs,
    "Loadings" = est,
    "Standard Error" = se,
    "zvalue" = z) %>%
  mutate(across(c("Loadings", "Standard Error", "zvalue"), round, 3))

params1$Scale <- NA
params1$Scale[1:3] <- "Individualism"
params1$Scale[4:7] <- "Hierarchy-Egalitarianism"
params1 <- params1[, c("Scale", setdiff(names(params1), "Scale"))]
CFAkahan <- knitr::kable(params1, booktabs = TRUE, caption = "Confirmatory Factor Analysis(CFA) on Kahan et al(2007) scale adapted to India") %>%
  kable_styling(latex_options = c("striped","scale_down", "hold_position", "repeat_header"))%>%
  column_spec(column = 2, width = "4cm")

# Print the table
CFAkahan %>%
  landscape()

```

\newpage

# Factor Analysis: New Eco-political Scale

```{r, echo=FALSE}

library(psych)
library(tidyverse)

ecopolall <- fascales %>%
        dplyr::select("WEALTHLIM","MECHANISATION","DECISIONDECEN","DECISIONCEN","INDUSTRYLARGE","INDUSTRYSMALL","ECONOMYGLOBAL","ECONOMYLOCAL", "ENVOVERDEV","DEVOVERENV","OWNERPVT","OWNERNOREG","OWNERPUB", "OWNERREG","DISPLACENUCLEAR",  "POLLUTENUCLEAR", "HEALTHNUCLEAR", "JOBSNUCLEAR", "BEAUTYNUCLEAR", "PRIDENUCLEAR", "NPRIDENUCLEAR", "DEVNUCLEAR", "PROSPERNUCLEAR", "RELYNUCLEAR")

ecopolall_alpha <- psych::alpha(ecopolall, check.keys = TRUE)

#summary(ecopolall_alpha)

# when we combine both the scales the alpha is high at 0.785 (almost 0.8)

# FA on all ecopol variables except systems of governance

ecopolall_cor <- cor(ecopolall)
PCA24 <- principal(ecopolall, nfactors = 24, rotate = "none") # no rotation to start with.

#PCA24$loadings

#fa.parallel(ecopolall_cor)

# this suggests 2 factors if we go by FA or 8 by PCA. let's try both and see how they look conceptually

# running EFA using psych package

ecopolall2.fa <- fa(ecopolall, nfactors = 2, rotate = "varimax")

# I don't know still doesn't make a lot of sense to me- the two factors make more sense - 1) people centered development 2) nationalist development
# So I am going to extract factor scores and bind them to the dataset

ecopolall_scores <- psych::factor.scores(ecopolall, ecopolall2.fa)

# The factor scores will be stored in the scores element of the returned list
scores1 <- ecopolall_scores$scores

#binding factor scores to the dataset

fascale_scores <- cbind(Kahan_Kscores, scores1) %>%
  rename(Pdevelop = MR1)%>%
  rename(Ndevelop = MR2) %>%
  dplyr::select(!starts_with(c("K_","WEALTHLIM","MECHANISATION","DECISIONDECEN","DECISIONCEN","INDUSTRYLARGE","INDUSTRYSMALL","ECONOMYGLOBAL","ECONOMYLOCAL", "ENVOVERDEV","DEVOVERENV","OWNERPVT","OWNERNOREG","OWNERPUB", "OWNERREG","DISPLACENUCLEAR",  "POLLUTENUCLEAR", "HEALTHNUCLEAR", "JOBSNUCLEAR", "BEAUTYNUCLEAR", "PRIDENUCLEAR", "NPRIDENUCLEAR", "DEVNUCLEAR", "PROSPERNUCLEAR", "RELYNUCLEAR")))

```

```{r, echo=FALSE}
library(ggplot2)
library(fmsb)

# Extract the loadings
loadings_matrix <- ecopolall2.fa$loadings[1:24, ]

loadings_df <- as.data.frame(as.matrix(loadings_matrix))

# Adding rownames to the dataframe as a new column
loadings_df <- loadings_df %>%
  rownames_to_column(var = "variable")

# Pivot to longer format first
long_loadings_df <- loadings_df %>%
  pivot_longer(cols = c(MR1, MR2), names_to = "Factors", values_to = "value")

# Pivot the longer format to wide format
wide_loadings_df <- long_loadings_df %>%
  pivot_wider(names_from = variable, values_from = value) %>%
  mutate(Factors = case_when(
    Factors == "MR1" ~ "People Centered Development (Pdevelop)",
    Factors == "MR2" ~ "Nationalist Development (Ndevelop)",
    TRUE ~ Factors
  )) %>%
  column_to_rownames("Factors")

# Define the variable ranges: maximum and minimum
max_min <- data.frame(
  WEALTHLIM = c(0.8, -0.6), MECHANISATION = c(0.8, -0.6), DECISIONDECEN = c(0.8, -0.6),
  DECISIONCEN = c(0.8, -0.6), INDUSTRYLARGE = c(0.8, -0.6), INDUSTRYSMALL = c(0.8, -0.6),
  ECONOMYGLOBAL = c(0.8, -0.6), ECONOMYLOCAL = c(0.8, -0.6), ENVOVERDEV = c(0.8, -0.6),
  DEVOVERENV = c(0.8, -0.6), OWNERPVT = c(0.8, -0.6), OWNERNOREG = c(0.8, -0.6),
  OWNERPUB = c(0.8, -0.6), OWNERREG = c(0.8, -0.6), DISPLACENUCLEAR = c(0.8, -0.6),
  POLLUTENUCLEAR = c(0.8, -0.6), HEALTHNUCLEAR = c(0.8, -0.6), JOBSNUCLEAR = c(0.8, -0.6),
  BEAUTYNUCLEAR = c(0.8, -0.6), PRIDENUCLEAR = c(0.8, -0.6), NPRIDENUCLEAR = c(0.8, -0.6),
  DEVNUCLEAR = c(0.8, -0.6), PROSPERNUCLEAR = c(0.8, -0.6), RELYNUCLEAR = c(0.8, -0.6)
)
rownames(max_min) <- c("Max", "Min")

# Assuming you have a dataframe 'exam_scores' with matching column names
# Bind the variable ranges to the data

radar2 <- rbind(max_min, wide_loadings_df) %>%
    rename(
      "Limits on Wealth " = "WEALTHLIM",
      "Anti Mechanisation of work" = "MECHANISATION",
      "Pro Decentralisation" = "DECISIONDECEN",
      "Pro Centralisation" = "DECISIONCEN",
      "Pro Large Industries" = "INDUSTRYLARGE",
      "Anti Large Industries " = "INDUSTRYSMALL",
      "Pro Globaleconomy" = "ECONOMYGLOBAL",
      "Pro Localeconomy" = "ECONOMYLOCAL",
      "Environment over Development" = "ENVOVERDEV",
      "Development over Environment" = "DEVOVERENV",
      "Pro Private ownership " = "OWNERPVT",
      "Anti Regulations" = "OWNERNOREG",
      "Pro Public ownership" = "OWNERPUB",
      "Pro Regulations" = "OWNERREG",
      "Displacement risk(nuclear)" = "DISPLACENUCLEAR",
      "Pollution risk(nuclear) " = "POLLUTENUCLEAR",
      "Health risk(nuclear)" = "HEALTHNUCLEAR",
      "Job opportunities(nuclear)" = "JOBSNUCLEAR",
      "Spoils Natural Beauty(nuclear)" = "BEAUTYNUCLEAR",
      "Community pride(nuclear)" = "PRIDENUCLEAR",
      "National pride(nuclear)" = "NPRIDENUCLEAR",
      "National development(nuclear)" = "DEVNUCLEAR",
      "Local prosperity(nuclear)" = "PROSPERNUCLEAR",
      "Reliance on government(nuclear)" = "RELYNUCLEAR"
    )

```

```{r, echo=FALSE, fig.width=20, fig.height=18, out.width='100%', out.height='100%'}
library(fmsb)

create_beautiful_radarchart <- function(data, color = "#00AFBB", 
                                        vlabels = colnames(data), vlcex = 0.7,
                                        caxislabels = NULL, title = NULL, ...){
  radarchart(
    data, axistype = 1,
    # Customize the polygon
    pcol = color, pfcol = scales::alpha(color, 0.5), plwd = 2, plty = 1,
    centerzero=FALSE,
    # Customize the grid
    cglcol = "grey", cglty = 1.5, cglwd = 1.5,
    # Customize the axis
    axislabcol = "black", 
     # Increase the axis label size
    cexc = 3,
    # Variable labels
    vlcex = vlcex, vlabels = vlabels, vlpos = 0.95,
    seg = 7, caxislabels = caxislabels, title = title, 
    # Increase title size
    cex.main = 2,
    ...
  )
}

# Reduce plot margin using par()
op <- par(mar = c(7, 3, 3, 3))
# Create the radar charts
create_beautiful_radarchart(
  data = radar2, caxislabels = c(-0.6,-0.4,-0.2, 0, 0.2, 0.4, 0.6, 0.8, 1),
  color = c("#00AFBB", "#F97D09"),
  vlcex = 1.1
)

# Add an horizontal legend
legend(
  x = "bottomright", legend = rownames(radar2[-c(1,2),]), horiz = TRUE,
  bty = "n", pch = 20 , col = c("#00AFBB", "#F97D09"),
  text.col = "black", cex = 1.5, pt.cex = 3,
  inset = c(0, 0)
  )
par(op)

```

\newpage

```{r, echo=FALSE}

library(flextable)

flex <- function(data, title=NULL) {
  # this grabs the data and converts it to a flextbale
  flextable(data) %>%
  # this makes the table fill the page width
  set_table_properties(layout = "autofit", width = 1) %>%
  # font size
  fontsize(size=10, part="all") %>%
    #this adds a ttitlecreates an automatic table number
      set_caption(title, 
                  autonum = officer::run_autonum(seq_id = "tab", 
                                                 pre_label = "Table ", 
                                                 post_label = "\n", 
                                                 bkm = "anytable")) #%>%
  # font type
  #font(fontname="Times New Roman", part="all")
}

rename_map <- c(
  "WEALTHLIM" = "Limits on Wealth",
  "MECHANISATION" = "Anti Mechanisation of work",
  "DECISIONDECEN" = "Pro Decentralisation",
  "DECISIONCEN" = "Pro Centralisation",
  "INDUSTRYLARGE" = "Pro Large Industries",
  "INDUSTRYSMALL" = "Anti Large Industries",
  "ECONOMYGLOBAL" = "Pro Globaleconomy",
  "ECONOMYLOCAL" = "Pro Localeconomy",
  "ENVOVERDEV" = "Environment over Development",
  "DEVOVERENV" = "Development over Environment",
  "OWNERPVT" = "Pro Private ownership",
  "OWNERNOREG" = "Anti Regulations",
  "OWNERPUB" = "Pro Public ownership",
  "OWNERREG" = "Pro Regulations",
  "DISPLACENUCLEAR" = "Displacement risk(nuclear)",
  "POLLUTENUCLEAR" = "Pollution risk(nuclear)",
  "HEALTHNUCLEAR" = "Health risk(nuclear)",
  "JOBSNUCLEAR" = "Job opportunities(nuclear)",
  "BEAUTYNUCLEAR" = "Spoils Natural Beauty(nuclear)",
  "PRIDENUCLEAR" = "Community pride(nuclear)",
  "NPRIDENUCLEAR" = "National pride(nuclear)",
  "DEVNUCLEAR" = "National development(nuclear)",
  "PROSPERNUCLEAR" = "Local prosperity(nuclear)",
  "RELYNUCLEAR" = "Reliance on government(nuclear)"
)

fa_table <- function(x, cut) {
  #get sorted loadings
  loadings <- fa.sort(x)$loadings %>% round(3)
  #supress loadings
  loadings[loadings > -cut & loadings < cut] <- ""
  #get additional info  
  add_info <- cbind(x$communalities, 
                    x$uniquenesses,
                    x$complexity) %>%
    # make it a data frame
    as.data.frame() %>%
    # column names
   rename("Communality" = V1,
           "Uniqueness" = V2,
           "Complexity" = V3) %>%
    #get the item names from the vector
    rownames_to_column("item")
  #build table
  loadings %>%
    unclass() %>%
    as.data.frame() %>%
    rownames_to_column("item") %>%
    left_join(add_info) %>%
    mutate(
      item = if_else(item %in% names(rename_map), rename_map[item], item),
      across(where(is.numeric), round, 3)
    )
}

fa_table(ecopolall2.fa, .4) %>%
  rename(Pdevelop = MR1, Ndevelop = MR2, Items= item) %>%
  flex("Eco-Pol Values Factor Analysis Table")

```

```{r, echo=FALSE}

ecopolall2.fa[["Vaccounted"]] %>%
  as.data.frame() %>%
  rownames_to_column("Property") %>%
  rename(Pdevelop = MR1, Ndevelop = MR2) %>%
    mutate(across(where(is.numeric), round, 3)) %>%
    flex("Eigenvalues and Variance Explained for Rotated Factor Solution")

```

\newpage

```{r, echo=FALSE}

ecopolall <- fascales %>%
        dplyr::select("WEALTHLIM","MECHANISATION","DECISIONDECEN","DECISIONCEN","INDUSTRYLARGE","INDUSTRYSMALL","ECONOMYGLOBAL","ECONOMYLOCAL", "ENVOVERDEV","DEVOVERENV","OWNERPVT","OWNERNOREG","OWNERPUB", "OWNERREG","DISPLACENUCLEAR",  "POLLUTENUCLEAR", "HEALTHNUCLEAR", "JOBSNUCLEAR", "BEAUTYNUCLEAR", "PRIDENUCLEAR", "NPRIDENUCLEAR", "DEVNUCLEAR", "PROSPERNUCLEAR", "RELYNUCLEAR")

pdevelop <- ecopolall %>%
  dplyr::select("POLLUTENUCLEAR", "HEALTHNUCLEAR", "BEAUTYNUCLEAR","DISPLACENUCLEAR","MECHANISATION","INDUSTRYSMALL","OWNERPUB")

ndevelop <- ecopolall %>%
  dplyr::select("JOBSNUCLEAR",  "PRIDENUCLEAR", "NPRIDENUCLEAR", "DEVNUCLEAR", "PROSPERNUCLEAR")

pdevelop_alpha <- psych::alpha(pdevelop, check.keys = TRUE)

ndevelop_alpha <- psych::alpha(ndevelop, check.keys = TRUE)

```

```{r, echo=FALSE}

data <- tibble(
  Scale = c(
    "People Centered Development (Pdevelop)", "", "", "", "", "", "",
    "Nationalist Development (Ndevelop)", "", "", "", ""
  ),
  Code = c(
    "Health risk(nuclear)", "Spoils Natural Beauty(nuclear)", "Anti Mechanisation of work",
    "Anti Large Industries", "Displacement risk(nuclear)", "Pollution risk(nuclear)",
    "Pro Regulations", "National development(nuclear)", "Community pride(nuclear)",
    "National pride(nuclear)", "Local prosperity(nuclear)", "Job opportunities(nuclear)"
  ),
  "Items and Loadings" = c(
    "Nuclear energy poses a great risk to the health of people living around it.(0.657)",
    "Nuclear energy spoils the natural beauty of the landscape.(0.638)",
    "Rapid mechanization of work is taking away jobs from workers in this country.(0.552)",
    "Large corporations are destroying the local industries in India and benefiting only a handful of people.(-0.532)",
    "Nuclear energy is leading to displacement of people from their land.(0.59)",
    "Nuclear energy increases pollution of air/water/land.(0.565)",
    "Regardless of ownership, the government should pass strong regulations and implement them.(-0.53)",
    "Nuclear energy pushes forward the country's development.(0.662)",
    "I would be proud if my community used nuclear energy.(0.623)",
    "Nuclear energy is a mark of pride for our nation.(0.605)",
    "Nuclear energy brings economic prosperity to the surrounding regions.(0.586)",
    "Nuclear energy will bring jobs to the local community.(0.427)"
  ),
  Alpha = c(0.757, "", "", "", "", "", "", 0.725, "", "", "", ""),
  Variance = c("", "", "", "", "", "", "", "", "", "", "", "")
)

```

```{r, echo=FALSE}

longvars <- knitr::kable(data, booktabs = TRUE, caption = "Two Factor Solution: Economic and Political Values Scale", 
                         longtable = TRUE, escape = FALSE) %>%
  kable_styling(latex_options = c("striped","scale_down", "hold_position", "repeat_header")) %>%
  column_spec(column = 2, width = "4cm") %>%
  column_spec(column = 1, width = "4cm") %>%
  column_spec(column = 3, width = "9cm")

# Print the table in landscape mode
longvars %>%
  landscape()

```

```{r, echo=FALSE}
# 3 linear models - 1) lm with demographic variables except age and State
m1_demo <- lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + Urban, data = fascale_scores)
# Print model summary
#summary(m1_demo)
#plot(m1_demo)

# lm suggests there is no significant white male effect or upper caste male effect for risk perception from nuclear energy in India. 
# the coefficients are very small and so is the r squared - rsquared is minimal.

# 2) lm with all demographic variables
m2_demo<- lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + Urban + age + State, data = fascale_scores)
# Print model summary
#summary(m2_demo)
#plot(m2_demo)

# States and some age have some impact on y variable.

# 3) lm with kahan 
m1_kahan <- lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + Urban + age + State + KahanS + KahanH , data = fascale_scores)
# Print model summary
#summary(m1_kahan)
#plot(m1_kahan)

# 4) lm with ecopolvars 
m1_allscales <- lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + Urban + age + State+ KahanS + KahanH + Pdevelop + Ndevelop, data = fascale_scores)
# Print model summary
#summary(m1_allscales)
#plot(m1_allscales)

```

\newpage

# H4 : Economic and Political Values will be important in explaining perceived risk from Nuclear Energy

```{r, echo=FALSE, results='asis'}
library(stargazer)

cat("\\begingroup")
cat("\\setlength{\\tabcolsep}{1pt}")
cat("\\renewcommand{\\arraystretch}{0.7}")
stargazer(m1_kahan, m1_allscales, 
          title = "Results from 2 linear regression models",
          type = "latex"
)
cat("\\endgroup")

```

\newpage

```{r, echo=FALSE}
# Q7 : political values scale - factor analysis - reduce dimensions and then run regression -logistic regression 
# Using logistic regression for the SYSTEM prefixed variables.

#codedmodule2
#codedmodule2$SYSTEMDEMO <- as.factor(codedmodule2$SYSTEMDEMO)
#levels(codedmodule2$SYSTEMDEMO)

#codedmodule2$SYSTEMRELIGION <- as.factor(codedmodule2$SYSTEMRELIGION)
#levels(codedmodule2$SYSTEMRELIGION)

#codedmodule2$SYSTEMTOTAL <- as.factor(codedmodule2$SYSTEMTOTAL)
#levels(codedmodule2$SYSTEMTOTAL)

#codedmodule2$SYSTEMTECHNO <- as.factor(codedmodule2$SYSTEMTECHNO)
#levels(codedmodule2$SYSTEMTECHNO)

#codedmodule2$Risky_Nuclear <- as.factor(codedmodule2$Risky_Nuclear)
#levels(codedmodule2$Risky_Nuclear)

#fitrisky_nucleargm <- with(codedmodule2, clm(Risky_Nuclear ~ 1 + SYSTEMDEMO + SYSTEMRELIGION + SYSTEMTOTAL + SYSTEMTECHNO))

```

\newpage

# Appendix: Characteristics of the Sample

The following graph shows that distribution of different demographic variables in our sample of 2,160 from the combined dataset from both surveys.
The percentages are rounded off to whole numbers.

```{r, echo=FALSE}

library(dplyr)
library(tidyr)

sampledemos <- joineddf1 %>% 
  dplyr::select("State", "Language", "age", "urban_rural", "gender", "caste" ,"religion") %>%
              mutate_all(~ifelse(. == "Rather not say/ Don't know", "Rather not say/Don't know", .))%>%
              mutate_all(~ifelse(. == "OBC", "Other Backward Classes (OBC)", .))%>%
              mutate_all(~ifelse(. == "SC", "Scheduled caste (SC)", .))%>%
              mutate_all(~ifelse(. == "ST", "Scheduled tribe (ST)", .))%>%
              mutate_all(~ifelse(. == "MAR", "Marathi", .))%>%
              mutate_all(~ifelse(. == "TA", "Tamil", .)) %>%
              mutate_all(~ifelse(. == "EN", "English", .))%>%
              mutate_all(~ifelse(. == "BN", "Bengali", .))%>%
              mutate_all(~ifelse(. == "MAR", "Marathi", .))%>%
              mutate_all(~ifelse(. == "HI", "Hindi", .))%>%
              na.omit()

#changing factor levels for each variable from high to low n

sampledemos$State <- factor(sampledemos$State, levels = c("Maharashtra","Rajasthan", "West Bengal" , "Tamil Nadu", "Uttar Pradesh"))
sampledemos$Language <- factor(sampledemos$Language, levels = c("Hindi","Marathi","Bengali","Tamil", "English"))
sampledemos$age <- factor(sampledemos$age, levels = c( "18-24 years old" ,  "25-34 years old","35-44 years old", "45-54 years old", "55-64 years old","65-74 years old","75 years or older"  ))
sampledemos$urban_rural <- factor(sampledemos$urban_rural, levels = c("Rural","Urban"))
sampledemos$gender <- factor(sampledemos$gender, levels = c("Male","Female"))
sampledemos$caste <- factor(sampledemos$caste, levels = c("Scheduled tribe (ST)", "Scheduled caste (SC)", "Other Backward Classes (OBC)", "General","Brahmin","Rather not say/Don't know"))
sampledemos$religion <- factor(sampledemos$religion, levels = c("Hinduism", "Islam", "Christanity","Buddhism","Agnostism/ Atheism","Other (Please mention)"))

```

```{r, echo=FALSE, fig.width=11, fig.height=14, out.width='80%', out.height='80%'}

# Loading the required libraries
library(tidyverse)
library(gridExtra)
library(scales)

# Function to create bar plot
create_bar_plot <- function(data, variable_name, palette_name) {
  factor_levels <- levels(sampledemos[[variable_name]])
  data %>%
    filter(Variable == variable_name) %>%
    mutate(Category = factor(Category, levels = factor_levels)) %>%
    ggplot(aes(y = "", x = Percentage, fill = Category)) +
    geom_bar(position = "fill", stat = "identity", width = 1) +
    geom_text(aes(label = paste0(round(Percentage, 0), "%")), 
              position = position_fill(vjust = 0.5)) +
    scale_x_continuous(labels = percent) +
    labs(y = NULL, x = "Percentage", fill = variable_name) +
    theme_classic() +
    theme(axis.text.y = element_blank(), strip.text = element_text(size = 10), legend.position = "top") +
    scale_fill_brewer(palette = palette_name)
}

# Gather data to long format and calculate counts and percentages
df_barplot <- sampledemos %>%
  gather(Variable, Category) %>%
  group_by(Variable, Category) %>%
  filter(!is.na(Category)) %>%
  summarise(n = n()) %>%
  mutate(Percentage = n / sum(n) * 100)

# Create separate plots for each variable
plot_state <- create_bar_plot(df_barplot, "State", "Spectral")
plot_language <- create_bar_plot(df_barplot, "Language", "Paired")
plot_age <- create_bar_plot(df_barplot, "age", "Set3")
plot_urban_rural <- create_bar_plot(df_barplot, "urban_rural", "Dark2")
plot_gender <- create_bar_plot(df_barplot, "gender", "Pastel1")
plot_caste <- create_bar_plot(df_barplot, "caste", "Set1")
plot_religion <- create_bar_plot(df_barplot, "religion", "Set2")

# Combine the plots with spacing
combined_plot <- gridExtra::arrangeGrob(plot_state, plot_language, plot_age, plot_urban_rural, plot_gender, plot_caste, plot_religion, ncol = 1, heights = c(1, 1, 1, 1, 1, 1, 1))

# Print the combined plot
grid::grid.newpage()
grid::grid.draw(combined_plot)

```