### 2023-09-20 module1/Extra.Rmd

---
title: "Extra"
output: html_document
---

```{r}

#FA on LSI/ISI components

df_FA <- addedcols %>% 
  dplyr::select(LIlenght,LIassociation, govtscheme,education, workknowledge, subsidyknow, drinkingwater, cookingwater,farwatermins, wateravail, toilet, house, houseown, polposition, polparty,  foodadequacy, foodtrouble, healthdist, illness, severeillness, stove, vehicle, food, income)%>%
  na.omit()

df_FA_numeric <- data.frame(lapply(df_FA, as.numeric))

# Print the new dataframe
print(df_FA_numeric)

lapply(df_FA, unique)

df_FA_cor <- cor(df_FA_numeric)
PCA_LI <- principal(df_FA_numeric, nfactors = 24, rotate = "none") # no rotation to start with.

print(PCA_LI)

PCA_LI$loadings

PCA_LI6 <- principal(df_FA_numeric, nfactors = 6, rotate = "varimax") # no rotation to start with.

print(PCA_LI6)

PCA_LI6$loadings

# Extract the loadings
loadings <- PCA_LI6$loadings

# Set loadings less than 0.4 to NA
loadings[abs(loadings) < 0.4] <- NA

# Print the updated loadings
print(loadings)
```

```{r}
## Now PCA on my SLI and LVI scales

SLI <-  addedcols %>%
  filter(Livelihood_Income == "LSI") %>%
  dplyr::select(Lconfidence, Llenght, Lassociation, animals, income, govtscheme,education, workknowledge, drinkingwater,
         cookingwater, farwatermins, wateravail,toilet, house, houseown, stove, vehicle, polposition, polparty, food, foodadequacy, foodtrouble, healthdist,          illness, severeillness) %>%
          na.omit()

# questions about coding the data - ask milind tomorrow

addedcols %>%
 filter(Livelihood_Income == "LSI") %>%
 summarise(n=n())

SLI %>%
  summarise_all(~sum(is.na(.)))

str(SLI)

# Convert all columns to numeric
SLI_numeric <- data.frame(lapply(SLI, function(x) as.numeric(as.factor(x))))

# Determine the number of factors to retain using parallel analysis

SLI_cor <- cor(SLI_numeric)
PCA25 <- principal(SLI_numeric, nfactors = 25 , rotate = "none") # no rotation to start with.

print(PCA25)

PCA25$loadings

fa.parallel(SLI_cor)

SLI.pca <- principal(SLI_numeric, nfactors = 9, rotate = "varimax")

print.psych(SLI.pca, cut = 0.40, sort = TRUE)

fa.diagram(SLI.fa)

```