### 2023-09-20 module2/factor analysis/extraFA.Rmd

---
title: "extraFA"
output: html_document
---

```{r load packages, warning = FALSE, message = FALSE, echo=FALSE}
  knitr::opts_chunk$set(warning = FALSE, message = FALSE)
suppressPackageStartupMessages(library(lavaan))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(readxl))
suppressPackageStartupMessages(library(tsibble))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(DT))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(psych)) # psychometrics

```

```{r, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
moduletwo <- read_excel(here::here("thesis_surveydata/political_factors_complete.xlsx"))
save(moduletwo, file = "moduletwo.rda")
module2 <- data.frame(moduletwo)
module2[module2 == "Rather not say/ Don't know"] <- "Rather not say/Don't know"
#module2 %>% filter(!row_number() %in% c(1,2))

# loaded data set for module 2 eco -pol variables and removes the rows with question number and question text
```

```{r, eval =TRUE, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
#This chunk contains the coding schemes of various scales used in survey one: eco-political factors, kahan scale and acceptance scale
codedmodule2 <- module2 %>%

#remove row 1
  filter(!row_number() %in% c(1,2)) %>%

# replace risky likert scale with numbers
  mutate_at(vars(starts_with("Risky")), funs(case_when(. =="Not at all risky" ~ 1, 
                                                       . =="Slightly risky" ~ 2, 
                                                       . =="Moderately risky" ~ 3, 
                                                       . =="Very risky" ~ 4, 
                                                       . =="Extremely risky" ~ 5))) %>%

# replace beneficial likert scale with numbers  
  mutate_at(vars(starts_with("Ben")), funs(case_when(. =="Not at all beneficial" ~ 1,
                                                     . =="Slightly beneficial" ~ 2,
                                                     . =="Moderately beneficial" ~ 3,
                                                     . =="Very beneficial" ~ 4,
                                                     . =="Extremely beneficial" ~ 5 ))) %>%

# replace nuclear acceptance likert scale with numbers
  mutate_at(vars(N_accept,N_reluctantlyaccept,N_reject), funs(case_when(. == "Strongly disagree" ~ 1, 
                                                                        . == "Somewhat disagree" ~ 2,
                                                                        . == "Neither agree nor disagree" ~3,
                                                                        . == "Somewhat agree" ~ 4,
                                                                        . == "Strongly agree" ~ 5))) %>%

# code likert scale for variables for Kahan scale into numbers
  mutate_at(vars(starts_with (c("K_I","K_H","DISPLACE", "POLLUTE", "HEALTH", "JOBS", "BEAUTY", "PRIDE", "NPRIDE", "DEV", "PROSPER", "RELY"))), funs(case_when(. == "Strongly disagree" ~ 1, 
               . == "Somewhat disagree" ~ 2,
               . == "Neither agree nor disagree" ~3,
               . == "Somewhat agree" ~ 4,
               . == "Strongly agree" ~ 5))) %>%

# reverse code for likert scale for variables for Kahan scale into numbers
mutate_at(vars(starts_with (c("K_S","K_E"))), funs(case_when(. == "Strongly disagree" ~ 5, 
                                                               . == "Somewhat disagree" ~ 4,
                                                               . == "Neither agree nor disagree" ~3,
                                                               . == "Somewhat agree" ~ 2,
                                                               . == "Strongly agree" ~ 1))) %>%

# code eco-pol scale variables into numbers
  mutate_at(vars(SYSTEMDEMO,SYSTEMRELIGION,SYSTEMTECHNO,SYSTEMTOTAL,WEALTHLIM,MECHANISATION,DECISIONDECEN,INDUSTRYSMALL,ECONOMYLOCAL,ENVOVERDEV,OWNERPUB, OWNERREG), funs(case_when(. == "Strongly disagree" ~ 1, 
                        . == "Somewhat disagree" ~ 2,
                        . == "Neither agree nor disagree" ~3,
                        . == "Somewhat agree" ~ 4,
                        . == "Strongly agree" ~ 5))) %>%

# reverse code eco-pol scale variables into numbers 
 mutate_at(vars(DECISIONCEN,INDUSTRYLARGE,ECONOMYGLOBAL,OWNERPVT,OWNERNOREG), funs(case_when(. == "Strongly disagree" ~ 5, 
                                                                                                         . == "Somewhat disagree" ~ 4,
                                                                                                         . == "Neither agree nor disagree" ~3,
                                                                                                         . == "Somewhat agree" ~ 2,
                                                                                                         . == "Strongly agree" ~ 1)))
```

```{r}
#scatter plot of Kahan scale scores around the median scores on Individualism and Hierarchy scales.

Kscalescores %>%
 ggplot(aes(Individualism_score, Hierarchy_score))+
  geom_count(aes(color = ..n.., size = ..n..))+
  guides(color = 'legend')+
  geom_hline(yintercept=2, colour="black", lwd=1)+
  geom_vline(xintercept=3, colour="black", lwd=1)+
  labs(x = " Communitarianism to Individualism", y = "Egalitarianism to Hierarchy")
```

```{r, echo=FALSE}

# checking assumptions before doing a factor analysis

# running Bartlett's test of sphericity on our scale

#ncol(ecopolall) # 24 variables in total 
#nrow(ecopolall) # 443

ecopolall_cor <- cor(ecopolall)

# conducting Bartlett's test of sphericity to check the assumption that the variables are correlated in the population
#ecopolall_bartlett <- cortest.bartlett(ecopolall_cor, n = nrow(ecopolall)) # p-value = 0

# Kaiser-Meyer-Olkin measure

#KMO(ecopolall) # above 0.7 ie MSA= 0.78 , the sampling adequacy was acceptable (KMO= 0.78) and the Bartlett's test of sphericity demonstrated that the correlations between the items were large enough for FA

```

```{r, echo=FALSE}
# checking normality

#longecopol <-ecopolall %>% 
  #gather(key = "variable", value = "score") %>% 
 # na.omit()

#ggplot(longecopol, aes(x = score, y = variable)) + 
  #geom_boxplot() # Q1 how do I deal with outliers here. Is this a problem ?

# Apply Shapiro-Wilk test to all columns
#shapiro_results <- apply(ecopolall, 2, shapiro.test)

# Extract p-values from test results
#p_values <- sapply(shapiro_results, function(x) x$p.value)

# Identify columns with non-normal distributions
#non_normal_columns <- names(p_values[p_values < 0.05])

# data is not normally distributed 
```

```{r}
##EXTRA :

  # nuclear 1 contains only those variables that were relevant in the EFA
Nuclear1 <- codedmodule2 %>%
dplyr::select("Risky_Nuclear", "HEALTHNUCLEAR", "POLLUTENUCLEAR", "BEAUTYNUCLEAR", "DISPLACENUCLEAR", "DEVNUCLEAR" , "PROSPERNUCLEAR", "NPRIDENUCLEAR", "PRIDENUCLEAR", "JOBSNUCLEAR", "RELYNUCLEAR") %>%
  na.omit()

nuclear_alpha1<- psych::alpha(Nuclear1, check.keys = TRUE)

summary(nuclear_alpha1)

#creating a condensed scale for the characteristics of the technology across two factors as identified with EFA - negative characteristics and positive characteristics.

Ncharac_scores<- Nuclear1 %>%
  rowwise()%>%
  na.omit()%>%
  dplyr::mutate(Negative_score = mean(c_across(HEALTHNUCLEAR:DISPLACENUCLEAR)))%>%
  dplyr::mutate(Positive_score = mean(c_across(DEVNUCLEAR:RELYNUCLEAR)))%>%
  dplyr::select(!starts_with(c("HEALTHNUCLEAR", "POLLUTENUCLEAR", "BEAUTYNUCLEAR", "DISPLACENUCLEAR", "DEVNUCLEAR" , "PROSPERNUCLEAR", "NPRIDENUCLEAR", "PRIDENUCLEAR", "JOBSNUCLEAR", "RELYNUCLEAR")))

Ncharac_lm <- lm(Risky_Nuclear ~ Negative_score + Positive_score , data = Ncharac_scores)
summary(Ncharac_lm )

kandecopol <- codedmodule2 %>%
  dplyr::select("Risky_Nuclear", "K_SHARM", "K_SLIMCHOI", "K_SPROTECT", "K_HEQUAL", "K_ERADEQ1",  "K_EWEALTH",  "K_ERADEQ2","State","HEALTHNUCLEAR", "POLLUTENUCLEAR", "BEAUTYNUCLEAR", "DISPLACENUCLEAR", "DEVNUCLEAR" , "PROSPERNUCLEAR", "NPRIDENUCLEAR", "PRIDENUCLEAR", "JOBSNUCLEAR", "RELYNUCLEAR") %>%
   rowwise()%>%
  na.omit()%>%
  dplyr::mutate(Individualism_score = mean(c_across(K_SHARM:K_SPROTECT)))%>%
  dplyr::mutate(Hierarchy_score = mean(c_across(K_HEQUAL:K_ERADEQ2)))%>%
  dplyr::mutate(Negative_score = mean(c_across(HEALTHNUCLEAR:DISPLACENUCLEAR)))%>%
  dplyr::mutate(Positive_score = mean(c_across(DEVNUCLEAR:RELYNUCLEAR)))%>%
  dplyr::select(!starts_with(c("HEALTHNUCLEAR", "POLLUTENUCLEAR", "BEAUTYNUCLEAR", "DISPLACENUCLEAR", "DEVNUCLEAR" , "PROSPERNUCLEAR", "NPRIDENUCLEAR", "PRIDENUCLEAR", "JOBSNUCLEAR", "RELYNUCLEAR", "K_I","K_S", "K_H", "K_E")))

```

```{r}

# CFA on the reduced ecopol value scale to check whether our reduced dimensions are good fit for the scale.

ecopolall

ecopolall_fas <- '
  Pdevelop =~ HEALTHNUCLEAR + BEAUTYNUCLEAR + POLLUTENUCLEAR
  Ndevelop =~ PRIDENUCLEAR + NPRIDENUCLEAR + PROSPERNUCLEAR
  Pdevelop ~~ Ndevelop 
'

# Run CFA with standardization all method - gives standardized loadings 
# The factors are assumed to be correlated

ecopolall_cfa <- cfa(ecopolall_fas, data = ecopolall, std.lv = TRUE)

# Print the summary of the CFA results, including fit measures
summary(ecopolall_cfa, fit.measure= TRUE, standardized = TRUE)

# RMSEA - between 0.05 and 0.08 (reasonable fit), less than 0.05 close fit and greater than 0.10 is poor fit 
# CFI - greater than 0.90 or 0.95 indicates good fit
#TFI - CFI>TFI >0.90 - indicates good fit

# cross validate with larger sample

# talk to terre about the two options - how much do I need to worry about CLI

```