### 2023-09-20 nuclear in comparison.Rmd

---
title: 'nuclear energy in comparison'
output:
  pdf_document:
    toc: yes
    toc_depth: 2
    keep_tex: yes
  header-includes:
  - \usepackage{dcolumn}
  - \usepackage{array}
  keep_md: yes
---

```{r load packages, warning = FALSE, message = FALSE, echo=FALSE}
  knitr::opts_chunk$set(warning = FALSE, message = FALSE)

suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(readxl))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(tidyr))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(lavaan))
suppressPackageStartupMessages(library(tsibble))
suppressPackageStartupMessages(library(DT))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(psych)) 
```

```{r setup, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

moduletwo <- read_excel(here::here("thesis_surveydata/political_factors_complete.xlsx"))
module2 <- data.frame(moduletwo)
module2[module2 == "Rather not say/ Don't know"] <- "Rather not say/Don't know"

# loaded data set for module 2 eco -pol variables and removes the rows with question number and question text
```

```{r, eval =TRUE, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
#This chunk contains the coding schemes of various scales used in survey one: eco-political factors, kahan scale and acceptance scale
codedmodule2 <- module2 %>%

#remove row 1
  filter(!row_number() %in% c(1,2)) %>%

# replace risky likert scale with numbers
  mutate_at(vars(starts_with("Risky")), funs(case_when(. =="Not at all risky" ~ 1, 
                                                       . =="Slightly risky" ~ 2, 
                                                       . =="Moderately risky" ~ 3, 
                                                       . =="Very risky" ~ 4, 
                                                       . =="Extremely risky" ~ 5))) %>%

# replace beneficial likert scale with numbers  
  mutate_at(vars(starts_with("Ben")), funs(case_when(. =="Not at all beneficial" ~ 1,
                                                     . =="Slightly beneficial" ~ 2,
                                                     . =="Moderately beneficial" ~ 3,
                                                     . =="Very beneficial" ~ 4,
                                                     . =="Extremely beneficial" ~ 5 ))) %>%

# replace nuclear acceptance likert scale with numbers
  mutate_at(vars(N_accept,N_reluctantlyaccept,N_reject), funs(case_when(. == "Strongly disagree" ~ 1, 
                                                                        . == "Somewhat disagree" ~ 2,
                                                                        . == "Neither agree nor disagree" ~3,
                                                                        . == "Somewhat agree" ~ 4,
                                                                        . == "Strongly agree" ~ 5))) %>%

# code likert scale for variables for Kahan scale into numbers
  mutate_at(vars(starts_with (c("K_I","K_H","DISPLACE", "POLLUTE", "HEALTH", "JOBS", "BEAUTY", "PRIDE", "NPRIDE", "DEV", "PROSPER", "RELY"))), funs(case_when(. == "Strongly disagree" ~ 1, 
               . == "Somewhat disagree" ~ 2,
               . == "Neither agree nor disagree" ~3,
               . == "Somewhat agree" ~ 4,
               . == "Strongly agree" ~ 5))) %>%

# reverse code for likert scale for variables for Kahan scale into numbers
  mutate_at(vars(starts_with (c("K_S","K_E"))), funs(case_when(. == "Strongly disagree" ~ 5, 
                                                               . == "Somewhat disagree" ~ 4,
                                                               . == "Neither agree nor disagree" ~3,
                                                               . == "Somewhat agree" ~ 2,
                                                               . == "Strongly agree" ~ 1))) %>%

# code eco-pol scale variables into numbers
  mutate_at(vars(SYSTEMDEMO,SYSTEMRELIGION,SYSTEMTECHNO,SYSTEMTOTAL,WEALTHLIM,MECHANISATION,DECISIONDECEN,INDUSTRYLARGE,ECONOMYGLOBAL,OWNERPVT,OWNERNOREG), funs(case_when(. == "Strongly disagree" ~ 1, 
                        . == "Somewhat disagree" ~ 2,
                        . == "Neither agree nor disagree" ~3,
                        . == "Somewhat agree" ~ 4,
                        . == "Strongly agree" ~ 5))) %>%

# reverse code eco-pol scale variables into numbers 
 mutate_at(vars(DECISIONCEN, INDUSTRYSMALL, ECONOMYLOCAL, ENVOVERDEV,OWNERPUB, OWNERREG), funs(case_when(. == "Strongly disagree" ~ 5, 
                                                                                                         . == "Somewhat disagree" ~ 4,
                                                                                                         . == "Neither agree nor disagree" ~3,
                                                                                                         . == "Somewhat agree" ~ 2,
                                                                                                         . == "Strongly agree" ~ 1))) %>%
  mutate_all(~ifelse(. == "Rather not say/ Don't know", "Rather not say/Don't know", .))%>%
  mutate_all(~ifelse(. == "West bengal", "West Bengal", .))%>%
  mutate_all(~ifelse(. == "WEST BENGAL", "West Bengal", .))%>%
  mutate_all(~ifelse(. == "Up", "Uttar Pradesh", .))%>%
  mutate_all(~ifelse(. == "Tamilnadu", "Tamil Nadu", .))%>%
  mutate_all(~ifelse(. == "UP", "Uttar Pradesh", .))

```

```{r, eval=TRUE, echo=FALSE}

moduleone <- read_excel(here::here("thesis_surveydata/livelihood_insecurity_complete.xlsx"))
module1 <- data.frame(moduleone)
# loaded data set for module 2 eco -pol variables and removes the rows with question number and question text
```

```{r, eval =TRUE, echo=FALSE}
#This chunk contains the coding schemes of various scales used in survey one: eco-political factors, kahan scale and acceptance scale
codedmodule1 <- module1 %>%

#remove row 1 and 2 
  filter(!row_number() %in% c(1,2)) %>%

# replace risky likert scale with numbers
  mutate_at(vars(starts_with("Risky")), funs(case_when(. =="Not at all risky" ~ 1, 
                                                       . =="Slightly risky" ~ 2, 
                                                       . =="Moderately risky" ~ 3, 
                                                       . =="Very risky" ~ 4, 
                                                       . =="Extremely risky" ~ 5))) %>%

# replace beneficial likert scale with numbers  
  mutate_at(vars(starts_with("Ben")), funs(case_when(. =="Not at all beneficial" ~ 1,
                                                     . =="Slightly beneficial" ~ 2,
                                                     . =="Moderately beneficial" ~ 3,
                                                     . =="Very beneficial" ~ 4,
                                                     . =="Extremely beneficial" ~ 5 ))) %>%

# replace nuclear acceptance likert scale with numbers
  mutate_at(vars(N_accept,N_reluctantlyaccept,N_reject), funs(case_when(. == "Strongly disagree" ~ 1, 
                                                                        . == "Somewhat disagree" ~ 2,
                                                                        . == "Neither agree nor disagree" ~3,
                                                                        . == "Somewhat agree" ~ 4,
                                                                        . == "Strongly agree" ~ 5)))%>%
  mutate_all(~ifelse(. == "Rather not say/ Don't know", "Rather not say/Don't know", .))%>%
  mutate_all(~ifelse(. == "West bengal", "West Bengal", .))%>%
  mutate_all(~ifelse(. == "WEST BENGAL", "West Bengal", .))%>%
  mutate_all(~ifelse(. == "Up", "Uttar Pradesh", .))%>%
  mutate_all(~ifelse(. == "Tamilnadu", "Tamil Nadu", .))%>%
  mutate_all(~ifelse(. == "UP", "Uttar Pradesh", .))

```

```{r, echo = FALSE}

suppressPackageStartupMessages(library(tidyselect))
# combine codedmodule1 and codedmodule2

numcol2 <- codedmodule2 %>% add_column(Respondent = 1:1099, .before = "Language") 
numcol1 <- codedmodule1 %>% add_column(Respondent = 1099:2160, .before = "Language")

joineddf <-  full_join(numcol2, numcol1, by = c("Respondent", "Language", "Survey_Date", "State", "Risky_Hydro", "Risky_Solar", "Risky_Wind", "Risky_Nuclear", "Risky_Coal", "Risky_Gas", "Risky_Oil", "Risky_INDHydro", "Risky_INDSolar", "Risky_INDWind", "Risky_INDBiogas", "Risky_INDDiesel", "Risky_INDKerosene", "Risky_INDFirewoodetc", "Risky_INDLPG", "Ben_Hydro", "Ben_Solar", "Ben_Wind", "Ben_Nuclear", "Ben_Coal", "Ben_Gas", "Ben_Oil", "Ben_INDHydro", "Ben_INDSolar", "Ben_INDWind", "Ben_INDDiesel", "Ben_INDBiogas", "Ben_INDFirewoodetc", "Ben_INDLPG", "Ben_INDKerosene", "N_accept", "N_reluctantlyaccept", "N_reject", "age", "urban_rural", "gender", "caste", "religion", "religiontext"))

numcolb <- module2 %>%   filter(!row_number() %in% c(1,2))%>% 
                        add_column(Respondent = 1:1099, .before = "Language") 
numcola <- module1 %>%   filter(!row_number() %in% c(1,2)) %>% 
                        add_column(Respondent = 1099:2160, .before = "Language")

uncodedjoineddf <- full_join(numcolb, numcola, by = c("Respondent", "Language", "Survey_Date", "State", "Risky_Hydro", "Risky_Solar", "Risky_Wind", "Risky_Nuclear", "Risky_Coal", "Risky_Gas", "Risky_Oil", "Risky_INDHydro", "Risky_INDSolar", "Risky_INDWind", "Risky_INDBiogas", "Risky_INDDiesel", "Risky_INDKerosene", "Risky_INDFirewoodetc", "Risky_INDLPG", "Ben_Hydro", "Ben_Solar", "Ben_Wind", "Ben_Nuclear", "Ben_Coal", "Ben_Gas", "Ben_Oil", "Ben_INDHydro", "Ben_INDSolar", "Ben_INDWind", "Ben_INDDiesel", "Ben_INDBiogas", "Ben_INDFirewoodetc", "Ben_INDLPG", "Ben_INDKerosene", "N_accept", "N_reluctantlyaccept", "N_reject", "age", "urban_rural", "gender", "caste", "religion", "religiontext")) %>%
  mutate_all(~ifelse(. == "Rather not say/ Don't know", "Rather not say/Don't know", .))%>%
  mutate_all(~ifelse(. == "West bengal", "West Bengal", .))%>%
  mutate_all(~ifelse(. == "WEST BENGAL", "West Bengal", .))%>%
  mutate_all(~ifelse(. == "Up", "Uttar Pradesh", .))%>%
  mutate_all(~ifelse(. == "Tamilnadu", "Tamil Nadu", .))%>%
  mutate_all(~ifelse(. == "UP", "Uttar Pradesh", .))

```

\newpage
# State Specific Differences.

We see huge differences for perceived risk and perceived benefit for each technology from state to state. These graphs explore that. 
The number of respondents from each state are also reported below.

## Perceived risk by State

```{r, echo=FALSE, fig.width=11, fig.height=14, out.width='80%', out.height='80%'}

# part 2: state specific data on perceived risk
state_risk <- uncodedjoineddf %>%
    dplyr::select(starts_with(c("RISKY", "State")))%>%
    gather(Technology, Perceived_Risk, Risky_Hydro:Risky_INDLPG, factor_key = TRUE)%>%
    separate(Technology, c("Risky", "Technology"), extra = "merge", fill = "left")%>%
    dplyr::select(-Risky) %>% 
    group_by(State, Technology, Perceived_Risk) %>%
    dplyr::summarise(n=n())

count<- uncodedjoineddf %>%
  dplyr::select(starts_with(c("RISKY", "State")))%>%
   group_by(State) %>%
  dplyr::summarise(n=n())

 state_risk_percent <- state_risk %>%
    group_by(Technology, State) %>%
    mutate(Percentage = round(n/sum(n)*100, 0)) %>%
    ungroup()

 # Create a named vector with the number of respondents for each state
respondents_per_state <- c("Maharashtra" = 653, "Rajasthan" = 677, "Tamil Nadu" = 320, "Uttar Pradesh" = 125, "West Bengal" = 386)

respondents_per_state

state_risk %>% ggplot(aes(fill= fct_rev(factor(Perceived_Risk, levels=c("Extremely risky", "Very risky","Moderately risky","Slightly risky","Not at all risky","Rather not say/Don't know"))), y=n , x=Technology)) + 
    geom_bar(position="fill",stat="identity")+
     geom_text(data = state_risk_percent, aes(label=paste0(Percentage, "%"), y=n), position = position_fill(vjust=0.5), size=5) +
    scale_x_discrete(limits = c("Solar", "Wind" ,"Hydro", "Oil", "Coal","Gas", "Nuclear"))+
    scale_fill_manual("legend", values = c("Extremely risky" = "#8C4646", "Very risky" = "#D9695F", "Moderately risky" = "#F2A679", "Slightly risky" = "#F2D091","Not at all risky" = "#5D8C7B", "Rather not say/Don't know" = "grey94"))+
    theme_classic()+ 
    facet_wrap(~ State, ncol = 2)+
    labs(x = "mean perceived risk", y = "energy technologies")+
    theme(axis.text.x = element_text(angle = 45, hjust = 1))+
    theme(axis.text.x = element_text(angle = 45, hjust = 1, size=12),  # Increase x-axis text size
        axis.text.y = element_text(size=12),  # Increase y-axis text size
        strip.text = element_text(size=12))  # Increase facet label text size

```

\newpage

## Perceived Benefit by State

```{r, echo=FALSE, fig.width=12, fig.height=15, out.width='100%', out.height='100%'}
# part 2: state specific data on perceived benefit

state_ben <- uncodedjoineddf %>%
  dplyr::select(starts_with(c("Ben", "State")))%>%
  gather(Technology, Perceived_Benefit, Ben_Hydro:Ben_INDKerosene, factor_key = TRUE)%>%
  separate(Technology, c("Ben", "Technology"), extra = "merge", fill = "left")%>%
  dplyr::select(-Ben) %>% 
  group_by(State, Technology, Perceived_Benefit) %>%
  dplyr::summarise(n=n())

countben<- uncodedjoineddf %>%
  dplyr::select(starts_with(c("Ben", "State")))%>%
   group_by(State) %>%
  dplyr::summarise(n=n())

 # Create a named vector with the number of respondents for each state
respondents_per_state <- c("Maharashtra" = 653, "Rajasthan" = 677, "Tamil Nadu" = 320, "Uttar Pradesh" = 125, "West Bengal" = 386)

respondents_per_state

# Calculate percentages in each group
state_ben_percent <- state_ben %>%
  group_by(Technology, State) %>%
  mutate(Percentage = round(n/sum(n)*100, 0)) %>%
  ungroup()

#perceived benefit centrally owned technologies by state
state_ben %>% ggplot(aes(fill= factor(Perceived_Benefit, levels=c("Rather not say/Don't know","Not at all beneficial","Slightly beneficial","Moderately beneficial","Very beneficial", "Extremely beneficial")), y=n , x=Technology)) + 
  geom_bar(position="fill", stat = "identity")+
  geom_text(data = state_ben_percent, aes(label=paste0(Percentage, "%"), y=n), position = position_fill(vjust=0.5), size=5) +
  scale_fill_manual("legend", values = c("Rather not say/Don't know" = "grey94" ,"Not at all beneficial" = "orange","Slightly beneficial"="yellow" ,"Moderately beneficial" = "palegreen","Very beneficial"= "green3", "Extremely beneficial"= "green4"))+
  scale_x_discrete(limits = c("Solar", "Wind", "Hydro", "Oil", "Coal","Gas","Nuclear"))+
  theme_classic()+
  facet_wrap(~ State, ncol = 2)+
  labs(x = "mean perceived benefit", y = "energy technologies")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size=12), 
        axis.text.y = element_text(size=12), 
        strip.text = element_text(size=12))

```

```{r}
staterisky <- joineddf %>%
  dplyr::select(!starts_with(c("Ben_IND", "Risky_IND")))%>%
  dplyr::select(starts_with("Risky"), State)

staterisky_mean <- staterisky %>%
  group_by(State) %>%
  summarise(
    Mean_Risky_Hydro = mean(Risky_Hydro, na.rm = TRUE),
    Mean_Risky_Solar = mean(Risky_Solar, na.rm = TRUE),
    Mean_Risky_Wind = mean(Risky_Wind, na.rm = TRUE),
    Mean_Risky_Nuclear = mean(Risky_Nuclear, na.rm = TRUE),
    Mean_Risky_Coal = mean(Risky_Coal, na.rm = TRUE),
    Mean_Risky_Gas = mean(Risky_Gas, na.rm = TRUE),
    Mean_Risky_Oil = mean(Risky_Oil, na.rm = TRUE)
  )

str(staterisky_mean)

# Load required packages
library(dplyr)
library(tidyr)
library(ggplot2)

# Convert to long format
long_data <- staterisky_mean %>%
  gather(Technology, Mean_Score, -State) %>%
  mutate(Technology = gsub("Mean_Risky_", "", Technology))

# Order Technologies
technology_order <- c("Nuclear", "Gas", "Coal", "Oil", "Hydro", "Wind", "Solar")
long_data$Technology <- factor(long_data$Technology, levels = technology_order)

# Plot the graph
ggplot(long_data, aes(x = Technology, y = Mean_Score, color = State, group = State)) +
  geom_line(size = 1) +
  geom_point(size = 3) +
  labs(title = "Mean Perceived Risk by State",
       x = "Technology", y = "Mean Score") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

```{r}
# Using dplyr and tidyr

standardized_data <- long_data %>%
  group_by(Technology) %>%
  mutate(
    Overall_Mean = mean(Mean_Score, na.rm = TRUE),
    Std_Dev = sd(Mean_Score, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  mutate(Standardized_Score = (Mean_Score - Overall_Mean) / Std_Dev) %>%
  dplyr::select(-Overall_Mean, -Std_Dev)  #

# Order Technologies
technology_order <- c("Nuclear", "Gas", "Coal", "Oil", "Hydro", "Wind", "Solar")
standardized_data$Technology <- factor(standardized_data$Technology, levels = technology_order)

# Plot the graph
ggplot(standardized_data, aes(x = Technology, y = Standardized_Score, color = State, group = State)) +
  geom_line(size = 1) +
  geom_point(size = 3) +
  labs(title = "Standardised Scores of Technologies by State",
       x = "Technology", y = "Z Score") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

```{r}
# Switch roles of State and Technology for plotting
ggplot(standardized_data, aes(x = State, y = Standardized_Score, color = Technology, group = Technology)) +
  geom_line(size = 1) +
  geom_point(size = 3) +
  labs(title = "Standardised Scores of Technologies by State",
       x = "State", y = "Z Score") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

```{r, echo = FALSE}

# combining risk and benefit from each energy technology

acceptance <- joineddf_subset %>%
  na.omit()%>%
  mutate(Nuclear = Ben_Nuclear - Risky_Nuclear,
         Hyrdo = Ben_Hydro - Risky_Hydro,
         Wind = Ben_Wind - Risky_Wind,
         Solar = Ben_Solar - Risky_Solar,
         Coal = Ben_Coal - Risky_Coal,
         Gas = Ben_Gas - Risky_Gas,
         Oil = Ben_Oil - Risky_Oil)%>%
  select(-starts_with(c("Risky", "Ben")))

#str(acceptance)

# Compute the mean of each column
mean_acceptance <- data.frame(col = names(acceptance), mean_val = sapply(acceptance, mean))

# Compute standard error (not deviation)(divided by root of n)
sem <- function(x) sqrt(var(x, na.rm = TRUE) / length(na.omit(x)))
sem_acceptance <- data.frame(col = names(acceptance), sem_val = sapply(acceptance, sem))

mean_acceptance <- merge(mean_acceptance, sem_acceptance, by = "col")

# Plot
ggplot(mean_acceptance, aes(x = col, y = mean_val, fill = ifelse(mean_val >= 0, 'Above 0', 'Below 0'))) + 
  geom_bar(stat = "identity") +
  geom_errorbar(aes(ymin = mean_val - sem_val, ymax = mean_val + sem_val), width = 0.2) +
  geom_text(aes(label = round(mean_val, 2)), vjust = -0.9) +
  scale_fill_manual(values = c('Above 0' = 'palegreen3', 'Below 0' = 'firebrick2')) +
  labs(fill = "Legend") +
  xlab("Energy Technologies") +
  ylab("Mean Value") +
  ggtitle("Benefit- Risk for Each Energy Technology") +
  theme_minimal()+
  ylim(-0.5, 1.75)

```
\newpage

# Solar Energy - Linear Regression models

```{r, echo=FALSE}

salldemos <- joineddf %>%
            dplyr::select(Risky_Solar,Uppercaste, Male, Hindu, Urban, age, State)%>%
            na.omit()

salldemoslm1 <- lm(Risky_Solar ~ Uppercaste + Male + Hindu + Urban, data = salldemos)
# Print model summary
#summary(salldemoslm1)
#plot(alldemoslm1)

# lm suggests there is no significant white male effect or upper caste male effect for risk perception from nuclear energy in India. 
# the coefficients are very small and so is the r squared - rsquared is minimal.

# 2) lm with all demographic variables
salldemoslm2<- lm(Risky_Solar ~ Uppercaste + Male + Hindu + Urban + age + State, data = salldemos)
# Print model summary
#summary(salldemoslm2)
#plot(alldemoslm1)

```

## Only demographic variables

```{r, echo=FALSE, results='asis'}
library(stargazer)

#stargazer code for all demographic variable lms

cat("\\begingroup")
cat("\\setlength{\\tabcolsep}{1pt}")
cat("\\renewcommand{\\arraystretch}{0.7}")
stargazer(salldemoslm1, salldemoslm2, 
          title = "Results from 2 linear regression models",
          type = "latex"
)
cat("\\endgroup")

```

## Kahan Scale

```{r, echo = FALSE}
# kahan scores for solar 
sfascales <- codedmodule2 %>%
  dplyr::select(starts_with(c("K_","Risky_Solar","WEALTHLIM","MECHANISATION","DECISIONDECEN","DECISIONCEN","INDUSTRYLARGE","INDUSTRYSMALL","ECONOMYGLOBAL","ECONOMYLOCAL", "ENVOVERDEV","DEVOVERENV","OWNERPVT","OWNERNOREG","OWNERPUB", "OWNERREG","DISPLACESOLAR",  "POLLUTESOLAR", "HEALTHSOLAR", "JOBSSOLAR", "BEAUTYSOLAR", "PRIDESOLAR", "NPRIDESOLAR", "DEVSOLAR", "PROSPERSOLAR", "RELYSOLAR", "Uppercaste", "Male", "Hindu", "Urban", "State","age")))%>%
  na.omit()

sKahan_fas <- '
  KahanS =~ K_SHARM + K_SLIMCHOI + K_SPROTECT 
  KahanH =~ K_HEQUAL+ K_ERADEQ1 + K_EWEALTH + K_ERADEQ2
'

# Run CFA
sKahan_cfa <- cfa(sKahan_fas, data = sfascales, std.lv = TRUE)

# Print the summary of the CFA results, including fit measures
#summary(sKahan_cfa, fit.measure= TRUE, standardized = TRUE)

# RMSEA - between 0.05 and 0.08 (reasonable fit), less than 0.05 close fit and greater than 0.10 is poor fit 
# CFI - greater than 0.90 or 0.95 indicates good fit
#TFI - CFI>TFI >0.90 - indicates good fit
#pvalue should be greater than 0.05 p > 0.0.5

# creating table of factor scores

sKahan_scores <- predict(sKahan_cfa, newdata = sfascales)

sKahan_Kscores <- cbind(sfascales, sKahan_scores)

# 3) lm with kahan 
sm1_kahan <- lm(Risky_Solar ~ Uppercaste + Male + Hindu + Urban + age + State + KahanS + KahanH , data = sKahan_Kscores)
# Print model summary
#summary(sm1_kahan)
#plot(sm1_kahan)

```

```{r, echo=FALSE, results='asis'}
library(stargazer)

cat("\\begingroup")
cat("\\setlength{\\tabcolsep}{1pt}")
cat("\\renewcommand{\\arraystretch}{0.7}")
stargazer(salldemoslm1, salldemoslm2, sm1_kahan,
          title = "Results from 3 linear regression models",
          type = "latex"
)
cat("\\endgroup")

```

\newpage

# Coal

```{r, echo=FALSE}

calldemos <- joineddf %>%
            dplyr::select(Risky_Coal,Uppercaste, Male, Hindu, Urban, age, State)%>%
            na.omit()

calldemoslm1 <- lm(Risky_Coal ~ Uppercaste + Male + Hindu + Urban, data = calldemos)
# Print model summary
#summary(salldemoslm1)
#plot(alldemoslm1)

# lm suggests there is no significant white male effect or upper caste male effect for risk perception from nuclear energy in India. 
# the coefficients are very small and so is the r squared - rsquared is minimal.

# 2) lm with all demographic variables
calldemoslm2<- lm(Risky_Coal ~ Uppercaste + Male + Hindu + Urban + age + State, data = calldemos)
# Print model summary
#summary(salldemoslm2)
#plot(alldemoslm1)

```

## Only demographic variables

```{r, echo=FALSE, results='asis'}
library(stargazer)

#stargazer code for all demographic variable lms

cat("\\begingroup")
cat("\\setlength{\\tabcolsep}{1pt}")
cat("\\renewcommand{\\arraystretch}{0.7}")
stargazer(calldemoslm1, calldemoslm2, 
          title = "Results from 2 linear regression models",
          type = "latex"
)
cat("\\endgroup")

```

## Kahan Scale

```{r, echo = FALSE}
# kahan scores for solar 
cfascales <- codedmodule2 %>%
  dplyr::select(starts_with(c("K_","Risky_Coal","WEALTHLIM","MECHANISATION","DECISIONDECEN","DECISIONCEN","INDUSTRYLARGE","INDUSTRYSMALL","ECONOMYGLOBAL","ECONOMYLOCAL", "ENVOVERDEV","DEVOVERENV","OWNERPVT","OWNERNOREG","OWNERPUB", "OWNERREG","DISPLACESOLAR",  "POLLUTESOLAR", "HEALTHSOLAR", "JOBSSOLAR", "BEAUTYSOLAR", "PRIDESOLAR", "NPRIDESOLAR", "DEVSOLAR", "PROSPERSOLAR", "RELYSOLAR", "Uppercaste", "Male", "Hindu", "Urban", "State","age")))%>%
  na.omit()

cKahan_fas <- '
  KahanS =~ K_SHARM + K_SLIMCHOI + K_SPROTECT 
  KahanH =~ K_HEQUAL+ K_ERADEQ1 + K_EWEALTH + K_ERADEQ2
'

# Run CFA
cKahan_cfa <- cfa(cKahan_fas, data = cfascales, std.lv = TRUE)

# Print the summary of the CFA results, including fit measures
#summary(sKahan_cfa, fit.measure= TRUE, standardized = TRUE)

# RMSEA - between 0.05 and 0.08 (reasonable fit), less than 0.05 close fit and greater than 0.10 is poor fit 
# CFI - greater than 0.90 or 0.95 indicates good fit
#TFI - CFI>TFI >0.90 - indicates good fit
#pvalue should be greater than 0.05 p > 0.0.5

# creating table of factor scores

cKahan_scores <- predict(cKahan_cfa, newdata = cfascales)

cKahan_Kscores <- cbind(cfascales, cKahan_scores)

# 3) lm with kahan 
cm1_kahan <- lm(Risky_Coal ~ Uppercaste + Male + Hindu + Urban + age + State + KahanS + KahanH , data = cKahan_Kscores)
# Print model summary
#summary(cm1_kahan)
#plot(cm1_kahan)

```

```{r, echo=FALSE, results='asis'}
library(stargazer)

cat("\\begingroup")
cat("\\setlength{\\tabcolsep}{1pt}")
cat("\\renewcommand{\\arraystretch}{0.7}")
stargazer(calldemoslm1, calldemoslm2, cm1_kahan,
          title = "Results from 3 linear regression models",
          type = "latex"
)
cat("\\endgroup")

```