### 2024-01-15 FAoptions.Rmd
commiting after a long time. oh god

---
title: "FAoptions"
header-includes:
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
output:
  pdf_document: 
    toc: yes
    toc_depth: 2
    keep_tex: yes
  keep_md: yes
---

```{r load packages, warning = FALSE, message = FALSE, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(readxl))
suppressPackageStartupMessages(library(tsibble))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(psych))
suppressPackageStartupMessages(library(lavaan))
library(officer)
library(magrittr)

```

```{r, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
moduletwo <- read_excel(here::here("thesis_surveydata/political_factors_complete.xlsx"))
save(moduletwo, file = "moduletwo.rda")
module2 <- data.frame(moduletwo)
module2[module2 == "Rather not say/ Don't know"] <- "Rather not say/Don't know"
#module2 %>% filter(!row_number() %in% c(1,2))

# loaded data set for module 2 eco -pol variables and removes the rows with question number and question text
```

```{r, eval =TRUE, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
#This chunk contains the coding schemes of various scales used in survey one: eco-political factors, kahan scale and acceptance scale
codedmodule2 <- module2 %>%

#remove row 1
  filter(!row_number() %in% c(1,2)) %>%

# replace risky likert scale with numbers
  mutate_at(vars(starts_with("Risky")), funs(case_when(. =="Not at all risky" ~ 1, 
                                                       . =="Slightly risky" ~ 2, 
                                                       . =="Moderately risky" ~ 3, 
                                                       . =="Very risky" ~ 4, 
                                                       . =="Extremely risky" ~ 5))) %>%

# replace beneficial likert scale with numbers  
  mutate_at(vars(starts_with("Ben")), funs(case_when(. =="Not at all beneficial" ~ 1,
                                                     . =="Slightly beneficial" ~ 2,
                                                     . =="Moderately beneficial" ~ 3,
                                                     . =="Very beneficial" ~ 4,
                                                     . =="Extremely beneficial" ~ 5 ))) %>%

# replace nuclear acceptance likert scale with numbers
  mutate_at(vars(N_accept,N_reluctantlyaccept,N_reject), funs(case_when(. == "Strongly disagree" ~ 1, 
                                                                        . == "Somewhat disagree" ~ 2,
                                                                        . == "Neither agree nor disagree" ~3,
                                                                        . == "Somewhat agree" ~ 4,
                                                                        . == "Strongly agree" ~ 5))) %>%

# code likert scale for variables for Kahan scale into numbers
  mutate_at(vars(starts_with (c("K_S","K_H","DISPLACE", "POLLUTE", "HEALTH", "JOBS", "BEAUTY", "PRIDE", "NPRIDE", "DEVNUCLEAR","DEVSOLAR", "DEVCOAL", "PROSPER", "RELY"))), funs(case_when(. == "Strongly disagree" ~ 1, 
               . == "Somewhat disagree" ~ 2,
               . == "Neither agree nor disagree" ~3,
               . == "Somewhat agree" ~ 4,
               . == "Strongly agree" ~ 5))) %>%

# reverse code for likert scale for variables for Kahan scale into numbers
mutate_at(vars(starts_with (c("K_I","K_E"))), funs(case_when(. == "Strongly disagree" ~ 5, 
                                                               . == "Somewhat disagree" ~ 4,
                                                               . == "Neither agree nor disagree" ~3,
                                                               . == "Somewhat agree" ~ 2,
                                                               . == "Strongly agree" ~ 1))) %>%

# code eco-pol scale variables into numbers
  mutate_at(vars(SYSTEMDEMO,SYSTEMRELIGION,SYSTEMTECHNO,SYSTEMTOTAL,WEALTHLIM,MECHANISATION,DECISIONDECEN,INDUSTRYSMALL,ECONOMYLOCAL,ENVOVERDEV,OWNERPUB, OWNERREG), funs(case_when(. == "Strongly disagree" ~ 1, 
                        . == "Somewhat disagree" ~ 2,
                        . == "Neither agree nor disagree" ~3,
                        . == "Somewhat agree" ~ 4,
                        . == "Strongly agree" ~ 5))) %>%

# no - reverse code eco-pol scale variables into numbers 
 mutate_at(vars(DECISIONCEN,INDUSTRYLARGE,ECONOMYGLOBAL,OWNERPVT,OWNERNOREG, DEVOVERENV), funs(case_when(. == "Strongly disagree" ~ 1, 
                                                                                                         . == "Somewhat disagree" ~ 2,
                                                                                                         . == "Neither agree nor disagree" ~3,
                                                                                                         . == "Somewhat agree" ~ 4,
                                                                                                         . == "Strongly agree" ~ 5)))%>%
  mutate_all(~ifelse(. == "Rather not say/ Don't know", "Rather not say/Don't know", .))%>%
  mutate_all(~ifelse(. == "West bengal", "West Bengal", .))%>%
  mutate_all(~ifelse(. == "WEST BENGAL", "West Bengal", .))%>%
  mutate_all(~ifelse(. == "Up", "Uttar Pradesh", .))%>%
  mutate_all(~ifelse(. == "Tamilnadu", "Tamil Nadu", .))%>%
  mutate_all(~ifelse(. == "UP", "Uttar Pradesh", .))%>%
  mutate_at(vars(age), funs(case_when(. == "18-24 years old" ~ 1, 
                                      . == "25-34 years old" ~ 2,
                                      . ==  "35-44 years old" ~3,
                                      . == "45-54 years old" ~ 4,
                                      . == "55-64 years old" ~ 5,
                                      . == "65-74 years old" ~ 6,
                                      . == "75 years or older" ~ 7)))
```

```{r, echo = FALSE}

#sort demo vars

codedmodule2$State <- as.factor(codedmodule2$State)
codedmodule2$gender <- as.factor(codedmodule2$gender)
codedmodule2$Urban <- as.factor(codedmodule2$urban_rural)

codedmodule2$caste <- as.factor(codedmodule2$caste)
codedmodule2$religion <- as.factor(codedmodule2$religion)
# # Convert `caste` into binary variable 'upper' (1) vs 'non-upper' (0)
codedmodule2$Uppercaste <- ifelse(codedmodule2$caste %in% c('Brahmin', 'General'), 1, 0)
# Convert `gender` into binary variable 'male' (1) vs 'female' (0)
codedmodule2$Male <- ifelse(codedmodule2$gender == 'Male', 1, 0)
# Convert `religion` into binary variable 'Hinduism' (1) vs 'Non-Hinduism' (0)
codedmodule2$Hindu <- ifelse(codedmodule2$religion == 'Hinduism', 1, 0)

codedmodule21 <- codedmodule2 %>%
            dplyr::select(Risky_Nuclear, Risky_Solar, Risky_Coal, Uppercaste, Male, Hindu, Urban, age, State) %>%
            na.omit()

# lm on nuclear energy with all demovars

alldemos <- codedmodule21 %>%
            dplyr::select(Risky_Nuclear, Uppercaste, Male, Hindu, Urban, age, State) %>%
  mutate_at(vars(age), funs(case_when(. == "18-24 years old" ~ 1, 
                                      . == "25-34 years old" ~ 2,
                                      . ==  "35-44 years old" ~3,
                                      . == "45-54 years old" ~ 4,
                                      . == "55-64 years old" ~ 5,
                                      . == "65-74 years old" ~ 6,
                                      . == "75 years or older" ~ 7)))
```

```{r, echo=FALSE}
fascales <- codedmodule2 %>%
  dplyr::select(starts_with(c("K_","Risky_Nuclear","WEALTHLIM","MECHANISATION","DECISIONDECEN","DECISIONCEN","INDUSTRYLARGE","INDUSTRYSMALL","ECONOMYGLOBAL","ECONOMYLOCAL", "ENVOVERDEV","DEVOVERENV","OWNERPVT","OWNERNOREG","OWNERPUB", "OWNERREG","DISPLACENUCLEAR",  "POLLUTENUCLEAR", "HEALTHNUCLEAR", "JOBSNUCLEAR", "BEAUTYNUCLEAR", "PRIDENUCLEAR", "NPRIDENUCLEAR", "DEVNUCLEAR", "PROSPERNUCLEAR", "RELYNUCLEAR","SYSTEMDEMO","SYSTEMRELIGION", "SYSTEMTECHNO","SYSTEMTOTAL", "Uppercaste", "Male", "Hindu", "Urban", "State","age")))%>%
  mutate_at(vars(age), funs(case_when(. == "18-24 years old" ~ 1, 
                                      . == "25-34 years old" ~ 2,
                                      . ==  "35-44 years old" ~3,
                                      . == "45-54 years old" ~ 4,
                                      . == "55-64 years old" ~ 5,
                                      . == "65-74 years old" ~ 6,
                                      . == "75 years or older" ~ 7))) %>%
  na.omit()
```

```{r, echo=FALSE}
# Q6: Does the Kahan scale explain perception of risk from nuclear energy and other DVs ?

# we conducted fa on the kahan scale 
# CFA for Kahan scale. We shouldn't do FA here since this a well used scale with theoretical basis for factor distinctions and also previous studies using the same scale

library(lavaan)
library(semTools)
library(kableExtra)

fascales <- codedmodule2 %>%
  filter(
    rowSums(is.na(dplyr::select(., starts_with("K_")))) == 0 &  # This line ensures that rows with any NA in columns starting with         "K_" are   filtered out.
    !is.na(WEALTHLIM) & 
    !is.na(MECHANISATION) & 
    !is.na(DECISIONDECEN) & 
    !is.na(DECISIONCEN) & 
    !is.na(INDUSTRYLARGE) & 
    !is.na(INDUSTRYSMALL) & 
    !is.na(ECONOMYGLOBAL) & 
    !is.na(ECONOMYLOCAL) & 
    !is.na(ENVOVERDEV) & 
    !is.na(DEVOVERENV) & 
    !is.na(OWNERPVT) & 
    !is.na(OWNERNOREG) & 
    !is.na(OWNERPUB) & 
    !is.na(OWNERREG) & 
    !is.na(DISPLACENUCLEAR) & 
    !is.na(POLLUTENUCLEAR) & 
    !is.na(HEALTHNUCLEAR) & 
    !is.na(JOBSNUCLEAR) & 
    !is.na(BEAUTYNUCLEAR) & 
    !is.na(PRIDENUCLEAR) & 
    !is.na(NPRIDENUCLEAR) & 
    !is.na(DEVNUCLEAR) & 
    !is.na(PROSPERNUCLEAR) & 
    !is.na(RELYNUCLEAR)
  )

fascales2 <- codedmodule2 %>%dplyr::select(starts_with(c("K_","Risky_Nuclear","WEALTHLIM","MECHANISATION","DECISIONDECEN","DECISIONCEN","INDUSTRYLARGE","INDUSTRYSMALL","ECONOMYGLOBAL","ECONOMYLOCAL", "ENVOVERDEV","DEVOVERENV","OWNERPVT","OWNERNOREG","OWNERPUB", "OWNERREG","DISPLACENUCLEAR",  "POLLUTENUCLEAR", "HEALTHNUCLEAR", "JOBSNUCLEAR", "BEAUTYNUCLEAR", "PRIDENUCLEAR", "NPRIDENUCLEAR", "DEVNUCLEAR", "PROSPERNUCLEAR", "RELYNUCLEAR", "Uppercaste", "Male", "Hindu", "Urban", "State","age")))%>%na.omit()

Kahan_fas <- '
  KahanS =~ K_SHARM + K_SLIMCHOI + K_SPROTECT 
  KahanH =~ K_HEQUAL+ K_ERADEQ1 + K_EWEALTH + K_ERADEQ2
'

# Run CFA
Kahan_cfa <- cfa(Kahan_fas, data = fascales, std.lv = TRUE)

# Print the summary of the CFA results, including fit measures
#summary(Kahan_cfa, fit.measure= TRUE, standardized = TRUE)

# RMSEA - between 0.05 and 0.08 (reasonable fit), less than 0.05 close fit and greater than 0.10 is poor fit 
# CFI - greater than 0.90 or 0.95 indicates good fit
#TFI - CFI>TFI >0.90 - indicates good fit
#pvalue should be greater than 0.05 p > 0.0.5

# after having removed a few variables that had very low loading, the model is now a good fit, we have two latent variables - Hierarchy- egalitarian and Communitarian. Again we see that individualistic variables just don't fit here.

# creating table of factor scores

Kahan_scores <- predict(Kahan_cfa, newdata = fascales)

Kahan_Kscores <- cbind(fascales, Kahan_scores)

## RMSEA is still pretty big.

```

```{r, echo = FALSE}

# the psych code for FA on eco-pol scale for nuclear energy - all vars combined

library(psych)
library(tidyverse)

ecopolall <- fascales %>%
        dplyr::select("WEALTHLIM","MECHANISATION","DECISIONDECEN","DECISIONCEN","INDUSTRYLARGE","INDUSTRYSMALL","ECONOMYGLOBAL","ECONOMYLOCAL", "ENVOVERDEV","DEVOVERENV","OWNERPVT","OWNERNOREG","OWNERPUB", "OWNERREG","DISPLACENUCLEAR",  "POLLUTENUCLEAR", "HEALTHNUCLEAR", "BEAUTYNUCLEAR", "PRIDENUCLEAR", "NPRIDENUCLEAR", "DEVNUCLEAR", "RELYNUCLEAR","PROSPERNUCLEAR", "JOBSNUCLEAR")%>%
    rename(
      "Limits on Wealth" = "WEALTHLIM",
      "Anti Mechanisation of work" = "MECHANISATION",
      "Pro Decentralisation" = "DECISIONDECEN",
      "Pro Centralisation" = "DECISIONCEN",
      "Pro Large Industries" = "INDUSTRYLARGE",
      "Anti Large Industries" = "INDUSTRYSMALL",
      "Pro Globaleconomy" = "ECONOMYGLOBAL",
      "Pro Localeconomy" = "ECONOMYLOCAL",
      "Environment over Development" = "ENVOVERDEV",
      "Development over Environment" = "DEVOVERENV",
      "Pro Private ownership" = "OWNERPVT",
      "Anti Regulations" = "OWNERNOREG",
      "Pro Public ownership" = "OWNERPUB",
      "Pro Regulations" = "OWNERREG",
      "Displacement risk(nuclear)" = "DISPLACENUCLEAR",
      "Pollution risk(nuclear)" = "POLLUTENUCLEAR",
      "Health risk(nuclear)" = "HEALTHNUCLEAR",
      "Job opportunities(nuclear)" = "JOBSNUCLEAR",
      "Spoils Natural Beauty(nuclear)" = "BEAUTYNUCLEAR",
      "Local prosperity(nuclear)" = "PROSPERNUCLEAR",
      "Community pride(nuclear)" = "PRIDENUCLEAR",
      "National pride(nuclear)" = "NPRIDENUCLEAR",
      "National development(nuclear)" = "DEVNUCLEAR",
      "Reliance on government(nuclear)" = "RELYNUCLEAR"
    )

ecopolall2.fa <- fa(ecopolall, nfactors = 2, rotate = "varimax")

#print.psych(ecopolall2.fa, cut = 0.4, sort = TRUE)

ecopol_fas <- '
  Pdevelop =~ HEALTHNUCLEAR + BEAUTYNUCLEAR + DISPLACENUCLEAR + POLLUTENUCLEAR + INDUSTRYSMALL + OWNERREG + MECHANISATION
  Ndevelop =~ DEVNUCLEAR + PROSPERNUCLEAR + PRIDENUCLEAR + NPRIDENUCLEAR + JOBSNUCLEAR
'
ecopol_cfa <- cfa(ecopol_fas, data = fascales, std.lv = TRUE)
#summary(ecopol_cfa, fit.measure= TRUE, standardized = TRUE)

```

\blandscape
# FA on Combined Scale (2 Factor Solution): People Centered Development and Nationalist Development 
## Nuclear Energy: EFA on Combined Scale (2 Factor Solution)

```{r, echo=FALSE}

library(flextable)

flex <- function(data, title=NULL, col_widths=NULL) {
  tbl <- flextable(data) %>%
           #padding(padding = 1, part = "all") %>%
           #set_table_properties(layout = "autofit") %>%
           fontsize(size=10, part="all") %>%
           set_caption(title, 
                       autonum = officer::run_autonum(seq_id = "tab", 
                                                      pre_label = "Table ", 
                                                      post_label = "\n", 
                                                      bkm = "anytable"))

  # If col_widths is provided, set column widths
  if(!is.null(col_widths)) {
    for(col in names(col_widths)){
      tbl <- width(tbl, j = col, width = col_widths[[col]])
    }
  }

  return(tbl)
}

rename_map <- c(
  "Limits on Wealth" = "A limit should be put to how much wealth a person can amass",
  "Anti Mechanisation of work" = "Rapid mechanization of work is taking away jobs from workers in this country",
  "Pro Decentralisation" = "Local politicians shouldn't have to ask permission from the central government to implement policies",
  "Pro Centralisation" = "Laws and policies would be implemented more smoothly if more power lay with the central government",
  "Pro Large Industries" = "Large scale industries are required for the development of the country that will benefit everyone",
  "Anti Large Industries" = "Large corporations are destroying the local industries in India and benefiting only a handful of people",
  "Pro Globaleconomy" = "Foreign companies have led to a range of benefits for the Indian people and society",
  "Pro Localeconomy" = "India would be better off if foreign companies didn't come to here",
  "Environment over Development" = "Polluting industries that spoil the environment should be shut down even if it costs people their jobs",
  "Development over Environment" = "Economic growth and creating jobs should be prioritized over environmental protection",
  "Pro Private ownership" = "All businesses and industries should be owned privately",
  "Anti Regulations" = "There is too much red-tape and the government should not interfere with businesses and industries",
  "Pro Public ownership" = "The government should own most large businesses and industries",
  "Pro Regulations" = "Regardless of ownership, the government should pass strong regulations and implement them",
  "Displacement risk(nuclear)" = "Nuclear energy is leading to displacement of people from their land",
  "Spoils Natural Beauty(nuclear)" = "Nuclear energy spoils the natural beauty of the landscape",
  "Pollution risk(nuclear)" = "Nuclear energy increases pollution of air/water/land",
  "Health risk(nuclear)" = "Nuclear energy poses a great risk to the health of people living around it",
  "Job opportunities(nuclear)" = "Nuclear energy will bring jobs to the local community",
  "Community pride(nuclear)" = "I would be proud if my community used nuclear energy",
  "National pride(nuclear)" = "Nuclear energy is a mark of pride for our nation",
  "National development(nuclear)" = "Nuclear energy pushes forward the country's development",
  "Local prosperity(nuclear)" = "Nuclear energy brings economic prosperity to the surrounding regions",
  "Reliance on government(nuclear)" = "I don't like the idea that I have to rely on the government for electricity from nuclear energy"
)

fa_table <- function(x, cut) {
  #get sorted loadings
  loadings <- fa.sort(x)$loadings %>% round(3)
  #suppress loadings
  loadings[loadings > -cut & loadings < cut] <- ""
  #get additional info  
  add_info <- cbind(x$communalities, 
                    x$uniquenesses,
                    x$complexity) %>%
    # make it a data frame
    as.data.frame() %>%
    # column names
    rename("Communality" = V1,
           "Uniqueness" = V2,
           "Complexity" = V3) %>%
    #get the item names from the vector
    rownames_to_column("item")

  #build table
  loadings %>%
    unclass() %>%
    as.data.frame() %>%
    rownames_to_column("item") %>%
    left_join(add_info) %>%
    mutate(
      item = if_else(item %in% names(rename_map), rename_map[item], item),
      code = names(rename_map)[match(item, rename_map)], # extracting short form
      across(where(is.numeric), round, 3)
    ) %>%
    dplyr::select(code, everything()) # reorder columns to bring code to the front
}

fa_table(ecopolall2.fa, .4) %>%
  rename(Pdevelop = MR1, Ndevelop = MR2, Items= item) %>%
  flex("2 factor solution", col_widths = list(Items = 4.5, code = 1.5))

```

\elandscape

```{r, echo = FALSE}
ecopolall2.fa[["Vaccounted"]] %>%
  as.data.frame() %>%
  rownames_to_column("Property") %>%
  rename(Pdevelop = MR1, Ndevelop = MR2) %>%
    mutate(across(where(is.numeric), round, 3)) %>%
    flex("Eigenvalues and Variance Explained for Combined Scale (2 Factor Solution)")
```

```{r, echo=FALSE}

## Nuclear - linear regression with 2 Factor solution

# Linear Model with 2 factor solution for  Risk perception from Nuclear Energy

ecopolall_scores <- psych::factor.scores(ecopolall, ecopolall2.fa)

# The factor scores will be stored in the scores element of the returned list
scores1 <- ecopolall_scores$scores

#binding factor scores to the dataset

fascale_scores <- cbind(Kahan_Kscores, scores1) %>%
  rename(Pdevelop = MR1)%>%
  rename(Ndevelop = MR2)

nuclear2fa <- lm(Risky_Nuclear ~ age + Uppercaste + Male + Hindu + Urban + State + KahanS + KahanH + Pdevelop + Ndevelop, data = fascale_scores)

#m2_allscales <- lm(Risky_Nuclear ~ age + Uppercaste + Male + Hindu + Urban +State+ KahanS + KahanH + NPRIDENUCLEAR + PRIDENUCLEAR + DEVNUCLEAR + PROSPERNUCLEAR, data = fascale_scores)
# Print model summary
#summary(nuclear2fa)

```
```{r,echo = FALSE}
acceptance2 <- fascale_scores %>%
  mutate(Nuclear = Ben_Nuclear - Risky_Nuclear,
         Hyrdo = Ben_Hydro - Risky_Hydro,
         Wind = Ben_Wind - Risky_Wind,
         Solar = Ben_Solar - Risky_Solar,
         Coal = Ben_Coal - Risky_Coal,
         Gas = Ben_Gas - Risky_Gas,
         Oil = Ben_Oil - Risky_Oil)

nuclearacceptlm <- lm(Nuclear ~ age + Uppercaste + Male + Hindu + Urban + State + KahanS + KahanH + Pdevelop + Ndevelop, data = acceptance2)

#m2_allscales <- lm(Risky_Nuclear ~ age + Uppercaste + Male + Hindu + Urban +State+ KahanS + KahanH + NPRIDENUCLEAR + PRIDENUCLEAR + DEVNUCLEAR + PROSPERNUCLEAR, data = fascale_scores)
# Print model summary
#summary(nuclearacceptlm)

ggplot(acceptance2, aes(x = Pdevelop, y = Nuclear)) +
  geom_point() +  # or geom_line() depending on how you want to visualize it
  facet_wrap(~ State) +  # This will create separate plots for each State
  labs(title = "Net benefit Nuclear vs Pdevelop by State",
       x = "Pdevelop",
       y = "net ben Nuclear") +
  theme_minimal()

```

```{r, echo=FALSE}
library(psych)
library(tidyverse)

sfascales <- codedmodule2 %>%
  dplyr::select(starts_with(c("K_","Risky_Solar","WEALTHLIM","MECHANISATION","DECISIONDECEN","DECISIONCEN","INDUSTRYLARGE","INDUSTRYSMALL","ECONOMYGLOBAL","ECONOMYLOCAL", "ENVOVERDEV","DEVOVERENV","OWNERPVT","OWNERNOREG","OWNERPUB", "OWNERREG","DISPLACESOLAR", "POLLUTESOLAR", "HEALTHSOLAR", "JOBSSOLAR", "BEAUTYSOLAR", "PRIDESOLAR", "NPRIDESOLAR", "DEVSOLAR", "PROSPERSOLAR", "RELYSOLAR", "Uppercaste", "Male", "Hindu", "Urban", "State","age")))%>%
  na.omit()

secopolall <- sfascales %>%
        dplyr::select("WEALTHLIM","MECHANISATION","DECISIONDECEN","DECISIONCEN","INDUSTRYLARGE","INDUSTRYSMALL","ECONOMYGLOBAL","ECONOMYLOCAL", "ENVOVERDEV","DEVOVERENV","OWNERPVT","OWNERNOREG","OWNERPUB", "OWNERREG","DISPLACESOLAR", "POLLUTESOLAR", "HEALTHSOLAR", "JOBSSOLAR", "BEAUTYSOLAR", "PRIDESOLAR", "NPRIDESOLAR", "DEVSOLAR", "PROSPERSOLAR", "RELYSOLAR")%>%
    rename(
      "Limits on Wealth " = "WEALTHLIM",
      "Anti Mechanisation of work" = "MECHANISATION",
      "Pro Decentralisation" = "DECISIONDECEN",
      "Pro Centralisation" = "DECISIONCEN",
      "Pro Large Industries" = "INDUSTRYLARGE",
      "Anti Large Industries " = "INDUSTRYSMALL",
      "Pro Globaleconomy" = "ECONOMYGLOBAL",
      "Pro Localeconomy" = "ECONOMYLOCAL",
      "Environment over Development" = "ENVOVERDEV",
      "Development over Environment" = "DEVOVERENV",
      "Pro Private ownership " = "OWNERPVT",
      "Anti Regulations" = "OWNERNOREG",
      "Pro Public ownership" = "OWNERPUB",
      "Pro Regulations" = "OWNERREG",
      "Displacement risk(solar)" = "DISPLACESOLAR",
      "Pollution risk(solar) " = "POLLUTESOLAR",
      "Health risk(solar)" = "HEALTHSOLAR",
      "Job opportunities(solar)" = "JOBSSOLAR",
      "Spoils Natural Beauty(solar)" = "BEAUTYSOLAR",
      "Community pride(solar)" = "PRIDESOLAR",
      "National pride(solar)" = "NPRIDESOLAR",
      "National development(solar)" = "DEVSOLAR",
      "Local prosperity(solar)" = "PROSPERSOLAR",
      "Reliance on government(solar)" = "RELYSOLAR"
    )

secopolall2.fa <- fa(secopolall, nfactors = 2, rotate = "varimax")

```

```{r, echo = FALSE}

# SOLAR : CFA on kahan scale and eco-pol scale

# kahan scores for solar 
sfascales <- codedmodule2 %>%
    filter(
    rowSums(is.na(dplyr::select(., starts_with("K_")))) == 0 &  # This line ensures that rows with any NA in columns starting with         "K_" are   filtered out.
    !is.na(WEALTHLIM) & 
    !is.na(MECHANISATION) & 
    !is.na(DECISIONDECEN) & 
    !is.na(DECISIONCEN) & 
    !is.na(INDUSTRYLARGE) & 
    !is.na(INDUSTRYSMALL) & 
    !is.na(ECONOMYGLOBAL) & 
    !is.na(ECONOMYLOCAL) & 
    !is.na(ENVOVERDEV) & 
    !is.na(DEVOVERENV) & 
    !is.na(OWNERPVT) & 
    !is.na(OWNERNOREG) & 
    !is.na(OWNERPUB) & 
    !is.na(OWNERREG) & 
    !is.na(DISPLACESOLAR) & 
    !is.na(POLLUTESOLAR) & 
    !is.na(HEALTHSOLAR) & 
    !is.na(JOBSSOLAR) & 
    !is.na(BEAUTYSOLAR) & 
    !is.na(PRIDESOLAR) & 
    !is.na(NPRIDESOLAR) & 
    !is.na(DEVSOLAR) & 
    !is.na(PROSPERSOLAR) & 
    !is.na(RELYSOLAR)
  )

  #dplyr::select(starts_with(c("K_","Risky_Solar","WEALTHLIM","MECHANISATION","DECISIONDECEN","DECISIONCEN","INDUSTRYLARGE","INDUSTRYSMALL","ECONOMYGLOBAL","ECONOMYLOCAL", "ENVOVERDEV","DEVOVERENV","OWNERPVT","OWNERNOREG","OWNERPUB", "OWNERREG","DISPLACESOLAR",  "POLLUTESOLAR", "HEALTHSOLAR", "JOBSSOLAR", "BEAUTYSOLAR", "PRIDESOLAR", "NPRIDESOLAR", "DEVSOLAR", "PROSPERSOLAR", "RELYSOLAR", "Uppercaste", "Male", "Hindu", "Urban", "State","age")))%>%na.omit()

sKahan_fas <- '
  KahanS =~ K_SHARM + K_SLIMCHOI + K_SPROTECT 
  KahanH =~ K_HEQUAL+ K_ERADEQ1 + K_EWEALTH + K_ERADEQ2
'

# Run CFA
sKahan_cfa <- cfa(sKahan_fas, data = sfascales, std.lv = TRUE)

# Print the summary of the CFA results, including fit measures
#summary(sKahan_cfa, fit.measure= TRUE, standardized = TRUE)

sKahan_scores <- predict(sKahan_cfa, newdata = sfascales)

sKahan_Kscores <- cbind(sfascales, sKahan_scores)

sm1_kahan <- lm(Risky_Solar ~ Uppercaste + Male + Hindu + Urban + age + State + KahanS + KahanH , data = sKahan_Kscores)

## CFA on ecopol vars for solar energy

secopol_fas <- '
  Pdevelop =~ HEALTHSOLAR + BEAUTYSOLAR + DISPLACESOLAR + POLLUTESOLAR + INDUSTRYSMALL + OWNERREG + MECHANISATION
  Ndevelop =~ DEVSOLAR + PRIDESOLAR + NPRIDESOLAR + PROSPERSOLAR
'

secopol_cfa <- cfa(secopol_fas, data = sfascales, std.lv = TRUE)
summary(secopol_cfa, fit.measure= TRUE, standardized = TRUE)

# Cronbach's alpha test on Pdevelop and Ndevelop for Solar energy

solarpdevelop <- sfascales %>%
  dplyr::select(HEALTHSOLAR, BEAUTYSOLAR, DISPLACESOLAR, POLLUTESOLAR, INDUSTRYSMALL, OWNERREG, MECHANISATION)%>%
  na.omit()

solarpdevelop.alpha <- psych::alpha(solarpdevelop, check.keys = TRUE)

# alpha is strong at c!

solarndevelop <- sfascales %>%
  dplyr::select(DEVSOLAR, PRIDESOLAR, NPRIDESOLAR, PROSPERSOLAR)%>%
  na.omit()

solarndevelop.alpha <- psych::alpha(solarndevelop, check.keys = TRUE)

#alpha is strong at 0.86!

```

\newpage

## Solar Energy : CFA on Combined Scale (2 Factor Solution)

Solar People Centered Development alpha = 0.75

Solar Nationalist Development alpha = 0.86

```{r, echo=FALSE}

# Extract parameter estimates
sparams <- parameterEstimates(secopol_cfa, standardized = TRUE)%>%
   slice(1:(n()-15))

sparams1 <- sparams %>%
  mutate(rhs = case_when(
  rhs == "DISPLACESOLAR" ~ "Solar energy is leading to displacement of people from their land",
  rhs == "BEAUTYSOLAR" ~ "Solar energy spoils the natural beauty of the landscape",
  rhs == "POLLUTESOLAR" ~ "Solar energy increases pollution of air/water/land",
  rhs == "HEALTHSOLAR" ~ "Solar energy poses a great risk to the health of people living around it",
  rhs == "JOBSSOLAR" ~ "Solar energy will bring jobs to the local community",
  rhs == "PRIDESOLAR" ~ "I would be proud if my community used Solar energy",
  rhs == "NPRIDESOLAR" ~ "Solar energy is a mark of pride for our nation",
  rhs == "DEVSOLAR" ~ "Solar energy pushes forward the country's development",
  rhs == "PROSPERSOLAR" ~ "Solar energy brings economic prosperity to the surrounding regions",
  rhs == "RELYSOLAR" ~ "I don't like the idea that I have to rely on the government for electricity from Solar energy",
  rhs == "INDUSTRYSMALL" ~ "Large corporations are destroying the local industries in India and benefiting only a handful of people",
  rhs == "OWNERREG" ~ "Regardless of ownership, the government should pass strong regulations and implement them",
  rhs == "MECHANISATION" ~ "Rapid mechanization of work is taking away jobs from workers in this country",

    TRUE ~ as.character(rhs)
  )) %>%
  dplyr::select(!c(lhs,op, std.nox)) %>%
    rename(
    "Items" = rhs,
    "Loadings" = est,
    "Standard Error" = se,
    "zvalue" = z) %>%
  mutate(across(c("Loadings", "Standard Error", "zvalue"), round, 3))

sparams1$Scale <- NA
sparams1$Scale[1:7] <- "People Centered Development"
sparams1$Scale[8:10] <- "Nationalist Development"
sparams1 <- sparams1[, c("Scale", setdiff(names(sparams1), "Scale"))]
CFAsolar <- knitr::kable(sparams1, booktabs = TRUE, caption = "Confirmatory Factor Analysis(CFA) on Solar Energy") %>%
  kable_styling(latex_options = c("striped","scale_down", "hold_position", "repeat_header"))%>%
  column_spec(column = 2, width = "4cm")

# Print the table
CFAsolar %>%
  landscape()

```

```{r, echo=FALSE}

# Solar - linear regression with 2 Factor solution

# Linear Model with 2 factor solution for  Risk perception from Solar Energy

# scores from EFA
#secopolall_scores2 <- psych::factor.scores(secopolall, secopolall2.fa)

# The factor scores will be stored in the scores element of the returned list
#scores2 <- secopolall_scores2$scores

#binding factor scores to the dataset

#sfascale_scores2 <- cbind(sKahan_Kscores, scores2) %>%
  #rename(Pdevelop = MR1)%>%
  #rename(Ndevelop = MR2)
#solar2fa2 <- lm(Risky_Solar ~ age + Uppercaste + Male + Hindu + Urban + State + KahanS + KahanH + Pdevelop + Ndevelop, data = sfascale_scores2)

#summary(solar2fa2)
# lesson: doesn't matter if I use CFA or EFA scores- the difference is small and is probably happening due to slightly different samples - this brings to mind the problem of large missing values - maybe it's worth talking to a stat consultant about this issue

#scores from CFA 
#binding factor scores to the dataset
secopolall_scores <- predict(secopol_cfa, newdata = sfascales)

sfascale_scores <- cbind(sKahan_Kscores, secopolall_scores)

solar2fa <- lm(Risky_Solar ~ age + Uppercaste + Male + Hindu + Urban + State + KahanS + KahanH + Pdevelop + Ndevelop, data = sfascale_scores)

# Print model summary
summary(solar2fa)

```

```{r,echo = FALSE}
sacceptance2 <- sfascale_scores %>%
  mutate(Solar = Ben_Solar - Risky_Solar)

solaracceptlm <- lm(Solar ~ age + Uppercaste + Male + Hindu + Urban + State + KahanS + KahanH + Pdevelop + Ndevelop, data = sacceptance2)

#m2_allscales <- lm(Risky_Nuclear ~ age + Uppercaste + Male + Hindu + Urban +State+ KahanS + KahanH + NPRIDENUCLEAR + PRIDENUCLEAR + DEVNUCLEAR + PROSPERNUCLEAR, data = fascale_scores)
# Print model summary
#summary(solaracceptlm)

```

```{r, echo = FALSE}

cecopolall <- codedmodule2 %>%
        dplyr::select("WEALTHLIM","MECHANISATION","DECISIONDECEN","DECISIONCEN","INDUSTRYLARGE","INDUSTRYSMALL","ECONOMYGLOBAL","ECONOMYLOCAL", "ENVOVERDEV","DEVOVERENV","OWNERPVT","OWNERNOREG","OWNERPUB", "OWNERREG","DISPLACECOAL", "POLLUTECOAL", "HEALTHCOAL", "JOBSCOAL", "BEAUTYCOAL", "PRIDECOAL", "NPRIDECOAL", "DEVCOAL", "PROSPERCOAL", "RELYCOAL")%>%
      rename(
      "Limits on Wealth " = "WEALTHLIM",
      "Anti Mechanisation of work" = "MECHANISATION",
      "Pro Decentralisation" = "DECISIONDECEN",
      "Pro Centralisation" = "DECISIONCEN",
      "Pro Large Industries" = "INDUSTRYLARGE",
      "Anti Large Industries " = "INDUSTRYSMALL",
      "Pro Globaleconomy" = "ECONOMYGLOBAL",
      "Pro Localeconomy" = "ECONOMYLOCAL",
      "Environment over Development" = "ENVOVERDEV",
      "Development over Environment" = "DEVOVERENV",
      "Pro Private ownership " = "OWNERPVT",
      "Anti Regulations" = "OWNERNOREG",
      "Pro Public ownership" = "OWNERPUB",
      "Pro Regulations" = "OWNERREG",
      "Displacement risk(coal)" = "DISPLACECOAL",
      "Pollution risk(coal) " = "POLLUTECOAL",
      "Health risk(coal)" = "HEALTHCOAL",
      "Job opportunities(coal)" = "JOBSCOAL",
      "Spoils Natural Beauty(coal)" = "BEAUTYCOAL",
      "Community pride(coal)" = "PRIDECOAL",
      "National pride(coal)" = "NPRIDECOAL",
      "National development(coal)" = "DEVCOAL",
      "Local prosperity(coal)" = "PROSPERCOAL",
      "Reliance on government(coal)" = "RELYCOAL"
    )

cecopolall2.fa <- fa(cecopolall, nfactors = 2, rotate = "varimax")

#print.psych(cecopolall2.fa, cut = 0.5, sort = TRUE)

```

```{r, echo = FALSE}

## COAL: CFA on kahan scale and eco-pol scale

# kahan scores for coal and CFA on ecopol vars  
cfascales <- codedmodule2 %>%
      filter(
    rowSums(is.na(dplyr::select(., starts_with("K_")))) == 0 &  # This line ensures that rows with any NA in columns starting with         "K_" are   filtered out.
    !is.na(WEALTHLIM) & 
    !is.na(MECHANISATION) & 
    !is.na(DECISIONDECEN) & 
    !is.na(DECISIONCEN) & 
    !is.na(INDUSTRYLARGE) & 
    !is.na(INDUSTRYSMALL) & 
    !is.na(ECONOMYGLOBAL) & 
    !is.na(ECONOMYLOCAL) & 
    !is.na(ENVOVERDEV) & 
    !is.na(DEVOVERENV) & 
    !is.na(OWNERPVT) & 
    !is.na(OWNERNOREG) & 
    !is.na(OWNERPUB) & 
    !is.na(OWNERREG) & 
    !is.na(DISPLACECOAL) & 
    !is.na(POLLUTECOAL) & 
    !is.na(HEALTHCOAL) & 
    !is.na(JOBSCOAL) & 
    !is.na(BEAUTYCOAL) & 
    !is.na(PRIDECOAL) & 
    !is.na(NPRIDECOAL) & 
    !is.na(DEVCOAL) & 
    !is.na(PROSPERCOAL) & 
    !is.na(RELYCOAL)
  )

  #dplyr::select(starts_with(c("K_","Risky_Coal","WEALTHLIM","MECHANISATION","DECISIONDECEN","DECISIONCEN","INDUSTRYLARGE","INDUSTRYSMALL","ECONOMYGLOBAL","ECONOMYLOCAL", "ENVOVERDEV","DEVOVERENV","OWNERPVT","OWNERNOREG","OWNERPUB", "OWNERREG","DISPLACECOAL",  "POLLUTECOAL", "HEALTHCOAL", "JOBSCOAL", "BEAUTYCOAL", "PRIDECOAL", "NPRIDECOAL", "DEVCOAL", "PROSPERCOAL", "RELYCOAL", "Uppercaste", "Male", "Hindu", "Urban", "State","age"))) %>%na.omit()

cKahan_fas <- '
  KahanS =~ K_SHARM + K_SLIMCHOI + K_SPROTECT 
  KahanH =~ K_HEQUAL+ K_ERADEQ1 + K_EWEALTH + K_ERADEQ2
'

# Run CFA
cKahan_cfa <- cfa(cKahan_fas, data = cfascales, std.lv = TRUE)

cKahan_scores <- predict(cKahan_cfa, newdata = cfascales)

cKahan_Kscores <- cbind(cfascales, cKahan_scores)

cecopol_fas <- '
  Pdevelop =~ HEALTHCOAL + BEAUTYCOAL + DISPLACECOAL + POLLUTECOAL + INDUSTRYSMALL + OWNERREG + MECHANISATION
  Ndevelop =~ DEVCOAL + PROSPERCOAL + PRIDECOAL + NPRIDECOAL 
'

cecopol_cfa <- cfa(cecopol_fas, data = cfascales, std.lv = TRUE)
#summary(cecopol_cfa, fit.measure= TRUE, standardized = TRUE)

# 3) lm with kahan 
cm1_kahan <- lm(Risky_Coal ~ Uppercaste + Male + Hindu + Urban + age + State + KahanS + KahanH , data = cKahan_Kscores)
# Print model summary
#summary(cm1_kahan)
#plot(cm1_kahan)

# Cronbach's alpha test on Pdevelop and Ndevelop for Solar energy

coalpdevelop <- cfascales %>%
  dplyr::select(HEALTHCOAL, BEAUTYCOAL, DISPLACECOAL, POLLUTECOAL, INDUSTRYSMALL, OWNERREG, MECHANISATION)%>%
  na.omit()

coalpdevelop.alpha <- psych::alpha(coalpdevelop, check.keys = TRUE)

# alpha is strong at 0.78!

coalndevelop <- cfascales %>%
  dplyr::select(DEVCOAL, PRIDECOAL, NPRIDECOAL, PROSPERCOAL)%>%
  na.omit()

coalndevelop.alpha <- psych::alpha(coalndevelop, check.keys = TRUE)

#alpha n develop = 0.6

```

\newpage
## Coal : CFA on Combined Scale 2 Factor Solution (Pdevelop and Ndevelop)

Coal People Centered Development Alpha = 0.78

Coal Nationalist Development Alpha = 0.6

```{r, echo=FALSE}

# Extract parameter estimates
cparams <- parameterEstimates(cecopol_cfa, standardized = TRUE)%>%
   slice(1:(n()-15))

cparams1 <- cparams %>%
  mutate(rhs = case_when(
  rhs == "DISPLACECOAL" ~ "Coal powered plants is leading to displacement of people from their land",
  rhs == "BEAUTYCOAL" ~ "Coal powered plants spoils the natural beauty of the landscape",
  rhs == "POLLUTECOAL" ~ "Coal powered plants increases pollution of air/water/land",
  rhs == "HEALTHCOAL" ~ "Coal powered plants poses a great risk to the health of people living around it",
  rhs == "JOBSCOAL" ~ "Coal powered plants will bring jobs to the local community",
  rhs == "PRIDECOAL" ~ "I would be proud if my community used Coal powered plants",
  rhs == "NPRIDECOAL" ~ "Coal powered plants is a mark of pride for our nation",
  rhs == "DEVCOAL" ~ "Coal powered plants pushes forward the country's development",
  rhs == "PROSPERCOAL" ~ "Coal powered plants brings economic prosperity to the surrounding regions",
  rhs == "RELYCOAL" ~ "I don't like the idea that I have to rely on the government for electricity from Coal powered plants",
  rhs == "INDUSTRYSMALL" ~ "Large corporations are destroying the local industries in India and benefiting only a handful of people",
  rhs == "OWNERREG" ~ "Regardless of ownership, the government should pass strong regulations and implement them",
  rhs == "MECHANISATION" ~ "Rapid mechanization of work is taking away jobs from workers in this country",

    TRUE ~ as.character(rhs)
  )) %>%
  dplyr::select(!c(lhs,op, std.nox)) %>%
    rename(
    "Items" = rhs,
    "Loadings" = est,
    "Standard Error" = se,
    "zvalue" = z) %>%
  mutate(across(c("Loadings", "Standard Error", "zvalue"), round, 3))

cparams1$Scale <- NA
cparams1$Scale[1:7] <- "People Centered Development"
cparams1$Scale[8:10] <- "Nationalist Development"
cparams1 <- cparams1[, c("Scale", setdiff(names(sparams1), "Scale"))]
CFAcoal <- knitr::kable(cparams1, booktabs = TRUE, caption = "Confirmatory Factor Analysis(CFA) on Coal") %>%
  kable_styling(latex_options = c("striped","scale_down", "hold_position", "repeat_header"))%>%
  column_spec(column = 2, width = "4cm")

# Print the table
CFAcoal %>%
  landscape()

```

```{r, echo=FALSE}

## Coal linear regression - 2 FA solution

# Linear Model with 2 factor solution for  Risk perception from coal Energy

#binding factor scores to the dataset
cecopolall_scores <- predict(cecopol_cfa, newdata = cfascales)

cfascale_scores <- cbind(cKahan_Kscores, cecopolall_scores)# %>%
  #rename(Pdevelop = MR1)%>%
  #rename(Ndevelop = MR2)

coal2fa <- lm(Risky_Coal ~ age + Uppercaste + Male + Hindu + Urban + State + KahanS + KahanH + Pdevelop + Ndevelop, data = cfascale_scores)

# Print model summary
#summary(coal2fa)

```

```{r,echo = FALSE}
cacceptance2 <- cfascale_scores %>%
  mutate(Coal = Ben_Coal - Risky_Coal)

coalacceptlm <- lm(Coal ~ age + Uppercaste + Male + Hindu + Urban + State + KahanS + KahanH + Pdevelop + Ndevelop, data = cacceptance2)

#m2_allscales <- lm(Risky_Nuclear ~ age + Uppercaste + Male + Hindu + Urban +State+ KahanS + KahanH + NPRIDENUCLEAR + PRIDENUCLEAR + DEVNUCLEAR + PROSPERNUCLEAR, data = fascale_scores)
# Print model summary
#summary(coalacceptlm)

```

\newpage

## Linear Models: Perceived risk X Pdevelop and Ndevelop

```{r, echo=FALSE, results='asis'}
library(stargazer)

cat("\\begingroup")
cat("\\setlength{\\tabcolsep}{1pt}")
cat("\\renewcommand{\\arraystretch}{0.7}")
stargazer(nuclear2fa, solar2fa, coal2fa, 
         title = "Perceived Risk X Pdevelop and Ndevelope",
        type = "latex"
)
cat("\\endgroup")

```

```{r, echo =FALSE}

# same things with benefit from Nuclear Energy
fascalesben <- codedmodule2 %>%
  dplyr::select(starts_with(c("K_","Ben_Nuclear","WEALTHLIM","MECHANISATION","DECISIONDECEN","DECISIONCEN","INDUSTRYLARGE","INDUSTRYSMALL","ECONOMYGLOBAL","ECONOMYLOCAL", "ENVOVERDEV","DEVOVERENV","OWNERPVT","OWNERNOREG","OWNERPUB", "OWNERREG","DISPLACENUCLEAR",  "POLLUTENUCLEAR", "HEALTHNUCLEAR", "JOBSNUCLEAR", "BEAUTYNUCLEAR", "PRIDENUCLEAR", "NPRIDENUCLEAR", "DEVNUCLEAR", "PROSPERNUCLEAR", "RELYNUCLEAR", "Uppercaste", "Male", "Hindu", "Urban", "State","age")))%>%
  na.omit()

Kahan_fas <- '
  KahanS =~ K_SHARM + K_SLIMCHOI + K_SPROTECT 
  KahanH =~ K_HEQUAL+ K_ERADEQ1 + K_EWEALTH + K_ERADEQ2
'

# Run CFA
Kahan_cfaben <- cfa(Kahan_fas, data = fascalesben, std.lv = TRUE)

Kahan_scoresben <- predict(Kahan_cfaben, newdata = fascalesben)

Kahan_Kscoresben <- cbind(fascalesben, Kahan_scoresben)

ecopolallben <- fascalesben %>%
        dplyr::select("WEALTHLIM","MECHANISATION","DECISIONDECEN","DECISIONCEN","INDUSTRYLARGE","INDUSTRYSMALL","ECONOMYGLOBAL","ECONOMYLOCAL", "ENVOVERDEV","DEVOVERENV","OWNERPVT","OWNERNOREG","OWNERPUB", "OWNERREG","DISPLACENUCLEAR",  "POLLUTENUCLEAR", "HEALTHNUCLEAR", "JOBSNUCLEAR", "BEAUTYNUCLEAR", "PRIDENUCLEAR", "NPRIDENUCLEAR", "DEVNUCLEAR", "PROSPERNUCLEAR", "RELYNUCLEAR")

ecopolallben_alpha <- psych::alpha(ecopolallben, check.keys = TRUE)

#summary(ecopolall_alpha)

# when we combine both the scales the alpha is high at 0.785 (almost 0.8)

# FA on all ecopol variables except systems of governance

ecopolallben_cor <- cor(ecopolallben)
PCA24 <- principal(ecopolallben, nfactors = 24, rotate = "none") # no rotation to start with.

ecopolallben2.fa <- fa(ecopolallben, nfactors = 2, rotate = "varimax")

# I don't know still doesn't make a lot of sense to me- the two factors make more sense - 1) people centered development 2) nationalist development
# So I am going to extract factor scores and bind them to the dataset

ecopolallben_scores <- psych::factor.scores(ecopolallben, ecopolallben2.fa)

# The factor scores will be stored in the scores element of the returned list
scores2 <- ecopolallben_scores$scores

#binding factor scores to the dataset

fascale_scoresben <- cbind(Kahan_Kscoresben, scores2) %>%
  rename(Pdevelop = MR1)%>%
  rename(Ndevelop = MR2) #%>%
  #dplyr::select(!starts_with(c("K_","WEALTHLIM","MECHANISATION","DECISIONDECEN","DECISIONCEN","INDUSTRYLARGE","INDUSTRYSMALL","ECONOMYGLOBAL","ECONOMYLOCAL", "ENVOVERDEV","DEVOVERENV","OWNERPVT","OWNERNOREG","OWNERPUB", "OWNERREG","DISPLACENUCLEAR",  "POLLUTENUCLEAR", "HEALTHNUCLEAR", "JOBSNUCLEAR", "BEAUTYNUCLEAR", "PRIDENUCLEAR", "NPRIDENUCLEAR", "DEVNUCLEAR", "PROSPERNUCLEAR", "RELYNUCLEAR")))

mbennuclear <- lm(Ben_Nuclear ~ Uppercaste + Male + Hindu + Urban + age + State+ KahanS + KahanH + Pdevelop + Ndevelop , data = fascale_scoresben)
# Print model summary
#summary(mbennuclear)

```

```{r, echo =FALSE}

# same things with benefit from Solar Energy
sfascalesben <- codedmodule2 %>%
  dplyr::select(starts_with(c("K_","Ben_Solar","WEALTHLIM","MECHANISATION","DECISIONDECEN","DECISIONCEN","INDUSTRYLARGE","INDUSTRYSMALL","ECONOMYGLOBAL","ECONOMYLOCAL", "ENVOVERDEV","DEVOVERENV","OWNERPVT","OWNERNOREG","OWNERPUB", "OWNERREG","DISPLACESOLAR",  "POLLUTESOLAR", "HEALTHSOLAR", "JOBSSOLAR", "BEAUTYSOLAR", "PRIDESOLAR", "NPRIDESOLAR", "DEVSOLAR", "PROSPERSOLAR", "RELYSOLAR", "Uppercaste", "Male", "Hindu", "Urban", "State","age")))%>%
  na.omit()

Kahan_fas <- '
  KahanS =~ K_SHARM + K_SLIMCHOI + K_SPROTECT 
  KahanH =~ K_HEQUAL+ K_ERADEQ1 + K_EWEALTH + K_ERADEQ2
'

# Run CFA
sKahan_cfaben <- cfa(Kahan_fas, data = sfascalesben, std.lv = TRUE)

sKahan_scoresben <- predict(sKahan_cfaben, newdata = sfascalesben)

sKahan_Kscoresben <- cbind(sfascalesben, sKahan_scoresben)

secopolallben <- sfascalesben %>%
        dplyr::select("WEALTHLIM","MECHANISATION","DECISIONDECEN","DECISIONCEN","INDUSTRYLARGE","INDUSTRYSMALL","ECONOMYGLOBAL","ECONOMYLOCAL", "ENVOVERDEV","DEVOVERENV","OWNERPVT","OWNERNOREG","OWNERPUB", "OWNERREG","DISPLACESOLAR",  "POLLUTESOLAR", "HEALTHSOLAR", "JOBSSOLAR", "BEAUTYSOLAR", "PRIDESOLAR", "NPRIDESOLAR", "DEVSOLAR", "PROSPERSOLAR", "RELYSOLAR")

secopolallben_alpha <- psych::alpha(secopolallben, check.keys = TRUE)

#summary(ecopolall_alpha)

# when we combine both the scales the alpha is high at 0.785 (almost 0.8)

# FA on all ecopol variables except systems of governance

secopolallben_cor <- cor(secopolallben)
PCA24 <- principal(secopolallben, nfactors = 24, rotate = "none") # no rotation to start with.

secopolallben2.fa <- fa(secopolallben, nfactors = 2, rotate = "varimax")

# I don't know still doesn't make a lot of sense to me- the two factors make more sense - 1) people centered development 2) nationalist development
# So I am going to extract factor scores and bind them to the dataset

secopolallben_scores <- psych::factor.scores(secopolallben, secopolallben2.fa)

# The factor scores will be stored in the scores element of the returned list
sscores2 <- secopolallben_scores$scores

#binding factor scores to the dataset

sfascale_scoresben <- cbind(sKahan_Kscoresben, sscores2) %>%
  rename(Pdevelop = MR1)%>%
  rename(Ndevelop = MR2) #%>%
  #dplyr::select(!starts_with(c("K_","WEALTHLIM","MECHANISATION","DECISIONDECEN","DECISIONCEN","INDUSTRYLARGE","INDUSTRYSMALL","ECONOMYGLOBAL","ECONOMYLOCAL", "ENVOVERDEV","DEVOVERENV","OWNERPVT","OWNERNOREG","OWNERPUB", "OWNERREG","DISPLACENUCLEAR",  "POLLUTENUCLEAR", "HEALTHNUCLEAR", "JOBSNUCLEAR", "BEAUTYNUCLEAR", "PRIDENUCLEAR", "NPRIDENUCLEAR", "DEVNUCLEAR", "PROSPERNUCLEAR", "RELYNUCLEAR")))

mbensolar <- lm(Ben_Solar ~ Uppercaste + Male + Hindu + Urban + age + State+ KahanS + KahanH + Pdevelop + Ndevelop , data = sfascale_scoresben)
# Print model summary
#summary(mbensolar)

```

```{r, echo =FALSE}

# same things with benefit from COAL Energy
cfascalesben <- codedmodule2 %>%
  dplyr::select(starts_with(c("K_","Ben_Coal","WEALTHLIM","MECHANISATION","DECISIONDECEN","DECISIONCEN","INDUSTRYLARGE","INDUSTRYSMALL","ECONOMYGLOBAL","ECONOMYLOCAL", "ENVOVERDEV","DEVOVERENV","OWNERPVT","OWNERNOREG","OWNERPUB", "OWNERREG","DISPLACECOAL",  "POLLUTECOAL", "HEALTHCOAL", "JOBSCOAL", "BEAUTYCOAL", "PRIDECOAL", "NPRIDECOAL", "DEVCOAL", "PROSPERCOAL", "RELYCOAL", "Uppercaste", "Male", "Hindu", "Urban", "State","age")))%>%
  na.omit()

Kahan_fas <- '
  KahanS =~ K_SHARM + K_SLIMCHOI + K_SPROTECT 
  KahanH =~ K_HEQUAL+ K_ERADEQ1 + K_EWEALTH + K_ERADEQ2
'

# Run CFA
cKahan_cfaben <- cfa(Kahan_fas, data = cfascalesben, std.lv = TRUE)

cKahan_scoresben <- predict(cKahan_cfaben, newdata = cfascalesben)

cKahan_Kscoresben <- cbind(cfascalesben, cKahan_scoresben)

cecopolallben <- cfascalesben %>%
        dplyr::select("WEALTHLIM","MECHANISATION","DECISIONDECEN","DECISIONCEN","INDUSTRYLARGE","INDUSTRYSMALL","ECONOMYGLOBAL","ECONOMYLOCAL", "ENVOVERDEV","DEVOVERENV","OWNERPVT","OWNERNOREG","OWNERPUB", "OWNERREG","DISPLACECOAL",  "POLLUTECOAL", "HEALTHCOAL", "JOBSCOAL", "BEAUTYCOAL", "PRIDECOAL", "NPRIDECOAL", "DEVCOAL", "PROSPERCOAL", "RELYCOAL")

cecopolallben_alpha <- psych::alpha(cecopolallben, check.keys = TRUE)

#summary(ecopolall_alpha)

# when we combine both the scales the alpha is high at 0.785 (almost 0.8)

# FA on all ecopol variables except systems of governance

cecopolallben_cor <- cor(cecopolallben)
PCA24 <- principal(cecopolallben, nfactors = 24, rotate = "none") # no rotation to start with.

cecopolallben2.fa <- fa(cecopolallben, nfactors = 2, rotate = "varimax")

# I don't know still doesn't make a lot of sense to me- the two factors make more sense - 1) people centered development 2) nationalist development
# So I am going to extract factor scores and bind them to the dataset

cecopolallben_scores <- psych::factor.scores(cecopolallben, cecopolallben2.fa)

# The factor scores will be stored in the scores element of the returned list
sscores2 <- cecopolallben_scores$scores

#binding factor scores to the dataset

cfascale_scoresben <- cbind(cKahan_Kscoresben, sscores2) %>%
  rename(Pdevelop = MR1)%>%
  rename(Ndevelop = MR2) #%>%
  #dplyr::select(!starts_with(c("K_","WEALTHLIM","MECHANISATION","DECISIONDECEN","DECISIONCEN","INDUSTRYLARGE","INDUSTRYSMALL","ECONOMYGLOBAL","ECONOMYLOCAL", "ENVOVERDEV","DEVOVERENV","OWNERPVT","OWNERNOREG","OWNERPUB", "OWNERREG","DISPLACENUCLEAR",  "POLLUTENUCLEAR", "HEALTHNUCLEAR", "JOBSNUCLEAR", "BEAUTYNUCLEAR", "PRIDENUCLEAR", "NPRIDENUCLEAR", "DEVNUCLEAR", "PROSPERNUCLEAR", "RELYNUCLEAR")))

mbencoal <- lm(Ben_Coal ~ Uppercaste + Male + Hindu + Urban + age + State+ KahanS + KahanH + Pdevelop + Ndevelop , data = cfascale_scoresben)
# Print model summary
#summary(mbencoal)

```

\newpage

## Linear Models: Perceived Benefit X Pdevelop and Ndevelop

```{r, echo=FALSE, results='asis'}
library(stargazer)

cat("\\begingroup")
cat("\\setlength{\\tabcolsep}{1pt}")
cat("\\renewcommand{\\arraystretch}{0.7}")
stargazer(mbennuclear,mbensolar , mbencoal, 
         title = "Perceived Benefit X Pdevelop and Ndevelop",
        type = "latex"
)
cat("\\endgroup")

```

\newpage

## Linear Models: Acceptance (Risk - Benefit) X Pdevelop & Ndevelop

```{r, echo=FALSE, results='asis'}
library(stargazer)

cat("\\begingroup")
cat("\\setlength{\\tabcolsep}{1pt}")
cat("\\renewcommand{\\arraystretch}{0.7}")
stargazer(nuclearacceptlm, solaracceptlm , coalacceptlm, 
         title = "Acceptance (Risk - Benefit) X Pdevelop and Ndevelop",
        type = "latex"
)
cat("\\endgroup")

```

```{r, echo = FALSE}

# below are only eco-pol values of the perceiver scale.

ecopolonly <- codedmodule2 %>%
            filter(!is.na(WEALTHLIM) & !is.na(MECHANISATION) & !is.na(INDUSTRYSMALL)  & !is.na(ECONOMYGLOBAL)  & !is.na(ENVOVERDEV)  & !is.na(OWNERNOREG) & !is.na(OWNERPUB) & !is.na(OWNERREG) & !is.na(DECISIONDECEN) & !is.na(DECISIONCEN) & !is.na(INDUSTRYLARGE) & !is.na(ECONOMYLOCAL) & !is.na(DEVOVERENV) & !is.na(OWNERPVT))
        #dplyr::select("WEALTHLIM","MECHANISATION","INDUSTRYSMALL","ECONOMYGLOBAL", "ENVOVERDEV","OWNERNOREG","OWNERPUB", "OWNERREG", "DECISIONDECEN","DECISIONCEN","INDUSTRYLARGE","ECONOMYLOCAL", "DEVOVERENV","OWNERPVT") %>%
  #na.omit()

ecopolval <- codedmodule2 %>%
        dplyr::select("WEALTHLIM","MECHANISATION","INDUSTRYSMALL","ECONOMYGLOBAL", "ENVOVERDEV","OWNERNOREG","OWNERPUB", "OWNERREG", "DECISIONDECEN","DECISIONCEN","INDUSTRYLARGE","ECONOMYLOCAL", "DEVOVERENV","OWNERPVT")%>%
  na.omit() %>%
    rename(
      "Limits on Wealth" = "WEALTHLIM",
      "Anti Mechanisation of work" = "MECHANISATION",
      "Pro Decentralisation" = "DECISIONDECEN",
      "Pro Centralisation" = "DECISIONCEN",
      "Pro Large Industries" = "INDUSTRYLARGE",
      "Anti Large Industries" = "INDUSTRYSMALL",
      "Pro Globaleconomy" = "ECONOMYGLOBAL",
      "Pro Localeconomy" = "ECONOMYLOCAL",
      "Environment over Development" = "ENVOVERDEV",
      "Development over Environment" = "DEVOVERENV",
      "Pro Private ownership" = "OWNERPVT",
      "Anti Regulations" = "OWNERNOREG",
      "Pro Public ownership" = "OWNERPUB",
      "Pro Regulations" = "OWNERREG"
    )

ecopolval2.fa <- fa(ecopolval, nfactors = 2, rotate = "varimax")
#print.psych(ecopolval2.fa, cut = 0.4, sort = TRUE)

ecopolval4.fa <- fa(ecopolval, nfactors = 4, rotate = "varimax")
#print.psych(ecopolval4.fa, cut = 0.4, sort = TRUE)

# 4 fA extraction

ecopolval_scores <- psych::factor.scores(ecopolval, ecopolval4.fa)

# The factor scores will be stored in the scores element of the returned list
ecopolonlyscores <- ecopolval_scores$scores

#binding factor scores to the dataset

ecopolonlywscores <- cbind(ecopolonly, ecopolonlyscores) %>%
  rename(regulateindustry = MR1)%>%
  rename(localisation = MR3)%>%
   rename(freemarket = MR4)%>%
  rename(pvtdevelopment = MR2)

# 2 factor

ecopolval2_scores <- psych::factor.scores(ecopolval, ecopolval2.fa)

# The factor scores will be stored in the scores element of the returned list
ecopolonlyscores2 <- ecopolval2_scores$scores

#binding factor scores to the dataset

ecopolonlywscores2 <- cbind(ecopolonly, ecopolonlyscores2)

# checking alpha

regulated_industry <- ecopolonly  %>%
  dplyr::select(OWNERREG, MECHANISATION, WEALTHLIM)%>%
  na.omit()

regulated_industry.alpha <- psych::alpha(regulated_industry, check.keys = TRUE)

#alpha 0.59

localisation <- ecopolonly %>%
  dplyr::select(INDUSTRYSMALL, ECONOMYLOCAL)%>%
  na.omit()

localisation.alpha <- psych::alpha(localisation, check.keys = TRUE)

# alpha 0.52

freemarket <- ecopolonly %>%
  dplyr::select(DECISIONCEN, ECONOMYGLOBAL,OWNERNOREG,INDUSTRYLARGE)%>%
  na.omit()

freemarket.alpha <- psych::alpha(freemarket, check.keys = TRUE)

# alpha = 0.57

pvtdevelopment <- ecopolonly %>%
  dplyr::select(DEVOVERENV,OWNERPVT, ENVOVERDEV, OWNERPUB)%>%
  na.omit()

pvtdevelopment.alpha <- psych::alpha(pvtdevelopment, check.keys = TRUE)

# alpha =0.58

#cor(ecopolval)

# questions for Terre , 2) 4 factors/3 factors/2 factors - variation explained how should i deal with that. 
# four factor analysis - MR1 -regulated industrialization MR3 - localisation MR4 - free market MR2- private development

```

\newpage
\blandscape
# FA on Separate scales: Scale 1:Eco-political values of the perceiver

## 4 factor solution (Characteristics of the perceiver)

Cronbach's alpha values: Regulated Industrialisation = 0.59, localisation = 0.52 , Free market = 0.57, Private Development = 0.58

```{r, echo=FALSE}

library(flextable)

flex <- function(data, title=NULL, col_widths=NULL) {
  tbl <- flextable(data) %>%
         #set_table_properties(layout = "autofit", width = 1) %>%
         fontsize(size=10, part="all") %>%
         set_caption(title, 
                     autonum = officer::run_autonum(seq_id = "tab", 
                                                    pre_label = "Table ", 
                                                    post_label = "\n", 
                                                    bkm = "anytable"))

  # If col_widths is provided, set column widths
  if(!is.null(col_widths)) {
    for(col in names(col_widths)){
      tbl <- width(tbl, j = col, width = col_widths[[col]])
    }
  }

  return(tbl)
}

rename_map <- c(
  "Limits on Wealth" = "A limit should be put to how much wealth a person can amass",
  "Anti Mechanisation of work" = "Rapid mechanization of work is taking away jobs from workers in this country",
  "Pro Decentralisation" = "Local politicians shouldn't have to ask permission from the central government to implement policies",
  "Pro Centralisation" = "Laws and policies would be implemented more smoothly if more power lay with the central government",
  "Pro Large Industries" = "Large scale industries are required for the development of the country that will benefit everyone",
  "Anti Large Industries" = "Large corporations are destroying the local industries in India and benefiting only a handful of people",
  "Pro Globaleconomy" = "Foreign companies have led to a range of benefits for the Indian people and society",
  "Pro Localeconomy" = "India would be better off if foreign companies didn't come to here",
  "Environment over Development" = "Polluting industries that spoil the environment should be shut down even if it costs people their jobs",
  "Development over Environment" = "Economic growth and creating jobs should be prioritized over environmental protection",
  "Pro Private ownership" = "All businesses and industries should be owned privately",
  "Anti Regulations" = "There is too much red-tape and the government should not interfere with businesses and industries",
  "Pro Public ownership" = "The government should own most large businesses and industries",
  "Pro Regulations" = "Regardless of ownership, the government should pass strong regulations and implement them"
)

fa_table <- function(x, cut) {
  #get sorted loadings
  loadings <- fa.sort(x)$loadings %>% round(3)
  #suppress loadings
  loadings[loadings > -cut & loadings < cut] <- ""
  #get additional info  
  add_info <- cbind(x$communalities, 
                    x$uniquenesses,
                    x$complexity) %>%
    # make it a data frame
    as.data.frame() %>%
    # column names
    rename("Communality" = V1,
           "Uniqueness" = V2,
           "Complexity" = V3) %>%
    #get the item names from the vector
    rownames_to_column("item")

  #build table
  loadings %>%
    unclass() %>%
    as.data.frame() %>%
    rownames_to_column("item") %>%
    left_join(add_info) %>%
    mutate(
      item = if_else(item %in% names(rename_map), rename_map[item], item),
      code = names(rename_map)[match(item, rename_map)], # extracting short form
      across(where(is.numeric), round, 3)
    ) %>%
    select(code, everything()) # reorder columns to bring code to the front
}

fa_table(ecopolval4.fa, .4) %>%
  rename("regulated industrialization" = MR1, "private development" = MR2, "localisation" = MR3, "free market" = MR4, Items= item) %>%
  flex("4 factor solution same for all technologies", col_widths = list(Items = 3.5, code = 1.5, "regulated industrialization" = 0.5, "private development" = 0.5, "localisation" = 0.5, "free market" = 0.5))

```

\elandscape

```{r, echo=FALSE}

library(flextable)

flex <- function(data, title=NULL) {
  # this grabs the data and converts it to a flextbale
  flextable(data) %>%
  # this makes the table fill the page width
  set_table_properties(layout = "autofit", width = 1) %>%
  # font size
  fontsize(size=10, part="all") %>%
    #this adds a ttitlecreates an automatic table number
      set_caption(title, 
                  autonum = officer::run_autonum(seq_id = "tab", 
                                                 pre_label = "Table ", 
                                                 post_label = "\n", 
                                                 bkm = "anytable")) #%>%
  # font type
  #font(fontname="Times New Roman", part="all")
}

ecopolval4.fa[["Vaccounted"]] %>%
  as.data.frame() %>%
  rownames_to_column("Property") %>%
  #rename(Pdevelop = MR1, Ndevelop = MR2) %>%
    mutate(across(where(is.numeric), round, 3)) %>%
    flex("Eigenvalues and Variance Explained for Rotated Factor Solution")

```

```{r, echo = FALSE}
# linear regression with 4 factor solution

#Risky_Nuclear

nuclear4fa <- lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + Urban + age + State + regulateindustry +  localisation + freemarket + pvtdevelopment , data = ecopolonlywscores)
# Print model summary
#summary(nuclear4fa)

#Ben_Nuclear
bennuclear4fa <- lm(Ben_Nuclear ~ Uppercaste + Male + Hindu + Urban + age + State + regulateindustry +  localisation + freemarket + pvtdevelopment , data = ecopolonlywscores)
# Print model summary
#summary(bennuclear4fa)

#Risky_Solar
solar4fa <- lm(Risky_Solar ~ Uppercaste + Male + Hindu + Urban + age + State + regulateindustry +  localisation + freemarket + pvtdevelopment , data = ecopolonlywscores)
# Print model summary
#summary(solar4fa)

#Ben_Solar

bensolar4fa <- lm(Ben_Solar ~ Uppercaste + Male + Hindu + Urban + age + State + regulateindustry +  localisation + freemarket + pvtdevelopment , data = ecopolonlywscores)
# Print model summary
#summary(bensolar4fa)

#Risky_Coal

coal4fa <- lm(Risky_Coal ~ Uppercaste + Male + Hindu + Urban + age + State + regulateindustry +  localisation + freemarket + pvtdevelopment , data = ecopolonlywscores)
# Print model summary
#summary(coal4fa)

#Ben_Coal

bencoal4fa <- lm(Ben_Coal ~ Uppercaste + Male + Hindu + Urban + age + State + regulateindustry +  localisation + freemarket + pvtdevelopment , data = ecopolonlywscores)
# Print model summary
#summary(bencoal4fa)

```

```{r,echo = FALSE}
acceptance3 <- ecopolonlywscores %>%
   mutate(Nuclear = Ben_Nuclear - Risky_Nuclear,
         Hyrdo = Ben_Hydro - Risky_Hydro,
         Wind = Ben_Wind - Risky_Wind,
         Solar = Ben_Solar - Risky_Solar,
         Coal = Ben_Coal - Risky_Coal,
         Gas = Ben_Gas - Risky_Gas,
         Oil = Ben_Oil - Risky_Oil)

naccept2 <- lm(Nuclear ~ age + Uppercaste + Male + Hindu + Urban + State  + regulateindustry +  localisation + freemarket + pvtdevelopment, data = acceptance3)

#summary(naccept2)

saccept2 <- lm(Solar ~ age + Uppercaste + Male + Hindu + Urban + State  + regulateindustry +  localisation + freemarket + pvtdevelopment, data = acceptance3)

#summary(saccept2)

caccept2 <- lm(Coal ~ age + Uppercaste + Male + Hindu + Urban + State  + regulateindustry +  localisation + freemarket + pvtdevelopment, data = acceptance3)

#summary(caccept2)
```

\newpage
## Linear Models: Perceived Risk X 4 Factor Solution(separate scale)

```{r, echo=FALSE, results='asis'}
library(stargazer)

cat("\\begingroup")
cat("\\setlength{\\tabcolsep}{1pt}")
cat("\\renewcommand{\\arraystretch}{0.7}")
stargazer(nuclear4fa, solar4fa, coal4fa, 
          title = "Perceived Risk X 4 Factor Solution(separate scale)",
          type = "latex"
)
cat("\\endgroup")

```

\newpage

##  Linear Models: : Perceived Risk X 4 Factor Solution (separate scale)

```{r, echo=FALSE, results='asis'}
library(stargazer)

cat("\\begingroup")
cat("\\setlength{\\tabcolsep}{1pt}")
cat("\\renewcommand{\\arraystretch}{0.7}")
stargazer(bennuclear4fa, bensolar4fa, bencoal4fa, 
          title = "Perceived Risk X 4 Factor Solution (separate scale)",
          type = "latex"
)
cat("\\endgroup")

```

\newpage

##  Linear Models: Acceptance(Risk-Benefit) X 4 Factor Solution (separate scale)

```{r, echo=FALSE, results='asis'}
library(stargazer)

cat("\\begingroup")
cat("\\setlength{\\tabcolsep}{1pt}")
cat("\\renewcommand{\\arraystretch}{0.7}")
stargazer(naccept2, saccept2, caccept2, 
          title = "Acceptance(Risk-Benefit) X 4 Factor Solution (separate scale)",
          type = "latex"
)
cat("\\endgroup")

```

```{r, echo = FALSE}
# linear regression with 2 factor solution

#Risky_Nuclear

nuclear2fa2 <- lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + Urban + age + State + MR1 +MR2 , data = ecopolonlywscores2)
# Print model summary
#summary(nuclear2fa2)

#Ben_Nuclear
bennuclear2fa2 <- lm(Ben_Nuclear ~ Uppercaste + Male + Hindu + Urban + age + State + MR1 +MR2 , data = ecopolonlywscores2)
# Print model summary
#summary(bennuclear2fa2)

#Risky_Solar
solar2fa2 <- lm(Risky_Solar ~ Uppercaste + Male + Hindu + Urban + age + State + MR1 +MR2 , data = ecopolonlywscores2)
# Print model summary
#summary(solar2fa2)

#Ben_Solar

bensolar2fa2 <- lm(Ben_Solar ~ Uppercaste + Male + Hindu + Urban + age + State + MR1 +MR2 , data = ecopolonlywscores2)
# Print model summary
#summary(bensolar2fa2)

#Risky_Coal

coal2fa2 <- lm(Risky_Coal ~ Uppercaste + Male + Hindu + Urban + age + State + MR1 +MR2 , data = ecopolonlywscores2)
# Print model summary
#summary(coal2fa2)

#Ben_Coal

bencoal2fa2 <- lm(Ben_Coal ~ Uppercaste + Male + Hindu + Urban + age + State + MR1 +MR2 , data = ecopolonlywscores2)
# Print model summary
#summary(bencoal2fa2)

```

\newpage
\blandscape

## 2 factor solution (Characteristics of the perceiver)

```{r, echo=FALSE}

library(flextable)

flex <- function(data, title=NULL, col_widths=NULL) {
  tbl <- flextable(data) %>%
         #set_table_properties(layout = "autofit", width = 1) %>%
         fontsize(size=10, part="all") %>%
         set_caption(title, 
                     autonum = officer::run_autonum(seq_id = "tab", 
                                                    pre_label = "Table ", 
                                                    post_label = "\n", 
                                                    bkm = "anytable"))

  # If col_widths is provided, set column widths
  if(!is.null(col_widths)) {
    for(col in names(col_widths)){
      tbl <- width(tbl, j = col, width = col_widths[[col]])
    }
  }

  return(tbl)
}

rename_map <- c(
  "Limits on Wealth" = "A limit should be put to how much wealth a person can amass",
  "Anti Mechanisation of work" = "Rapid mechanization of work is taking away jobs from workers in this country",
  "Pro Decentralisation" = "Local politicians shouldn't have to ask permission from the central government to implement policies",
  "Pro Centralisation" = "Laws and policies would be implemented more smoothly if more power lay with the central government",
  "Pro Large Industries" = "Large scale industries are required for the development of the country that will benefit everyone",
  "Anti Large Industries" = "Large corporations are destroying the local industries in India and benefiting only a handful of people",
  "Pro Globaleconomy" = "Foreign companies have led to a range of benefits for the Indian people and society",
  "Pro Localeconomy" = "India would be better off if foreign companies didn't come to here",
  "Environment over Development" = "Polluting industries that spoil the environment should be shut down even if it costs people their jobs",
  "Development over Environment" = "Economic growth and creating jobs should be prioritized over environmental protection",
  "Pro Private ownership" = "All businesses and industries should be owned privately",
  "Anti Regulations" = "There is too much red-tape and the government should not interfere with businesses and industries",
  "Pro Public ownership" = "The government should own most large businesses and industries",
  "Pro Regulations" = "Regardless of ownership, the government should pass strong regulations and implement them"
)

fa_table <- function(x, cut) {
  #get sorted loadings
  loadings <- fa.sort(x)$loadings %>% round(3)
  #suppress loadings
  loadings[loadings > -cut & loadings < cut] <- ""
  #get additional info  
  add_info <- cbind(x$communalities, 
                    x$uniquenesses,
                    x$complexity) %>%
    # make it a data frame
    as.data.frame() %>%
    # column names
    rename("Communality" = V1,
           "Uniqueness" = V2,
           "Complexity" = V3) %>%
    #get the item names from the vector
    rownames_to_column("item")

  #build table
  loadings %>%
    unclass() %>%
    as.data.frame() %>%
    rownames_to_column("item") %>%
    left_join(add_info) %>%
    mutate(
      item = if_else(item %in% names(rename_map), rename_map[item], item),
      code = names(rename_map)[match(item, rename_map)], # extracting short form
      across(where(is.numeric), round, 3)
    ) %>%
    select(code, everything()) # reorder columns to bring code to the front
}

fa_table(ecopolval2.fa, .4) %>%
  rename( Items= item) %>%
  flex("4 factor solution same for all technologies", col_widths = list(Items = 3.5, code = 1.5))

```
\elandscape

\newpage

## Linear Models: Perceived Risk X 2 factor solution (separate scale)
```{r, echo=FALSE, results='asis'}
library(stargazer)

cat("\\begingroup")
cat("\\setlength{\\tabcolsep}{1pt}")
cat("\\renewcommand{\\arraystretch}{0.7}")
stargazer(nuclear2fa2, solar2fa2, coal2fa2, 
          title = "Perceived Risk X 2 factor solution (separate scale)",
          type = "latex"
)
cat("\\endgroup")

```

\newpage

## Linear Models: Perceived Benefit X 2 FA solution (separate scales)
```{r, echo=FALSE, results='asis'}
library(stargazer)

cat("\\begingroup")
cat("\\setlength{\\tabcolsep}{1pt}")
cat("\\renewcommand{\\arraystretch}{0.7}")
stargazer(bennuclear2fa2, bensolar2fa2, bencoal2fa2, 
          title = "Perceived Benefit X 2 FA solution (separate scales)",
          type = "latex"
)
cat("\\endgroup")

```

```{r, echo = FALSE}
 # Adjust as per the number of factors

#fa.diagram(ecopolval4.fa, marg=c(.5,5,1,.5),size =1, adj = 0.25, e.size=.05,rsize=.35, main = "Four Factor Solution")

```

```{r, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

#alpha test on Eco-pol characteristics of the technology

Nuclear <- codedmodule2 %>%
  filter(rowSums(is.na(dplyr::select(., starts_with("K_")))) == 0)%>%
  filter(complete.cases(dplyr::select(., DISPLACENUCLEAR:RELYNUCLEAR)))

Nuclear2 <- Nuclear  %>%
  dplyr::select("DISPLACENUCLEAR","POLLUTENUCLEAR", "HEALTHNUCLEAR", "BEAUTYNUCLEAR", "PRIDENUCLEAR", "NPRIDENUCLEAR", "DEVNUCLEAR","PROSPERNUCLEAR")%>%    # alternate select("DISPLACENUCLEAR":"RELYNUCLEAR")
    rename(
   "Displacement risk(nuclear)" = "DISPLACENUCLEAR",
  "Pollution risk(nuclear)" = "POLLUTENUCLEAR",
  "Health risk(nuclear)" = "HEALTHNUCLEAR",
  #"Job opportunities(nuclear)" = "JOBSNUCLEAR",
  "Spoils Natural Beauty(nuclear)" = "BEAUTYNUCLEAR",
  "Community pride(nuclear)" = "PRIDENUCLEAR",
  "National pride(nuclear)" = "NPRIDENUCLEAR",
  "National development(nuclear)" = "DEVNUCLEAR",
  "Local prosperity(nuclear)" = "PROSPERNUCLEAR",
  #"Reliance on government(nuclear)" = "RELYNUCLEAR"
    )

Nuclear.fa <- fa(Nuclear2, nfactors = 2, rotate = "varimax")

print.psych(Nuclear.fa, cut = 0.5, sort = TRUE)

#fa.diagram(Nuclear.fa)

# CFA Kahan

Kahan_fas <- '
  KahanS =~ K_SHARM + K_SLIMCHOI + K_SPROTECT 
  KahanH =~ K_HEQUAL+ K_ERADEQ1 + K_EWEALTH + K_ERADEQ2
'

# Run CFA
n2Kahan_cfa <- cfa(Kahan_fas, data = Nuclear, std.lv = TRUE)

# creating table of factor scores

n2Kahan_scores <- predict(n2Kahan_cfa, newdata = Nuclear)

n2Kahan_Kscores <- cbind(Nuclear, n2Kahan_scores)

# checking cronbach's alpha 
nnegative <- Nuclear %>%
  dplyr::select(HEALTHNUCLEAR,POLLUTENUCLEAR,BEAUTYNUCLEAR, DISPLACENUCLEAR)%>%
  na.omit()

nnegative.alpha <- psych::alpha(nnegative, check.keys = TRUE)

# alpha = 0.79

npositive <- Nuclear %>%
  dplyr::select(DEVNUCLEAR,PROSPERNUCLEAR,NPRIDENUCLEAR,PRIDENUCLEAR)%>% #excluded JOBSnuclear 
  na.omit()

npositive.alpha <- psych::alpha(npositive, check.keys = TRUE)

# alpha = 0.72

```

\newpage
\blandscape
# FA on Separate scale: Scale 2: Characteristics of the Technology(2 factor solution)
## Nuclear Energy : positive (alpha = 0.73) and negative (alpha = 0.79) characteristics (2 factor solution)

```{r, echo=FALSE}

library(flextable)

set_flextable_defaults(docx.landscape = TRUE)

flex <- function(data, title=NULL, col_widths=NULL) {
  tbl <- flextable(data) %>%
           #padding(padding = 1, part = "all") %>%
           #set_table_properties(layout = "autofit") %>%
           fontsize(size=10, part="all") %>%
           set_caption(title, 
                       autonum = officer::run_autonum(seq_id = "tab", 
                                                      pre_label = "Table ", 
                                                      post_label = "\n", 
                                                      bkm = "anytable"))

  # If col_widths is provided, set column widths
  if(!is.null(col_widths)) {
    for(col in names(col_widths)){
      tbl <- width(tbl, j = col, width = col_widths[[col]])
    }
  }

  return(tbl)
}

rename_map <- c(
  "Displacement risk(nuclear)" = "Nuclear energy is leading to displacement of people from their land",
  "Spoils Natural Beauty(nuclear)" = "Nuclear energy spoils the natural beauty of the landscape",
  "Pollution risk(nuclear)" = "Nuclear energy increases pollution of air/water/land",
  "Health risk(nuclear)" = "Nuclear energy poses a great risk to the health of people living around it",
  "Community pride(nuclear)" = "I would be proud if my community used nuclear energy",
  "National pride(nuclear)" = "Nuclear energy is a mark of pride for our nation",
  "National development(nuclear)" = "Nuclear energy pushes forward the country's development",
  "Local prosperity(nuclear)" = "Nuclear energy brings economic prosperity to the surrounding regions"
)

fa_table <- function(x, cut) {
  #get sorted loadings
  loadings <- fa.sort(x)$loadings %>% round(3)
  #suppress loadings
  loadings[loadings > -cut & loadings < cut] <- ""
  #get additional info  
  add_info <- cbind(x$communalities, 
                    x$uniquenesses,
                    x$complexity) %>%
    # make it a data frame
    as.data.frame() %>%
    # column names
    rename("Communality" = V1,
           "Uniqueness" = V2,
           "Complexity" = V3) %>%
    #get the item names from the vector
    rownames_to_column("item")

  #build table
  loadings %>%
    unclass() %>%
    as.data.frame() %>%
    rownames_to_column("item") %>%
    left_join(add_info) %>%
    mutate(
      item = if_else(item %in% names(rename_map), rename_map[item], item),
      code = names(rename_map)[match(item, rename_map)], # extracting short form
      across(where(is.numeric), round, 3)
    ) %>%
    select(code, everything()) # reorder columns to bring code to the front
}

fa_table(Nuclear.fa, .4) %>%
  rename("Social Costs" = MR1, "Pride" = MR2, Items= item) %>%
  flex("Separate scale 2 factor solution: Nuclear energy ", col_widths = list(Items = 4.5, code = 1.5))

```
\elandscape

```{r, echo=FALSE}

library(flextable)

flex <- function(data, title=NULL) {
  # this grabs the data and converts it to a flextbale
  flextable(data) %>%
  # this makes the table fill the page width
  set_table_properties(layout = "autofit", width = 1) %>%
  # font size
  fontsize(size=10, part="all") %>%
    #this adds a ttitlecreates an automatic table number
      set_caption(title, 
                  autonum = officer::run_autonum(seq_id = "tab", 
                                                 pre_label = "Table ", 
                                                 post_label = "\n", 
                                                 bkm = "anytable")) #%>%
  # font type
  #font(fontname="Times New Roman", part="all")
}

Nuclear.fa[["Vaccounted"]] %>%
  as.data.frame() %>%
  rownames_to_column("Property") %>%
  rename(Social_Costs = MR1, Pride = MR2) %>%
    mutate(across(where(is.numeric), round, 3)) %>%
    flex("Eigenvalues and Variance Explained for Rotated Factor Solution")

```

```{r, echo=FALSE}

# 2 factor

Nuclear_scores <- psych::factor.scores(Nuclear2, Nuclear.fa)

# The factor scores will be stored in the scores element of the returned list
Nuclear_scores2 <- Nuclear_scores$scores

#binding factor scores to the dataset

Nuclearwscores2 <- cbind(Nuclear, Nuclear_scores2)%>%
    dplyr::rename(pride = MR2)%>%
    dplyr::rename(socialcosts = MR1)

nuclearfa <- lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + Urban + age + State +  socialcosts + pride , data = Nuclearwscores2)
# Print model summary
summary(nuclearfa)

bennuclearfa <- lm(Ben_Nuclear ~ Uppercaste + Male + Hindu + Urban + age + State +  socialcosts + pride , data = Nuclearwscores2)
# Print model summary
summary(bennuclearfa)

#n2Kahan_Kscores

Nuclearwscores3<- cbind(n2Kahan_Kscores, Nuclear_scores2)%>%
    rename(socialcosts = MR1)%>%
  rename(pride = MR2) %>%
  select(Risky_Nuclear, Uppercaste, Male, Hindu, Urban, age, State, socialcosts, pride) %>%
  na.omit()

nuclearfa2 <- lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + Urban + age + State + socialcosts + pride , data = Nuclearwscores3)
# Print model summary
summary(nuclearfa2)

#bennuclearfa2 <- lm(Ben_Nuclear ~ Uppercaste + Male + Hindu + Urban + age + State + KahanH + KahanS+ socialcosts + pride , data = Nuclearwscores3)
# Print model summary
#summary(bennuclearfa2)

# Selecting the specific variables from the dataset
#selected_data <- Nuclearwscores3[, c("KahanH", "KahanS", "socialcosts", "pride")]

# Calculating the correlation matrix
#corkahannew <- cor(selected_data, use = "complete.obs")  # 'use = "complete.obs"' handles missing values by excluding them

# Printing the correlation matrix
#print(corkahannew)

```

```{r,echo = FALSE}
acceptance4 <- Nuclearwscores2 %>%
   mutate(Nuclear = Ben_Nuclear - Risky_Nuclear,
         Hyrdo = Ben_Hydro - Risky_Hydro,
         Wind = Ben_Wind - Risky_Wind,
         Solar = Ben_Solar - Risky_Solar,
         Coal = Ben_Coal - Risky_Coal,
         Gas = Ben_Gas - Risky_Gas,
         Oil = Ben_Oil - Risky_Oil)

naccept3 <- lm(Nuclear ~ age + Uppercaste + Male + Hindu + Urban + State  +  socialcosts + pride, data = acceptance4)

#summary(naccept3)

```

\blandscape
\newpage
## Solar Energy : positive (alpha =0.85) and negative (alpha = 0.81)characteristics (2 factor solution)

```{r, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

Solar <- codedmodule2 %>%
  filter(rowSums(is.na(dplyr::select(., starts_with("K_")))) == 0)%>%
  filter(complete.cases(dplyr::select(., DISPLACESOLAR:RELYSOLAR)))

Solar2 <- Solar  %>%
  dplyr::select("DISPLACESOLAR","POLLUTESOLAR","HEALTHSOLAR","BEAUTYSOLAR","PRIDESOLAR","NPRIDESOLAR", "DEVSOLAR", "PROSPERSOLAR")%>%
    rename(
   "Displacement risk(Solar)" = "DISPLACESOLAR",
  "Pollution risk(Solar)" = "POLLUTESOLAR",
  "Health risk(Solar)" = "HEALTHSOLAR",
  #"Job opportunities(Solar)" = "JOBSSOLAR",
  "Spoils Natural Beauty(Solar)" = "BEAUTYSOLAR",
  "Community pride(Solar)" = "PRIDESOLAR",
  "National pride(Solar)" = "NPRIDESOLAR",
  "National development(Solar)" = "DEVSOLAR",
  "Local prosperity(Solar)" = "PROSPERSOLAR",
  #"Reliance on government(Solar)" = "RELYSOLAR"
    )

Solar.fa <- fa(Solar2, nfactors = 2, rotate = "varimax")

print.psych(Solar.fa, cut = 0.40, sort = TRUE)

#fa.diagram(Solar.fa)

# CFA Kahan

snegative <- Solar %>%
  dplyr::select(HEALTHSOLAR, POLLUTESOLAR,BEAUTYSOLAR, DISPLACESOLAR)%>%
  na.omit()

snegative.alpha <- psych::alpha(snegative, check.keys = TRUE)

# alpha = 0.81

spositive <- Solar %>%
  dplyr::select(DEVSOLAR, PROSPERSOLAR,NPRIDESOLAR,PRIDESOLAR)%>%
  na.omit()

spositive.alpha <- psych::alpha(spositive, check.keys = TRUE)

# alpha = 0.89

Kahan_fas <- '
  KahanS =~ K_SHARM + K_SLIMCHOI + K_SPROTECT 
  KahanH =~ K_HEQUAL+ K_ERADEQ1 + K_EWEALTH + K_ERADEQ2
'

# Run CFA
s2Kahan_cfa <- cfa(Kahan_fas, data = Solar, std.lv = TRUE)

# creating table of factor scores

s2Kahan_scores <- predict(s2Kahan_cfa, newdata = Solar)

s2Kahan_Kscores <- cbind(Solar, s2Kahan_scores)

```

```{r, echo=FALSE}

library(flextable)

set_flextable_defaults(docx.landscape = TRUE)

flex <- function(data, title=NULL, col_widths=NULL) {
  tbl <- flextable(data) %>%
           #padding(padding = 1, part = "all") %>%
           #set_table_properties(layout = "autofit") %>%
           fontsize(size=10, part="all") %>%
           set_caption(title, 
                       autonum = officer::run_autonum(seq_id = "tab", 
                                                      pre_label = "Table ", 
                                                      post_label = "\n", 
                                                      bkm = "anytable"))

  # If col_widths is provided, set column widths
  if(!is.null(col_widths)) {
    for(col in names(col_widths)){
      tbl <- width(tbl, j = col, width = col_widths[[col]])
    }
  }

  return(tbl)
}

rename_map <- c(
  "Displacement risk(Solar)" = "Solar energy is leading to displacement of people from their land",
  "Spoils Natural Beauty(Solar)" = "Solar energy spoils the natural beauty of the landscape",
  "Pollution risk(Solar)" = "Solar energy increases pollution of air/water/land",
  "Health risk(Solar)" = "Solar energy poses a great risk to the health of people living around it",
  "Community pride(Solar)" = "I would be proud if my community used solar energy",
  "National pride(Solar)" = "Solar energy is a mark of pride for our nation",
  "National development(Solar)" = "Solar energy pushes forward the country's development",
  "Local prosperity(Solar)" = "Solar energy brings economic prosperity to the surrounding regions"
)

fa_table <- function(x, cut) {
  #get sorted loadings
  loadings <- fa.sort(x)$loadings %>% round(3)
  #suppress loadings
  loadings[loadings > -cut & loadings < cut] <- ""
  #get additional info  
  add_info <- cbind(x$communalities, 
                    x$uniquenesses,
                    x$complexity) %>%
    # make it a data frame
    as.data.frame() %>%
    # column names
    rename("Communality" = V1,
           "Uniqueness" = V2,
           "Complexity" = V3) %>%
    #get the item names from the vector
    rownames_to_column("item")

  #build table
  loadings %>%
    unclass() %>%
    as.data.frame() %>%
    rownames_to_column("item") %>%
    left_join(add_info) %>%
    mutate(
      item = if_else(item %in% names(rename_map), rename_map[item], item),
      code = names(rename_map)[match(item, rename_map)], # extracting short form
      across(where(is.numeric), round, 3)
    ) %>%
    select(code, everything()) # reorder columns to bring code to the front
}

fa_table(Solar.fa, .4) %>%
  rename("Pride" = MR1, "Social Costs" = MR2, Items= item) %>%
  flex("Separate scale 2 factor solution: Solar energy ", col_widths = list(Items = 4.5, code = 1.5))

```
\elandscape

```{r, echo=FALSE}

library(flextable)

flex <- function(data, title=NULL) {
  # this grabs the data and converts it to a flextbale
  flextable(data) %>%
  # this makes the table fill the page width
  set_table_properties(layout = "autofit", width = 1) %>%
  # font size
  fontsize(size=10, part="all") %>%
    #this adds a ttitlecreates an automatic table number
      set_caption(title, 
                  autonum = officer::run_autonum(seq_id = "tab", 
                                                 pre_label = "Table ", 
                                                 post_label = "\n", 
                                                 bkm = "anytable")) #%>%
  # font type
  #font(fontname="Times New Roman", part="all")
}

Solar.fa[["Vaccounted"]] %>%
  as.data.frame() %>%
  rownames_to_column("Property") %>%
  rename(Pride = MR1, Social_Costs = MR2) %>%
    mutate(across(where(is.numeric), round, 3)) %>%
    flex("Eigenvalues and Variance Explained for Rotated Factor Solution")

```

```{r, echo=FALSE}

# 2 factor

Solar_scores <- psych::factor.scores(Solar2, Solar.fa)

# The factor scores will be stored in the scores element of the returned list
Solar_scores2 <- Solar_scores$scores

#binding factor scores to the dataset

Solarwscores2 <- cbind(s2Kahan_Kscores, Solar_scores2)%>%
    rename(pride = MR1)%>%
  rename(socialcosts = MR2)

Solarfa <- lm(Risky_Solar ~  Uppercaste + Male + Hindu + Urban + age + State + pride + socialcosts, data = Solarwscores2)
# Print model summary
summary(Solarfa)

benSolarfa <- lm(Ben_Solar ~  Uppercaste + Male + Hindu + Urban + age + State +  KahanH + KahanS + pride + socialcosts, data = Solarwscores2)
# Print model summary
summary(benSolarfa)

```

```{r,echo = FALSE}
sacceptance4 <- Solarwscores2 %>%
   mutate(Nuclear = Ben_Nuclear - Risky_Nuclear,
         Hyrdo = Ben_Hydro - Risky_Hydro,
         Wind = Ben_Wind - Risky_Wind,
         Solar = Ben_Solar - Risky_Solar,
         Coal = Ben_Coal - Risky_Coal,
         Gas = Ben_Gas - Risky_Gas,
         Oil = Ben_Oil - Risky_Oil)

saccept3 <- lm(Solar ~ age + Uppercaste + Male + Hindu + Urban + State  +  socialcosts + pride, data = sacceptance4)

#summary(saccept3)

```
\newpage
\blandscape

## Coal : positive (alpha = 0.67) and negative (alphs = 0.72) characteristics (2 factor solution)

```{r, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

#alpha test on Eco-pol characteristics of the technology

Coal <- codedmodule2 %>%
  filter(complete.cases(dplyr::select(., DISPLACECOAL:RELYCOAL)))

Coal2 <- Coal %>%
  dplyr::select("DISPLACECOAL","POLLUTECOAL","HEALTHCOAL","BEAUTYCOAL","PRIDECOAL","NPRIDECOAL","DEVCOAL","PROSPERCOAL")%>%
    rename(
   "Displacement risk(Coal)" = "DISPLACECOAL",
  "Pollution risk(Coal)" = "POLLUTECOAL",
  "Health risk(Coal)" = "HEALTHCOAL",
  #"Job opportunities(Coal)" = "JOBSCOAL",
  "Spoils Natural Beauty(Coal)" = "BEAUTYCOAL",
  "Community pride(Coal)" = "PRIDECOAL",
  "National pride(Coal)" = "NPRIDECOAL",
  "National development(Coal)" = "DEVCOAL",
  "Local prosperity(Coal)" = "PROSPERCOAL"
  #"Reliance on government(Coal)" = "RELYCOAL"
    )

Coal.fa <- fa(Coal2, nfactors = 2, rotate = "varimax")

print.psych(Coal.fa, cut = 0.40, sort = TRUE)

#fa.diagram(Coal.fa)

cnegative <- Coal %>%
  dplyr::select(HEALTHCOAL, POLLUTECOAL,BEAUTYCOAL, DISPLACECOAL)%>%
  na.omit()

cnegative.alpha <- psych::alpha(cnegative, check.keys = TRUE)

# alpha = 0.72

cpositive <- Coal %>%
  dplyr::select(DEVCOAL, PROSPERCOAL,NPRIDECOAL,PRIDECOAL)%>%
  na.omit()

cpositive.alpha <- psych::alpha(cpositive, check.keys = TRUE)

# alpha = 0.68

```

```{r, echo=FALSE}

library(flextable)

set_flextable_defaults(docx.landscape = TRUE)

flex <- function(data, title=NULL, col_widths=NULL) {
  tbl <- flextable(data) %>%
           #padding(padding = 1, part = "all") %>%
           #set_table_properties(layout = "autofit") %>%
           fontsize(size=10, part="all") %>%
           set_caption(title, 
                       autonum = officer::run_autonum(seq_id = "tab", 
                                                      pre_label = "Table ", 
                                                      post_label = "\n", 
                                                      bkm = "anytable"))

  # If col_widths is provided, set column widths
  if(!is.null(col_widths)) {
    for(col in names(col_widths)){
      tbl <- width(tbl, j = col, width = col_widths[[col]])
    }
  }

  return(tbl)
}

rename_map <- c(
  "Displacement risk(Coal)" = "Coal powered plants is leading to displacement of people from their land",
  "Spoils Natural Beauty(Coal)" = "Coal powered plants spoils the natural beauty of the landscape",
  "Pollution risk(Coal)" = "Coal powered plants increases pollution of air/water/land",
  "Health risk(Coal)" = "Coal powered plants poses a great risk to the health of people living around it",
  "Community pride(Coal)" = "I would be proud if my community used Coal powered plants",
  "National pride(Coal)" = "Coal powered plants is a mark of pride for our nation",
  "National development(Coal)" = "Coal powered plants pushes forward the country's development",
  "Local prosperity(Coal)" = "Coal powered plants brings economic prosperity to the surrounding regions"
)

fa_table <- function(x, cut) {
  #get sorted loadings
  loadings <- fa.sort(x)$loadings %>% round(3)
  #suppress loadings
  loadings[loadings > -cut & loadings < cut] <- ""
  #get additional info  
  add_info <- cbind(x$communalities, 
                    x$uniquenesses,
                    x$complexity) %>%
    # make it a data frame
    as.data.frame() %>%
    # column names
    rename("Communality" = V1,
           "Uniqueness" = V2,
           "Complexity" = V3) %>%
    #get the item names from the vector
    rownames_to_column("item")

  #build table
  loadings %>%
    unclass() %>%
    as.data.frame() %>%
    rownames_to_column("item") %>%
    left_join(add_info) %>%
    mutate(
      item = if_else(item %in% names(rename_map), rename_map[item], item),
      code = names(rename_map)[match(item, rename_map)], # extracting short form
      across(where(is.numeric), round, 3)
    ) %>%
    select(code, everything()) # reorder columns to bring code to the front
}

fa_table(Coal.fa, .4) %>%
  rename("Social Costs" = MR1, "Pride" = MR2, Items= item) %>%
  flex("Separate scale 2 factor solution: Coal ", col_widths = list(Items = 4.5, code = 1.5))

```
\elandscape

```{r, echo=FALSE}

library(flextable)

flex <- function(data, title=NULL) {
  # this grabs the data and converts it to a flextbale
  flextable(data) %>%
  # this makes the table fill the page width
  set_table_properties(layout = "autofit", width = 1) %>%
  # font size
  fontsize(size=10, part="all") %>%
    #this adds a ttitlecreates an automatic table number
      set_caption(title, 
                  autonum = officer::run_autonum(seq_id = "tab", 
                                                 pre_label = "Table ", 
                                                 post_label = "\n", 
                                                 bkm = "anytable")) #%>%
  # font type
  #font(fontname="Times New Roman", part="all")
}

Coal.fa[["Vaccounted"]] %>%
  as.data.frame() %>%
  rownames_to_column("Property") %>%
  rename(Social_Costs = MR1, Pride = MR2) %>%
    mutate(across(where(is.numeric), round, 3)) %>%
    flex("Eigenvalues and Variance Explained for Rotated Factor Solution")

```

```{r, echo=FALSE}

# 2 factor

Coal_scores <- psych::factor.scores(Coal2, Coal.fa)

# The factor scores will be stored in the scores element of the returned list
Coal_scores2 <- Coal_scores$scores

#binding factor scores to the dataset

Coalwscores2 <- cbind(Coal, Coal_scores2) %>%
  rename(socialcosts = MR1)%>%
  rename(pride = MR2)

Coalfa <- lm(Risky_Coal ~  Uppercaste + Male + Hindu + Urban + age + State + socialcosts + pride , data = Coalwscores2)
# Print model summary
#summary(Coalfa)

benCoalfa <- lm(Ben_Coal ~  Uppercaste + Male + Hindu + Urban + age + State + pride + socialcosts, data = Coalwscores2)
# Print model summary
#summary(benCoalfa)
```

```{r,echo = FALSE}
cacceptance4 <- Coalwscores2 %>%
   mutate(Nuclear = Ben_Nuclear - Risky_Nuclear,
         Hyrdo = Ben_Hydro - Risky_Hydro,
         Wind = Ben_Wind - Risky_Wind,
         Solar = Ben_Solar - Risky_Solar,
         Coal = Ben_Coal - Risky_Coal,
         Gas = Ben_Gas - Risky_Gas,
         Oil = Ben_Oil - Risky_Oil)

caccept3 <- lm(Coal ~ age + Uppercaste + Male + Hindu + Urban + State  + socialcosts + pride, data = cacceptance4)

#summary(caccept3)

```

\newpage

## Linear Models: Perceived Risk X positive and negative characteristics of the technology
```{r, echo=FALSE, results='asis'}
library(stargazer)

cat("\\begingroup")
cat("\\setlength{\\tabcolsep}{1pt}")
cat("\\renewcommand{\\arraystretch}{0.7}")
stargazer(nuclearfa, Solarfa, Coalfa, 
          title = "LMs with separate scale 2 FA solution: Perceived Risk",
          type = "latex"
)
cat("\\endgroup")

```

\newpage

## Linear Models: Perceived Benefit X positive and negative characteristics of the technology
```{r, echo=FALSE, results='asis'}
library(stargazer)

cat("\\begingroup")
cat("\\setlength{\\tabcolsep}{1pt}")
cat("\\renewcommand{\\arraystretch}{0.7}")
stargazer(bennuclearfa, benSolarfa, benCoalfa, 
          title = "LMs with separate scale 2 FA solution: Perceived Benefit",
          type = "latex"
)
cat("\\endgroup")

```

\newpage
## Linear Models: Acceptance(Benefit - Risk) X positive and negative characteristics of the technology
```{r, echo=FALSE, results='asis'}
library(stargazer)

cat("\\begingroup")
cat("\\setlength{\\tabcolsep}{1pt}")
cat("\\renewcommand{\\arraystretch}{0.7}")
stargazer(naccept3,saccept3, caccept3, 
          title = "LMs with separate scale 2 FA solution: Perceived Benefit",
          type = "latex"
)
cat("\\endgroup")

```