### 2024-01-15 corstuff.Rmd

---
title: "correlationstuff"
output:
  pdf_document:
    toc: yes
    toc_depth: 2
    keep_tex: yes
  header-includes:
  - \usepackage{dcolumn}
  - \usepackage{array}
  - "\\usepackage{placeins}"
  keep_md: yes
---

```{r load packages, warning = FALSE, message = FALSE, echo=FALSE}
  knitr::opts_chunk$set(warning = FALSE, message = FALSE)

suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(readxl))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(tidyr))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(lavaan))
suppressPackageStartupMessages(library(tsibble))
suppressPackageStartupMessages(library(DT))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(psych)) 
suppressPackageStartupMessages(library(foreign))
suppressPackageStartupMessages(library(MASS))
suppressPackageStartupMessages(library(Hmisc))
suppressPackageStartupMessages(library(reshape2))

```

```{r setup, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

moduletwo <- read_excel(here::here("thesis_surveydata/political_factors_complete.xlsx"))
module2 <- data.frame(moduletwo)
module2[module2 == "Rather not say/ Don't know"] <- "Rather not say/Don't know"

# loaded data set for module 2 eco -pol variables and removes the rows with question number and question text
```

```{r, eval =TRUE, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
#This chunk contains the coding schemes of various scales used in survey one: eco-political factors, kahan scale and acceptance scale
codedmodule2 <- module2 %>%

#remove row 1
  filter(!row_number() %in% c(1,2)) %>%

# replace risky likert scale with numbers
  mutate_at(vars(starts_with("Risky")), funs(case_when(. =="Not at all risky" ~ 1, 
                                                       . =="Slightly risky" ~ 2, 
                                                       . =="Moderately risky" ~ 3, 
                                                       . =="Very risky" ~ 4, 
                                                       . =="Extremely risky" ~ 5))) %>%

# replace beneficial likert scale with numbers  
  mutate_at(vars(starts_with("Ben")), funs(case_when(. =="Not at all beneficial" ~ 1,
                                                     . =="Slightly beneficial" ~ 2,
                                                     . =="Moderately beneficial" ~ 3,
                                                     . =="Very beneficial" ~ 4,
                                                     . =="Extremely beneficial" ~ 5 ))) %>%

# replace nuclear acceptance likert scale with numbers
  mutate_at(vars(N_accept,N_reluctantlyaccept,N_reject), funs(case_when(. == "Strongly disagree" ~ 1, 
                                                                        . == "Somewhat disagree" ~ 2,
                                                                        . == "Neither agree nor disagree" ~3,
                                                                        . == "Somewhat agree" ~ 4,
                                                                        . == "Strongly agree" ~ 5))) %>%

# code likert scale for variables for Kahan scale into numbers
  mutate_at(vars(starts_with (c("K_S","K_E","DISPLACE", "POLLUTE", "HEALTH", "JOBS", "BEAUTY", "PRIDE", "NPRIDE","DEVNUCLEAR","PROSPER", "RELY"))), funs(case_when(. == "Strongly disagree" ~ 1, 
               . == "Somewhat disagree" ~ 2,
               . == "Neither agree nor disagree" ~3,
               . == "Somewhat agree" ~ 4,
               . == "Strongly agree" ~ 5))) %>%

# reverse code for likert scale for variables for Kahan scale into numbers
  mutate_at(vars(starts_with (c("K_H","K_I"))), funs(case_when(. == "Strongly disagree" ~ 5, 
                                                               . == "Somewhat disagree" ~ 4,
                                                               . == "Neither agree nor disagree" ~3,
                                                               . == "Somewhat agree" ~ 2,
                                                               . == "Strongly agree" ~ 1))) %>%

# code eco-pol scale variables into numbers
mutate_at(vars(SYSTEMDEMO,SYSTEMRELIGION,SYSTEMTECHNO,SYSTEMTOTAL,WEALTHLIM,MECHANISATION,DECISIONDECEN,INDUSTRYSMALL,ECONOMYLOCAL,ENVOVERDEV,OWNERPUB, OWNERREG), funs(case_when(. == "Strongly disagree" ~ 1, 
                        . == "Somewhat disagree" ~ 2,
                        . == "Neither agree nor disagree" ~3,
                        . == "Somewhat agree" ~ 4,
                        . == "Strongly agree" ~ 5))) %>%

# reverse code eco-pol scale variables into numbers 
 mutate_at(vars(DECISIONCEN,INDUSTRYLARGE,ECONOMYGLOBAL,OWNERPVT,OWNERNOREG, DEVOVERENV), funs(case_when(. == "Strongly disagree" ~ 5, 
                                                                                                         . == "Somewhat disagree" ~ 4,
                                                                                                         . == "Neither agree nor disagree" ~3,
                                                                                                         . == "Somewhat agree" ~ 2,
                                                                                                         . == "Strongly agree" ~ 1))) %>%
  mutate_all(~ifelse(. == "Rather not say/ Don't know", "Rather not say/Don't know", .))%>%
  mutate_all(~ifelse(. == "West bengal", "West Bengal", .))%>%
  mutate_all(~ifelse(. == "WEST BENGAL", "West Bengal", .))%>%
  mutate_all(~ifelse(. == "Up", "Uttar Pradesh", .))%>%
  mutate_all(~ifelse(. == "Tamilnadu", "Tamil Nadu", .))%>%
  mutate_all(~ifelse(. == "UP", "Uttar Pradesh", .))

```

```{r, eval=TRUE, echo=FALSE}

moduleone <- read_excel(here::here("thesis_surveydata/livelihood_insecurity_complete.xlsx"))
module1 <- data.frame(moduleone)
# loaded data set for module 2 eco -pol variables and removes the rows with question number and question text
```

```{r, eval =TRUE, echo=FALSE}
#This chunk contains the coding schemes of various scales used in survey one: eco-political factors, kahan scale and acceptance scale
codedmodule1 <- module1 %>%

#remove row 1 and 2 
  filter(!row_number() %in% c(1,2)) %>%

# replace risky likert scale with numbers
  mutate_at(vars(starts_with("Risky")), funs(case_when(. =="Not at all risky" ~ 1, 
                                                       . =="Slightly risky" ~ 2, 
                                                       . =="Moderately risky" ~ 3, 
                                                       . =="Very risky" ~ 4, 
                                                       . =="Extremely risky" ~ 5))) %>%

# replace beneficial likert scale with numbers  
  mutate_at(vars(starts_with("Ben")), funs(case_when(. =="Not at all beneficial" ~ 1,
                                                     . =="Slightly beneficial" ~ 2,
                                                     . =="Moderately beneficial" ~ 3,
                                                     . =="Very beneficial" ~ 4,
                                                     . =="Extremely beneficial" ~ 5 ))) %>%

# replace nuclear acceptance likert scale with numbers
  mutate_at(vars(N_accept,N_reluctantlyaccept,N_reject), funs(case_when(. == "Strongly disagree" ~ 1, 
                                                                        . == "Somewhat disagree" ~ 2,
                                                                        . == "Neither agree nor disagree" ~3,
                                                                        . == "Somewhat agree" ~ 4,
                                                                        . == "Strongly agree" ~ 5)))%>%
  mutate_all(~ifelse(. == "Rather not say/ Don't know", "Rather not say/Don't know", .))%>%
  mutate_all(~ifelse(. == "West bengal", "West Bengal", .))%>%
  mutate_all(~ifelse(. == "WEST BENGAL", "West Bengal", .))%>%
  mutate_all(~ifelse(. == "Up", "Uttar Pradesh", .))%>%
  mutate_all(~ifelse(. == "Tamilnadu", "Tamil Nadu", .))%>%
  mutate_all(~ifelse(. == "UP", "Uttar Pradesh", .))

```

```{r, echo = FALSE}

suppressPackageStartupMessages(library(tidyselect))
# combine codedmodule1 and codedmodule2

numcol2 <- codedmodule2 %>% add_column(Respondent = 1:1099, .before = "Language") 
numcol1 <- codedmodule1 %>% add_column(Respondent = 1099:2160, .before = "Language")

joineddf <-  full_join(numcol2, numcol1, by = c("Respondent", "Language", "Survey_Date", "State", "Risky_Hydro", "Risky_Solar", "Risky_Wind", "Risky_Nuclear", "Risky_Coal", "Risky_Gas", "Risky_Oil", "Risky_INDHydro", "Risky_INDSolar", "Risky_INDWind", "Risky_INDBiogas", "Risky_INDDiesel", "Risky_INDKerosene", "Risky_INDFirewoodetc", "Risky_INDLPG", "Ben_Hydro", "Ben_Solar", "Ben_Wind", "Ben_Nuclear", "Ben_Coal", "Ben_Gas", "Ben_Oil", "Ben_INDHydro", "Ben_INDSolar", "Ben_INDWind", "Ben_INDDiesel", "Ben_INDBiogas", "Ben_INDFirewoodetc", "Ben_INDLPG", "Ben_INDKerosene", "N_accept", "N_reluctantlyaccept", "N_reject", "age", "urban_rural", "gender", "caste", "religion", "religiontext")) %>%
  mutate_at(vars(age), funs(case_when(. == "18-24 years old" ~ 1, 
                                      . == "25-34 years old" ~ 2,
                                      . ==  "35-44 years old" ~3,
                                      . == "45-54 years old" ~ 4,
                                      . == "55-64 years old" ~ 5,
                                      . == "65-74 years old" ~ 6,
                                      . == "75 years or older" ~ 7))) # converting age to continuous variable

joineddf1 <-  full_join(numcol2, numcol1, by = c("Respondent", "Language", "Survey_Date", "State", "Risky_Hydro", "Risky_Solar", "Risky_Wind", "Risky_Nuclear", "Risky_Coal", "Risky_Gas", "Risky_Oil", "Risky_INDHydro", "Risky_INDSolar", "Risky_INDWind", "Risky_INDBiogas", "Risky_INDDiesel", "Risky_INDKerosene", "Risky_INDFirewoodetc", "Risky_INDLPG", "Ben_Hydro", "Ben_Solar", "Ben_Wind", "Ben_Nuclear", "Ben_Coal", "Ben_Gas", "Ben_Oil", "Ben_INDHydro", "Ben_INDSolar", "Ben_INDWind", "Ben_INDDiesel", "Ben_INDBiogas", "Ben_INDFirewoodetc", "Ben_INDLPG", "Ben_INDKerosene", "N_accept", "N_reluctantlyaccept", "N_reject", "age", "urban_rural", "gender", "caste", "religion", "religiontext"))

numcolb <- module2 %>%   filter(!row_number() %in% c(1,2))%>% 
                        add_column(Respondent = 1:1099, .before = "Language") 
numcola <- module1 %>%   filter(!row_number() %in% c(1,2)) %>% 
                        add_column(Respondent = 1099:2160, .before = "Language")

uncodedjoineddf <- full_join(numcolb, numcola, by = c("Respondent", "Language", "Survey_Date", "State", "Risky_Hydro", "Risky_Solar", "Risky_Wind", "Risky_Nuclear", "Risky_Coal", "Risky_Gas", "Risky_Oil", "Risky_INDHydro", "Risky_INDSolar", "Risky_INDWind", "Risky_INDBiogas", "Risky_INDDiesel", "Risky_INDKerosene", "Risky_INDFirewoodetc", "Risky_INDLPG", "Ben_Hydro", "Ben_Solar", "Ben_Wind", "Ben_Nuclear", "Ben_Coal", "Ben_Gas", "Ben_Oil", "Ben_INDHydro", "Ben_INDSolar", "Ben_INDWind", "Ben_INDDiesel", "Ben_INDBiogas", "Ben_INDFirewoodetc", "Ben_INDLPG", "Ben_INDKerosene", "N_accept", "N_reluctantlyaccept", "N_reject", "age", "urban_rural", "gender", "caste", "religion", "religiontext")) %>%
  mutate_all(~ifelse(. == "Rather not say/ Don't know", "Rather not say/Don't know", .))%>%
  mutate_all(~ifelse(. == "West bengal", "West Bengal", .))%>%
  mutate_all(~ifelse(. == "WEST BENGAL", "West Bengal", .))%>%
  mutate_all(~ifelse(. == "Up", "Uttar Pradesh", .))%>%
  mutate_all(~ifelse(. == "Tamilnadu", "Tamil Nadu", .))%>%
  mutate_all(~ifelse(. == "UP", "Uttar Pradesh", .))

```

\newpage

## Mean Perceived Risk and Mean Perceived Benefit for all energy technologies.

```{r, echo=FALSE}
library(tidyverse)

# Filter variables of interest
joineddf_subset <- joineddf %>%
  dplyr::select(!starts_with(c("Ben_IND", "Risky_IND")))%>%
  dplyr::select(starts_with("Risky"), 
         starts_with("Ben"))

joineddf_subset_cor <- joineddf %>%
  dplyr::select(!starts_with(c("Ben_IND", "Risky_IND")))%>%
  dplyr::select(starts_with("Risky"), 
         starts_with("Ben"))%>%
  na.omit()

#cor(joineddf_subset_cor)

# Assuming your data is named joineddf_subset
long_df <- joineddf_subset %>%
  gather(key = "Type", value = "Value")

# Separate out Risky and Ben columns
risky_df <- long_df %>% filter(str_detect(Type, "^Risky"))
ben_df <- long_df %>% filter(str_detect(Type, "^Ben"))

# Extract the technology type (Hydro, Solar, etc.)
risky_df$Tech <- str_replace(risky_df$Type, "Risky_", "")
ben_df$Tech <- str_replace(ben_df$Type, "Ben_", "")

# Calculate the mean for each technology for Risky
risky_mean <- risky_df %>%
  group_by(Tech) %>%
  dplyr::summarize(Mean_Value = mean(Value, na.rm = TRUE),
          sd = sd(Value, na.rm = TRUE),
          se = sd(Value, na.rm = TRUE) / sqrt(n()),
          .groups = "drop")

# Do the same for Ben technologies
ben_mean <- ben_df %>%
  group_by(Tech) %>%
  dplyr::summarize(Mean_Value = mean(Value, na.rm = TRUE),
          sd = sd(Value, na.rm = TRUE),
          se = sd(Value, na.rm = TRUE) / sqrt(n()),
          .groups = "drop")

# Combine the two data frames into one for plotting
combined_mean <- bind_rows(
  mutate(risky_mean, Category = "Risky"),
  mutate(ben_mean, Category = "Beneficial")
)
# Reorder the factor levels of the Tech variable by Mean_Value
risky_mean$Tech <- factor(risky_mean$Tech, levels = risky_mean$Tech[order(risky_mean$Mean_Value, decreasing = FALSE)])

# Split the data into Risky and Ben subsets
risky_data <- joineddf_subset_cor %>% dplyr::select(starts_with("Risky"))
ben_data <- joineddf_subset_cor %>% dplyr::select(starts_with("Ben"))

# Rename the columns
names(risky_data) <- gsub("^Risky_", "Perceived Risk (", names(risky_data))
names(risky_data) <- paste0(names(risky_data), ")")
names(ben_data) <- gsub("^Ben_", "Perceived Benefit (", names(ben_data))
names(ben_data) <- paste0(names(ben_data), ")")

# Compute the correlation between Risky and Ben variables
correlation_matrix <- cor(risky_data, ben_data)

correlation_df <- as.data.frame(correlation_matrix)

```

```{r, echo=FALSE}
library(ggplot2)
library(gridExtra)

mean_risk <- mean(combined_mean$Mean_Value[combined_mean$Category == "Risky"], na.rm = TRUE)
mean_benefit <- mean(combined_mean$Mean_Value[combined_mean$Category == "Beneficial"], na.rm = TRUE)

# Order Technologies
technology_order <- c("Solar","Wind","Hydro","Oil","Coal", "Gas","Nuclear")
ben_mean$Tech <- factor(ben_mean$Tech, levels = technology_order)

# Generate the Ben plot with flipped y-axis
ben_plot <- ggplot(ben_mean, aes(y = Tech, x = Mean_Value)) +
  geom_bar(stat = "identity", fill = "palegreen3") +
  geom_errorbar(aes(xmin = Mean_Value - se, xmax = Mean_Value + se), width = 0.2) +
  coord_cartesian(xlim = c(5, 1)) + 
  labs(title = "Perceived Benefit", x = "Mean Values") +
  #scale_x_reverse() +
  theme(  # remove y-axis labels
        axis.title.y = element_blank()) + # remove the "Tech" axis title
  geom_vline(data = subset(combined_mean, Category == "Beneficial"), 
             aes(xintercept = mean_benefit), linetype="dotted", color = "black", size = 0.5)

# Generate the Risky plot
risky_plot <- ggplot(risky_mean, aes(y = Tech, x = Mean_Value)) +
  geom_bar(stat = "identity", fill = "firebrick2") +
  geom_errorbar(aes(xmin = Mean_Value - se, xmax = Mean_Value + se), width = 0.2) +
  coord_cartesian(xlim = c(1, 5)) + 
  labs(title = "Perceived Risk", x = "Mean Values") +
  theme(axis.text.y = element_blank(),
        axis.title.y = element_blank())+
  geom_vline(data = subset(combined_mean, Category == "Risky"), 
             aes(xintercept = mean_risk), linetype="dotted", color = "black", size = 0.5)
 # remove the "Tech" axis title

# Arrange plots side by side
#grid.arrange(ben_plot, risky_plot, ncol=2)

```

```{r, echo=FALSE}

## Paired T-test: Comparing mean perceived risk and mean perceived benefit for each technology.

#Null hypothesis (H0): The mean difference between perceived risk and perceived benefit from each technology is zero.

#All p values are less than 0.05 suggesting that the differences observed in the bar graph are statistically significant.

# Melt the data frame to long format
long_format <- joineddf_subset %>% 
  pivot_longer(everything(), names_to = "technology", values_to = "value") %>%
  mutate(risk_or_benefit = if_else(str_detect(technology, "Risky"), "risk", "benefit"),
         technology = str_replace(technology, "Risky_|Ben_", ""))

# Convert character columns to factors
long_format$technology <- factor(long_format$technology)
long_format$risk_or_benefit <- factor(long_format$risk_or_benefit)

# Perform the t- Test for each technology
p_values <- long_format %>%
  split(.$technology) %>%
  lapply(function(x) t.test(x$value[x$risk_or_benefit == "risk"], x$value[x$risk_or_benefit == "benefit"], paired = TRUE)) %>%
  lapply(function(x) x$p.value)

```

\newpage

```{r, echo=FALSE}

# # Convert `caste` into binary variable 'upper' (1) vs 'non-upper' (0)
joineddf$Uppercaste <- ifelse(joineddf$caste %in% c('Brahmin', 'General'), 1, 0)
# Convert `gender` into binary variable 'male' (1) vs 'female' (0)
joineddf$Male <- ifelse(joineddf$gender == 'Male', 1, 0)
# Convert `religion` into binary variable 'Hinduism' (1) vs 'Non-Hinduism' (0)
joineddf$Hindu <- ifelse(joineddf$religion == 'Hinduism', 1, 0)
codedmodule2$State <- as.factor(codedmodule2$State)
codedmodule2$gender <- as.factor(codedmodule2$gender)
codedmodule2$Urban <- as.factor(codedmodule2$urban_rural)
joineddf$Urban <- as.factor(joineddf$urban_rural)
codedmodule2$caste <- as.factor(codedmodule2$caste)
codedmodule2$religion <- as.factor(codedmodule2$religion)
# # Convert `caste` into binary variable 'upper' (1) vs 'non-upper' (0)
codedmodule2$Uppercaste <- ifelse(codedmodule2$caste %in% c('Brahmin', 'General'), 1, 0)
# Convert `gender` into binary variable 'male' (1) vs 'female' (0)
codedmodule2$Male <- ifelse(codedmodule2$gender == 'Male', 1, 0)
# Convert `religion` into binary variable 'Hinduism' (1) vs 'Non-Hinduism' (0)
codedmodule2$Hindu <- ifelse(codedmodule2$religion == 'Hinduism', 1, 0)

alldemos <- joineddf %>%
            dplyr::select(Risky_Nuclear, Uppercaste, Male, Hindu, Urban, age, State)

alldemoslm1 <- lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + Urban + age , data = alldemos)
# Print model summary
#summary(alldemoslm1)
#plot(alldemoslm1)

# 2) lm with all demographic variables
alldemoslm2<- lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + Urban + age + State, data = alldemos)
# Print model summary
#summary(alldemoslm2)
#plot(alldemoslm1)

```

## Demographic IVs - two linear regression models
```{r, echo=FALSE, results='asis'}
library(stargazer)

cat("\\begingroup")
cat("\\setlength{\\tabcolsep}{1pt}")
cat("\\renewcommand{\\arraystretch}{0.7}")
stargazer(alldemoslm1, alldemoslm2, 
          title = "Results from 2 linear regression models",
          type = "latex"
)
cat("\\endgroup")

```

\newpage

## Linear regression where Mean value is the intercept
Same model with mean value as intercept.

```{r, echo=FALSE}
# Centering the numeric predictor variables
alldemos$Uppercaste_centered <- alldemos$Uppercaste - mean(alldemos$Uppercaste, na.rm = TRUE)
alldemos$Male_centered      <- alldemos$Male - mean(alldemos$Male, na.rm = TRUE)
alldemos$Hindu_centered     <- alldemos$Hindu - mean(alldemos$Hindu, na.rm = TRUE)
alldemos$age_centered       <- alldemos$age - mean(alldemos$age, na.rm = TRUE)

# For factor variable 'Urban', we'll convert it to numeric (0/1) and then center
# Assuming "Rural" is the reference category
alldemos$Urban_numeric      <- ifelse(alldemos$Urban == "Urban", 1, 0)
alldemos$Urban_centered     <- alldemos$Urban_numeric - mean(alldemos$Urban_numeric, na.rm = TRUE)

# Now, run the regression with centered variables
alldemoslm1_centered <- lm(Risky_Nuclear ~ Uppercaste_centered + Male_centered + Hindu_centered + Urban_centered + age_centered, data = alldemos)
#summary(alldemoslm1_centered)

alldemos$State <- as.factor(alldemos$State)
alldemos$State <- relevel(alldemos$State, ref = "Rajasthan")

alldemoslm2_centered <- lm(Risky_Nuclear ~ Uppercaste_centered + Male_centered + Hindu_centered + Urban_centered + age_centered + State, data = alldemos)
#summary(alldemoslm2_centered)

```

```{r, echo=FALSE, results='asis'}
library(stargazer)

cat("\\begingroup")
cat("\\setlength{\\tabcolsep}{1pt}")
cat("\\renewcommand{\\arraystretch}{0.7}")
stargazer(alldemoslm1_centered, alldemoslm2_centered, 
          title = "Results from 2 linear regression models",
          type = "latex"
)
cat("\\endgroup")

```

\newpage

##  Kahan scale Alpha values 
Hierarchy - Egalitarianism scale alpha  =  0.7
Individualism - Communitarianism scale alpha= 0.52
Only Individualism scale alpha = 0.31 
Only Communitarianism scale alpha = 0.72

```{r, echo=FALSE}
# Q6: Does the Kahan scale explain perception of risk from nuclear energy and other DVs ?

# we conducted fa on the kahan scale 
# CFA for Kahan scale. We shouldn't do FA here since this a well used scale with theoretical basis for factor distinctions and also previous studies using the same scale

library(lavaan)
library(semTools)
library(kableExtra)

fascales <- codedmodule2 %>%
  dplyr::select(starts_with(c("K_","Risky_Nuclear","WEALTHLIM","MECHANISATION","DECISIONDECEN","DECISIONCEN","INDUSTRYLARGE","INDUSTRYSMALL","ECONOMYGLOBAL","ECONOMYLOCAL", "ENVOVERDEV","DEVOVERENV","OWNERPVT","OWNERNOREG","OWNERPUB", "OWNERREG","DISPLACENUCLEAR",  "POLLUTENUCLEAR", "HEALTHNUCLEAR", "JOBSNUCLEAR", "BEAUTYNUCLEAR", "PRIDENUCLEAR", "NPRIDENUCLEAR", "DEVNUCLEAR", "PROSPERNUCLEAR", "RELYNUCLEAR", "Uppercaste", "Male", "Hindu", "Urban", "State","age")))%>%
  mutate_at(vars(age), funs(case_when(. == "18-24 years old" ~ 1, 
                                      . == "25-34 years old" ~ 2,
                                      . ==  "35-44 years old" ~3,
                                      . == "45-54 years old" ~ 4,
                                      . == "55-64 years old" ~ 5,
                                      . == "65-74 years old" ~ 6,
                                      . == "75 years or older" ~ 7))) %>%
  na.omit()

Kahan_fas <- '
  KahanS =~ K_SHARM + K_SLIMCHOI + K_SPROTECT 
  KahanH =~ K_HEQUAL+ K_ERADEQ1 + K_EWEALTH + K_ERADEQ2
'

# Run CFA
Kahan_cfa <- cfa(Kahan_fas, data = fascales, std.lv = TRUE)

Kahan_scale <- codedmodule2 %>%
  dplyr::select(starts_with("K_"))

KahanHE<-Kahan_scale %>%
  dplyr::select(starts_with(c("K_H", "K_E")))

KahanHE_alpha<- psych::alpha(KahanHE, check.keys = TRUE) # alpha = 0.7

KahanSI<-Kahan_scale %>%
  dplyr::select(starts_with(c("K_S", "K_I")))

KahanSI_alpha<- psych::alpha(KahanSI, check.keys = TRUE) # alpha =0.52

KahanS<-Kahan_scale %>%
  dplyr::select(starts_with(c("K_S")))

KahanS_alpha<- psych::alpha(KahanS, check.keys = TRUE) # alpha = 0.72

KahanI<-Kahan_scale %>%
  dplyr::select(starts_with(c("K_I")))

KahanI_alpha<- psych::alpha(KahanI, check.keys = TRUE) # alpha = 0.31

Kahan_scores <- predict(Kahan_cfa, newdata = fascales)

Kahan_Kscores <- cbind(fascales, Kahan_scores)

```

# EDIT :  Kahan scale EFA  
```{r, echo=FALSE}

Kahan_scale <- codedmodule2 %>%
  dplyr::select(starts_with("K_"))

kahan.fa <- fa(Kahan_scale, nfactors = 2, rotate = "varimax")

#print.psych(kahan.fa, cut = 0.4, sort = TRUE)

```

```{r, echo=FALSE}

library(flextable)

flex <- function(data, title=NULL) {
  # this grabs the data and converts it to a flextbale
  flextable(data) %>%
  # this makes the table fill the page width
  set_table_properties(layout = "autofit", width = 1) %>%
  # font size
  fontsize(size=10, part="all") %>%
    #this adds a ttitlecreates an automatic table number
      set_caption(title, 
                  autonum = officer::run_autonum(seq_id = "tab", 
                                                 pre_label = "Table ", 
                                                 post_label = "\n", 
                                                 bkm = "anytable")) #%>%
  # font type
  #font(fontname="Times New Roman", part="all")
}

rename_map <- c(
  "K_SHARM" = "(C)Sometimes the government needs to make laws that keep people from hurting themselves.",
  "K_SLIMCHOI" = "(C)The government should put limits on the choices individuals can make so they don’t get in the way of what’s good for society.",
  "K_SPROTECT" = "(C)The government should do more to advance society’s goals, even if that means limiting the freedom and choices of individuals.",
  "K_HEQUAL" = "(H)We have gone too far in pushing equal rights in this country.",
  "K_ERADEQ1" = "(E)We need to dramatically reduce inequalities between the rich and the poor.",
  "K_EWEALTH" = "(E)Our society would be better off if the distribution of wealth was more equal.",
  "K_ERADEQ2" = "(E)We need to dramatically reduce inequalities between men and women."
)

fa_table <- function(x) {
  #get sorted loadings
  loadings <- fa.sort(x)$loadings %>% round(3)
  #supress loadings
  #loadings[loadings < cut] <- ""  # alternative loadings > -cut & loadings < cut
  #get additional info  
  add_info <- cbind(x$communalities, 
                    x$uniquenesses,
                    x$complexity) %>%
    # make it a data frame
    as.data.frame() %>%
    # column names
   rename("Communality" = V1,
           "Uniqueness" = V2,
           "Complexity" = V3) %>%
    #get the item names from the vector
    rownames_to_column("item")
  #build table
  loadings %>%
    unclass() %>%
    as.data.frame() %>%
    rownames_to_column("item") %>%
    left_join(add_info) %>%
    mutate(
      item = if_else(item %in% names(rename_map), rename_map[item], item),
      across(where(is.numeric), round, 3)
    )
}

fa_table(kahan.fa) %>%
  #rename(Pdevelop = MR1, Ndevelop = MR2, Items= item) %>%
  flex("Kahan Exploratory Factor Analysis Table")

```

## alternative Kahan scores
```{r, echo=FALSE}

Kscalescores <- fascales %>%
  dplyr::select(!starts_with(("K_I")))%>%
  rowwise()%>%
  dplyr::mutate(Communitarianism_score = mean(c_across(K_SHARM:K_SPROTECT)))%>%
  dplyr::mutate(Egalitarianism_score = mean(c_across(K_HEQUAL:K_ERADEQ2)))%>%
  dplyr::select(!starts_with(c("K_I","K_S", "K_H", "K_E")))

```

```{r fit-measures-display, echo = FALSE}
# Extract fit measures
fit_measures <- fitMeasures(Kahan_cfa, fit.measures = c("CFI", "TLI", "RMSEA", "rmsea.ci.upper", "rmsea.ci.lower"))
fit_measures2 <- fitMeasures(Kahan_cfa)
# Convert to a data frame for presentation
fit_df <- data.frame(Measure = names(fit_measures), Value = fit_measures) %>%
  mutate(Measure = case_when(
    Measure == "cfi" ~ "Comparative Fit Index (CFI) ",
    Measure == "tli" ~ "Tucker-Lewis Index (TLI)",
    Measure == "rmsea" ~ "Root Mean Square Error of Approximation(RMSEA)",
    Measure == "rmsea.ci.upper" ~ "RMSEA 90 Percent confidence interval - lower",
    Measure == "rmsea.ci.lower" ~ "RMSEA 90 Percent confidence interval - upper",
    TRUE ~ as.character(Measure)
  )) %>%
  mutate(Value = round(Value, 3)) %>%
  remove_rownames()

#RMSEA - between 0.05 and 0.08 (reasonable fit), less than 0.05 close fit and greater than 0.10 is poor fit 
#CFI - greater than 0.90 or 0.95 indicates good fit
#TFI - CFI>TFI >0.90 - indicates good fit

```
# Confirmatory Factor Analysis(CFA): Kahan Scale

**Cronbach's Alpha on Kahan et al(2007) Scale: A Note**

The Individualism items (indicated by K_I) were bringing down the Cronbach's alpha values in the Kahan scale. The Alpha for Individualism- Communitarian scale was 0.49. After removing the Individualism items (K_I) the alpha for this factor was 0.71. The reasons for this could be that the individualism items are not well adapted to the Indian population.

```{r, echo =FALSE}
# Display fit measures using kable
knitr::kable(fit_df, caption = "Fit Measures from the CFA", booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped", "hold_position"))

```

\newpage

```{r, echo=FALSE}

# Extract parameter estimates
params <- parameterEstimates(Kahan_cfa, standardized = TRUE)%>%
   slice(1:(n()-10))

params1 <- params %>%
  mutate(rhs = case_when(
    rhs == "K_SHARM" ~ "Sometimes the government needs to make laws that keep people from hurting themselves.",
    rhs == "K_SLIMCHOI" ~ "The government should put limits on the choices individuals can make so they don’t get in the way of what’s good for society.",
    rhs == "K_SPROTECT" ~ "The government should do more to advance society’s goals, even if that means limiting the freedom and choices of individuals.",
    rhs == "K_HEQUAL" ~ "We have gone too far in pushing equal rights in this country.",
    rhs == "K_ERADEQ1" ~ "We need to dramatically reduce inequalities between the rich and the poor.",
    rhs == "K_EWEALTH" ~ "Our society would be better off if the distribution of wealth was more equal.",
    rhs == "K_ERADEQ2" ~ "We need to dramatically reduce inequalities between men and women.",
    TRUE ~ as.character(rhs)
  )) %>%
  dplyr::select(!c(lhs,op, std.nox)) %>%
    rename(
    "Items" = rhs,
    "Loadings" = est,
    "Standard Error" = se,
    "zvalue" = z) %>%
  mutate(across(c("Loadings", "Standard Error", "zvalue"), round, 3))

params1$Scale <- NA
params1$Scale[1:3] <- "Communitarian"
params1$Scale[4:7] <- "Hierarchy-Egalitarianism"
params1 <- params1[, c("Scale", setdiff(names(params1), "Scale"))]
CFAkahan <- knitr::kable(params1, booktabs = TRUE, caption = "Confirmatory Factor Analysis(CFA) on Kahan et al(2007) scale adapted to India") %>%
  kable_styling(latex_options = c("striped","scale_down", "hold_position", "repeat_header"))%>%
  column_spec(column = 2, width = "4cm")

# Print the table
CFAkahan %>%
  landscape()

```

## Kahan scater plot from the CFA scores 
```{r, echo = FALSE}

Kahan_Kscores %>%
  ggplot(aes(KahanS, KahanH))+
  geom_count(aes(color = ..n.., size = ..n..))+
  guides(color = 'legend')+
  geom_hline(yintercept=0, colour="black", lwd=0.5)+
  geom_vline(xintercept=0, colour="black", lwd=0.5)+
  labs(x = "Individualism to Communitarianism", y = "Hierarchy to Egalitarianism")+
  ylim(-2,3)+
  xlim(-2,3)+
  theme_minimal() 
```

\newpage

# Factor Analysis: New Eco-political Scale

```{r, echo=FALSE}

library(psych)
library(tidyverse)

ecopolall <- fascales %>%
        dplyr::select("WEALTHLIM","MECHANISATION","DECISIONDECEN","DECISIONCEN","INDUSTRYLARGE","INDUSTRYSMALL","ECONOMYGLOBAL","ECONOMYLOCAL", "ENVOVERDEV","DEVOVERENV","OWNERPVT","OWNERNOREG","OWNERPUB", "OWNERREG","DISPLACENUCLEAR",  "POLLUTENUCLEAR", "HEALTHNUCLEAR", "JOBSNUCLEAR", "BEAUTYNUCLEAR", "PRIDENUCLEAR", "NPRIDENUCLEAR", "DEVNUCLEAR", "PROSPERNUCLEAR", "RELYNUCLEAR")

ecopolall_alpha <- psych::alpha(ecopolall, check.keys = TRUE)

#summary(ecopolall_alpha)

# when we combine both the scales the alpha is high at 0.785 (almost 0.8)

# FA on all ecopol variables except systems of governance

ecopolall_cor <- cor(ecopolall)
PCA24 <- principal(ecopolall, nfactors = 24, rotate = "none") # no rotation to start with.

#PCA24$loadings

#fa.parallel(ecopolall_cor)
# this suggests 2 factors if we go by FA or 8 by PCA. let's try both and see how they look conceptually 
# running EFA using psych package

ecopolall2.fa <- fa(ecopolall, nfactors = 2, rotate = "varimax")

print.psych(ecopolall2.fa, sort = TRUE)

# I don't know still doesn't make a lot of sense to me- the two factors make more sense - 1) people centered development 2) nationalist development
# So I am going to extract factor scores and bind them to the dataset

ecopolall_scores <- psych::factor.scores(ecopolall, ecopolall2.fa)

# The factor scores will be stored in the scores element of the returned list
scores1 <- ecopolall_scores$scores

#binding factor scores to the dataset

fascale_scores <- cbind(Kahan_Kscores, scores1) %>%
  rename(Pdevelop = MR1)%>%
  rename(Ndevelop = MR2) #%>%
  #dplyr::select(!starts_with(c("K_","WEALTHLIM","MECHANISATION","DECISIONDECEN","DECISIONCEN","INDUSTRYLARGE","INDUSTRYSMALL","ECONOMYGLOBAL","ECONOMYLOCAL", "ENVOVERDEV","DEVOVERENV","OWNERPVT","OWNERNOREG","OWNERPUB", "OWNERREG","DISPLACENUCLEAR",  "POLLUTENUCLEAR", "HEALTHNUCLEAR", "JOBSNUCLEAR", "BEAUTYNUCLEAR", "PRIDENUCLEAR", "NPRIDENUCLEAR", "DEVNUCLEAR", "PROSPERNUCLEAR", "RELYNUCLEAR")))

```

```{r, echo=FALSE}
library(ggplot2)
library(fmsb)

# Extract the loadings
loadings_matrix <- ecopolall2.fa$loadings[1:24, ]

loadings_df <- as.data.frame(as.matrix(loadings_matrix))

# Adding rownames to the dataframe as a new column
loadings_df <- loadings_df %>%
  rownames_to_column(var = "variable")

# Pivot to longer format first
long_loadings_df <- loadings_df %>%
  pivot_longer(cols = c(MR1, MR2), names_to = "Factors", values_to = "value")

# Pivot the longer format to wide format
wide_loadings_df <- long_loadings_df %>%
  pivot_wider(names_from = variable, values_from = value) %>%
  mutate(Factors = case_when(
    Factors == "MR1" ~ "People Centered Development (Pdevelop)",
    Factors == "MR2" ~ "Nationalist Development (Ndevelop)",
    TRUE ~ Factors
  )) %>%
  column_to_rownames("Factors")

# Define the variable ranges: maximum and minimum
max_min <- data.frame(
  WEALTHLIM = c(0.8, -0.6), MECHANISATION = c(0.8, -0.6), DECISIONDECEN = c(0.8, -0.6),
  DECISIONCEN = c(0.8, -0.6), INDUSTRYLARGE = c(0.8, -0.6), INDUSTRYSMALL = c(0.8, -0.6),
  ECONOMYGLOBAL = c(0.8, -0.6), ECONOMYLOCAL = c(0.8, -0.6), ENVOVERDEV = c(0.8, -0.6),
  DEVOVERENV = c(0.8, -0.6), OWNERPVT = c(0.8, -0.6), OWNERNOREG = c(0.8, -0.6),
  OWNERPUB = c(0.8, -0.6), OWNERREG = c(0.8, -0.6), DISPLACENUCLEAR = c(0.8, -0.6),
  POLLUTENUCLEAR = c(0.8, -0.6), HEALTHNUCLEAR = c(0.8, -0.6), JOBSNUCLEAR = c(0.8, -0.6),
  BEAUTYNUCLEAR = c(0.8, -0.6), PRIDENUCLEAR = c(0.8, -0.6), NPRIDENUCLEAR = c(0.8, -0.6),
  DEVNUCLEAR = c(0.8, -0.6), PROSPERNUCLEAR = c(0.8, -0.6), RELYNUCLEAR = c(0.8, -0.6)
)
rownames(max_min) <- c("Max", "Min")

# Assuming you have a dataframe 'exam_scores' with matching column names
# Bind the variable ranges to the data

radar2 <- rbind(max_min, wide_loadings_df) %>%
    rename(
      "Limits on Wealth " = "WEALTHLIM",
      "Anti Mechanisation of work" = "MECHANISATION",
      "Pro Decentralisation" = "DECISIONDECEN",
      "Pro Centralisation" = "DECISIONCEN",
      "Pro Large Industries" = "INDUSTRYLARGE",
      "Anti Large Industries " = "INDUSTRYSMALL",
      "Pro Globaleconomy" = "ECONOMYGLOBAL",
      "Pro Localeconomy" = "ECONOMYLOCAL",
      "Environment over Development" = "ENVOVERDEV",
      "Development over Environment" = "DEVOVERENV",
      "Pro Private ownership " = "OWNERPVT",
      "Anti Regulations" = "OWNERNOREG",
      "Pro Public ownership" = "OWNERPUB",
      "Pro Regulations" = "OWNERREG",
      "Displacement risk(nuclear)" = "DISPLACENUCLEAR",
      "Pollution risk(nuclear) " = "POLLUTENUCLEAR",
      "Health risk(nuclear)" = "HEALTHNUCLEAR",
      "Job opportunities(nuclear)" = "JOBSNUCLEAR",
      "Spoils Natural Beauty(nuclear)" = "BEAUTYNUCLEAR",
      "Community pride(nuclear)" = "PRIDENUCLEAR",
      "National pride(nuclear)" = "NPRIDENUCLEAR",
      "National development(nuclear)" = "DEVNUCLEAR",
      "Local prosperity(nuclear)" = "PROSPERNUCLEAR",
      "Reliance on government(nuclear)" = "RELYNUCLEAR"
    )

```

```{r, echo=FALSE, fig.width=20, fig.height=18, out.width='100%', out.height='100%'}
library(fmsb)

create_beautiful_radarchart <- function(data, color = "#00AFBB", 
                                        vlabels = colnames(data), vlcex = 0.7,
                                        caxislabels = NULL, title = NULL, ...){
  radarchart(
    data, axistype = 1,
    # Customize the polygon
    pcol = color, pfcol = scales::alpha(color, 0.5), plwd = 2, plty = 1,
    centerzero=FALSE,
    # Customize the grid
    cglcol = "black", cglty = 1.5, cglwd = 1.5,
    # Customize the axis
    axislabcol = "black", 
     # Increase the axis label size
    cexc = 3,
    # Variable labels
    vlcex = vlcex, vlabels = vlabels, vlpos = 0.98,
    seg = 6, caxislabels = caxislabels, title = title, 
    # Increase title size
    cex.main = 2,
    ...
  )
}

# Reduce plot margin using par()
op <- par(mar = c(7, 3, 3, 3))
# Create the radar charts
create_beautiful_radarchart(
  data = radar2, caxislabels = c(-0.6,-0.4,-0.2, 0, 0.2, 0.4, 0.6, 0.8,1),
  color = c("#00AFBB", "#F97D09"),
  vlcex = 1.1
)

# Add an horizontal legend
legend(
  x = "bottomright", legend = rownames(radar2[-c(1,2),]), horiz = TRUE,
  bty = "n", pch = 20 , col = c("#00AFBB", "#F97D09"),
  text.col = "black", cex = 1.5, pt.cex = 3,
  inset = c(0, 0)
  )
par(op)

```

\newpage

```{r, echo=FALSE}

library(flextable)

flex <- function(data, title=NULL) {
  # this grabs the data and converts it to a flextbale
  flextable(data) %>%
  # this makes the table fill the page width
  set_table_properties(layout = "autofit", width = 1) %>%
  # font size
  fontsize(size=10, part="all") %>%
    #this adds a ttitlecreates an automatic table number
      set_caption(title, 
                  autonum = officer::run_autonum(seq_id = "tab", 
                                                 pre_label = "Table ", 
                                                 post_label = "\n", 
                                                 bkm = "anytable")) #%>%
  # font type
  #font(fontname="Times New Roman", part="all")
}

rename_map <- c(
  "WEALTHLIM" = "Limits on Wealth",
  "MECHANISATION" = "Anti Mechanisation of work",
  "DECISIONDECEN" = "Pro Decentralisation",
  "DECISIONCEN" = "Pro Centralisation",
  "INDUSTRYLARGE" = "Pro Large Industries",
  "INDUSTRYSMALL" = "Anti Large Industries",
  "ECONOMYGLOBAL" = "Pro Globaleconomy",
  "ECONOMYLOCAL" = "Pro Localeconomy",
  "ENVOVERDEV" = "Environment over Development",
  "DEVOVERENV" = "Development over Environment",
  "OWNERPVT" = "Pro Private ownership",
  "OWNERNOREG" = "Anti Regulations",
  "OWNERPUB" = "Pro Public ownership",
  "OWNERREG" = "Pro Regulations",
  "DISPLACENUCLEAR" = "Displacement risk(nuclear)",
  "POLLUTENUCLEAR" = "Pollution risk(nuclear)",
  "HEALTHNUCLEAR" = "Health risk(nuclear)",
  "JOBSNUCLEAR" = "Job opportunities(nuclear)",
  "BEAUTYNUCLEAR" = "Spoils Natural Beauty(nuclear)",
  "PRIDENUCLEAR" = "Community pride(nuclear)",
  "NPRIDENUCLEAR" = "National pride(nuclear)",
  "DEVNUCLEAR" = "National development(nuclear)",
  "PROSPERNUCLEAR" = "Local prosperity(nuclear)",
  "RELYNUCLEAR" = "Reliance on government(nuclear)"
)

fa_table <- function(x) {
  #get sorted loadings
  loadings <- fa.sort(x)$loadings %>% round(3)
  #supress loadings
  #loadings[loadings < cut] <- ""  # alternative loadings > -cut & loadings < cut
  #get additional info  
  add_info <- cbind(x$communalities, 
                    x$uniquenesses,
                    x$complexity) %>%
    # make it a data frame
    as.data.frame() %>%
    # column names
   rename("Communality" = V1,
           "Uniqueness" = V2,
           "Complexity" = V3) %>%
    #get the item names from the vector
    rownames_to_column("item")
  #build table
  loadings %>%
    unclass() %>%
    as.data.frame() %>%
    rownames_to_column("item") %>%
    left_join(add_info) %>%
    mutate(
      item = if_else(item %in% names(rename_map), rename_map[item], item),
      across(where(is.numeric), round, 3)
    )
}

fa_table(ecopolall2.fa) %>%
  rename(Pdevelop = MR1, Ndevelop = MR2, Items= item) %>%
  flex("Eco-Pol Values Factor Analysis Table")

```

```{r, echo=FALSE}

ecopolall2.fa[["Vaccounted"]] %>%
  as.data.frame() %>%
  rownames_to_column("Property") %>%
  rename(Pdevelop = MR1, Ndevelop = MR2) %>%
    mutate(across(where(is.numeric), round, 3)) %>%
    flex("Eigenvalues and Variance Explained for Rotated Factor Solution")

```

\newpage

```{r, echo=FALSE}

ecopolall <- fascales %>%
        dplyr::select("WEALTHLIM","MECHANISATION","DECISIONDECEN","DECISIONCEN","INDUSTRYLARGE","INDUSTRYSMALL","ECONOMYGLOBAL","ECONOMYLOCAL", "ENVOVERDEV","DEVOVERENV","OWNERPVT","OWNERNOREG","OWNERPUB", "OWNERREG","DISPLACENUCLEAR",  "POLLUTENUCLEAR", "HEALTHNUCLEAR", "JOBSNUCLEAR", "BEAUTYNUCLEAR", "PRIDENUCLEAR", "NPRIDENUCLEAR", "DEVNUCLEAR", "PROSPERNUCLEAR", "RELYNUCLEAR")

pdevelop <- ecopolall %>%
  dplyr::select("POLLUTENUCLEAR", "HEALTHNUCLEAR", "BEAUTYNUCLEAR","DISPLACENUCLEAR","MECHANISATION","INDUSTRYSMALL","OWNERPUB")

ndevelop <- ecopolall %>%
  dplyr::select( "PRIDENUCLEAR", "NPRIDENUCLEAR", "DEVNUCLEAR", "PROSPERNUCLEAR")

pdevelop_alpha <- psych::alpha(pdevelop, check.keys = TRUE)

ndevelop_alpha <- psych::alpha(ndevelop, check.keys = TRUE)

```

```{r, echo=FALSE}

data <- tibble(
  Scale = c(
    "People Centered Development (Pdevelop)", "", "", "", "", "", "",
    "Nationalist Development (Ndevelop)", "", "", ""
  ),
  Code = c(
    "Health risk(nuclear)", "Spoils Natural Beauty(nuclear)", "Anti Mechanisation of work",
    "Anti Large Industries", "Displacement risk(nuclear)", "Pollution risk(nuclear)",
    "Pro Regulations", "National development(nuclear)", "Community pride(nuclear)",
    "National pride(nuclear)", "Local prosperity(nuclear)"
  ),
  "Items and Loadings" = c(
    "Nuclear energy poses a great risk to the health of people living around it.(0.657)",
    "Nuclear energy spoils the natural beauty of the landscape.(0.638)",
    "Rapid mechanization of work is taking away jobs from workers in this country.(0.552)",
    "Large corporations are destroying the local industries in India and benefiting only a handful of people.(0.532)",
    "Nuclear energy is leading to displacement of people from their land.(0.59)",
    "Nuclear energy increases pollution of air/water/land.(0.565)",
    "Regardless of ownership, the government should pass strong regulations and implement them.(0.53)",
    "Nuclear energy pushes forward the country's development.(0.662)",
    "I would be proud if my community used nuclear energy.(0.623)",
    "Nuclear energy is a mark of pride for our nation.(0.605)",
    "Nuclear energy brings economic prosperity to the surrounding regions.(0.586)"
  ),
  Alpha = c(0.757, "", "", "", "", "", "", 0.725, "", "", ""),
  Variance = c(0.13, "", "", "", "", "", "", 0.10, "", "", "")
)

```

```{r, echo=FALSE}

longvars <- knitr::kable(data, booktabs = TRUE, caption = "Two Factor Solution: Economic and Political Values Scale", 
                         longtable = TRUE, escape = FALSE) %>%
  kable_styling(latex_options = c("striped","scale_down", "hold_position", "repeat_header")) %>%
  column_spec(column = 2, width = "4cm") %>%
  column_spec(column = 1, width = "4cm") %>%
  column_spec(column = 3, width = "9cm")

# Print the table in landscape mode
longvars %>%
  landscape()

```

# all lms after FA
```{r, echo=FALSE}
# 3 linear models - 1) lm with demographic variables except age and State
m1_demo <- lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + Urban, data = fascale_scores)

# 2) lm with all demographic variables
m2_demo<- lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + Urban + age + State, data = fascale_scores)

# 3) lm with kahan 
m1_kahan <- lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + Urban + age + State + KahanS + KahanH , data = fascale_scores)

# 4) lm with ecopolvars 
m1_allscales <- lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + Urban + age + State+ Pdevelop + Ndevelop + KahanS + KahanH , data = fascale_scores)

m1_ecopol <- lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + Urban + age + State + Pdevelop + Ndevelop,  data = fascale_scores)

summary(m1_ecopol)

m1_allscales2 <- lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + Urban + age + KahanS*State + KahanH*State + Pdevelop*State + Ndevelop*State , data = fascale_scores)

summary(m1_allscales2)

```

# Correlation Eco-pol and alt Kahan(mean across)

```{r}
# binding ecopol with alternative Kahan scale score 
fascale_scores2 <-cbind(Kscalescores, scores1)  %>%
  rename(Pdevelop = MR1)%>%
  rename(Ndevelop = MR2)

fascale_scores2

fascale_scores2cor <- fascale_scores2 %>%
  dplyr::select(Communitarianism_score, Egalitarianism_score, Pdevelop, Ndevelop)

cor_kecopol <- cor(fascale_scores2cor)

print(cor_kecopol)
```

# Lms with alt Kahan scores 
```{r, echo = FALSE}

# 3) lm with kahan 
m1_kahan2 <- lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + Urban + age + State + Communitarianism_score + Egalitarianism_score , data = fascale_scores2)
# Print model summary
summary(m1_kahan2)
#plot(m1_kahan)

# 4) lm with ecopolvars 
m1_allscales2 <- lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + Urban + age + State+ Pdevelop + Ndevelop + Communitarianism_score + Egalitarianism_score , data = fascale_scores2)
# Print model summary
summary(m1_allscales2)
#plot(m1_allscales)

m1_ecopol2 <- lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + Urban + age + State + Pdevelop + Ndevelop,  data = fascale_scores2)

summary(m1_ecopol2)

```

```{r, echo=FALSE, results='asis'}
library(stargazer)

cat("\\begingroup")
cat("\\setlength{\\tabcolsep}{1pt}")
cat("\\renewcommand{\\arraystretch}{0.7}")
stargazer(m1_kahan2, m1_ecopol2, m1_allscales2, 
          title = "Results from 3 linear regression models",
          type = "latex"
)
cat("\\endgroup")

```

\newpage

# LMs : eco-pol and kahan scale

```{r, echo=FALSE, results='asis'}
library(stargazer)

cat("\\begingroup")
cat("\\setlength{\\tabcolsep}{1pt}")
cat("\\renewcommand{\\arraystretch}{0.7}")
stargazer(m1_kahan, m1_allscales, 
          title = "Results from 2 linear regression models",
          type = "latex"
)
cat("\\endgroup")

```

\newpage
```{r, echo=FALSE, results='asis'}
library(stargazer)

cat("\\begingroup")
cat("\\setlength{\\tabcolsep}{1pt}")
cat("\\renewcommand{\\arraystretch}{0.7}")
stargazer(m1_ecopol, m1_allscales, 
          title = "Results from 2 linear regression models",
          type = "latex"
)
cat("\\endgroup")

```
\newpage

# LMs by State

```{r, echo=FALSE}

fascale_scores

fascale_scores2

# make state wise dts

# Uttar Pradesh
UP<- fascale_scores %>%
  filter(State == "Uttar Pradesh")

Raj<- fascale_scores %>%
  filter(State == "Rajasthan")

TM<- fascale_scores %>%
  filter(State == "Tamil Nadu")

Mah<- fascale_scores %>%
  filter(State == "Maharasthra")

Bengal <- fascale_scores %>%
  filter(State == "Bengal")

# 3 linear models - 1) lm with demographic variables except age and State
TM_demo <- lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + Urban, data = TM)

# 3) lm with kahan 
TM1_kahan <- lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + Urban + age  + KahanS + KahanH , data = TM)

# 4) lm with ecopolvars 
TM1_allscales <- lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + Urban + age + Pdevelop + Ndevelop + KahanS + KahanH , data = TM)

TM1_ecopol <- lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + Urban + age +  Pdevelop + Ndevelop,  data = TM)

summary(TM1_ecopol)

```

```{r, echo=FALSE, results='asis'}
library(stargazer)

cat("\\begingroup")
cat("\\setlength{\\tabcolsep}{1pt}")
cat("\\renewcommand{\\arraystretch}{0.7}")
stargazer(TM_demo, TM1_kahan, TM1_allscales, TM1_ecopol,
          title = "Results from 2 linear regression models",
          type = "latex"
)
cat("\\endgroup")

```