### 2024-01-15 nuclear in comparison.Rmd

---
title: "nuclear energy in comparison"
output:
  pdf_document:
    toc: yes
    toc_depth: 2
    keep_tex: yes
    keep_md: yes
  header-includes:
  - \usepackage{dcolumn}
  - \usepackage{array}
  - \usepackage{pdflscape}
  - \newcommand{\blandscape}{\begin{landscape}}
  - \newcommand{\elandscape}{\end{landscape}}
  html_document:
    toc: yes
    toc_depth: '2'
    df_print: paged
---

# Abstract

```{r load packages, warning = FALSE, message = FALSE, echo=FALSE}
  knitr::opts_chunk$set(warning = FALSE, message = FALSE)

suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(readxl))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(tidyr))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(lavaan))
suppressPackageStartupMessages(library(tsibble))
suppressPackageStartupMessages(library(DT))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(psych)) 
```

```{r setup, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

moduletwo <- read_excel(here::here("thesis_surveydata/political_factors_complete.xlsx"))
module2 <- data.frame(moduletwo)
module2[module2 == "Rather not say/ Don't know"] <- "Rather not say/Don't know"

# loaded data set for module 2 eco -pol variables and removes the rows with question number and question text
```

```{r, eval =TRUE, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
#This chunk contains the coding schemes of various scales used in survey one: eco-political factors, kahan scale and acceptance scale
codedmodule2 <- module2 %>%

#remove row 1
  filter(!row_number() %in% c(1,2)) %>%

# replace risky likert scale with numbers
  mutate_at(vars(starts_with("Risky")), funs(case_when(. =="Not at all risky" ~ 1, 
                                                       . =="Slightly risky" ~ 2, 
                                                       . =="Moderately risky" ~ 3, 
                                                       . =="Very risky" ~ 4, 
                                                       . =="Extremely risky" ~ 5))) %>%

# replace beneficial likert scale with numbers  
  mutate_at(vars(starts_with("Ben")), funs(case_when(. =="Not at all beneficial" ~ 1,
                                                     . =="Slightly beneficial" ~ 2,
                                                     . =="Moderately beneficial" ~ 3,
                                                     . =="Very beneficial" ~ 4,
                                                     . =="Extremely beneficial" ~ 5 ))) %>%

# replace nuclear acceptance likert scale with numbers
  mutate_at(vars(N_accept,N_reluctantlyaccept,N_reject), funs(case_when(. == "Strongly disagree" ~ 1, 
                                                                        . == "Somewhat disagree" ~ 2,
                                                                        . == "Neither agree nor disagree" ~3,
                                                                        . == "Somewhat agree" ~ 4,
                                                                        . == "Strongly agree" ~ 5))) %>%

# code likert scale for variables for Kahan scale into numbers
  mutate_at(vars(starts_with (c("K_I","K_H","DISPLACE", "POLLUTE", "HEALTH", "JOBS", "BEAUTY", "PRIDE", "NPRIDE", "DEV", "PROSPER", "RELY"))), funs(case_when(. == "Strongly disagree" ~ 1, 
               . == "Somewhat disagree" ~ 2,
               . == "Neither agree nor disagree" ~3,
               . == "Somewhat agree" ~ 4,
               . == "Strongly agree" ~ 5))) %>%

# reverse code for likert scale for variables for Kahan scale into numbers
  mutate_at(vars(starts_with (c("K_S","K_E"))), funs(case_when(. == "Strongly disagree" ~ 5, 
                                                               . == "Somewhat disagree" ~ 4,
                                                               . == "Neither agree nor disagree" ~3,
                                                               . == "Somewhat agree" ~ 2,
                                                               . == "Strongly agree" ~ 1))) %>%

# code eco-pol scale variables into numbers
  mutate_at(vars(SYSTEMDEMO,SYSTEMRELIGION,SYSTEMTECHNO,SYSTEMTOTAL,WEALTHLIM,MECHANISATION,DECISIONDECEN,INDUSTRYLARGE,ECONOMYGLOBAL,OWNERPVT,OWNERNOREG), funs(case_when(. == "Strongly disagree" ~ 1, 
                        . == "Somewhat disagree" ~ 2,
                        . == "Neither agree nor disagree" ~3,
                        . == "Somewhat agree" ~ 4,
                        . == "Strongly agree" ~ 5))) %>%

# reverse code eco-pol scale variables into numbers 
 mutate_at(vars(DECISIONCEN, INDUSTRYSMALL, ECONOMYLOCAL, ENVOVERDEV,OWNERPUB, OWNERREG), funs(case_when(. == "Strongly disagree" ~ 5, 
                                                                                                         . == "Somewhat disagree" ~ 4,
                                                                                                         . == "Neither agree nor disagree" ~3,
                                                                                                         . == "Somewhat agree" ~ 2,
                                                                                                         . == "Strongly agree" ~ 1))) %>%
  mutate_all(~ifelse(. == "Rather not say/ Don't know", "Rather not say/Don't know", .))%>%
  mutate_all(~ifelse(. == "West bengal", "West Bengal", .))%>%
  mutate_all(~ifelse(. == "WEST BENGAL", "West Bengal", .))%>%
  mutate_all(~ifelse(. == "Up", "Uttar Pradesh", .))%>%
  mutate_all(~ifelse(. == "Tamilnadu", "Tamil Nadu", .))%>%
  mutate_all(~ifelse(. == "UP", "Uttar Pradesh", .))

```

```{r, eval=TRUE, echo=FALSE}

moduleone <- read_excel(here::here("thesis_surveydata/livelihood_insecurity_complete.xlsx"))
module1 <- data.frame(moduleone)
# loaded data set for module 2 eco -pol variables and removes the rows with question number and question text
```

```{r, eval =TRUE, echo=FALSE}
#This chunk contains the coding schemes of various scales used in survey one: eco-political factors, kahan scale and acceptance scale
codedmodule1 <- module1 %>%

#remove row 1 and 2 
  filter(!row_number() %in% c(1,2)) %>%

# replace risky likert scale with numbers
  mutate_at(vars(starts_with("Risky")), funs(case_when(. =="Not at all risky" ~ 1, 
                                                       . =="Slightly risky" ~ 2, 
                                                       . =="Moderately risky" ~ 3, 
                                                       . =="Very risky" ~ 4, 
                                                       . =="Extremely risky" ~ 5))) %>%

# replace beneficial likert scale with numbers  
  mutate_at(vars(starts_with("Ben")), funs(case_when(. =="Not at all beneficial" ~ 1,
                                                     . =="Slightly beneficial" ~ 2,
                                                     . =="Moderately beneficial" ~ 3,
                                                     . =="Very beneficial" ~ 4,
                                                     . =="Extremely beneficial" ~ 5 ))) %>%

# replace nuclear acceptance likert scale with numbers
  mutate_at(vars(N_accept,N_reluctantlyaccept,N_reject), funs(case_when(. == "Strongly disagree" ~ 1, 
                                                                        . == "Somewhat disagree" ~ 2,
                                                                        . == "Neither agree nor disagree" ~3,
                                                                        . == "Somewhat agree" ~ 4,
                                                                        . == "Strongly agree" ~ 5)))%>%
  mutate_all(~ifelse(. == "Rather not say/ Don't know", "Rather not say/Don't know", .))%>%
  mutate_all(~ifelse(. == "West bengal", "West Bengal", .))%>%
  mutate_all(~ifelse(. == "WEST BENGAL", "West Bengal", .))%>%
  mutate_all(~ifelse(. == "Up", "Uttar Pradesh", .))%>%
  mutate_all(~ifelse(. == "Tamilnadu", "Tamil Nadu", .))%>%
  mutate_all(~ifelse(. == "UP", "Uttar Pradesh", .))

```

```{r, echo = FALSE}

suppressPackageStartupMessages(library(tidyselect))
# combine codedmodule1 and codedmodule2

numcol2 <- codedmodule2 %>% add_column(Respondent = 1:1099, .before = "Language") 
numcol1 <- codedmodule1 %>% add_column(Respondent = 1099:2160, .before = "Language")

joineddf <-  full_join(numcol2, numcol1, by = c("Respondent", "Language", "Survey_Date", "State", "Risky_Hydro", "Risky_Solar", "Risky_Wind", "Risky_Nuclear", "Risky_Coal", "Risky_Gas", "Risky_Oil", "Risky_INDHydro", "Risky_INDSolar", "Risky_INDWind", "Risky_INDBiogas", "Risky_INDDiesel", "Risky_INDKerosene", "Risky_INDFirewoodetc", "Risky_INDLPG", "Ben_Hydro", "Ben_Solar", "Ben_Wind", "Ben_Nuclear", "Ben_Coal", "Ben_Gas", "Ben_Oil", "Ben_INDHydro", "Ben_INDSolar", "Ben_INDWind", "Ben_INDDiesel", "Ben_INDBiogas", "Ben_INDFirewoodetc", "Ben_INDLPG", "Ben_INDKerosene", "N_accept", "N_reluctantlyaccept", "N_reject", "age", "urban_rural", "gender", "caste", "religion", "religiontext"))

numcolb <- module2 %>%   filter(!row_number() %in% c(1,2))%>% 
                        add_column(Respondent = 1:1099, .before = "Language") 
numcola <- module1 %>%   filter(!row_number() %in% c(1,2)) %>% 
                        add_column(Respondent = 1099:2160, .before = "Language")

uncodedjoineddf <- full_join(numcolb, numcola, by = c("Respondent", "Language", "Survey_Date", "State", "Risky_Hydro", "Risky_Solar", "Risky_Wind", "Risky_Nuclear", "Risky_Coal", "Risky_Gas", "Risky_Oil", "Risky_INDHydro", "Risky_INDSolar", "Risky_INDWind", "Risky_INDBiogas", "Risky_INDDiesel", "Risky_INDKerosene", "Risky_INDFirewoodetc", "Risky_INDLPG", "Ben_Hydro", "Ben_Solar", "Ben_Wind", "Ben_Nuclear", "Ben_Coal", "Ben_Gas", "Ben_Oil", "Ben_INDHydro", "Ben_INDSolar", "Ben_INDWind", "Ben_INDDiesel", "Ben_INDBiogas", "Ben_INDFirewoodetc", "Ben_INDLPG", "Ben_INDKerosene", "N_accept", "N_reluctantlyaccept", "N_reject", "age", "urban_rural", "gender", "caste", "religion", "religiontext")) %>%
  mutate_all(~ifelse(. == "Rather not say/ Don't know", "Rather not say/Don't know", .))%>%
  mutate_all(~ifelse(. == "West bengal", "West Bengal", .))%>%
  mutate_all(~ifelse(. == "WEST BENGAL", "West Bengal", .))%>%
  mutate_all(~ifelse(. == "Up", "Uttar Pradesh", .))%>%
  mutate_all(~ifelse(. == "Tamilnadu", "Tamil Nadu", .))%>%
  mutate_all(~ifelse(. == "UP", "Uttar Pradesh", .))

```

\newpage

# H1: Nuclear Energy will be seen as riskier than other energy technologies in India.  this paper explores nuclear energy in comparison with other technologies

## Likert Responses (n= 2160)
The percentages are rounded off to whole numbers.

```{r, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
# Q1: how much is the awareness of energy technologies in India ? How does it relate to state specific data ? 
# summary of NA results from the combined sample of 2000 responses (simple percentage bar graphs)
# part 1- how much is the awareness of energy technologies in India -

#plotting perceived risk from different technologies including rather not say don't know
colorsequence<- c("#8C4646","#D9695F","#F2A679","#F2D091","#5D8C7B","#808080")

#following is the data frame for making the barplot 
df_barplot <- uncodedjoineddf %>%
  dplyr::select(starts_with("RISKY"))%>%
  gather(Technology, Perceived_Risk, Risky_Hydro:Risky_INDLPG, factor_key = TRUE)%>%
  separate(Technology, c("Risky", "Technology"), extra = "merge", fill = "left")%>%
  dplyr::select(-Risky) %>% 
  group_by(Technology, Perceived_Risk) %>%
  dplyr::summarise(n=n())

#following data frame gives us percentage for each response 
df_barplot_percent<- df_barplot%>%
  group_by(Technology) %>%
  mutate(Percentage = round(n/sum(n)*100, 0)) %>%
  ungroup()

#plotting perceived benefit from different technologies including rather not say don't know

df_barplotben <- uncodedjoineddf %>%
  dplyr::select(starts_with("Ben"))%>%
  gather(Technology, Perceived_Benefit, Ben_Hydro:Ben_INDKerosene, factor_key = TRUE)%>%
  separate(Technology, c("Ben", "Technology"), extra = "merge", fill = "left")%>%
  dplyr::select(-Ben) %>% 
  group_by(Technology, Perceived_Benefit) %>%
  dplyr::summarise(n=n())

df_barplotben_percent<- df_barplotben%>%
  group_by(Technology) %>%
  mutate(Percentage = round(n/sum(n)*100, 0)) %>%
  ungroup()

#bar plot for likert responses perceived risk - centrally owned technologies 
df_barplot %>% ggplot(aes(fill= fct_rev(factor(Perceived_Risk, levels=c("Extremely risky", "Very risky","Moderately risky","Slightly risky","Not at all risky","Rather not say/Don't know"))), y=n , x=Technology)) + 
  geom_bar(position="fill",stat="identity")+
  geom_text(data = df_barplot_percent, aes(label=paste0(Percentage, "%"), y=n), position = position_fill(vjust=0.5), size=3) +
  scale_x_discrete(limits = c("Solar", "Wind" ,"Hydro", "Oil", "Coal","Gas", "Nuclear"))+
  scale_fill_manual("legend", values = c("Extremely risky" = "#8C4646", "Very risky" = "#D9695F", "Moderately risky" = "#F2A679", "Slightly risky" = "#F2D091","Not at all risky" = "#5D8C7B", "Rather not say/Don't know" = "grey94"))+
  labs( y = "Perceived Risk", x ="Centrally Managed Energy Technologies")+
  theme_classic()

#perceived benefit centrally owned technologies

df_barplotben %>% ggplot(aes(fill= factor(Perceived_Benefit, levels=c("Rather not say/Don't know","Not at all beneficial","Slightly beneficial","Moderately beneficial","Very beneficial", "Extremely beneficial")), y=n , x=Technology)) + 
  geom_bar(position="fill",stat="identity")+
   geom_text(data = df_barplotben_percent, aes(label=paste0(Percentage, "%"), y=n), position = position_fill(vjust=0.5), size=3) +
  scale_fill_manual("legend", values = c("Rather not say/Don't know" = "grey94" ,"Not at all beneficial" = "orange","Slightly beneficial"="yellow" ,"Moderately beneficial" = "palegreen","Very beneficial"= "green3", "Extremely beneficial"= "green4"))+
  scale_x_discrete(limits = c("Solar", "Wind", "Hydro", "Oil", "Coal","Gas","Nuclear"))+
  labs( y = "Perceived Benefit", x ="Centrally Managed Energy Technologies")+
  theme_classic()

```

\newpage
## Introduction: Mean Perceived Risk and Mean Perceived Benefit for all energy technologies.

```{r, echo=FALSE}
library(tidyverse)

# Filter variables of interest
joineddf_subset <- joineddf %>%
  dplyr::select(!starts_with(c("Ben_IND", "Risky_IND")))%>%
  dplyr::select(starts_with("Risky"), 
         starts_with("Ben"))

# Assuming your data is named joineddf_subset
long_df <- joineddf_subset %>%
  gather(key = "Type", value = "Value")

# Separate out Risky and Ben columns
risky_df <- long_df %>% filter(str_detect(Type, "^Risky"))
ben_df <- long_df %>% filter(str_detect(Type, "^Ben"))

# Extract the technology type (Hydro, Solar, etc.)
risky_df$Tech <- str_replace(risky_df$Type, "Risky_", "")
ben_df$Tech <- str_replace(ben_df$Type, "Ben_", "")

# Calculate the mean for each technology for Risky
risky_mean <- risky_df %>%
  group_by(Tech) %>%
  dplyr::summarize(Mean_Value = mean(Value, na.rm = TRUE),
          sd = sd(Value, na.rm = TRUE),
          se = sd(Value, na.rm = TRUE) / sqrt(n()),
          .groups = "drop")

# Do the same for Ben technologies
ben_mean <- ben_df %>%
  group_by(Tech) %>%
  dplyr::summarize(Mean_Value = mean(Value, na.rm = TRUE),
          sd = sd(Value, na.rm = TRUE),
          se = sd(Value, na.rm = TRUE) / sqrt(n()),
          .groups = "drop")

# Combine the two data frames into one for plotting
combined_mean <- bind_rows(
  mutate(risky_mean, Category = "Risky"),
  mutate(ben_mean, Category = "Beneficial")
)
# Reorder the factor levels of the Tech variable by Mean_Value
risky_mean$Tech <- factor(risky_mean$Tech, levels = risky_mean$Tech[order(risky_mean$Mean_Value, decreasing = FALSE)])

```

```{r, echo=FALSE}
library(ggplot2)
library(gridExtra)

mean_risk <- mean(combined_mean$Mean_Value[combined_mean$Category == "Risky"], na.rm = TRUE)
mean_benefit <- mean(combined_mean$Mean_Value[combined_mean$Category == "Beneficial"], na.rm = TRUE)

# Order Technologies
technology_order <- c("Solar","Wind","Hydro","Oil","Coal", "Gas","Nuclear")
ben_mean$Tech <- factor(ben_mean$Tech, levels = technology_order)

# Generate the Ben plot with flipped y-axis
ben_plot <- ggplot(ben_mean, aes(y = Tech, x = Mean_Value)) +
  geom_bar(stat = "identity", fill = "palegreen3") +
  geom_errorbar(aes(xmin = Mean_Value - se, xmax = Mean_Value + se), width = 0.2) +
  coord_cartesian(xlim = c(5, 1)) + 
  labs(title = "Perceived Benefit", x = "Mean Values") +
  #scale_x_reverse() +
  theme(  # remove y-axis labels
        axis.title.y = element_blank()) + # remove the "Tech" axis title
  geom_vline(data = subset(combined_mean, Category == "Beneficial"), 
             aes(xintercept = mean_benefit), linetype="dotted", color = "black", size = 0.5)

# Generate the Risky plot
risky_plot <- ggplot(risky_mean, aes(y = Tech, x = Mean_Value)) +
  geom_bar(stat = "identity", fill = "firebrick2") +
  geom_errorbar(aes(xmin = Mean_Value - se, xmax = Mean_Value + se), width = 0.2) +
  coord_cartesian(xlim = c(1, 5)) + 
  labs(title = "Perceived Risk", x = "Mean Values") +
  theme(axis.text.y = element_blank(),
        axis.title.y = element_blank())+
  geom_vline(data = subset(combined_mean, Category == "Risky"), 
             aes(xintercept = mean_risk), linetype="dotted", color = "black", size = 0.5)
 # remove the "Tech" axis title

# Arrange plots side by side
grid.arrange(ben_plot, risky_plot, ncol=2)

```

## Boxplot for perceived benefit and perceived risk by technology

```{r, echo = FALSE, fig.width=16, fig.height=6}

#Box plot to see the distribution

# Loading the required libraries
library(tidyverse)
library(stringr)
library(gridExtra)

# Reshape the data to a long format
joineddf_subset_long <- joineddf_subset %>%
  pivot_longer(cols = everything(), 
               names_to = "variable",
               values_to = "value")

# Separate the data into two: one for "Risky_" variables and one for "Ben_" variables
risky_df <- joineddf_subset_long %>% 
  filter(str_detect(variable, "^Risky_")) %>%
  mutate(variable = str_remove(variable, "Risky_"))  # Remove the "Risky_" prefix

ben_df <- joineddf_subset_long %>% 
  filter(str_detect(variable, "^Ben_")) %>%
  mutate(variable = str_remove(variable, "Ben_"))  # Remove the "Ben_" prefix

# Remove NA values
risky_df <- risky_df %>% drop_na()
ben_df <- ben_df %>% drop_na()

# Create box plots
risky_plot <- ggplot(risky_df, aes(x = variable, y = value, fill = variable)) +
  geom_boxplot() +
  scale_fill_manual(values = rep("firebrick2", length(unique(risky_df$variable)))) +
  labs(x = "Energy Technologies", y = "Likert", fill = "Technologies") +
  ggtitle("Perceived risk") +
  theme_minimal()+
  theme(axis.text.y = element_text(size=20), 
        axis.text.x = element_text(size=20),
        axis.title.y = element_text(size=20),
        axis.title.x = element_text(size=20),
        title = element_text(size=20))+
  guides(fill = FALSE)

ben_plot <- ggplot(ben_df, aes(x = variable, y = value, fill = variable)) +
  geom_boxplot() +
  scale_fill_manual(values = rep("palegreen3", length(unique(ben_df$variable)))) +
  labs(x = "Energy Technologies", y = "Likert Responses", fill = "Technologies") +
  ggtitle("Perceived Benefit") +
  theme_minimal()+
  theme(axis.text.y = element_text(size=20), 
        axis.text.x = element_text(size=20),
        axis.title.y = element_text(size=20),
        axis.title.x = element_text(size=20),
        title = element_text(size=20))+
  guides(fill = FALSE)

# Put the two plots next to each other
grid.arrange(risky_plot, ben_plot, ncol = 2)

```

\newpage

## Pairwise T-test: Mean perceived risk and mean perceived benefit (all energy technologies)

The red and green pairs indicate that there is a statistically significant difference between the means of the two groups. 
White and grey indicate - no differences between the means of the two groups.

```{r, echo = FALSE, fig.width=14, fig.height=6}
#Q2 :How are the perceptions of the risk of producing energy from nuclear fission different from that of other conventional sources of energy like coal or hydro power in India?
#H : Nuclear energy will be perceived as riskier than other sources of energy (coal or hydro) in India like in the west.
perceivedriskc <-joineddf %>% 
  dplyr::select("Risky_Solar", "Risky_Hydro", "Risky_Wind", "Risky_Nuclear", "Risky_Coal", "Risky_Gas", "Risky_Oil") %>%
  gather(key = "technology", value = "perceived_risk") %>% 
  mutate(technology = gsub("Risky_", "", technology)) %>%
  na.omit()

result_ttest <- pairwise.t.test(perceivedriskc$perceived_risk, perceivedriskc$technology, p.adjust.method = "BH")

# visualising the pairwaise t test results. the non significant pairs are grey

tidy_result_ttest <- broom::tidy(result_ttest)

heatmap1 <- ggplot(tidy_result_ttest, aes(x = group1, y = group2, fill = p.value)) + 
  geom_tile(color = "white") + 
  scale_fill_gradient2(low = "firebrick2", high = "white", midpoint = 0.025, 
                       limits = c(0, 0.05), space = "Lab", 
                       name = "P-value") + 
  theme_classic() + 
  labs(x = "", y = "") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
   theme(axis.text.y = element_text(size=20), 
        axis.text.x = element_text(size=20),
        axis.title.y = element_text(size=20),
        axis.title.x = element_text(size=20),
        title = element_text(size=20),
        legend.text = element_text(size=12))

#Question- Why there is significant difference statistically - if the medians are the same in the boxplot.
#Q3: How do perceptions of benefit from nuclear energy compare with other source of energy?

# H : benefits from all energy technologies will be similar 
# include regression, heat maps, bar graphs? - decide which ones are good to include

perceivedbenefitc <-joineddf %>% 
  dplyr::select("Ben_Solar", "Ben_Hydro", "Ben_Wind", "Ben_Nuclear", "Ben_Coal", "Ben_Gas", "Ben_Oil") %>%
  gather(key = "technology", value = "perceived_benefit") %>% 
  mutate(technology = gsub("Ben_", "", technology)) %>%
  na.omit()

result_ttestben <- pairwise.t.test(perceivedbenefitc$perceived_benefit, perceivedbenefitc$technology, p.adjust.method = "BH")

# visualising the pairwaise t test results. the non significant pairs are grey

tidy_resultbc <- broom::tidy(result_ttestben)

library(ggplot2)
heatmap2 <- ggplot(tidy_resultbc, aes(x = group1, y = group2, fill = p.value)) + 
  geom_tile(color = "white") + 
  scale_fill_gradient2(low = "palegreen3", high = "white", midpoint = 0.025, 
                       limits = c(0, 0.05), space = "Lab", 
                       name = "P-value") + 
  theme_classic() + 
  labs(x = "", y = "") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
   theme(axis.text.y = element_text(size=20), 
        axis.text.x = element_text(size=20),
        axis.title.y = element_text(size=20),
        axis.title.x = element_text(size=20),
        title = element_text(size=20),
        legend.text = element_text(size=12))

grid.arrange(heatmap1, heatmap2, ncol=2)

```

\newpage

## Benefit - Risk for each technology 
```{r, echo = FALSE}

# combining risk and benefit from each energy technology 
# Filter variables of interest
joineddf_subset <- joineddf %>%
  dplyr::select(!starts_with(c("Ben_IND", "Risky_IND")))%>%
  dplyr::select(starts_with("Risky"), 
         starts_with("Ben"))

acceptance <- joineddf_subset %>%
  na.omit()%>%
  mutate(Nuclear = Ben_Nuclear - Risky_Nuclear,
         Hyrdo = Ben_Hydro - Risky_Hydro,
         Wind = Ben_Wind - Risky_Wind,
         Solar = Ben_Solar - Risky_Solar,
         Coal = Ben_Coal - Risky_Coal,
         Gas = Ben_Gas - Risky_Gas,
         Oil = Ben_Oil - Risky_Oil)%>%
  dplyr::select(-starts_with(c("Risky", "Ben")))

#str(acceptance)

# Compute the mean of each column
mean_acceptance <- data.frame(col = names(acceptance), mean_val = sapply(acceptance, mean))

# Compute standard error (not deviation)(divided by root of n)
sem <- function(x) sqrt(var(x, na.rm = TRUE) / length(na.omit(x)))
sem_acceptance <- data.frame(col = names(acceptance), sem_val = sapply(acceptance, sem))

mean_acceptance <- merge(mean_acceptance, sem_acceptance, by = "col")

# Plot
ggplot(mean_acceptance, aes(x = col, y = mean_val, fill = ifelse(mean_val >= 0, 'Above 0', 'Below 0'))) + 
  geom_bar(stat = "identity") +
  geom_errorbar(aes(ymin = mean_val - sem_val, ymax = mean_val + sem_val), width = 0.2) +
  geom_text(aes(label = round(mean_val, 2)), vjust = -0.9) +
  scale_fill_manual(values = c('Above 0' = 'palegreen3', 'Below 0' = 'firebrick2')) +
  labs(fill = "Legend") +
  xlab("Energy Technologies") +
  ylab("Mean Value") +
  ggtitle("Benefit- Risk for Each Energy Technology (n = 1184)") +
  theme_minimal()+
  ylim(-0.5, 1.75)

```

\newpage

# Claim 1: Demographic variables: nuclear, solar, coal

```{r, echo=FALSE}

# # Convert `caste` into binary variable 'upper' (1) vs 'non-upper' (0)
joineddf$Uppercaste <- ifelse(joineddf$caste %in% c('Brahmin', 'General'), 1, 0)
# Convert `gender` into binary variable 'male' (1) vs 'female' (0)
joineddf$Male <- ifelse(joineddf$gender == 'Male', 1, 0)
# Convert `religion` into binary variable 'Hinduism' (1) vs 'Non-Hinduism' (0)
joineddf$Hindu <- ifelse(joineddf$religion == 'Hinduism', 1, 0)
codedmodule2$State <- as.factor(codedmodule2$State)
codedmodule2$gender <- as.factor(codedmodule2$gender)
codedmodule2$Urban <- as.factor(codedmodule2$urban_rural)
joineddf$Urban <- as.factor(joineddf$urban_rural)
codedmodule2$caste <- as.factor(codedmodule2$caste)
codedmodule2$religion <- as.factor(codedmodule2$religion)
# # Convert `caste` into binary variable 'upper' (1) vs 'non-upper' (0)
codedmodule2$Uppercaste <- ifelse(codedmodule2$caste %in% c('Brahmin', 'General'), 1, 0)
# Convert `gender` into binary variable 'male' (1) vs 'female' (0)
codedmodule2$Male <- ifelse(codedmodule2$gender == 'Male', 1, 0)
# Convert `religion` into binary variable 'Hinduism' (1) vs 'Non-Hinduism' (0)
codedmodule2$Hindu <- ifelse(codedmodule2$religion == 'Hinduism', 1, 0)

joineddf1 <- joineddf %>%
            dplyr::select(Risky_Nuclear, Risky_Solar, Risky_Coal, Uppercaste, Male, Hindu, Urban, age, State) %>%
            na.omit()

# lm on nuclear energy with all demovars

alldemos <- joineddf1 %>%
            dplyr::select(Risky_Nuclear, Uppercaste, Male, Hindu, Urban, age, State) %>%
            mutate(RiskNuclear = Risky_Nuclear)%>%
  mutate_at(vars(age), funs(case_when(. == "18-24 years old" ~ 1, 
                                      . == "25-34 years old" ~ 2,
                                      . ==  "35-44 years old" ~3,
                                      . == "45-54 years old" ~ 4,
                                      . == "55-64 years old" ~ 5,
                                      . == "65-74 years old" ~ 6,
                                      . == "75 years or older" ~ 7)))

alldemoslm1 <- lm(RiskNuclear ~ Uppercaste + Male + Hindu + Urban + age , data = alldemos)

alldemoslm2<- lm(RiskNuclear ~ Uppercaste + Male + Hindu + Urban + age + State, data = alldemos)

#lm on solar with all demovars

salldemos <- joineddf1 %>%
            dplyr::select(Risky_Solar,Uppercaste, Male, Hindu, Urban, age, State)%>%
            mutate(RiskSolar = Risky_Solar)%>%
  mutate_at(vars(age), funs(case_when(. == "18-24 years old" ~ 1, 
                                      . == "25-34 years old" ~ 2,
                                      . ==  "35-44 years old" ~3,
                                      . == "45-54 years old" ~ 4,
                                      . == "55-64 years old" ~ 5,
                                      . == "65-74 years old" ~ 6,
                                      . == "75 years or older" ~ 7)))

salldemoslm1 <- lm(RiskSolar ~ Uppercaste + Male + Hindu + Urban + age , data = salldemos)

salldemoslm2<- lm(RiskSolar ~ Uppercaste + Male + Hindu + Urban + age + State, data = salldemos)

#lm on coal with all demovars 
calldemos <- joineddf1 %>%
            dplyr::select(Risky_Coal,Uppercaste, Male, Hindu, Urban, age, State)%>%
            mutate(RiskCoal = Risky_Coal)%>%
            mutate_at(vars(age), funs(case_when(. == "18-24 years old" ~ 1, 
                                      . == "25-34 years old" ~ 2,
                                      . ==  "35-44 years old" ~3,
                                      . == "45-54 years old" ~ 4,
                                      . == "55-64 years old" ~ 5,
                                      . == "65-74 years old" ~ 6,
                                      . == "75 years or older" ~ 7)))

calldemoslm1 <- lm(RiskCoal ~ Uppercaste + Male + Hindu + Urban + age, data = calldemos)

calldemoslm2<- lm(RiskCoal ~ Uppercaste + Male + Hindu + Urban + age + State, data = calldemos)

```

```{r, echo=FALSE}

# trying linear regression with acceptance scale

joineddf_subset1 <- joineddf %>%
  dplyr::select(!starts_with(c("Ben_IND", "Risky_IND")))%>%
  dplyr::select(starts_with("Risky"), 
         starts_with("Ben"), Uppercaste, Male, Hindu, Urban, age)

acceptance1 <- joineddf_subset1 %>%
  na.omit()%>%
  mutate(NetBenNuclear = Ben_Nuclear - Risky_Nuclear,
         NetBeneHyrdo = Ben_Hydro - Risky_Hydro,
         NetBeneWind = Ben_Wind - Risky_Wind,
         NetBenSolar = Ben_Solar - Risky_Solar,
         NetBenCoal = Ben_Coal - Risky_Coal,
         NetBenGas = Ben_Gas - Risky_Gas,
         NetBenOil = Ben_Oil - Risky_Oil)%>%
  dplyr::select(-starts_with(c("Risky", "Ben")))%>%
  mutate_at(vars(age), funs(case_when(. == "18-24 years old" ~ 1, 
                                      . == "25-34 years old" ~ 2,
                                      . ==  "35-44 years old" ~3,
                                      . == "45-54 years old" ~ 4,
                                      . == "55-64 years old" ~ 5,
                                      . == "65-74 years old" ~ 6,
                                      . == "75 years or older" ~ 7)))

alldemoslm3<- lm(NetBenNuclear ~ Uppercaste + Male + Hindu + Urban + age, data = acceptance1)
#summary(alldemoslm3)

salldemoslm3<- lm(NetBenSolar ~ Uppercaste + Male + Hindu + Urban + age, data = acceptance1)
calldemoslm3<- lm(NetBenCoal ~ Uppercaste + Male + Hindu + Urban + age, data = acceptance1)

```

\blandscape

## Linear Regression with demographic variables (caste, religion, gender and age)

```{r, echo=FALSE, results='asis'}
library(stargazer)

#stargazer code for all demographic variable lms

cat("\\begingroup")
cat("\\small")  # Adjusts the font size
cat("\\setlength{\\tabcolsep}{1pt}")
cat("\\renewcommand{\\arraystretch}{0.7}")
stargazer(alldemoslm1, salldemoslm1, calldemoslm1,alldemoslm3, salldemoslm3, calldemoslm3,
          title = "Linear Models: Perceived Risk and Net Perceived Benefit(Nuclear, Solar and Coal)",
          type = "latex"
)
cat("\\endgroup")

```

\elandscape

```{r, echo=FALSE}

stargazer(alldemoslm1, salldemoslm1, calldemoslm1, alldemoslm3, salldemoslm3, calldemoslm3, 
          type = "text", # Choose "html" or "latex" for different formats
          title = "Regression Models Summary",
          align = TRUE)

```

\newpage

# Claim 2: Regional differences

We see huge differences for perceived risk and perceived benefit for each technology from state to state. These graphs explore that. The number of respondents from each state are also reported below.

## Perceived risk by State

```{r, echo=FALSE, fig.width=11, fig.height=14, out.width='80%', out.height='80%'}

# part 2: state specific data on perceived risk
state_risk <- uncodedjoineddf %>%
    dplyr::select(starts_with(c("RISKY", "State")))%>%
    gather(Technology, Perceived_Risk, Risky_Hydro:Risky_INDLPG, factor_key = TRUE)%>%
    separate(Technology, c("Risky", "Technology"), extra = "merge", fill = "left")%>%
    dplyr::select(-Risky) %>% 
    group_by(State, Technology, Perceived_Risk) %>%
    dplyr::summarise(n=n())

count<- uncodedjoineddf %>%
  dplyr::select(starts_with(c("RISKY", "State")))%>%
   group_by(State) %>%
  dplyr::summarise(n=n())

 state_risk_percent <- state_risk %>%
    group_by(Technology, State) %>%
    mutate(Percentage = round(n/sum(n)*100, 0)) %>%
    ungroup()

 # Create a named vector with the number of respondents for each state
respondents_per_state <- c("Maharashtra" = 653, "Rajasthan" = 677, "Tamil Nadu" = 320, "Uttar Pradesh" = 125, "West Bengal" = 386)

respondents_per_state

state_risk %>% ggplot(aes(fill= fct_rev(factor(Perceived_Risk, levels=c("Extremely risky", "Very risky","Moderately risky","Slightly risky","Not at all risky","Rather not say/Don't know"))), y=n , x=Technology)) + 
    geom_bar(position="fill",stat="identity")+
     geom_text(data = state_risk_percent, aes(label=paste0(Percentage, "%"), y=n), position = position_fill(vjust=0.5), size=5) +
    scale_x_discrete(limits = c("Solar", "Wind" ,"Hydro", "Oil", "Coal","Gas", "Nuclear"))+
    scale_fill_manual("legend", values = c("Extremely risky" = "#8C4646", "Very risky" = "#D9695F", "Moderately risky" = "#F2A679", "Slightly risky" = "#F2D091","Not at all risky" = "#5D8C7B", "Rather not say/Don't know" = "grey94"))+
    theme_classic()+ 
    facet_wrap(~ State, ncol = 2)+
    labs(x = "mean perceived risk", y = "energy technologies")+
    theme(axis.text.x = element_text(angle = 45, hjust = 1))+
    theme(axis.text.x = element_text(angle = 45, hjust = 1, size=12),  # Increase x-axis text size
        axis.text.y = element_text(size=12),  # Increase y-axis text size
        strip.text = element_text(size=12))  # Increase facet label text size

```

\newpage

## Perceived Benefit by State

```{r, echo=FALSE, fig.width=12, fig.height=15, out.width='100%', out.height='100%'}
# part 2: state specific data on perceived benefit

state_ben <- uncodedjoineddf %>%
  dplyr::select(starts_with(c("Ben", "State")))%>%
  gather(Technology, Perceived_Benefit, Ben_Hydro:Ben_INDKerosene, factor_key = TRUE)%>%
  separate(Technology, c("Ben", "Technology"), extra = "merge", fill = "left")%>%
  dplyr::select(-Ben) %>% 
  group_by(State, Technology, Perceived_Benefit) %>%
  dplyr::summarise(n=n())

countben<- uncodedjoineddf %>%
  dplyr::select(starts_with(c("Ben", "State")))%>%
   group_by(State) %>%
  dplyr::summarise(n=n())

 # Create a named vector with the number of respondents for each state
respondents_per_state <- c("Maharashtra" = 653, "Rajasthan" = 677, "Tamil Nadu" = 320, "Uttar Pradesh" = 125, "West Bengal" = 386)

respondents_per_state

# Calculate percentages in each group
state_ben_percent <- state_ben %>%
  group_by(Technology, State) %>%
  mutate(Percentage = round(n/sum(n)*100, 0)) %>%
  ungroup()

#perceived benefit centrally owned technologies by state
state_ben %>% ggplot(aes(fill= factor(Perceived_Benefit, levels=c("Rather not say/Don't know","Not at all beneficial","Slightly beneficial","Moderately beneficial","Very beneficial", "Extremely beneficial")), y=n , x=Technology)) + 
  geom_bar(position="fill", stat = "identity")+
  geom_text(data = state_ben_percent, aes(label=paste0(Percentage, "%"), y=n), position = position_fill(vjust=0.5), size=5) +
  scale_fill_manual("legend", values = c("Rather not say/Don't know" = "grey94" ,"Not at all beneficial" = "orange","Slightly beneficial"="yellow" ,"Moderately beneficial" = "palegreen","Very beneficial"= "green3", "Extremely beneficial"= "green4"))+
  scale_x_discrete(limits = c("Solar", "Wind", "Hydro", "Oil", "Coal","Gas","Nuclear"))+
  theme_classic()+
  facet_wrap(~ State, ncol = 2)+
  labs(x = "mean perceived benefit", y = "energy technologies")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size=12), 
        axis.text.y = element_text(size=12), 
        strip.text = element_text(size=12))

```
## Radar chart - risk and benefit by state

```{r, echo=FALSE}
library(tidyverse)

# Filter variables of interest
joineddf_subset2 <- joineddf %>%
  dplyr::select(!starts_with(c("Ben_IND", "Risky_IND")))%>%
  dplyr::select(starts_with("Risky"), 
         starts_with("Ben"), starts_with("State"))

#long_df2 <- joineddf_subset %>%
#gather(key = "Type", value = "Value")
# Reshape data to long format
long_df2 <- joineddf_subset2 %>%
  pivot_longer(cols = -State, names_to = "variable", values_to = "value")

# Separate out Risky and Ben columns
risky_df2 <- long_df2 %>% filter(str_detect(variable, "^Risky"))
ben_df2 <- long_df2 %>% filter(str_detect(variable, "^Ben"))

# Extract the technology type (Hydro, Solar, etc.)
risky_df2$Tech <- str_replace(risky_df2$variable, "Risky_", "")
ben_df2$Tech <- str_replace(ben_df2$variable, "Ben_", "")

# Calculate the mean for each technology for Risky
risky_mean2 <- risky_df2 %>%
  group_by(State,Tech) %>%
  dplyr::summarize(Mean_Value = mean(value, na.rm = TRUE),
          sd = sd(value, na.rm = TRUE),
          se = sd(value, na.rm = TRUE) / sqrt(n()),
          .groups = "drop")

# Do the same for Ben technologies
ben_mean2 <- ben_df2 %>%
  group_by(State,Tech) %>%
  dplyr::summarize(Mean_Value = mean(value, na.rm = TRUE),
          sd = sd(value, na.rm = TRUE),
          se = sd(value, na.rm = TRUE) / sqrt(n()),
          .groups = "drop")

# Combine the two data frames into one for plotting
combined_mean2 <- bind_rows(
  mutate(risky_mean2, Category = "Risky"),
  mutate(ben_mean2, Category = "Beneficial")
)

# New order of technologies
new_tech_order <- c("Solar", "Wind", "Hydro", "Oil", "Coal", "Gas","Nuclear")

# Apply this new order to the 'Tech' column in your combined_mean dataset
combined_mean2$Tech <- factor(combined_mean2$Tech, levels = new_tech_order)

# Split the data by state
data_by_state <- split(combined_mean2, combined_mean2$State)

```

```{r, echo=FALSE}
# Create a list to store plots
plots_list <- list()

for (state in names(data_by_state)) {
  # Subset data for each state
  state_data <- data_by_state[[state]]

  # Split into beneficial and risky data
  ben_mean <- subset(state_data, Category == "Beneficial")
  risky_mean <- subset(state_data, Category == "Risky")

  # Generate plots for each category
  ben_plot <- ggplot(ben_mean, aes(y = Tech, x = Mean_Value)) +
    geom_bar(stat = "identity", fill = "palegreen3") +
    geom_errorbar(aes(xmin = Mean_Value - se, xmax = Mean_Value + se), width = 0.2) +
    coord_cartesian(xlim = c(5, 1)) + 
    labs(title = paste("Perceived Benefit in", state), x = "Mean Values") +
    theme(axis.title.y = element_blank()) +
    geom_vline(xintercept = mean_benefit, linetype="dotted", color = "black", size = 0.5)

  risky_plot <- ggplot(risky_mean, aes(y = Tech, x = Mean_Value)) +
    geom_bar(stat = "identity", fill = "firebrick2") +
    geom_errorbar(aes(xmin = Mean_Value - se, xmax = Mean_Value + se), width = 0.2) +
    coord_cartesian(xlim = c(1, 5)) + 
    labs(title = paste("Perceived Risk in", state), x = "Mean Values") +
    theme(axis.text.y = element_blank(), axis.title.y = element_blank())+
    geom_vline(xintercept = mean_risk, linetype="dotted", color = "black", size = 0.5)

  # Combine the plots side by side
  combined_plot <- grid.arrange(ben_plot, risky_plot, ncol=2)
  plots_list[[state]] <- combined_plot
}

```

## Net benefit by state

```{r, echo=FALSE}
library(dplyr)
library(ggplot2)

# Including the State variable in the dataset
joineddf_subset_state <- joineddf %>%
  dplyr::select(starts_with("Risky"), 
         starts_with("Ben"), 
         State)

acceptance_state <- joineddf_subset_state %>%
  na.omit() %>%
  mutate(Nuclear = Ben_Nuclear - Risky_Nuclear,
         Hydro = Ben_Hydro - Risky_Hydro,
         Wind = Ben_Wind - Risky_Wind,
         Solar = Ben_Solar - Risky_Solar,
         Coal = Ben_Coal - Risky_Coal,
         Gas = Ben_Gas - Risky_Gas,
         Oil = Ben_Oil - Risky_Oil) %>%
  dplyr::select(-starts_with(c("Risky", "Ben")))

# Calculate mean and standard error for each state and technology
mean_acceptance_state <- acceptance_state %>%
  group_by(State) %>%
  summarise(across(Nuclear:Oil, list(mean = ~mean(., na.rm = TRUE), 
                                     sem = ~sqrt(var(., na.rm = TRUE) / length(na.omit(.))))))

```

```{r, echo=FALSE}

# Reshape data for plotting
long_mean <- mean_acceptance_state %>%
  pivot_longer(cols = -State, names_to = "Technology_SEM", values_to = "mean_val") %>%
  separate(Technology_SEM, into = c("Technology", "Metric"), sep = "_") %>%
  filter(Metric == "mean") %>%
  select(-Metric)

long_sem <- mean_acceptance_state %>%
  pivot_longer(cols = -State, names_to = "Technology_SEM", values_to = "sem_val") %>%
  separate(Technology_SEM, into = c("Technology", "Metric"), sep = "_") %>%
  filter(Metric == "sem") %>%
  select(-Metric)

# Merge mean and SEM data
plot_data <- left_join(long_mean, long_sem, by = c("State", "Technology"))

# Create the plot
ggplot(plot_data, aes(x = Technology, y = mean_val, fill = ifelse(mean_val >= 0, 'Above 0', 'Below 0'))) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = mean_val - sem_val, ymax = mean_val + sem_val), 
                position = position_dodge(), width = 0.25) +
  facet_wrap(~ State) +
  scale_fill_manual(values = c('Above 0' = 'palegreen3', 'Below 0' = 'firebrick2')) +
  labs(title = "Net Benefit (Perceived Risk - Perceived Benefit) by State",
       x = "Energy Technologies", y = "Net Benefit", fill = "Legend") +
  theme_minimal() +
  ylim(-1.75, 3.5)

```

```{r, echo=FALSE}

staterisky <- joineddf %>%
  dplyr::select(!starts_with(c("Ben_IND", "Risky_IND")))%>%
  dplyr::select(starts_with("Risky"), State)

staterisky_mean <- staterisky %>%
  group_by(State) %>%
  summarise(
    Mean_Risky_Hydro = mean(Risky_Hydro, na.rm = TRUE),
    Mean_Risky_Solar = mean(Risky_Solar, na.rm = TRUE),
    Mean_Risky_Wind = mean(Risky_Wind, na.rm = TRUE),
    Mean_Risky_Nuclear = mean(Risky_Nuclear, na.rm = TRUE),
    Mean_Risky_Coal = mean(Risky_Coal, na.rm = TRUE),
    Mean_Risky_Gas = mean(Risky_Gas, na.rm = TRUE),
    Mean_Risky_Oil = mean(Risky_Oil, na.rm = TRUE)
  )

# Convert to long format
long_data <- staterisky_mean %>%
  gather(Technology, Mean_Score, -State) %>%
  mutate(Technology = gsub("Mean_Risky_", "", Technology))

# Order Technologies
technology_order <- c("Coal", "Gas", "Nuclear", "Oil", "Hydro", "Wind", "Solar")
long_data$Technology <- factor(long_data$Technology, levels = technology_order)

cbbPalette <- c("#000000","#CC79A7","#F0E442","#E69F00","#56B4E9","#009E73","#D55E00")

names(cbbPalette) <- technology_order

# Order States (assuming you want an alphabetical order, modify as needed)
states_order <- c( "Rajasthan", "Maharashtra", "Tamil Nadu", "Uttar Pradesh",  "West Bengal")
long_data$State <- factor(long_data$State, levels = states_order)

# Plot the graph
stateplot1 <- ggplot(long_data, aes(x = State, y = Mean_Score, color = Technology, group = Technology)) +
  geom_line(size = 0.7) +
  geom_point(size = 2) +
  labs(title = " ",
       x = "State", y = "Means") +
  ylim(1,5)+
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_color_manual(values=cbbPalette)

```

```{r, echo= FALSE}

# Reshape the data from wide to long format
staterisky_long <- staterisky %>%
  pivot_longer(
    cols = starts_with("Risky_"),
    names_to = "Technology",
    values_to = "Riskiness"
  ) %>%
  mutate(Technology = sub("Risky_", "", Technology))

# Calculate the sample size for each state

state_sample_size <- staterisky %>%
  group_by(State) %>%
  summarize(n = n())

# Join this back to the main data frame
staterisky_long <- staterisky_long %>%
  left_join(state_sample_size, by = "State")

# Calculate the weighted mean for each technology in each state
weighted_means <- staterisky_long %>%
  group_by(State, Technology) %>%
  summarize(weighted_mean = sum(Riskiness / n, na.rm = TRUE) / sum(1 / n, na.rm = TRUE), .groups = 'drop')

```

```{r, echo=FALSE}

weighted_means$Technology <- factor(weighted_means$Technology, levels = technology_order)
states_order <- c( "Rajasthan", "Maharashtra", "Tamil Nadu", "Uttar Pradesh",  "West Bengal")
weighted_means$State <- factor(weighted_means$State, levels = states_order)

stateplot2 <- ggplot(weighted_means, aes(x = State, y = weighted_mean, color = Technology, group = Technology)) +
  geom_line(size = 0.7) +
  geom_point(size = 2) +
  labs(title = " ",
       x = "State", y = "Weighted Means") +
  ylim(0.5,4.5)+
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_color_manual(values = cbbPalette)

```

\newpage 
## Mean Perceived Risk and Weighted Mean

```{r, echo=FALSE, fig.width=8, fig.height=4,out.width='100%', out.height='100%'}
library(gridExtra)
# mean perceived risk and zscores by state side by side

spacer <- grid::rectGrob(gp = grid::gpar(col = "white"))

grid.arrange(stateplot1, spacer, stateplot2, ncol=3, widths=c(0.8, 0.06, 0.8))
```

\newpage
## Mean Perceived risk and weighten means by Technologies- solar, coal and nuclear

```{r, echo=FALSE}

staterisky3t <- joineddf %>%
  dplyr::select(!starts_with(c("Ben_IND", "Risky_IND")))%>%
  dplyr::select(Risky_Solar, Risky_Nuclear, Risky_Coal, State)

staterisky3t_mean <- staterisky3t %>%
  group_by(State) %>%
  summarise(
    Mean_Risky_Solar = mean(Risky_Solar, na.rm = TRUE),
    Mean_Risky_Nuclear = mean(Risky_Nuclear, na.rm = TRUE),
    Mean_Risky_Coal = mean(Risky_Coal, na.rm = TRUE)
  )

# Convert to long format
long_data3t <- staterisky3t_mean %>%
  gather(Technology, Mean_Score, -State) %>%
  mutate(Technology = gsub("Mean_Risky_", "", Technology))

# Order Technologies
technology_order <- c("Coal", "Gas", "Nuclear", "Oil", "Hydro", "Wind", "Solar")
long_data3t$Technology <- factor(long_data3t$Technology, levels = technology_order)

cbbPalette <- c("#000000","#CC79A7","#F0E442","#E69F00","#56B4E9","#009E73","#D55E00")

names(cbbPalette) <- technology_order

# Order States (assuming you want an alphabetical order, modify as needed)
states_order <- c( "Rajasthan", "Maharashtra", "Tamil Nadu", "Uttar Pradesh",  "West Bengal")
long_data3t$State <- factor(long_data3t$State, levels = states_order)

# Plot the graph
stateplot3t <- ggplot(long_data3t, aes(x = State, y = Mean_Score, color = Technology, group = Technology)) +
  geom_line(size = 0.7) +
  geom_point(size = 2) +
  labs(title = " ",
       x = "State", y = "Means") +
  ylim(1,5)+
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_color_manual(values=cbbPalette)

```

```{r, echo= FALSE}

# Reshape the data from wide to long format
staterisky3t_long <- staterisky3t %>%
  pivot_longer(
    cols = starts_with("Risky_"),
    names_to = "Technology",
    values_to = "Riskiness"
  ) %>%
  mutate(Technology = sub("Risky_", "", Technology))

# Calculate the sample size for each state

state_sample_size3t <- staterisky3t %>%
  group_by(State) %>%
  summarize(n = n())

# Join this back to the main data frame
staterisky3t_long <- staterisky3t_long %>%
  left_join(state_sample_size3t, by = "State")

# Calculate the weighted mean for each technology in each state
weighted_means3t <- staterisky3t_long %>%
  group_by(State, Technology) %>%
  summarize(weighted_mean = sum(Riskiness / n, na.rm = TRUE) / sum(1 / n, na.rm = TRUE), .groups = 'drop')

```

```{r, echo=FALSE}

weighted_means3t$Technology <- factor(weighted_means3t$Technology, levels = technology_order)
states_order <- c( "Rajasthan", "Maharashtra", "Tamil Nadu", "Uttar Pradesh",  "West Bengal")
weighted_means3t$State <- factor(weighted_means3t$State, levels = states_order)

stateplot3t2 <- ggplot(weighted_means3t, aes(x = State, y = weighted_mean, color = Technology, group = Technology)) +
  geom_line(size = 0.7) +
  geom_point(size = 2) +
  labs(title = " ",
       x = "State", y = "Weighted Means") +
  ylim(0.5,4.5)+
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_color_manual(values = cbbPalette)

```

```{r, echo=FALSE, fig.width=8, fig.height=4,out.width='100%', out.height='100%'}
library(gridExtra)
# mean perceived risk and zscores by state side by side

spacer <- grid::rectGrob(gp = grid::gpar(col = "white"))

grid.arrange(stateplot3t, spacer, stateplot3t2, ncol=3, widths=c(0.8, 0.06, 0.8))
```

## Zscores - 3 techs - solar, nuclear and coal

```{r, echo = FALSE, fig.width=8, fig.height=4}

# Extract only Nuclear
nuclear_data <- staterisky3t_mean %>%
  dplyr::select(State, Mean_Risky_Nuclear)%>%
  mutate(
    Z_Risky_Nuclear = (Mean_Risky_Nuclear - mean(Mean_Risky_Nuclear, na.rm = TRUE)) / 
                      sd(Mean_Risky_Nuclear, na.rm = TRUE)
  )

mean_nuclear_state <- ggplot(nuclear_data, aes(x = State, y = Mean_Risky_Nuclear, group = 1)) +
  geom_line(color = "#F0BA32", size = 1) +
  geom_point(color = "#F0BA32", size = 3) +
  labs(title = "Means: Perceived Risk of Nuclear Energy",
       x = "State", y = "Mean Perceived Risk") +
  theme_minimal() +
  ylim(1,5)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot the z-scores
zscore_nuclear_state <- ggplot(nuclear_data, aes(x = State, y = Z_Risky_Nuclear, group = 1)) +
  geom_line(color = "#F0BA32", size = 1) +
  geom_point(color = "#F0BA32", size = 3) +
  labs(title = "Z-scores: Perceived Risk of Nuclear Energy",
       x = "State", y = "Z-Score") +
  ylim(-1 ,2)+
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Extract only Solar
solar_data <- staterisky3t_mean %>%
  dplyr::select(State, Mean_Risky_Solar)%>%
  mutate(
    Z_Risky_Solar = (Mean_Risky_Solar - mean(Mean_Risky_Solar, na.rm = TRUE)) / 
                      sd(Mean_Risky_Solar, na.rm = TRUE)
  )

mean_solar_state <- ggplot(solar_data, aes(x = State, y = Mean_Risky_Solar, group = 1)) +
  geom_line(color = "#D55E00", size = 1) +
  geom_point(color = "#D55E00", size = 3) +
  labs(title = "Means: Perceived Risk of Solar Energy",
       x = "State", y = "Mean Perceived Risk") +
  theme_minimal() +
  ylim(1,5)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot the z-scores
zscore_solar_state <- ggplot(solar_data, aes(x = State, y = Z_Risky_Solar, group = 1)) +
  geom_line(color = "#D55E00", size = 1) +
  geom_point(color = "#D55E00", size = 3) +
  labs(title = "Z-scores: Perceived Risk of Solar Energy",
       x = "State", y = "Z-Score") +
  ylim(-1 ,2)+
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Extract only Coal
coal_data <- staterisky3t_mean %>%
  dplyr::select(State, Mean_Risky_Coal)%>%
  mutate(
    Z_Risky_Coal = (Mean_Risky_Coal - mean(Mean_Risky_Coal, na.rm = TRUE)) / 
                      sd(Mean_Risky_Coal, na.rm = TRUE)
  )

mean_coal_state <- ggplot(coal_data, aes(x = State, y = Mean_Risky_Coal, group = 1)) +
  geom_line(color = "#000000", size = 1) +
  geom_point(color = "#000000", size = 3) +
  labs(title = "Means: Perceived Risk of Coal",
       x = "State", y = "Mean Perceived Risk") +
  theme_minimal() +
  ylim(1,5)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot the z-scores
zscore_coal_state <- ggplot(coal_data, aes(x = State, y = Z_Risky_Coal, group = 1)) +
  geom_line(color = "#000000", size = 1) +
  geom_point(color = "#000000", size = 3) +
  labs(title = "Z-scores: Perceived Risk of Coal",
       x = "State", y = "Z-Score") +
  ylim(-2 ,1)+
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

library(gridExtra)

# Assuming mean_nuclear_state and zscore_nuclear_state are both ggplot objects

# We'll add a spacer grob for spacing between plots
spacer <- grid::rectGrob(gp = grid::gpar(col = "white"))

grid.arrange(stateplot3t, spacer, zscore_nuclear_state, ncol=3, widths=c(0.8, 0.06, 0.8))

grid.arrange(zscore_solar_state, spacer, zscore_coal_state, ncol=3, widths=c(0.8, 0.06, 0.8))

```

```{r, echo=FALSE}
stateben <- joineddf %>%
  dplyr::select(!starts_with(c("Ben_IND", "Risky_IND")))%>%
  dplyr::select(starts_with("Ben"), State)

stateben_mean <- stateben %>%
  group_by(State) %>%
  summarise(
    Mean_Ben_Hydro = mean(Ben_Hydro, na.rm = TRUE),
    Mean_Ben_Solar = mean(Ben_Solar, na.rm = TRUE),
    Mean_Ben_Wind = mean(Ben_Wind, na.rm = TRUE),
    Mean_Ben_Nuclear = mean(Ben_Nuclear, na.rm = TRUE),
    Mean_Ben_Coal = mean(Ben_Coal, na.rm = TRUE),
    Mean_Ben_Gas = mean(Ben_Gas, na.rm = TRUE),
    Mean_Ben_Oil = mean(Ben_Oil, na.rm = TRUE)
  )

# Convert to long format
long_data2 <- stateben_mean %>%
  gather(Technology, Mean_Score, -State) %>%
  mutate(Technology = gsub("Mean_Ben_", "", Technology))

# Order Technologies
technology_order <- c("Coal", "Gas", "Nuclear", "Oil", "Hydro", "Wind", "Solar")
long_data2$Technology <- factor(long_data$Technology, levels = technology_order)

cbbPalette <- c("#000000","#CC79A7","#F0E442","#E69F00","#56B4E9","#009E73","#D55E00")

names(cbbPalette) <- technology_order

# Order States (assuming you want an alphabetical order, modify as needed)
states_order <- c( "Rajasthan", "Maharashtra", "Tamil Nadu", "Uttar Pradesh",  "West Bengal")
long_data2$State <- factor(long_data2$State, levels = states_order)

# Plot the graph
stateplotben1 <- ggplot(long_data2, aes(x = State, y = Mean_Score, color = Technology, group = Technology)) +
  geom_line(size = 0.7) +
  geom_point(size = 2) +
  labs(title = "Mean Perceived Benefit by State",
       x = "State", y = "Mean Score") +
  ylim(1,5)+
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_color_manual(values=cbbPalette)
```

\newpage

## Mean Perceived Benefit and Z-scores by State

```{r, echo=FALSE}

standardized_data2 <- long_data2 %>%
  group_by(State) %>% #z scores are by state- that means mean_score - overall mean by state (including all technologies) divided by sd by state
  mutate(
    Overall_Mean = mean(Mean_Score, na.rm = TRUE),
    Std_Dev = sd(Mean_Score, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  mutate(Standardized_Score = (Mean_Score - Overall_Mean) / Std_Dev) #%>%
  #dplyr::select(-Overall_Mean, -Std_Dev)  #

# Order Technologies
standardized_data2$Technology <- factor(standardized_data2$Technology, levels = technology_order)

stateplotben2 <- ggplot(standardized_data2, aes(x = State, y = Standardized_Score, color = Technology, group = Technology)) +
  geom_line(size = 0.7) +
  geom_point(size = 2) +
  labs(title = "Z Scores for Perceived Benefit by State",
       x = "State", y = "Z Score") +
  ylim(-3,2)+
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_color_manual(values = cbbPalette)

```

```{r, echo=FALSE, fig.width=8, fig.height=4, out.width='100%', out.height='100%'}
library(gridExtra)
# mean perceived risk and zscores by state side by side

spacer <- grid::rectGrob(gp = grid::gpar(col = "white"))

grid.arrange(stateplotben1, spacer, stateplotben2, ncol=3, widths=c(0.8, 0.06, 0.8))
```

\newpage

## Mean Perceived Benefit and Z-scores by State - coal solar and nuclear

```{r, echo = FALSE, fig.width=8, fig.height=4}

staterisky_mean <- staterisky %>%
  group_by(State) %>%
  summarise(
    Mean_Risky_Hydro = mean(Risky_Hydro, na.rm = TRUE),
    Mean_Risky_Solar = mean(Risky_Solar, na.rm = TRUE),
    Mean_Risky_Wind = mean(Risky_Wind, na.rm = TRUE),
    Mean_Risky_Nuclear = mean(Risky_Nuclear, na.rm = TRUE),
    Mean_Risky_Coal = mean(Risky_Coal, na.rm = TRUE),
    Mean_Risky_Gas = mean(Risky_Gas, na.rm = TRUE),
    Mean_Risky_Oil = mean(Risky_Oil, na.rm = TRUE)
  )

# Extract only Nuclear
nuclear_data <- staterisky_mean %>%
  dplyr::select(State, Mean_Risky_Nuclear)%>%
  mutate(
    Z_Risky_Nuclear = (Mean_Risky_Nuclear - mean(Mean_Risky_Nuclear, na.rm = TRUE)) / 
                      sd(Mean_Risky_Nuclear, na.rm = TRUE)
  )

mean_nuclear_state <- ggplot(nuclear_data, aes(x = State, y = Mean_Risky_Nuclear, group = 1)) +
  geom_line(color = "#F0BA32", size = 1) +
  geom_point(color = "#F0BA32", size = 3) +
  labs(title = "Means: Perceived Risk of Nuclear Energy",
       x = "State", y = "Mean Perceived Risk") +
  theme_minimal() +
  ylim(1,5)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot the z-scores
zscore_nuclear_state <- ggplot(nuclear_data, aes(x = State, y = Z_Risky_Nuclear, group = 1)) +
  geom_line(color = "#F0BA32", size = 1) +
  geom_point(color = "#F0BA32", size = 3) +
  labs(title = "Z-scores: Perceived Risk of Nuclear Energy",
       x = "State", y = "Z-Score") +
  ylim(-1 ,2)+
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

library(gridExtra)

# Assuming mean_nuclear_state and zscore_nuclear_state are both ggplot objects

# We'll add a spacer grob for spacing between plots
spacer <- grid::rectGrob(gp = grid::gpar(col = "white"))

grid.arrange(mean_nuclear_state, spacer, zscore_nuclear_state, ncol=3, widths=c(0.8, 0.06, 0.8))

```

\newpage

```{r, echo=FALSE}
# Assuming you've already processed staterisky_mean as in your given code

staterisky3 <- joineddf %>%
  dplyr::select(!starts_with(c("Ben_IND", "Risky_IND")))%>%
  dplyr::select(Risky_Solar,Risky_Nuclear, Risky_Coal, State)

staterisky3_mean <- staterisky3 %>%
  group_by(State) %>%
  summarise(
    Mean_Risky_Solar = mean(Risky_Solar, na.rm = TRUE),
    Mean_Risky_Nuclear = mean(Risky_Nuclear, na.rm = TRUE),
    Mean_Risky_Coal = mean(Risky_Coal, na.rm = TRUE),
  )

# Convert to long format
long_data3 <- staterisky3_mean %>%
  gather(Technology, Mean_Score, -State) %>%
  mutate(Technology = gsub("Mean_Risky_", "", Technology))

# Order Technologies
technology_order <- c("Coal", "Gas", "Nuclear", "Oil", "Hydro", "Wind", "Solar")
long_data3$Technology <- factor(long_data3$Technology, levels = technology_order)

cbbPalette <- c("#000000","#CC79A7","#F0E442","#E69F00","#56B4E9","#009E73","#D55E00")

names(cbbPalette) <- technology_order

# Order States (assuming you want an alphabetical order, modify as needed)
states_order <- c( "Rajasthan", "Maharashtra", "Tamil Nadu", "Uttar Pradesh",  "West Bengal")
long_data3$State <- factor(long_data3$State, levels = states_order)

# Plot the graph
stateplot3 <- ggplot(long_data3, aes(x = State, y = Mean_Score, color = Technology, group = Technology)) +
  geom_line(size = 0.7) +
  geom_point(size = 2) +
  labs(title = "Mean Perceived Risk",
       x = "State", y = "Mean Score") +
  ylim(1,5)+
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_color_manual(values=cbbPalette)

```

```{r, echo = FALSE}

staterisky_long <- staterisky_long %>%
  group_by(State) %>%
  mutate(Riskiness_Z = scale(Riskiness, center = TRUE, scale = TRUE))

ggplot(staterisky_long, aes(x = Technology, y = Riskiness_Z, fill = State)) +
  geom_boxplot() +
  facet_wrap(~ State) +
  theme_minimal() +
  labs(title = "Boxplot of Riskiness Z-scores by State and Technology", 
       x = "Technology", 
       y = "Z-score")

ggplot(staterisky_long, aes(x = Riskiness_Z, fill = Technology)) +
  geom_density(alpha = 0.7) +
  facet_wrap(~ State) +
  theme_minimal() +
  labs(title = "Density Plot of Riskiness Z-scores by State and Technology", 
       x = "Z-score", 
       y = "Density")

technology_order <- c("Coal", "Gas", "Nuclear", "Oil", "Hydro", "Wind", "Solar")
staterisky_long$Technology <- factor(staterisky_long$Technology, levels = technology_order)
cbbPalette <- c("#000000","#CC79A7","#F0E442","#E69F00","#56B4E9","#009E73","#D55E00")

names(cbbPalette) <- technology_order

ggplot(staterisky_long, aes(x = Riskiness_Z, color = Technology)) +
  geom_density() +
  facet_wrap(~ State) +
  theme_minimal() +
  labs(title = "Density Plot of Riskiness Z-scores by State and Technology", 
       x = "Z-score", 
       y = "Density")+
  scale_color_manual(values=cbbPalette)

# Assuming you have the ggplot2 package loaded
# Filter the dataset for only Coal, Nuclear, and Solar technologies
filtered_data <- staterisky_long %>%
  filter(Technology %in% c("Coal", "Nuclear", "Solar"))

# Define your color palette (replace 'cbbPalette' with actual color values if needed)
cbbPalette <- c("Coal" = "#000000", "Nuclear" = "#F0E442", "Solar" = "#D55E00")

# Generate the plot
ggplot(filtered_data, aes(x = Riskiness_Z, color = Technology)) +
  geom_density() +
  facet_wrap(~ State) +
  theme_minimal() +
  labs(title = "Density Plot of Riskiness Z-scores by State and Technology", 
       x = "Z-score", 
       y = "Density") +
  scale_color_manual(values = cbbPalette)

```

```{r, echo=FALSE}

# trying linear regression with acceptance scale

joineddf_subset2 <- joineddf %>%
  dplyr::select(!starts_with(c("Ben_IND", "Risky_IND")))%>%
  dplyr::select(starts_with("Risky"), 
         starts_with("Ben"), Uppercaste, Male, Hindu, Urban, age, State)

acceptance2 <- joineddf_subset2 %>%
  na.omit()%>%
  mutate(NetBenNuclear = Ben_Nuclear - Risky_Nuclear,
         NetBenHyrdo = Ben_Hydro - Risky_Hydro,
         NetBenWind = Ben_Wind - Risky_Wind,
         NetBenSolar = Ben_Solar - Risky_Solar,
         NetBenCoal = Ben_Coal - Risky_Coal,
         NetBenGas = Ben_Gas - Risky_Gas,
         NetBenOil = Ben_Oil - Risky_Oil)%>%
  dplyr::select(-starts_with(c("Risky", "Ben")))%>%
  mutate_at(vars(age), funs(case_when(. == "18-24 years old" ~ 1, 
                                      . == "25-34 years old" ~ 2,
                                      . ==  "35-44 years old" ~3,
                                      . == "45-54 years old" ~ 4,
                                      . == "55-64 years old" ~ 5,
                                      . == "65-74 years old" ~ 6,
                                      . == "75 years or older" ~ 7)))

alldemoslm4<- lm(NetBenNuclear ~ Uppercaste + Male + Hindu + Urban + age + State, data = acceptance2)
#summary(alldemoslm3)

salldemoslm4<- lm(NetBenSolar ~ Uppercaste + Male + Hindu + Urban + age + State, data = acceptance2)
calldemoslm4<- lm(NetBenCoal ~ Uppercaste + Male + Hindu + Urban + age + State, data = acceptance2)

```

\blandscape
\newpage

## Linear Regression : Perceived Risk and Net Benefit- demographic variables including State

```{r, echo=FALSE, results='asis'}
library(stargazer)

#stargazer code for all demographic variable lms

cat("\\begingroup")
cat("\\small")  # Adjusts the font size
cat("\\setlength{\\tabcolsep}{1pt}")
cat("\\renewcommand{\\arraystretch}{0.7}")
stargazer(alldemoslm2, salldemoslm2, calldemoslm2,alldemoslm4, salldemoslm4, calldemoslm4,
          title = "Linear Models: Perceived Risk and Net Perceived Benefit(Nuclear, Solar and Coal)",
          type = "latex"
)
cat("\\endgroup")

```

```{r, echo=FALSE, results='asis'}
library(stargazer)

#stargazer code for all demographic variable lms

# stargazer code for all demographic variable linear models

stargazer(alldemoslm2, salldemoslm2, calldemoslm2, alldemoslm4, salldemoslm4, calldemoslm4,
          title = "Linear Models: Perceived Risk and Net Perceived Benefit (Nuclear, Solar and Coal)",
          type = "html"
          )

```

\elandscape

```{r, echo=FALSE}

stargazer(alldemoslm2, salldemoslm2, calldemoslm2, alldemoslm4, salldemoslm4, calldemoslm4, 
          type = "text", # Choose "html" or "latex" for different formats
          title = "Regression Models Summary",
          align = TRUE)

```

\newpage

# Claim 3: Cultural, Political and Economic Values

```{r, echo = FALSE}

library(lavaan)
library(semTools)
library(kableExtra)

fascales <- codedmodule2 %>%
  dplyr::select(starts_with(c("K_","Risky_Nuclear","WEALTHLIM","MECHANISATION","DECISIONDECEN","DECISIONCEN","INDUSTRYLARGE","INDUSTRYSMALL","ECONOMYGLOBAL","ECONOMYLOCAL", "ENVOVERDEV","DEVOVERENV","OWNERPVT","OWNERNOREG","OWNERPUB", "OWNERREG","DISPLACENUCLEAR",  "POLLUTENUCLEAR", "HEALTHNUCLEAR", "JOBSNUCLEAR", "BEAUTYNUCLEAR", "PRIDENUCLEAR", "NPRIDENUCLEAR", "DEVNUCLEAR", "PROSPERNUCLEAR", "RELYNUCLEAR", "Uppercaste", "Male", "Hindu", "Urban", "State","age")))%>%
  mutate_at(vars(age), funs(case_when(. == "18-24 years old" ~ 1, 
                                      . == "25-34 years old" ~ 2,
                                      . ==  "35-44 years old" ~3,
                                      . == "45-54 years old" ~ 4,
                                      . == "55-64 years old" ~ 5,
                                      . == "65-74 years old" ~ 6,
                                      . == "75 years or older" ~ 7))) %>%
  na.omit()

Kahan_fas <- '
  KahanS =~ K_SHARM + K_SLIMCHOI + K_SPROTECT 
  KahanH =~ K_HEQUAL+ K_ERADEQ1 + K_EWEALTH + K_ERADEQ2
'
Kahan_cfa <- cfa(Kahan_fas, data = fascales, std.lv = TRUE)

Kahan_scores <- predict(Kahan_cfa, newdata = fascales)

Kahan_Kscores <- cbind(fascales, Kahan_scores)

```

## CFA on eco-pol scale for Solar Energy

```{r, echo = FALSE}
# kahan scores for solar 
sfascales <- codedmodule2 %>%
  dplyr::select(starts_with(c("K_","Risky_Solar","WEALTHLIM","MECHANISATION","DECISIONDECEN","DECISIONCEN","INDUSTRYLARGE","INDUSTRYSMALL","ECONOMYGLOBAL","ECONOMYLOCAL", "ENVOVERDEV","DEVOVERENV","OWNERPVT","OWNERNOREG","OWNERPUB", "OWNERREG","DISPLACESOLAR",  "POLLUTESOLAR", "HEALTHSOLAR", "JOBSSOLAR", "BEAUTYSOLAR", "PRIDESOLAR", "NPRIDESOLAR", "DEVSOLAR", "PROSPERSOLAR", "RELYSOLAR", "Uppercaste", "Male", "Hindu", "Urban", "State","age")))%>%
  na.omit()

sKahan_fas <- '
  KahanS =~ K_SHARM + K_SLIMCHOI + K_SPROTECT 
  KahanH =~ K_HEQUAL+ K_ERADEQ1 + K_EWEALTH + K_ERADEQ2
'

# Run CFA
sKahan_cfa <- cfa(sKahan_fas, data = sfascales, std.lv = TRUE)

# Print the summary of the CFA results, including fit measures
#summary(sKahan_cfa, fit.measure= TRUE, standardized = TRUE)

sKahan_scores <- predict(sKahan_cfa, newdata = sfascales)

sKahan_Kscores <- cbind(sfascales, sKahan_scores)

sm1_kahan <- lm(Risky_Solar ~ Uppercaste + Male + Hindu + Urban + age + State + KahanS + KahanH , data = sKahan_Kscores)

## CFA on ecopol vars for solar energy

secopol_fas <- '
  Pdevelop =~ HEALTHSOLAR + BEAUTYSOLAR + DISPLACESOLAR + POLLUTESOLAR + INDUSTRYSMALL + OWNERREG + MECHANISATION
  Ndevelop =~ DEVSOLAR + PRIDESOLAR + NPRIDESOLAR + PROSPERSOLAR + JOBSSOLAR
'

secopol_cfa <- cfa(secopol_fas, data = sfascales, std.lv = TRUE)
#summary(secopol_cfa, fit.measure= TRUE, standardized = TRUE)

# Cronbach's alpha test on Pdevelop and Ndevelop for Solar energy

solarpdevelop <- sfascales %>%
  dplyr::select(HEALTHSOLAR, BEAUTYSOLAR, DISPLACESOLAR, POLLUTESOLAR, INDUSTRYSMALL, OWNERREG, MECHANISATION)%>%
  na.omit()

solarpdevelop.alpha <- psych::alpha(solarpdevelop, check.keys = TRUE)

# alpha is strong at 0.75!

solarndevelop <- sfascales %>%
  dplyr::select(DEVSOLAR, PRIDESOLAR, NPRIDESOLAR, PROSPERSOLAR, JOBSSOLAR)%>%
  na.omit()

solarndevelop.alpha <- psych::alpha(solarndevelop, check.keys = TRUE)

#alpha is strong at 0.86!

```

```{r, echo = FALSE}

# Extract fit measures
fitsecopol <- fitMeasures(secopol_cfa, fit.measures = c("CFI", "TLI", "RMSEA", "rmsea.ci.upper", "rmsea.ci.lower"))
fitsecopol2 <- fitMeasures(secopol_cfa)
# Convert to a data frame for presentation
fitsecopoldf <- data.frame(Measure = names(fitsecopol), Value = fitsecopol) %>%
  mutate(Measure = case_when(
    Measure == "cfi" ~ "Comparative Fit Index (CFI) ",
    Measure == "tli" ~ "Tucker-Lewis Index (TLI)",
    Measure == "rmsea" ~ "Root Mean Square Error of Approximation(RMSEA)",
    Measure == "rmsea.ci.upper" ~ "RMSEA 90 Percent confidence interval - lower",
    Measure == "rmsea.ci.lower" ~ "RMSEA 90 Percent confidence interval - upper",
    TRUE ~ as.character(Measure)
  )) %>%
  mutate(Value = round(Value, 3)) %>%
  remove_rownames()

#RMSEA - between 0.05 and 0.08 (reasonable fit), less than 0.05 close fit and greater than 0.10 is poor fit 
#CFI - greater than 0.90 or 0.95 indicates good fit
#TFI - CFI>TFI >0.90 - indicates good fit

```

```{r, echo =FALSE}
# Display fit measures using kable
#knitr::kable(fitsecopoldf, caption = "Fit Measures: CFA on eco-pol scale for Solar Energy", booktabs = TRUE) %>%
  #kable_styling(latex_options = c("striped", "hold_position"))

```

```{r, echo=FALSE}

# Extract parameter estimates
paramssolar <- parameterEstimates(secopol_cfa, standardized = TRUE)%>%
   slice(1:(n()-14))

paramssolar1 <- paramssolar %>%
  mutate(rhs = case_when(
    rhs == "HEALTHSOLAR" ~ "Solar energy poses a great risk to the health of people living around it.",
    rhs == "BEAUTYSOLAR" ~ "Solar energy spoils the natural beauty of the landscape.",
    rhs == "DISPLACESOLAR" ~ "Solar energy is leading to displacement of people from their land.",
    rhs == "POLLUTESOLAR" ~ "Solar energy increases pollution of air/water/land.",
    rhs == "INDUSTRYSMALL" ~ "Large corporations are destroying the local industries in India and benefiting only a handful of people.",
    rhs == "OWNERREG" ~ "Regardless of ownership, the government should pass strong regulations and implement them.",
    rhs == "DEVSOLAR" ~ "Solar energy pushes forward the country's development.",
    rhs == "PRIDESOLAR" ~ "I would be proud if my community used solar energy",
    rhs == "NPRIDESOLAR" ~ "Solar energy is a mark of pride for our nation.",
    rhs == "PROSPERSOLAR" ~ "Solar energy brings economic prosperity to the surrounding regions.",
    rhs == "JOBSSOLAR" ~ "Solar energy will bring jobs to the local community.",
    TRUE ~ as.character(rhs)
  )) %>%
  dplyr::select(!c(lhs,op, std.nox)) %>%
    rename(
    "Items" = rhs,
    "Loadings" = est,
    "Standard Error" = se,
    "zvalue" = z) %>%
  mutate(across(c("Loadings", "Standard Error", "zvalue"), round, 3))

paramssolar1$Scale <- NA
paramssolar1$Scale[1:6] <- "People Centered Development"
paramssolar1$Scale[7:11] <- "Nationalist Development"
paramssolar1<- paramssolar1[, c("Scale", setdiff(names(paramssolar1), "Scale"))]
CFAsolar <- knitr::kable(paramssolar1, booktabs = TRUE, caption = "Confirmatory Factor Analysis(CFA) on newly developed eco-pol scale") %>%
  kable_styling(latex_options = c("striped","scale_down", "hold_position", "repeat_header"))%>%
  column_spec(column = 2, width = "4cm")

# Print the table
CFAsolar %>%
  landscape()

```

\newpage
## CFA on eco-pol scale for Coal

```{r, echo = FALSE}
# kahan scores for coal and CFA on ecopol vars  
cfascales <- codedmodule2 %>%
  dplyr::select(starts_with(c("K_","Risky_Coal","WEALTHLIM","MECHANISATION","DECISIONDECEN","DECISIONCEN","INDUSTRYLARGE","INDUSTRYSMALL","ECONOMYGLOBAL","ECONOMYLOCAL", "ENVOVERDEV","DEVOVERENV","OWNERPVT","OWNERNOREG","OWNERPUB", "OWNERREG","DISPLACECOAL",  "POLLUTECOAL", "HEALTHCOAL", "JOBSCOAL", "BEAUTYCOAL", "PRIDECOAL", "NPRIDECOAL", "DEVCOAL", "PROSPERCOAL", "RELYCOAL", "Uppercaste", "Male", "Hindu", "Urban", "State","age"))) %>%
  na.omit()

cKahan_fas <- '
  KahanS =~ K_SHARM + K_SLIMCHOI + K_SPROTECT 
  KahanH =~ K_HEQUAL+ K_ERADEQ1 + K_EWEALTH + K_ERADEQ2
'

# Run CFA
cKahan_cfa <- cfa(cKahan_fas, data = cfascales, std.lv = TRUE)

cKahan_scores <- predict(cKahan_cfa, newdata = cfascales)

cKahan_Kscores <- cbind(cfascales, cKahan_scores)

cecopol_fas <- '
  Pdevelop =~ HEALTHCOAL + BEAUTYCOAL + DISPLACECOAL + POLLUTECOAL + INDUSTRYSMALL + OWNERREG + MECHANISATION
  Ndevelop =~ DEVCOAL + PROSPERCOAL + JOBSCOAL + PRIDECOAL + NPRIDECOAL 
'

cecopol_cfa <- cfa(cecopol_fas, data = cfascales, std.lv = TRUE)
#summary(cecopol_cfa, fit.measure= TRUE, standardized = TRUE)

# 3) lm with kahan 
cm1_kahan <- lm(Risky_Coal ~ Uppercaste + Male + Hindu + Urban + age + State + KahanS + KahanH , data = cKahan_Kscores)
# Print model summary
#summary(cm1_kahan)
#plot(cm1_kahan)

# Cronbach's alpha test on Pdevelop and Ndevelop for Solar energy

coalpdevelop <- cfascales %>%
  dplyr::select(HEALTHCOAL, BEAUTYCOAL, DISPLACECOAL, POLLUTECOAL, INDUSTRYSMALL, OWNERREG, MECHANISATION)%>%
  na.omit()

coalpdevelop.alpha <- psych::alpha(coalpdevelop, check.keys = TRUE)

# alpha is strong at 0.78!

coalndevelop <- cfascales %>%
  dplyr::select(DEVCOAL, PRIDECOAL, NPRIDECOAL, PROSPERCOAL, JOBSCOAL,)%>%
  na.omit()

coalndevelop.alpha <- psych::alpha(coalndevelop, check.keys = TRUE)

#alpha

```

```{r, echo = FALSE}

# Extract fit measures
fitcecopol <- fitMeasures(cecopol_cfa, fit.measures = c("CFI", "TLI", "RMSEA", "rmsea.ci.upper", "rmsea.ci.lower"))
fitcecopol2 <- fitMeasures(cecopol_cfa)
# Convert to a data frame for presentation
fitcecopoldf <- data.frame(Measure = names(fitcecopol), Value = fitcecopol) %>%
  mutate(Measure = case_when(
    Measure == "cfi" ~ "Comparative Fit Index (CFI) ",
    Measure == "tli" ~ "Tucker-Lewis Index (TLI)",
    Measure == "rmsea" ~ "Root Mean Square Error of Approximation(RMSEA)",
    Measure == "rmsea.ci.upper" ~ "RMSEA 90 Percent confidence interval - lower",
    Measure == "rmsea.ci.lower" ~ "RMSEA 90 Percent confidence interval - upper",
    TRUE ~ as.character(Measure)
  )) %>%
  mutate(Value = round(Value, 3)) %>%
  remove_rownames()

#RMSEA - between 0.05 and 0.08 (reasonable fit), less than 0.05 close fit and greater than 0.10 is poor fit 
#CFI - greater than 0.90 or 0.95 indicates good fit
#TFI - CFI>TFI >0.90 - indicates good fit

```

```{r, echo =FALSE}
# Display fit measures using kable
#knitr::kable(fitcecopoldf, caption = "Fit Measures: CFA on eco-pol scale for Coal", booktabs = TRUE) %>%
  #kable_styling(latex_options = c("striped", "hold_position"))

```

```{r, echo=FALSE}

# Extract parameter estimates
paramscoal <- parameterEstimates(cecopol_cfa, standardized = TRUE)%>%
   slice(1:(n()-12))

paramscoal1 <- paramscoal %>%
  mutate(rhs = case_when(
    rhs == "HEALTHCOAL" ~ "Coal powered plants poses a great risk to the health of people living around it.",
    rhs == "BEAUTYCOAL" ~ "Coal powered plants spoils the natural beauty of the landscape.",
    rhs == "DISPLACECOAL" ~ "Coal powered plants is leading to displacement of people from their land.",
    rhs == "POLLUTECOAL" ~ "Coal powered plants increases pollution of air/water/land.",
    rhs == "INDUSTRYSMALL" ~ "Large corporations are destroying the local industries in India and benefiting only a handful of people.",
    rhs == "OWNERREG" ~ "Regardless of ownership, the government should pass strong regulations and implement them.",
    rhs == "DEVCOAL" ~ "Coal powered plants pushes forward the country's development.",
    rhs == "PROSPERCOAL" ~ "Coal powered plants brings economic prosperity to the surrounding regions.",
    rhs == "JOBSCOAL" ~ "Coal powered plants will bring jobs to the local community.",
    TRUE ~ as.character(rhs)
  )) %>%
  dplyr::select(!c(lhs,op, std.nox)) %>%
    rename(
    "Items" = rhs,
    "Loadings" = est,
    "Standard Error" = se,
    "zvalue" = z) %>%
  mutate(across(c("Loadings", "Standard Error", "zvalue"), round, 3))

paramscoal1$Scale <- NA
paramscoal1$Scale[1:6] <- "People Centered Development"
paramscoal1$Scale[7:9] <- "Nationalist Development"
paramscoal1<- paramscoal1[, c("Scale", setdiff(names(paramscoal1), "Scale"))]
CFAcoal <- knitr::kable(paramscoal1, booktabs = TRUE, caption = "Confirmatory Factor Analysis(CFA) on newly developed eco-pol scale") %>%
  kable_styling(latex_options = c("striped","scale_down", "hold_position", "repeat_header"))%>%
  column_spec(column = 2, width = "4cm")

# Print the table
CFAcoal %>%
  landscape()

```