### 2024-03-10 withmice.Rmd

---
title: "models with mice"
output:
  pdf_document:
    toc: yes
    toc_depth: 2
    keep_tex: yes
  header-includes:
  - \usepackage{dcolumn}
  - \usepackage{array}
  - "\\usepackage{placeins}"
  keep_md: yes
---

```{r load packages, warning = FALSE, message = FALSE, echo=FALSE}
  knitr::opts_chunk$set(warning = FALSE, message = FALSE)

suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(readxl))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(tidyr))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(lavaan))
suppressPackageStartupMessages(library(tsibble))
suppressPackageStartupMessages(library(DT))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(psych)) 
suppressPackageStartupMessages(library(foreign))
suppressPackageStartupMessages(library(MASS))
suppressPackageStartupMessages(library(Hmisc))
suppressPackageStartupMessages(library(reshape2))
suppressPackageStartupMessages(library(naniar))
suppressPackageStartupMessages(library(mice))

```

```{r setup, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

moduletwo <- read_excel(here::here("thesis_surveydata/political_factors_complete.xlsx"))
module2 <- data.frame(moduletwo)
module2[module2 == "Rather not say/ Don't know"] <- "Rather not say/Don't know"

# loaded data set for module 2 eco -pol variables and removes the rows with question number and question text
```

```{r, eval =TRUE, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
#This chunk contains the coding schemes of various scales used in survey one: eco-political factors, kahan scale and acceptance scale
codedmodule2 <- module2 %>%

#remove row 1
  filter(!row_number() %in% c(1,2)) %>%

# replace risky likert scale with numbers
  mutate_at(vars(starts_with("Risky")), funs(case_when(. =="Not at all risky" ~ 1, 
                                                       . =="Slightly risky" ~ 2, 
                                                       . =="Moderately risky" ~ 3, 
                                                       . =="Very risky" ~ 4, 
                                                       . =="Extremely risky" ~ 5))) %>%

# replace beneficial likert scale with numbers  
  mutate_at(vars(starts_with("Ben")), funs(case_when(. =="Not at all beneficial" ~ 1,
                                                     . =="Slightly beneficial" ~ 2,
                                                     . =="Moderately beneficial" ~ 3,
                                                     . =="Very beneficial" ~ 4,
                                                     . =="Extremely beneficial" ~ 5 ))) %>%

# replace nuclear acceptance likert scale with numbers
  mutate_at(vars(N_accept,N_reluctantlyaccept,N_reject), funs(case_when(. == "Strongly disagree" ~ 1, 
                                                                        . == "Somewhat disagree" ~ 2,
                                                                        . == "Neither agree nor disagree" ~3,
                                                                        . == "Somewhat agree" ~ 4,
                                                                        . == "Strongly agree" ~ 5))) %>%

# code likert scale for variables for Kahan scale into numbers
  mutate_at(vars(starts_with (c("K_S","K_E","DISPLACE", "POLLUTE", "HEALTH", "JOBS", "BEAUTY", "PRIDE", "NPRIDE","DEV","PROSPER", "RELY"))), funs(case_when(. == "Strongly disagree" ~ 1, 
               . == "Somewhat disagree" ~ 2,
               . == "Neither agree nor disagree" ~3,
               . == "Somewhat agree" ~ 4,
               . == "Strongly agree" ~ 5))) %>%

# NO reverse code for likert scale for variables for Kahan scale into numbers
  mutate_at(vars(starts_with (c("K_H","K_I"))), funs(case_when(. == "Strongly disagree" ~ 1, 
                                                               . == "Somewhat disagree" ~ 2,
                                                               . == "Neither agree nor disagree" ~3,
                                                               . == "Somewhat agree" ~ 4,
                                                               . == "Strongly agree" ~ 5))) %>%

# code eco-pol scale variables into numbers
mutate_at(vars(SYSTEMDEMO,SYSTEMRELIGION,SYSTEMTECHNO,SYSTEMTOTAL,WEALTHLIM,MECHANISATION,DECISIONDECEN,INDUSTRYSMALL,ECONOMYLOCAL,ENVOVERDEV,OWNERPUB, OWNERREG), funs(case_when(. == "Strongly disagree" ~ 1, 
                        . == "Somewhat disagree" ~ 2,
                        . == "Neither agree nor disagree" ~3,
                        . == "Somewhat agree" ~ 4,
                        . == "Strongly agree" ~ 5))) %>%

# NO reverse code eco-pol scale variables into numbers 
 mutate_at(vars(DECISIONCEN,INDUSTRYLARGE,ECONOMYGLOBAL,OWNERPVT,OWNERNOREG), funs(case_when(. == "Strongly disagree" ~ 1, 
                                                                                                         . == "Somewhat disagree" ~ 2,
                                                                                                         . == "Neither agree nor disagree" ~3,
                                                                                                         . == "Somewhat agree" ~ 4,
                                                                                                         . == "Strongly agree" ~ 5))) %>%
  mutate_all(~ifelse(. == "Rather not say/ Don't know", "Rather not say/Don't know", .))%>%
  mutate_all(~ifelse(. == "West bengal", "West Bengal", .))%>%
  mutate_all(~ifelse(. == "WEST BENGAL", "West Bengal", .))%>%
  mutate_all(~ifelse(. == "Up", "Uttar Pradesh", .))%>%
  mutate_all(~ifelse(. == "Tamilnadu", "Tamil Nadu", .))%>%
  mutate_all(~ifelse(. == "UP", "Uttar Pradesh", .))

```

## Can do something with accept, reject and reluctantly accept.

```{r}

## Can do something with accept, reject and reluctantly accept.

n_acceptreject<- codedmodule2 %>%
                dplyr::select("N_accept", "N_reluctantlyaccept", "N_reject")

summary(n_acceptreject)
```

# missing data pattern and MCAR

```{r, echo=FALSE}

formice1 <- codedmodule2[, grep("State|age|urban_rural|gender|caste|religion|Solar|SOLAR|Nuclear|NUCLEAR|Coal|COAL|K_S|K_E|K_H|K_I|WEALTHLIM|MECHANISATION|DECISIONDECEN|INDUSTRYSMALL|ECONOMYLOCAL|ENVOVERDEV|OWNERPUB|OWNERREG|DECISIONCEN|INDUSTRYLARGE|ECONOMYGLOBAL|OWNERPVT|OWNERNOREG", names(codedmodule2))]%>%
        dplyr::select(-c("Language", "Risky_INDSolar", "Ben_INDSolar", "religiontext")) %>%
          mutate(across(age, ~factor(.x, ordered = TRUE)))%>%  
          mutate(across(urban_rural:religion, ~factor(.x, ordered = FALSE)))%>%
          mutate(across(State, ~factor(.x, ordered = FALSE))) %>%
          mutate(Uppercaste = ifelse(caste %in% c('Brahmin', 'General'), 1, 0)) %>%
          mutate(Male = ifelse(gender == 'Male', 1, 0))%>%
          mutate(Hindu = ifelse(religion == 'Hinduism', 1, 0))%>%
           mutate(Age = case_when(
                                  age == "18-24 years old" ~ 1,
                                  age == "25-34 years old" ~ 2,
                                  age == "35-44 years old" ~ 3,
                                  age == "45-54 years old" ~ 4,
                                  age == "55-64 years old" ~ 5,
                                  age == "65-74 years old" ~ 6,
                                  age == "75 years or older" ~ 7))

#dplyr:: select(starts_with (c("State", "age", "urban_rural", "gender", "caste", "religion","Risky_Solar", "Risky_Nuclear", "Risky_Coal", "Ben_Solar","Ben_Nuclear","Ben_Coal","K_S","K_E","K_H","K_I", "DISPLACE", "POLLUTE", "HEALTH", "JOBS", "BEAUTY", "PRIDE", "NPRIDE","DEV", "PROSPER", "RELY", "WEALTHLIM", "MECHANISATION", "DECISIONDECEN", "INDUSTRYSMALL", "ECONOMYLOCAL", "ENVOVERDEV", "OWNERPUB", "OWNERREG", "DECISIONCEN","INDUSTRYLARGE","ECONOMYGLOBAL","OWNERPVT","OWNERNOREG" )))

library(dplyr)

formice <- formice1 %>%
  mutate(across(Risky_Solar:age, ~factor(.x, ordered = TRUE)))%>%
  mutate(across(urban_rural:religion, ~factor(.x, ordered = FALSE)))%>%
  mutate(across(State, ~factor(.x, ordered = FALSE)))

#str(formice)
#summary(formice)

# litte's MCAR test

mcar_test(data=formice)

# p value = 1, which means MAR condition is met.

```

# mice on entire dataset

```{r, echo=FALSE}

#str(formice1)

#formice1.imp <- mice(formice1, m =5, maxit =5,  seed = 430, print = TRUE)

#formice1.imp$chainMean

#plot(formice1.imp)

```

# Diagnostic checking
```{r, echo= FALSE}

formice1.imp$imp$Risky_Nuclear

complete(formice1.imp)

library(mice)

# Select the variables you want to plot
selected_vars <- c("Risky_Nuclear", "Risky_Solar", "Risky_Coal")

# Use stripplot to plot only the selected variables

stripplot(formice1.imp, scales = list(x = list(relation = "free")),layout = c(3, 1), pch = 20, cex = 1.2)

```

# Density plot - entire dataset 
```{r, echo=FALSE}

densityplot(formice1.imp, scales = list(x = list(relation = "free")), layout = c(5, 1))

```

# propensity score check imputations

```{r}

formice1.imp
#hc.na <- is.na(boys$hc)

#fit.hc <- with(imp.kendall, glm(hc.na ~ age + wgt + hgt + reg,
#family = binomial))
#ps <- rep(rowMeans(sapply(fit.hc$analyses, fitted.values)), 6)
```

# Checking MAR condition and missing data pattern for the Kahan Scale.
```{r, echo=FALSE}

kahanall <- codedmodule2 %>%
          dplyr::select(c(starts_with("K_"), "age", "gender", "caste", "State", "urban_rural", "religion")) %>%
          mutate(across(age:religion, ~factor(.x, ordered = FALSE)))

md.pattern(kahanall) #the variables with missing values: K_SHARM K_IPROTECT K_IPRIVACY K_SPROTECT K_SLIMCHOI K_IINTRFER K_HEQUAL K_HREVDIS1

mcar_test(data=kahanall) # p values is less than 0.05 - the condition for MCAR is not met

# dataset with missing variables to compare and test group comaprisons

MARkahan <- codedmodule2 %>%
          dplyr::select(c("K_SHARM", "K_IPROTECT", "K_IPRIVACY", "K_SPROTECT", "K_SLIMCHOI", "K_IINTRFER",  "K_HEQUAL", "K_HREVDIS1" , "age", "gender", "caste", "State", "urban_rural", "religion"))

MARkahan$K_SHARM_MAR <- MARkahan$K_SHARM

MARkahan$K_IPROTECT_MAR <- MARkahan$K_IPROTECT

MARkahan$K_SPROTECT_MAR <- MARkahan$K_SPROTECT

MARkahan$K_SLIMCHOI_MAR <- MARkahan$K_SLIMCHOI

library(finalfit)

# "K_SHARM" = "(C)Sometimes the government needs to make laws that keep people from hurting themselves.",
explanatory = c("age", "gender", "caste", "State", "urban_rural", "religion")
dependent = "K_SHARM_MAR"
MARkahan %>% 
  missing_compare(dependent, explanatory) %>% 
    knitr::kable(row.names=FALSE, align = c("l", "l", "r", "r", "r"), 
        caption = "comparisons between values of responders (Not missing) and 
        non-responders (Missing) on the K_SHARM variable.")

#   "K_IPROTECT" = "(I)It’s not the government’s business to try to protect people from themselves."
explanatory = c("age", "gender", "caste", "State", "urban_rural", "religion")
dependent = "K_IPROTECT_MAR"
MARkahan %>% 
  missing_compare(dependent, explanatory) %>% 
    knitr::kable(row.names=FALSE, align = c("l", "l", "r", "r", "r"), 
        caption = "comparisons between values of responders (Not missing) and 
        non-responders (Missing) on the K_IPROTECT variable.")

#  "K_SPROTECT" = "(C)The government should do more to advance society’s goals, even if that means limiting the freedom and choices of individuals.",
explanatory = c("age", "gender", "caste", "State", "urban_rural", "religion")
dependent = "K_SPROTECT_MAR"
MARkahan %>% 
  missing_compare(dependent, explanatory) %>% 
    knitr::kable(row.names=FALSE, align = c("l", "l", "r", "r", "r"), 
        caption = "comparisons between values of responders (Not missing) and 
        non-responders (Missing) on the K_SPROTECT variable.")

#  "K_SLIMCHOI" = "(C)The government should put limits on the choices individuals can make so they don’t get in the way of what’s good for society.",
explanatory = c("age", "gender", "caste", "State", "urban_rural", "religion")
dependent = "K_SLIMCHOI_MAR"
MARkahan %>% 
  missing_compare(dependent, explanatory) %>% 
    knitr::kable(row.names=FALSE, align = c("l", "l", "r", "r", "r"), 
        caption = "comparisons between values of responders (Not missing) and 
        non-responders (Missing) on the K_SPROTECT variable.")

```

## Mifa for Kahan scale 
```{r, echo =FALSE}

library(mifa)
library(mice)
library(psych)

 #cov_vars = -c(State, Risky_Solar, Risky_Nuclear, Risky_Coal, Ben_Solar, Ben_Nuclear, Ben_Coal, DECISIONDECEN, DECISIONCEN, INDUSTRYSMALL, INDUSTRYLARGE, ECONOMYLOCAL, ECONOMYGLOBAL, ENVOVERDEV, OWNERPVT, OWNERPUB, OWNERREG, OWNERNOREG, WEALTHLIM, MECHANISATION, DISPLACESOLAR, POLLUTESOLAR, HEALTHSOLAR, JOBSSOLAR, BEAUTYSOLAR, PRIDESOLAR, NPRIDESOLAR, DEVSOLAR, PROSPERSOLAR, RELYSOLAR, DISPLACECOAL, POLLUTECOAL, HEALTHCOAL, JOBSCOAL, BEAUTYCOAL, PRIDECOAL, NPRIDECOAL, DEVCOAL, PROSPERCOAL, RELYCOAL, DISPLACENUCLEAR, POLLUTENUCLEAR, HEALTHNUCLEAR, JOBSNUCLEAR, BEAUTYNUCLEAR, PRIDENUCLEAR, NPRIDENUCLEAR, DEVNUCLEAR, PROSPERNUCLEAR, RELYNUCLEAR, age, urban_rural, gender, caste, religion),

mikahan <- mifa(data= kahanall,
                cov_vars  = -c("age", "gender", "caste", "State", "urban_rural", "religion"),
                m = 5,
                n_pc= 2:4,
                ci = "fieller", 
                n_boot=1000,
                print = FALSE,
                progress= TRUE, 
                seed = 427,)

mikahan

library(psych)

fit <- psych::fa(mikahan$cov_combined, nfactors = 2, rotate = "varimax")

print.psych(fit, sort= TRUE, cut= 0.4)

```

# checking merging of imputations  : kahan scale

```{r}

kahan_imp <- mice(kahanall, m =5, maxit =30,  seed = 427, print = TRUE)

kahan_imp$chainMean

data_imp <- mice::complete(mice::mice(kahanall, 1, print = FALSE))

fct_scores <- data.frame(factor.scores(data_imp[, 1:12], fit)$scores)

plot(kahan_imp)

```

#checking imputations: Kahan scale - are values plausible ?

yes all value are whole numbers between 1 and 5

```{r, echo=FALSE}

orgscores <- data.frame(kahanefascores)

summary(mikahan)

str(data_imp)
names(mikahan)

mikahan$mids

complete_data_first <- complete(mikahan$mids, 1)

summary(complete_data_first)

complete_data_2 <- complete(mikahan$mids, 2)

summary(complete_data_2)

complete_data_3 <- complete(mikahan$mids, 3)

summary(complete_data_3)

complete_data_4 <- complete(mikahan$mids, 4)

summary(complete_data_4)

complete_data_5 <- complete(mikahan$mids, 5)

summary(complete_data_5)
```

## Distribution check: kahan Scale

```{r, echo = False }

densityplot(kahan_imp, scales = list(x = list(relation = "free")), layout = c(5, 1))

```

```{r, echo=FALSE}
# Add an identifier column to each dataset
fct_scores$Dataset <- 'impscores'
orgscores$Dataset <- 'orgscores'

# Combine the datasets for plotting
combined_data <- rbind(fct_scores, orgscores)

# Select only necessary columns (assuming 'Index' isn't needed anymore)
combined_data <- combined_data[, c("MR1", "MR2", "Dataset")]

ggplot(combined_data, aes(x = MR1, color = Dataset)) +
  geom_density(fill = NA, alpha = 0.5) +
  labs(title = "Density Plot of Hierachy- Egalitarianism for Each Dataset", x = "HE Value", y = "Density") +
  scale_color_manual(values = c("impscores" = "blue", "orgscores" = "red")) +
  theme_minimal()

```

```{r}
ggplot(combined_data, aes(x = MR2, color = Dataset)) +
  geom_density(fill = NA, alpha = 0.5) +
  labs(title = "Density Plot of Communitarianism for Each Dataset", x = "Communitarianism Value", y = "Density") +
  scale_color_manual(values = c("impscores" = "blue", "orgscores" = "red")) +
  theme_minimal()

```

# ecopol scale - checking MAR

```{r}

ecopoln <- formice1 %>%
  dplyr::select(
    DECISIONDECEN, DECISIONCEN, INDUSTRYSMALL, INDUSTRYLARGE,
    ECONOMYLOCAL, ECONOMYGLOBAL, ENVOVERDEV, OWNERPVT, OWNERPUB,
    OWNERREG, OWNERNOREG, WEALTHLIM, MECHANISATION,
    DISPLACENUCLEAR,  POLLUTENUCLEAR, HEALTHNUCLEAR, JOBSNUCLEAR, BEAUTYNUCLEAR,
    PRIDENUCLEAR, NPRIDENUCLEAR, DEVNUCLEAR, PROSPERNUCLEAR,
    RELYNUCLEAR,age, urban_rural, gender, caste, religion, State
  )

md.pattern(ecopoln) #the variables with missing values: PRIDENUCLEAR PROSPERNUCLEAR JOBSNUCLEAR HEALTHNUCLEAR BEAUTYNUCLEAR POLLUTENUCLEAR DISPLACENUCLEAR RELYNUCLEAR

mcar_test(data=ecopoln) # p values is less than 0.76 - the condition for MCAR is met

# dataset with missing variables to compare and test group comaprisons

MARecopol <- codedmodule2 %>%
          dplyr::select(c(DECISIONDECEN, DECISIONCEN, INDUSTRYSMALL, INDUSTRYLARGE,
    ECONOMYLOCAL, ECONOMYGLOBAL, ENVOVERDEV, OWNERPVT, OWNERPUB,
    OWNERREG, OWNERNOREG, WEALTHLIM, MECHANISATION,
    DISPLACENUCLEAR,  POLLUTENUCLEAR, HEALTHNUCLEAR, JOBSNUCLEAR, BEAUTYNUCLEAR,
    PRIDENUCLEAR, NPRIDENUCLEAR, DEVNUCLEAR, PROSPERNUCLEAR,
    RELYNUCLEAR,age, urban_rural, gender, caste, religion, State))

MARecopol$JOBSNUCLEAR_MAR <- MARecopol$JOBSNUCLEAR

MARecopol$PRIDENUCLEAR_MAR <- MARecopol$PRIDENUCLEAR

library(finalfit)

# JOBSNUCLEAR
explanatory = c("age", "gender", "caste", "State", "urban_rural", "religion")
dependent = "JOBSNUCLEAR_MAR"
MARecopol %>% 
  missing_compare(dependent, explanatory) %>% 
    knitr::kable(row.names=FALSE, align = c("l", "l", "r", "r", "r"), 
        caption = "comparisons between values of responders (Not missing) and 
        non-responders (Missing) on the JOBSNUCLEAR variable.")

# PRIDENUCLEAR
explanatory = c("age", "gender", "caste", "State", "urban_rural", "religion")
dependent = "PRIDENUCLEAR_MAR"
MARecopol %>% 
  missing_compare(dependent, explanatory) %>% 
    knitr::kable(row.names=FALSE, align = c("l", "l", "r", "r", "r"), 
        caption = "comparisons between values of responders (Not missing) and 
        non-responders (Missing) on the PRIDENUCLEAR variable.")

```

# Mifa for new eco-pol value scale : Nuclear energy

```{r}

#ecopoln <- codedmodule2[, grep("NUCLEAR|WEALTHLIM|MECHANISATION|DECISIONDECEN|INDUSTRYSMALL|ECONOMYLOCAL|ENVOVERDEV|OWNERPUB|OWNERREG|DECISIONCEN|INDUSTRYLARGE|ECONOMYGLOBAL|OWNERPVT|OWNERNOREG|", names(codedmodule2))]

summary(formice1)

miecopol <- mifa(data= ecopoln,
                cov_vars  = -c("age", "gender", "caste", "State", "urban_rural", "religion"),
                n_pc= 2:4,
                ci = "fieller", 
                n_boot=1000,
                print = TRUE,
                progress= TRUE, 
                seed = 429,)

miecopol

fit2 <- psych::fa(miecopol$cov_combined, nfactors = 2, rotate = "varimax")

print.psych(fit2, sort= TRUE, cut= 0.4)

data_imp2 <- mice::complete(mice::mice(ecopoln, 1, print = FALSE))

fct_scores2 <- data.frame(factor.scores(data_imp2[, 1:23], fit2)$scores)

```

## Checking imputations : Eco-pol scale

imputations are pluasible- all natural numbers between 1 and 5

```{r, echo=FALSE}

summary(miecopol)
names(miecopol)

miecopol$mids

complete_ecopol_first <- complete(miecopol$mids, 1)

summary(complete_ecopol_first)

complete_ecopol_2 <- complete(miecopol$mids, 2)

summary(complete_ecopol_2)

complete_ecopol_3 <- complete(miecopol$mids, 3)

summary(complete_ecopol_3)

complete_ecopol_4 <- complete(miecopol$mids, 4)

summary(complete_ecopol_4)

complete_ecopol_5 <- complete(miecopol$mids, 5)

summary(complete_ecopol_5)

```

# Density plots - 2 factors from eco-pol scale 
```{r}

nscores <- data.frame(ecopolall_scores$scores)

# Add an identifier column to each dataset
fct_scores2$Dataset <- 'imputed'
nscores$Dataset <- 'original'

# Combine the datasets for plotting
compare2 <- rbind(fct_scores2, nscores)

# Select only necessary columns (assuming 'Index' isn't needed anymore)
compare2<- compare2[, c("MR1", "MR2", "Dataset")]

```

```{r}

library(ggplot2)

ggplot(compare2, aes(x = MR2, color = Dataset)) +
  geom_density(fill = NA, alpha = 0.5) +
  labs(title = "Density Plot of MR2 for Each Dataset", x = "MR2 Value", y = "Density") +
  scale_color_manual(values = c("imputed" = "blue", "original" = "red")) +
  theme_minimal()

```

# Binding imputed values to formice dataset

```{r, echo = FALSE}

# binding the imputed  factor analysis scores for kahan scale and eco-pol value scale to the larger dataset codedmodule2

ecopolnuclear <- fct_scores2 %>%
  rename(Ndevelop = MR1, Pdevelop = MR2)

kahannuclear <- fct_scores%>%
  rename(Hierarchy_Egalitarian = MR1, Communitarian = MR2)

formice1

allformice<- cbind(formice1,ecopolnuclear, kahannuclear)

#summary(formice)

#method_list <- rep("polr", ncol(formice))
#names(method_list) <- colnames(formice)

#skip_vars <- c("age",  "urban_rural", "gender",  "caste", "religion", "K_ERADEQ2","K_ERADEQ1", "INDUSTRYSMALL","INDUSTRYLARGE","WEALTHLIM", "MECHANISATION", "State","K_IINTRFER",  "K_IPRIVACY",  "K_SHARM",  "K_IPROTECT",  "K_SLIMCHOI",  "K_SPROTECT",  "K_HEQUAL",  "K_HREVDIS1",  "K_EDISCRIM",    "K_EWEALTH",  "DECISIONDECEN",  "DECISIONCEN",     "ECONOMYLOCAL",  "ECONOMYGLOBAL",  "ENVOVERDEV",  "OWNERPVT",  "OWNERPUB",  "OWNERREG",  "DISPLACESOLAR",  "POLLUTESOLAR",  "HEALTHSOLAR",  "JOBSSOLAR",  "BEAUTYSOLAR",  "PRIDESOLAR",  "NPRIDESOLAR",  "DEVSOLAR",  "PROSPERSOLAR",  "RELYSOLAR",  "DISPLACECOAL",  "POLLUTECOAL",  "HEALTHCOAL",  "JOBSCOAL",  "BEAUTYCOAL",  "PRIDECOAL",  "NPRIDECOAL",  "DEVCOAL",  "PROSPERCOAL",  "RELYCOAL",  "DISPLACENUCLEAR",  "POLLUTENUCLEAR",  "HEALTHNUCLEAR",  "JOBSNUCLEAR",  "BEAUTYNUCLEAR",  "PRIDENUCLEAR",  "NPRIDENUCLEAR",  "DEVNUCLEAR",  "PROSPERNUCLEAR",  "RELYNUCLEAR", "OWNERNOREG")

#Set method to "" for variables to skip
#method_list[skip_vars] <- ""

#nuclear_imp1 <- mice(formice, method=method_list, maxit = 5, m =5,  seed = 123, print = TRUE)

print(nuclear_imp1)
summary(nuclear_imp1)
nuclear_imp1$imp$Risky_Nuclear

```

```{r, echo=FALSE}

str(allformice)
library(mice)
library(dplyr)

# Initialize the mice process to create the predictor matrix
ini <- mice(allformice, maxit = 0, pri = FALSE)

# Extract the predictor matrix from the initialized object
pred <- ini$predictorMatrix

# Set the columns for "Uppercaste", "Male", "Hindu", and "Age" to 0
pred[, "Uppercaste"] <- 0
pred[, "Male"] <- 0
pred[, "Hindu"] <- 0
pred[, "Age"] <- 0
pred[, "Ndevelop"] <- 0
pred[, "Pdevelop"] <- 0
pred[, "Hierarchy_Egalitarian"] <- 0
pred[, "Communitarian"] <- 0
# Identify columns that start with "K_" and set them to 0
k_columns <- grep("^K_", colnames(formice1))
pred[, k_columns] <- 0

# Perform the mice imputation with the updated predictor matrix
allformice.imp <- mice(allformice, m = 5, maxit = 10, predictorMatrix = pred, seed = 430, print = TRUE)

summary(allformice.imp)
allformice.imp$chainMean

plot(allformice.imp)

```

```{r}

nuclearfit1 <- with(allformice.imp, lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + urban_rural + Age))

nuclearfit2 <- with(allformice.imp, lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + urban_rural + Age + State))

nuclearfit3 <- with(allformice.imp, lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + urban_rural + Age + State+ Communitarian + Hierarchy_Egalitarian))

nuclearfit4 <- with(allformice.imp, lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + urban_rural + Age + State+ Communitarian + Hierarchy_Egalitarian + Pdevelop + Ndevelop))

summary(pool(nuclearfit1))
summary(pool(nuclearfit2))
summary(pool(nuclearfit3))
summary(pool(nuclearfit4))

```

# LM with demographic variables

```{r}

# The pool() function from the mice package creates a list containing the results,
# so we need to access the summary of the model directly.

pooled_fit <- pool(nuclearfit1)

pooled_summary <- summary(pooled_fit)

coefficients_df <- pooled_summary
# To ensure the numeric values are rounded to two decimal places, you can use:
coefficients_df <- round(coefficients_df[, -1], 2)  # Skipping the first column as it's factors (term names)
coefficients_df$term <- pooled_summary$term  # Adding back the unrounded terms

# Print the dataframe
data.frame(coefficients_df) %>%
  dplyr::select(term, estimate, std.error, statistic, df, p.value)
```

# LM with demographic variables and State

```{r}

# The pool() function from the mice package creates a list containing the results,
# so we need to access the summary of the model directly.

pooled_fit2 <- pool(nuclearfit2)

pooled_summary2 <- summary(pooled_fit2)

coefficients_df2 <- pooled_summary2
# To ensure the numeric values are rounded to two decimal places, you can use:
coefficients_df2 <- round(coefficients_df2[, -1], 2)  # Skipping the first column as it's factors (term names)
coefficients_df2$term <- pooled_summary2$term  # Adding back the unrounded terms

# Print the dataframe
data.frame(coefficients_df2) %>%
  dplyr::select(term, estimate, std.error, statistic, df, p.value)
```

# LM with demographic variables and State

```{r}
# The pool() function from the mice package creates a list containing the results,
# so we need to access the summary of the model directly.

pooled_fit3 <- pool(nuclearfit3)

pooled_summary3 <- summary(pooled_fit3)

coefficients_df3 <- pooled_summary3
# To ensure the numeric values are rounded to two decimal places, you can use:
coefficients_df3 <- round(coefficients_df3[, -1], 2)  # Skipping the first column as it's factors (term names)
coefficients_df3$term <- pooled_summary3$term  # Adding back the unrounded terms

# Print the dataframe
data.frame(coefficients_df3) %>%
  dplyr::select(term, estimate, std.error, statistic, df, p.value)
```

# LM with demographic variables and State

```{r}

# The pool() function from the mice package creates a list containing the results,
# so we need to access the summary of the model directly.

pooled_fit4 <- pool(nuclearfit4)

pooled_summary4 <- summary(pooled_fit4)

coefficients_df4 <- pooled_summary4
# To ensure the numeric values are rounded to two decimal places, you can use:
coefficients_df4 <- round(coefficients_df4[, -1], 2)  # Skipping the first column as it's factors (term names)
coefficients_df4$term <- pooled_summary4$term  # Adding back the unrounded terms

# Print the dataframe
data.frame(coefficients_df4) %>%
  dplyr::select(term, estimate, std.error, statistic, df, p.value)
```