### 2024-03-13 withmice.Rmd
making pretty Lm tables

---
title: "models with mice"
header-includes:
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
output:
  pdf_document: 
    toc: yes
    toc_depth: 2
    keep_tex: yes
  keep_md: yes
---

```{r load packages, warning = FALSE, message = FALSE, echo=FALSE}
  knitr::opts_chunk$set(warning = FALSE, message = FALSE)

suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(readxl))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(tidyr))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(lavaan))
suppressPackageStartupMessages(library(tsibble))
suppressPackageStartupMessages(library(DT))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(psych)) 
suppressPackageStartupMessages(library(foreign))
suppressPackageStartupMessages(library(MASS))
suppressPackageStartupMessages(library(Hmisc))
suppressPackageStartupMessages(library(reshape2))
suppressPackageStartupMessages(library(naniar))
suppressPackageStartupMessages(library(mice))

```

```{r setup, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

moduletwo <- read_excel(here::here("thesis_surveydata/political_factors_complete.xlsx"))
module2 <- data.frame(moduletwo)
module2[module2 == "Rather not say/ Don't know"] <- "Rather not say/Don't know"

# loaded data set for module 2 eco -pol variables and removes the rows with question number and question text
```

```{r, eval =TRUE, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
#This chunk contains the coding schemes of various scales used in survey one: eco-political factors, kahan scale and acceptance scale
codedmodule2 <- module2 %>%

#remove row 1
  filter(!row_number() %in% c(1,2)) %>%

# replace risky likert scale with numbers
  mutate_at(vars(starts_with("Risky")), funs(case_when(. =="Not at all risky" ~ 1, 
                                                       . =="Slightly risky" ~ 2, 
                                                       . =="Moderately risky" ~ 3, 
                                                       . =="Very risky" ~ 4, 
                                                       . =="Extremely risky" ~ 5))) %>%

# replace beneficial likert scale with numbers  
  mutate_at(vars(starts_with("Ben")), funs(case_when(. =="Not at all beneficial" ~ 1,
                                                     . =="Slightly beneficial" ~ 2,
                                                     . =="Moderately beneficial" ~ 3,
                                                     . =="Very beneficial" ~ 4,
                                                     . =="Extremely beneficial" ~ 5 ))) %>%

# replace nuclear acceptance likert scale with numbers
  mutate_at(vars(N_accept,N_reluctantlyaccept,N_reject), funs(case_when(. == "Strongly disagree" ~ 1, 
                                                                        . == "Somewhat disagree" ~ 2,
                                                                        . == "Neither agree nor disagree" ~3,
                                                                        . == "Somewhat agree" ~ 4,
                                                                        . == "Strongly agree" ~ 5))) %>%

# code likert scale for variables for Kahan scale into numbers
  mutate_at(vars(starts_with (c("K_S","K_E","DISPLACE", "POLLUTE", "HEALTH", "JOBS", "BEAUTY", "PRIDE", "NPRIDE","DEV","PROSPER", "RELY"))), funs(case_when(. == "Strongly disagree" ~ 1, 
               . == "Somewhat disagree" ~ 2,
               . == "Neither agree nor disagree" ~3,
               . == "Somewhat agree" ~ 4,
               . == "Strongly agree" ~ 5))) %>%

# NO reverse code for likert scale for variables for Kahan scale into numbers
  mutate_at(vars(starts_with (c("K_H","K_I"))), funs(case_when(. == "Strongly disagree" ~ 1, 
                                                               . == "Somewhat disagree" ~ 2,
                                                               . == "Neither agree nor disagree" ~3,
                                                               . == "Somewhat agree" ~ 4,
                                                               . == "Strongly agree" ~ 5))) %>%

# code eco-pol scale variables into numbers
mutate_at(vars(SYSTEMDEMO,SYSTEMRELIGION,SYSTEMTECHNO,SYSTEMTOTAL,WEALTHLIM,MECHANISATION,DECISIONDECEN,INDUSTRYSMALL,ECONOMYLOCAL,ENVOVERDEV,OWNERPUB, OWNERREG), funs(case_when(. == "Strongly disagree" ~ 1, 
                        . == "Somewhat disagree" ~ 2,
                        . == "Neither agree nor disagree" ~3,
                        . == "Somewhat agree" ~ 4,
                        . == "Strongly agree" ~ 5))) %>%

# NO reverse code eco-pol scale variables into numbers 
 mutate_at(vars(DECISIONCEN,INDUSTRYLARGE,ECONOMYGLOBAL,OWNERPVT,OWNERNOREG), funs(case_when(. == "Strongly disagree" ~ 1, 
                                                                                                         . == "Somewhat disagree" ~ 2,
                                                                                                         . == "Neither agree nor disagree" ~3,
                                                                                                         . == "Somewhat agree" ~ 4,
                                                                                                         . == "Strongly agree" ~ 5))) %>%
  mutate_all(~ifelse(. == "Rather not say/ Don't know", "Rather not say/Don't know", .))%>%
  mutate_all(~ifelse(. == "West bengal", "West Bengal", .))%>%
  mutate_all(~ifelse(. == "WEST BENGAL", "West Bengal", .))%>%
  mutate_all(~ifelse(. == "Up", "Uttar Pradesh", .))%>%
  mutate_all(~ifelse(. == "Tamilnadu", "Tamil Nadu", .))%>%
  mutate_all(~ifelse(. == "UP", "Uttar Pradesh", .))

```

# Can do something with accept, reject and reluctantly accept.

```{r, echo=FALSE}

## Can do something with accept, reject and reluctantly accept.

n_acceptreject<- codedmodule2 %>%
                dplyr::select("N_accept", "N_reluctantlyaccept", "N_reject")

summary(n_acceptreject)
```

# missing data pattern and MCAR

```{r, echo=FALSE}

formice1 <- codedmodule2[, grep("State|age|urban_rural|gender|caste|religion|Solar|SOLAR|Nuclear|NUCLEAR|Coal|COAL|K_S|K_E|K_H|K_I|WEALTHLIM|MECHANISATION|DECISIONDECEN|INDUSTRYSMALL|ECONOMYLOCAL|ENVOVERDEV|DEVOVERENV|OWNERPUB|OWNERREG|DECISIONCEN|INDUSTRYLARGE|ECONOMYGLOBAL|OWNERPVT|OWNERNOREG", names(codedmodule2))]%>%
        dplyr::select(-c("Language", "Risky_INDSolar", "Ben_INDSolar", "religiontext")) %>%
          mutate(across(age, ~factor(.x, ordered = TRUE)))%>%  
          mutate(across(urban_rural:religion, ~factor(.x, ordered = FALSE)))%>%
          mutate(across(State, ~factor(.x, ordered = FALSE))) %>%
          mutate(Uppercaste = ifelse(caste %in% c('Brahmin', 'General'), 1, 0)) %>%
          mutate(Male = ifelse(gender == 'Male', 1, 0))%>%
          mutate(Hindu = ifelse(religion == 'Hinduism', 1, 0))%>%
           mutate(Age = case_when(
                                  age == "18-24 years old" ~ 1,
                                  age == "25-34 years old" ~ 2,
                                  age == "35-44 years old" ~ 3,
                                  age == "45-54 years old" ~ 4,
                                  age == "55-64 years old" ~ 5,
                                  age == "65-74 years old" ~ 6,
                                  age == "75 years or older" ~ 7))

#dplyr:: select(starts_with (c("State", "age", "urban_rural", "gender", "caste", "religion","Risky_Solar", "Risky_Nuclear", "Risky_Coal", "Ben_Solar","Ben_Nuclear","Ben_Coal","K_S","K_E","K_H","K_I", "DISPLACE", "POLLUTE", "HEALTH", "JOBS", "BEAUTY", "PRIDE", "NPRIDE","DEV", "PROSPER", "RELY", "WEALTHLIM", "MECHANISATION", "DECISIONDECEN", "INDUSTRYSMALL", "ECONOMYLOCAL", "ENVOVERDEV", "OWNERPUB", "OWNERREG", "DECISIONCEN","INDUSTRYLARGE","ECONOMYGLOBAL","OWNERPVT","OWNERNOREG" )))

library(dplyr)

formice <- formice1 %>%
  mutate(across(Risky_Solar:age, ~factor(.x, ordered = TRUE)))%>%
  mutate(across(urban_rural:religion, ~factor(.x, ordered = FALSE)))%>%
  mutate(across(State, ~factor(.x, ordered = FALSE)))

#str(formice)
summary(formice)

# litte's MCAR test

mcar_test(data= codedmodule2)

# p value = 1, which means MAR condition is met.

```

# Kahan scale: MAR condition and missing data pattern 
```{r, echo=FALSE}

kahanall <- codedmodule2 %>%
          dplyr::select(c(starts_with("K_"), "age", "gender", "caste", "State", "urban_rural", "religion")) %>%
          mutate(across(gender:religion, ~factor(.x, ordered = FALSE))) %>%
          mutate(across(age, ~factor(.x, ordered = TRUE)))

str(kahanall)

md.pattern(kahanall) #the variables with missing values: K_SHARM K_IPROTECT K_IPRIVACY K_SPROTECT K_SLIMCHOI K_IINTRFER K_HEQUAL K_HREVDIS1

mcar_test(data=kahanall) # p values is less than 0.05 - the condition for MCAR is not met

# dataset with missing variables to compare and test group comaprisons

MARkahan <- codedmodule2 %>%
          dplyr::select(c("K_SHARM", "K_IPROTECT", "K_IPRIVACY", "K_SPROTECT", "K_SLIMCHOI", "K_IINTRFER",  "K_HEQUAL", "K_HREVDIS1" , "age", "gender", "caste", "State", "urban_rural", "religion"))

MARkahan$K_SHARM_MAR <- MARkahan$K_SHARM

MARkahan$K_IPROTECT_MAR <- MARkahan$K_IPROTECT

MARkahan$K_SPROTECT_MAR <- MARkahan$K_SPROTECT

MARkahan$K_SLIMCHOI_MAR <- MARkahan$K_SLIMCHOI

library(finalfit)

# "K_SHARM" = "(C)Sometimes the government needs to make laws that keep people from hurting themselves.",
explanatory = c("age", "gender", "caste", "State", "urban_rural", "religion")
dependent = "K_SHARM_MAR"
MARkahan %>% 
  missing_compare(dependent, explanatory) %>% 
    knitr::kable(row.names=FALSE, align = c("l", "l", "r", "r", "r"), 
        caption = "comparisons between values of responders (Not missing) and 
        non-responders (Missing) on the K_SHARM variable.")

#   "K_IPROTECT" = "(I)It’s not the government’s business to try to protect people from themselves."
explanatory = c("age", "gender", "caste", "State", "urban_rural", "religion")
dependent = "K_IPROTECT_MAR"
MARkahan %>% 
  missing_compare(dependent, explanatory) %>% 
    knitr::kable(row.names=FALSE, align = c("l", "l", "r", "r", "r"), 
        caption = "comparisons between values of responders (Not missing) and 
        non-responders (Missing) on the K_IPROTECT variable.")

#  "K_SPROTECT" = "(C)The government should do more to advance society’s goals, even if that means limiting the freedom and choices of individuals.",
explanatory = c("age", "gender", "caste", "State", "urban_rural", "religion")
dependent = "K_SPROTECT_MAR"
MARkahan %>% 
  missing_compare(dependent, explanatory) %>% 
    knitr::kable(row.names=FALSE, align = c("l", "l", "r", "r", "r"), 
        caption = "comparisons between values of responders (Not missing) and 
        non-responders (Missing) on the K_SPROTECT variable.")

#  "K_SLIMCHOI" = "(C)The government should put limits on the choices individuals can make so they don’t get in the way of what’s good for society.",
explanatory = c("age", "gender", "caste", "State", "urban_rural", "religion")
dependent = "K_SLIMCHOI_MAR"
MARkahan %>% 
  missing_compare(dependent, explanatory) %>% 
    knitr::kable(row.names=FALSE, align = c("l", "l", "r", "r", "r"), 
        caption = "comparisons between values of responders (Not missing) and 
        non-responders (Missing) on the K_SPROTECT variable.")

```

## Mifa for Kahan scale 
```{r, echo =FALSE}

library(mifa)
library(mice)
library(psych)

mikahan <- mifa(data= kahanall,
                cov_vars  = -c("age", "gender", "caste", "State", "urban_rural", "religion"),
                m = 50,
                n_pc= 2:4,
                ci = "fieller", 
                n_boot=1000,
                print = FALSE,
                progress= TRUE, 
                seed = 427,)

mikahan

library(psych)

fit <- psych::fa(mikahan$cov_combined, nfactors = 2, rotate = "varimax")

print.psych(fit, sort= TRUE, cut= 0.4)

```

## checking merging of imputations  : kahan scale

```{r, echo=FALSE}

kahan_imp <- mice(kahanall, m =5, maxit =30,  seed = 427, print = FALSE)

kahan_imp$chainMean

```

#Kahan Factor scores: mean across all imputations

```{r, echo=FALSE}

kahanallscores <- mice::complete(mice::mice(kahanall, m = 5, seed = 427, print = FALSE), "broad")

kahanallscores <- kahanallscores %>%dplyr::select(-(contains(c("age", "urban_rural", "gender","caste","religion", "State"))))

str(kahanallscores)

nkahanall <- kahanallscores %>%
  rowwise() %>%
  mutate( K_EWEALTH_mean = mean(c_across(starts_with("K_EWEALTH"))),
    K_IINTRFER_mean = mean(c_across(starts_with("K_IINTRFER"))),
    K_IPRIVACY_mean = mean(c_across(starts_with("K_IPRIVACY"))),
    K_SHARM_mean = mean(c_across(starts_with("K_SHARM"))),
    K_IPROTECT_mean = mean(c_across(starts_with("K_IPROTECT"))),
    K_SLIMCHOI_mean = mean(c_across(starts_with("K_SLIMCHOI"))),
    K_SPROTECT_mean = mean(c_across(starts_with("K_SPROTECT"))),
    K_HEQUAL_mean = mean(c_across(starts_with("K_HEQUAL"))),
    K_HREVDIS1_mean = mean(c_across(starts_with("K_HREVDIS1"))),
    K_EDISCRIM_mean = mean(c_across(starts_with("K_EDISCRIM"))),
    K_ERADEQ1_mean = mean(c_across(starts_with("K_ERADEQ1"))),
    K_ERADEQ2_mean = mean(c_across(starts_with("K_ERADEQ2")))
  ) %>%
  ungroup()%>%
  dplyr::select(contains("_mean")) %>%
   rename_with(~ gsub("_mean$", "", .x))

nkahanfinal <- data.frame(factor.scores(nkahanall[, 1:12], fit)$scores)

```

## Kahan: factor scores from 1st imputed dataset
```{r, echo=FALSE}

# extracting factor scores from mifa for 1st imputed dataset

data_imp <- mice::complete(mice::mice(kahanall, 1, print = FALSE))

fct_scores <- data.frame(factor.scores(data_imp[, 1:12], fit)$scores)

plot(kahan_imp)

```

## Kahan: factor scores from second imputed dataset
```{r, echo=FALSE}

data_impk2 <- mice::complete(mice::mice(kahanall, 2, print = FALSE))

fct_scoresk2 <- data.frame(factor.scores(data_impk2[, 1:12], fit)$scores)

```

## Kahan: factor scores from third imputed dataset
```{r, echo=FALSE}

data_impk3 <- mice::complete(mice::mice(kahanall, 3, print = FALSE))

fct_scoresk3 <- data.frame(factor.scores(data_impk3[, 1:12], fit)$scores)

```

## Kahan: factor scores from fourth imputed dataset
```{r, echo=FALSE}

data_impk4 <- mice::complete(mice::mice(kahanall, 4, print = FALSE))

fct_scoresk4 <- data.frame(factor.scores(data_impk4[, 1:12], fit)$scores)

```

## Kahan: factor scores from fifth imputed dataset
```{r, echo=FALSE}

data_impk5 <- mice::complete(mice::mice(kahanall, 5, print = FALSE))

fct_scoresk5 <- data.frame(factor.scores(data_impk5[, 1:12], fit)$scores)

```

## are imputations plausible: Kahan scale - are values plausible ?

yes all value are whole numbers between 1 and 5

```{r, echo=FALSE}

summary(mikahan)

str(data_imp)
names(mikahan)

mikahan$mids

complete_data_first <- complete(mikahan$mids, 1)

summary(complete_data_first)

complete_data_2 <- complete(mikahan$mids, 2)

summary(complete_data_2)

complete_data_3 <- complete(mikahan$mids, 3)

summary(complete_data_3)

complete_data_4 <- complete(mikahan$mids, 4)

summary(complete_data_4)

complete_data_5 <- complete(mikahan$mids, 5)

summary(complete_data_5)
```

# Eco-pol scale: MAR condition and missing data pattern

```{r, echo=FALSE}

ecopoln <- formice1 %>%
  dplyr::select(
    DECISIONDECEN, DECISIONCEN, INDUSTRYSMALL, INDUSTRYLARGE,
    ECONOMYLOCAL, ECONOMYGLOBAL, ENVOVERDEV,DEVOVERENV, OWNERPVT, OWNERPUB,
    OWNERREG, OWNERNOREG, WEALTHLIM, MECHANISATION,
    DISPLACENUCLEAR,  POLLUTENUCLEAR, HEALTHNUCLEAR, JOBSNUCLEAR, BEAUTYNUCLEAR,
    PRIDENUCLEAR, NPRIDENUCLEAR, DEVNUCLEAR, PROSPERNUCLEAR,
    RELYNUCLEAR,age, urban_rural, gender, caste, religion, State
  )

md.pattern(ecopoln) #the variables with missing values: PRIDENUCLEAR PROSPERNUCLEAR JOBSNUCLEAR HEALTHNUCLEAR BEAUTYNUCLEAR POLLUTENUCLEAR DISPLACENUCLEAR RELYNUCLEAR

mcar_test(data=ecopoln) # p values is less than 0.76 - the condition for MCAR is met

# dataset with missing variables to compare and test group comaprisons

MARecopol <- codedmodule2 %>%
          dplyr::select(c(DECISIONDECEN, DECISIONCEN, INDUSTRYSMALL, INDUSTRYLARGE,
    ECONOMYLOCAL, ECONOMYGLOBAL, ENVOVERDEV, OWNERPVT, OWNERPUB,
    OWNERREG, OWNERNOREG, WEALTHLIM, MECHANISATION,
    DISPLACENUCLEAR,  POLLUTENUCLEAR, HEALTHNUCLEAR, JOBSNUCLEAR, BEAUTYNUCLEAR,
    PRIDENUCLEAR, NPRIDENUCLEAR, DEVNUCLEAR, PROSPERNUCLEAR,
    RELYNUCLEAR,age, urban_rural, gender, caste, religion, State))

MARecopol$JOBSNUCLEAR_MAR <- MARecopol$JOBSNUCLEAR

MARecopol$PRIDENUCLEAR_MAR <- MARecopol$PRIDENUCLEAR

library(finalfit)

# JOBSNUCLEAR
explanatory = c("age", "gender", "caste", "State", "urban_rural", "religion")
dependent = "JOBSNUCLEAR_MAR"
MARecopol %>% 
  missing_compare(dependent, explanatory) %>% 
    knitr::kable(row.names=FALSE, align = c("l", "l", "r", "r", "r"), 
        caption = "comparisons between values of responders (Not missing) and 
        non-responders (Missing) on the JOBSNUCLEAR variable.")

# PRIDENUCLEAR
explanatory = c("age", "gender", "caste", "State", "urban_rural", "religion")
dependent = "PRIDENUCLEAR_MAR"
MARecopol %>% 
  missing_compare(dependent, explanatory) %>% 
    knitr::kable(row.names=FALSE, align = c("l", "l", "r", "r", "r"), 
        caption = "comparisons between values of responders (Not missing) and 
        non-responders (Missing) on the PRIDENUCLEAR variable.") 
```

## Mifa for new eco-pol value scale : Nuclear energy

```{r, echo=FALSE}

miecopol <- mifa(data= ecopoln,
                cov_vars  = -c("age", "gender", "caste", "State", "urban_rural", "religion"),
                n_pc= 2:4,
                ci = "fieller", 
                n_boot=1000,
                print = FALSE,
                progress= TRUE, 
                seed = 429,)

miecopol

ecopolfit2 <- psych::fa(miecopol$cov_combined, nfactors = 2, rotate = "varimax")

print.psych(ecopolfit2, sort= TRUE, cut= 0.4)

```

#Ecopol Factor scores: mean across all imputations

```{r, echo=FALSE}

ecopolallscores <- mice::complete(mice::mice(ecopoln, m = 5, seed = 429, print = FALSE), "broad")

ecopolallscores <- ecopolallscores %>%dplyr::select(-(contains(c("age", "urban_rural", "gender","caste","religion", "State"))))

#str(ecopolallscores)

necopol <- ecopolallscores %>%
  rowwise() %>%
  mutate(DEVNUCLEAR_mean = mean(c_across(starts_with("DEVNUCLEAR"))),
         PROSPERNUCLEAR_mean = mean(c_across(starts_with("PROSPERNUCLEAR"))),
         NPRIDENUCLEAR_mean = mean(c_across(starts_with("NPRIDENUCLEAR"))),
         PRIDENUCLEAR_mean = mean(c_across(starts_with("PRIDENUCLEAR"))),
         JOBSNUCLEAR_mean = mean(c_across(starts_with("JOBSNUCLEAR"))),
         ECONOMYGLOBAL_mean = mean(c_across(starts_with("ECONOMYGLOBAL"))),
         WEALTHLIM_mean = mean(c_across(starts_with("WEALTHLIM"))),
         INDUSTRYLARGE_mean = mean(c_across(starts_with("INDUSTRYLARGE"))),
         DECISIONCEN_mean = mean(c_across(starts_with("DECISIONCEN"))),
         MECHANISATION_mean = mean(c_across(starts_with("MECHANISATION"))),
         OWNERNOREG_mean = mean(c_across(starts_with("OWNERNOREG"))),
         INDUSTRYSMALL_mean = mean(c_across(starts_with("INDUSTRYSMALL"))),
         DEVOVERENV_mean = mean(c_across(starts_with("DEVOVERENV"))),
         OWNERPVT_mean = mean(c_across(starts_with("OWNERPVT"))),
         RELYNUCLEAR_mean = mean(c_across(starts_with("RELYNUCLEAR"))),
         BEAUTYNUCLEAR_mean = mean(c_across(starts_with("BEAUTYNUCLEAR"))),
         HEALTHNUCLEAR_mean = mean(c_across(starts_with("HEALTHNUCLEAR"))),
         POLLUTENUCLEAR_mean = mean(c_across(starts_with("POLLUTENUCLEAR"))),
         DISPLACENUCLEAR_mean = mean(c_across(starts_with("DISPLACENUCLEAR"))),
         ENVOVERDEV_mean = mean(c_across(starts_with("ENVOVERDEV"))),
         OWNERPUB_mean = mean(c_across(starts_with("OWNERPUB"))),
         OWNERREG_mean = mean(c_across(starts_with("OWNERREG"))),
         DECISIONDECEN_mean = mean(c_across(starts_with("DECISIONDECEN"))),
         ECONOMYLOCAL_mean = mean(c_across(starts_with("ECONOMYLOCAL")))
  ) %>%
  ungroup()%>%
  dplyr::select(contains("_mean")) %>%
   rename_with(~ gsub("_mean$", "", .x))

necopolfinal <- data.frame(factor.scores(necopol[, 1:24], ecopolfit2)$scores)

```

# Eco-pol: Radar map

```{r, echo=FALSE}
library(ggplot2)
library(fmsb)

nrow(ecopolfit2$loadings)

# Extract the loadings
loadings_matrix4 <- ecopolfit2$loadings[1:24, ]

loadings_df2 <- as.data.frame(as.matrix(loadings_matrix4))

# Adding rownames to the dataframe as a new column
loadings_df2 <- loadings_df2 %>%
  rownames_to_column(var = "variable")

# Pivot to longer format first
long_loadings_df2 <- loadings_df2 %>%
  pivot_longer(cols = c(MR1, MR2), names_to = "Factors", values_to = "value")

# Pivot the longer format to wide format
wide_loadings_df2 <- long_loadings_df2 %>%
  pivot_wider(names_from = variable, values_from = value) %>%
  mutate(Factors = case_when(
    Factors == "MR1" ~ "Nationalist Development (Ndevelop)",
    Factors == "MR2" ~ "People Centered Development (Pdevelop)",
    TRUE ~ Factors
  )) %>%
  column_to_rownames("Factors")

# Define the variable ranges: maximum and minimum
max_min2 <- data.frame(
  DEVNUCLEAR = c(0.8, -0.6), PROSPERNUCLEAR = c(0.8, -0.6), NPRIDENUCLEAR = c(0.8, -0.6), PRIDENUCLEAR = c(0.8, -0.6),     JOBSNUCLEAR = c(0.8, -0.6), ECONOMYGLOBAL = c(0.8, -0.6), WEALTHLIM = c(0.8, -0.6), INDUSTRYLARGE = c(0.8, -0.6),   DECISIONCEN = c(0.8, -0.6), MECHANISATION = c(0.8, -0.6), OWNERNOREG = c(0.8, -0.6), INDUSTRYSMALL = c(0.8, -0.6), DEVOVERENV = c(0.8, -0.6), OWNERPVT = c(0.8, -0.6), RELYNUCLEAR = c(0.8, -0.6), BEAUTYNUCLEAR = c(0.8, -0.6), HEALTHNUCLEAR = c(0.8, -0.6), POLLUTENUCLEAR = c(0.8, -0.6), DISPLACENUCLEAR = c(0.8, -0.6), DECISIONDECEN = c(0.8, -0.6),
ECONOMYLOCAL = c(0.8, -0.6), ENVOVERDEV = c(0.8, -0.6), OWNERPUB = c(0.8, -0.6), OWNERREG = c(0.8, -0.6)   
)
rownames(max_min2) <- c("Max", "Min")

# Assuming you have a dataframe 'exam_scores' with matching column names
# Bind the variable ranges to the data

radar4 <- rbind(max_min2, wide_loadings_df2) %>%
    rename(
      "National development(nuclear)" = "DEVNUCLEAR",
      "Local prosperity(nuclear)" = "PROSPERNUCLEAR",
      "National pride(nuclear)" = "NPRIDENUCLEAR",
      "Community pride(nuclear)" = "PRIDENUCLEAR",
      "Job opportunities(nuclear)" = "JOBSNUCLEAR",
      "Pro Globaleconomy" = "ECONOMYGLOBAL",
      "Limits on Wealth " = "WEALTHLIM",
      "Pro Large Industries" = "INDUSTRYLARGE",
      "Pro Centralisation" = "DECISIONCEN",
      "Anti Mechanisation of work" = "MECHANISATION",
      "Anti Regulations" = "OWNERNOREG",
      "Anti Large Industries " = "INDUSTRYSMALL",
      "Development over Environment" = "DEVOVERENV",
      "Pro Private ownership " = "OWNERPVT",
      "Reliance on government(nuclear)" = "RELYNUCLEAR",
      "Spoils Natural Beauty(nuclear)" = "BEAUTYNUCLEAR",
      "Health risk(nuclear)" = "HEALTHNUCLEAR",
      "Pollution risk(nuclear) " = "POLLUTENUCLEAR",
      "Displacement risk(nuclear)" = "DISPLACENUCLEAR",
      "Environment over Development" = "ENVOVERDEV",
      "Pro Public ownership" = "OWNERPUB",
      "Pro Regulations" = "OWNERREG",
      "Pro Decentralisation" = "DECISIONDECEN",
      "Pro Localeconomy" = "ECONOMYLOCAL",
    )

```

```{r, echo=FALSE, fig.width=15, fig.height=7, out.width='70', out.height='70'}
library(fmsb)

create_beautiful_radarchart <- function(data, color = "#00AFBB", 
                                        vlabels = colnames(data), vlcex = 0.7,
                                        caxislabels = NULL, title = NULL, ...){
  radarchart(
    data, axistype = 1,
    # Customize the polygon
    pcol = color, pfcol = scales::alpha(color, 0.7), plwd = 2, plty = 1,
    centerzero=FALSE,
    # Customize the grid
    cglcol = "black", cglty = 1, cglwd = 2,
    # Customize the axis
    axislabcol = "black", cexc = 1.5, # Increase label size for axis
    # Variable labels
    vlcex = vlcex, vlabels = vlabels, vlpos = 0.98,
    seg = 7, caxislabels = caxislabels, title = title, 
    # Increase title size
    cex.main = 4,
    ...
  )
}

# Reduce plot margin using par()
op <- par(mar = c(7, 3, 3, 3))
# Create the radar charts
create_beautiful_radarchart(
  data = radar4, caxislabels = c(-0.6, -0.4, -0.2, 0, 0.2, 0.4, 0.6, 0.8),
  color = c( "#F97D09","#00AFBB"),
  vlcex = 1.1
)

# Add an horizontal legend
legend(
  x = "bottomright", legend = rownames(radar4[-c(1,2),]), horiz = TRUE,
  bty = "n", pch = 20 , col = c("#F97D09","#00AFBB"),
  text.col = "black", cex = 1.5, pt.cex = 3,
  inset = c(0, 0)
  )
par(op)

```

```{r, echo=FALSE}

# Extracting factors from first imputed dataset

data_imp2 <- mice::complete(mice::mice(ecopoln, 1, print = FALSE))

fct_scores2 <- data.frame(factor.scores(data_imp2[, 1:24], ecopolfit2)$scores)

```

##  eco-pol scale : Factor scores from 2nd imputation
```{r, echo=FALSE}

data_impe2 <- mice::complete(mice::mice(ecopoln, 2, print = FALSE))

fct_scorese2 <- data.frame(factor.scores(data_impe2[, 1:24], ecopolfit2)$scores)

```

##  eco-pol scale : Factor scores from 3rd imputation
```{r, echo=FALSE}

data_impe3 <- mice::complete(mice::mice(ecopoln, 3, print = FALSE))

fct_scorese3 <- data.frame(factor.scores(data_impe3[, 1:24], ecopolfit2)$scores)

```

##  eco-pol scale : Factor scores from 4th imputation
```{r, echo=FALSE}

data_impe4 <- mice::complete(mice::mice(ecopoln, 4, print = FALSE))

fct_scorese4 <- data.frame(factor.scores(data_impe4[, 1:24], ecopolfit2)$scores)

```

##  eco-pol scale : Factor scores from 5th imputation
```{r, echo=FALSE}

data_impe5 <- mice::complete(mice::mice(ecopoln, 5, print = FALSE))

fct_scorese5 <- data.frame(factor.scores(data_impe5[, 1:24], ecopolfit2)$scores)

```

## Are imputations plausible : Eco-pol scale

yes, imputations are pluasible- all natural numbers between 1 and 5

```{r, echo=FALSE}

summary(miecopol)
names(miecopol)

miecopol$mids

complete_ecopol_first <- complete(miecopol$mids, 1)

summary(complete_ecopol_first)

complete_ecopol_2 <- complete(miecopol$mids, 2)

summary(complete_ecopol_2)

complete_ecopol_3 <- complete(miecopol$mids, 3)

summary(complete_ecopol_3)

complete_ecopol_4 <- complete(miecopol$mids, 4)

summary(complete_ecopol_4)

complete_ecopol_5 <- complete(miecopol$mids, 5)

summary(complete_ecopol_5)

```

# Checking consistency of LM coefficients across 5 imputed datasets 
coefficients and LMs are fairly consistent

## Binding datasets: 1st imputation

```{r, echo = FALSE}

# binding the imputed  factor analysis scores for kahan scale and eco-pol value scale to the larger dataset codedmodule2

ecopolnuclear <- fct_scores2 %>%
  rename(Ndevelop = MR1, Pdevelop = MR2)

kahannuclear <- fct_scores%>%
  rename(Hierarchy_Egalitarian = MR1, Communitarian = MR2)

allformice<- cbind(formice1,ecopolnuclear, kahannuclear)

```

## State Ns before MIFA

```{r, echo=FALSE}
Statenacheck1 <- allformice%>%
  na.omit()%>%
  dplyr::select(State)

summary(Statenacheck1)
```

```{r, echo=FALSE}

# Allformice: Do imputations merge ?

# Initialize the mice process to create the predictor matrix
#ini <- mice(allformice, maxit = 0, pri = FALSE)

# Extract the predictor matrix from the initialized object
#pred <- ini$predictorMatrix

# Set the columns for "Uppercaste", "Male", "Hindu", and "Age" to 0
#pred[, "Uppercaste"] <- 0
#pred[, "Male"] <- 0
#pred[, "Hindu"] <- 0
#pred[, "Age"] <- 0
#pred[, "Ndevelop"] <- 0
#pred[, "Pdevelop"] <- 0
#pred[, "Hierarchy_Egalitarian"] <- 0
#pred[, "Communitarian"] <- 0
# Identify columns that start with "K_" and set them to 0
#k_columns <- grep("^K_", colnames(formice1))
#pred[, k_columns] <- 0

# Perform the mice imputation with the updated predictor matrix
#allformice.imp <- mice(allformice, m = 5, maxit = 10, predictorMatrix = pred, seed = 430, print = TRUE)

#summary(allformice.imp)
#allformice.imp$chainMean

#plot(allformice.imp)

```

## State Ns after MIFA on two scales

```{r, echo=FALSE}

nuclear_missing <- allformice %>% dplyr::select(Risky_Nuclear,Uppercaste, Male, Hindu, urban_rural, Age, State)%>%
  na.omit()

summary(nuclear_missing)

```

## LMs with 1st imputed datasets from FAs

```{r, echo=FALSE}

nuclearIVmice <- lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + urban_rural + Age + State, data = allformice)

nuclearIVmice2 <- lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + urban_rural + Age + State + Communitarian + Hierarchy_Egalitarian, data = allformice)

nuclearIVmice3 <- lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + urban_rural + Age + State + Communitarian + Hierarchy_Egalitarian + Pdevelop + Ndevelop, data = allformice)

summary(nuclearIVmice)
summary(nuclearIVmice2)
summary(nuclearIVmice3)

```

## binding dataset - 2nd imputation

```{r, echo=FALSE}

ecopolnuclear2 <- fct_scorese2 %>%
  rename(Ndevelop = MR1, Pdevelop = MR2)

kahannuclear2 <- fct_scoresk2%>%
  rename(Hierarchy_Egalitarian = MR1, Communitarian = MR2)

allformice2<- cbind(formice1,ecopolnuclear2, kahannuclear2)

```

## LMs with second imputation 
```{r, echo=FALSE}

nuclearIVmiceimp2 <- lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + urban_rural + Age + State, data = allformice2)

nuclearIVmiceimp22 <- lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + urban_rural + Age + State + Communitarian + Hierarchy_Egalitarian, data = allformice2)

nuclearIVmiceimp23 <- lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + urban_rural + Age + State + Communitarian + Hierarchy_Egalitarian + Pdevelop + Ndevelop, data = allformice2)

summary(nuclearIVmiceimp2)
summary(nuclearIVmiceimp22)
summary(nuclearIVmiceimp23)

```

## binding dataset - 3rd imputation

```{r, echo=FALSE}

ecopolnuclear3 <- fct_scorese3 %>%
  rename(Ndevelop = MR1, Pdevelop = MR2)

kahannuclear3 <- fct_scoresk3%>%
  rename(Hierarchy_Egalitarian = MR1, Communitarian = MR2)

allformice3<- cbind(formice1,ecopolnuclear3, kahannuclear3)

```

## LMs with third imputation 
```{r, echo=FALSE}

nuclearIVmiceimp3 <- lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + urban_rural + Age + State, data = allformice3)

nuclearIVmiceimp32 <- lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + urban_rural + Age + State + Communitarian + Hierarchy_Egalitarian, data = allformice3)

nuclearIVmiceimp33 <- lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + urban_rural + Age + State + Communitarian + Hierarchy_Egalitarian + Pdevelop + Ndevelop, data = allformice3)

summary(nuclearIVmiceimp3)
summary(nuclearIVmiceimp32)
summary(nuclearIVmiceimp33)

```

## binding dataset - 4th imputation

```{r, echo=FALSE}

ecopolnuclear4 <- fct_scorese4 %>%
  rename(Ndevelop = MR1, Pdevelop = MR2)

kahannuclear4 <- fct_scoresk4%>%
  rename(Hierarchy_Egalitarian = MR1, Communitarian = MR2)

allformice4<- cbind(formice1, ecopolnuclear4, kahannuclear4)

```

## LMs with fourth imputation 
```{r, echo=FALSE}

nuclearIVmiceimp4 <- lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + urban_rural + Age + State, data = allformice4)

nuclearIVmiceimp24 <- lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + urban_rural + Age + State + Communitarian + Hierarchy_Egalitarian, data = allformice4)

nuclearIVmiceimp34 <- lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + urban_rural + Age + State + Communitarian + Hierarchy_Egalitarian + Pdevelop + Ndevelop, data = allformice4)

summary(nuclearIVmiceimp4)
summary(nuclearIVmiceimp24)
summary(nuclearIVmiceimp34)

```

# Mean across imputations

## Binding datasets

```{r, echo = FALSE}

# binding the imputed  factor analysis scores for kahan scale and eco-pol value scale to the larger dataset codedmodule2

ecopolnuclearfinal <- necopolfinal %>%
  rename(Ndevelop = MR1, Pdevelop = MR2)

kahannuclearfinal <- nkahanfinal%>%
  rename(Hierarchy_Egalitarian = MR1, Communitarian = MR2)

allformicefinal<- cbind(formice1,ecopolnuclearfinal, kahannuclearfinal)

```

## LMs with final imputed datasets from FAs

```{r, echo=FALSE}

nuclearfinal <- lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + urban_rural + Age, data = allformicefinal)

nuclearfinal2 <- lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + urban_rural + Age + State, data = allformicefinal)

nuclearfinal3 <- lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + urban_rural + Age + State + Communitarian + Hierarchy_Egalitarian, data = allformicefinal)

nuclearfinal4 <- lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + urban_rural + Age + State + Communitarian + Hierarchy_Egalitarian + Pdevelop + Ndevelop, data = allformicefinal)

#summary(nuclearfinal)
#summary(nuclearfinal2)
#summary(nuclearfinal3)
#summary(nuclearfinal4)
```

\newpage
\blandscape

```{r, echo=FALSE, results='asis'}
library(stargazer)

cat("\\begingroup")
cat("\\setlength{\\tabcolsep}{1pt}")
cat("\\renewcommand{\\arraystretch}{0.7}")
stargazer(nuclearfinal, nuclearfinal2, nuclearfinal3, nuclearfinal4,
          title = "4 linear models: Perceived Risk of Nuclear Energy",
          type = "latex"
)
cat("\\endgroup")

```

\elandscape

```{r, echo=FALSE}

# LM with demographic variables and State 
# The pool() function from the mice package creates a list containing the results,
# so we need to access the summary of the model directly.

#pooled_fit4 <- pool(nuclearfit4)

#pooled_summary4 <- summary(pooled_fit4)

#coefficients_df4 <- pooled_summary4
# To ensure the numeric values are rounded to two decimal places, you can use:
#coefficients_df4 <- round(coefficients_df4[, -1], 2)  # Skipping the first column as it's factors (term names)
#coefficients_df4$term <- pooled_summary4$term  # Adding back the unrounded terms

# Print the dataframe
#data.frame(coefficients_df4) %>%
 # dplyr::select(term, estimate, std.error, statistic, df, p.value)
```