### 2024-03-23 miceEXTRA.Rmd
put extra imputation tests etc for mice into another file

---
title: "miceEXTRA"
output: html_document
---

# EXTRA

## Kahan: factor scores from 1st imputed dataset
```{r, echo=FALSE}

# extracting factor scores from mifa for 1st imputed dataset

data_imp <- mice::complete(mice::mice(kahanall, 1, print = FALSE))

fct_scores <- data.frame(factor.scores(data_imp[, 1:12], fit)$scores)

plot(kahan_imp)

```

## Kahan: factor scores from second imputed dataset
```{r, echo=FALSE}

data_impk2 <- mice::complete(mice::mice(kahanall, 2, print = FALSE))

fct_scoresk2 <- data.frame(factor.scores(data_impk2[, 1:12], fit)$scores)

```

## Kahan: factor scores from third imputed dataset
```{r, echo=FALSE}

data_impk3 <- mice::complete(mice::mice(kahanall, 3, print = FALSE))

fct_scoresk3 <- data.frame(factor.scores(data_impk3[, 1:12], fit)$scores)

```

## Kahan: factor scores from fourth imputed dataset
```{r, echo=FALSE}

data_impk4 <- mice::complete(mice::mice(kahanall, 4, print = FALSE))

fct_scoresk4 <- data.frame(factor.scores(data_impk4[, 1:12], fit)$scores)

```

## Kahan: factor scores from fifth imputed dataset
```{r, echo=FALSE}

data_impk5 <- mice::complete(mice::mice(kahanall, 5, print = FALSE))

fct_scoresk5 <- data.frame(factor.scores(data_impk5[, 1:12], fit)$scores)

```

## are imputations plausible: Kahan scale - are values plausible ?

yes all value are whole numbers between 1 and 5

```{r, echo=FALSE}

summary(mikahan)

str(data_imp)
names(mikahan)

mikahan$mids

complete_data_first <- complete(mikahan$mids, 1)

summary(complete_data_first)

complete_data_2 <- complete(mikahan$mids, 2)

summary(complete_data_2)

complete_data_3 <- complete(mikahan$mids, 3)

summary(complete_data_3)

complete_data_4 <- complete(mikahan$mids, 4)

summary(complete_data_4)

complete_data_5 <- complete(mikahan$mids, 5)

summary(complete_data_5)
```

```{r, echo=FALSE}

# Extracting factors from first imputed dataset

data_imp2 <- mice::complete(mice::mice(ecopoln, 1, print = FALSE))

fct_scores2 <- data.frame(factor.scores(data_imp2[, 1:24], ecopolfit2)$scores)

```

##  eco-pol scale : Factor scores from 2nd imputation
```{r, echo=FALSE}

data_impe2 <- mice::complete(mice::mice(ecopoln, 2, print = FALSE))

fct_scorese2 <- data.frame(factor.scores(data_impe2[, 1:24], ecopolfit2)$scores)

```

##  eco-pol scale : Factor scores from 3rd imputation
```{r, echo=FALSE}

data_impe3 <- mice::complete(mice::mice(ecopoln, 3, print = FALSE))

fct_scorese3 <- data.frame(factor.scores(data_impe3[, 1:24], ecopolfit2)$scores)

```

##  eco-pol scale : Factor scores from 4th imputation
```{r, echo=FALSE}

data_impe4 <- mice::complete(mice::mice(ecopoln, 4, print = FALSE))

fct_scorese4 <- data.frame(factor.scores(data_impe4[, 1:24], ecopolfit2)$scores)

```

##  eco-pol scale : Factor scores from 5th imputation
```{r, echo=FALSE}

data_impe5 <- mice::complete(mice::mice(ecopoln, 5, print = FALSE))

fct_scorese5 <- data.frame(factor.scores(data_impe5[, 1:24], ecopolfit2)$scores)

```

## Are imputations plausible : Eco-pol scale

yes, imputations are pluasible- all natural numbers between 1 and 5

```{r, echo=FALSE}

summary(miecopol)
names(miecopol)

miecopol$mids

complete_ecopol_first <- complete(miecopol$mids, 1)

summary(complete_ecopol_first)

complete_ecopol_2 <- complete(miecopol$mids, 2)

summary(complete_ecopol_2)

complete_ecopol_3 <- complete(miecopol$mids, 3)

summary(complete_ecopol_3)

complete_ecopol_4 <- complete(miecopol$mids, 4)

summary(complete_ecopol_4)

complete_ecopol_5 <- complete(miecopol$mids, 5)

summary(complete_ecopol_5)

```

# Checking consistency of LM coefficients across 5 imputed datasets 
coefficients and LMs are fairly consistent

## Binding datasets: 1st imputation

```{r, echo = FALSE}

# binding the imputed  factor analysis scores for kahan scale and eco-pol value scale to the larger dataset codedmodule2

ecopolnuclear <- fct_scores2 %>%
  rename(Ndevelop = MR1, Pdevelop = MR2)

kahannuclear <- fct_scores%>%
  rename(Egalitarian = MR1, Communitarian = MR2)

allformice<- cbind(formice1,ecopolnuclear, kahannuclear)

```

## LMs with 1st imputed datasets from FAs

```{r, echo=FALSE}

nuclearIVmice <- lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + urban_rural + Age + State, data = allformice)

nuclearIVmice2 <- lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + urban_rural + Age + State + Communitarian + Egalitarian, data = allformice)

nuclearIVmice3 <- lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + urban_rural + Age + State + Communitarian + Egalitarian + Pdevelop + Ndevelop, data = allformice)

summary(nuclearIVmice)
summary(nuclearIVmice2)
summary(nuclearIVmice3)

```

## binding dataset - 2nd imputation

```{r, echo=FALSE}

ecopolnuclear2 <- fct_scorese2 %>%
  rename(Ndevelop = MR1, Pdevelop = MR2)

kahannuclear2 <- fct_scoresk2%>%
  rename(Egalitarian = MR1, Communitarian = MR2)

allformice2<- cbind(formice1,ecopolnuclear2, kahannuclear2)

```

## LMs with second imputation 
```{r, echo=FALSE}

nuclearIVmiceimp2 <- lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + urban_rural + Age + State, data = allformice2)

nuclearIVmiceimp22 <- lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + urban_rural + Age + State + Communitarian + Egalitarian, data = allformice2)

nuclearIVmiceimp23 <- lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + urban_rural + Age + State + Communitarian + Egalitarian + Pdevelop + Ndevelop, data = allformice2)

summary(nuclearIVmiceimp2)
summary(nuclearIVmiceimp22)
summary(nuclearIVmiceimp23)

```

## binding dataset - 3rd imputation

```{r, echo=FALSE}

ecopolnuclear3 <- fct_scorese3 %>%
  rename(Ndevelop = MR1, Pdevelop = MR2)

kahannuclear3 <- fct_scoresk3%>%
  rename(Egalitarian = MR1, Communitarian = MR2)

allformice3<- cbind(formice1,ecopolnuclear3, kahannuclear3)

```

## LMs with third imputation 
```{r, echo=FALSE}

nuclearIVmiceimp3 <- lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + urban_rural + Age + State, data = allformice3)

nuclearIVmiceimp32 <- lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + urban_rural + Age + State + Communitarian + Egalitarian, data = allformice3)

nuclearIVmiceimp33 <- lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + urban_rural + Age + State + Communitarian + Egalitarian + Pdevelop + Ndevelop, data = allformice3)

summary(nuclearIVmiceimp3)
summary(nuclearIVmiceimp32)
summary(nuclearIVmiceimp33)

```

## binding dataset - 4th imputation

```{r, echo=FALSE}

ecopolnuclear4 <- fct_scorese4 %>%
  rename(Ndevelop = MR1, Pdevelop = MR2)

kahannuclear4 <- fct_scoresk4%>%
  rename(Egalitarian = MR1, Communitarian = MR2)

allformice4<- cbind(formice1, ecopolnuclear4, kahannuclear4)

```

## LMs with fourth imputation 
```{r, echo=FALSE}

nuclearIVmiceimp4 <- lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + urban_rural + Age + State, data = allformice4)

nuclearIVmiceimp24 <- lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + urban_rural + Age + State + Communitarian + Egalitarian, data = allformice4)

nuclearIVmiceimp34 <- lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + urban_rural + Age + State + Communitarian + Egalitarian + Pdevelop + Ndevelop, data = allformice4)

summary(nuclearIVmiceimp4)
summary(nuclearIVmiceimp24)
summary(nuclearIVmiceimp34)

```

```{r, echo=FALSE}

# Allformice: Do imputations merge ?

# Initialize the mice process to create the predictor matrix
#ini <- mice(allformice, maxit = 0, pri = FALSE)

# Extract the predictor matrix from the initialized object
#pred <- ini$predictorMatrix

# Set the columns for "Uppercaste", "Male", "Hindu", and "Age" to 0
#pred[, "Uppercaste"] <- 0
#pred[, "Male"] <- 0
#pred[, "Hindu"] <- 0
#pred[, "Age"] <- 0
#pred[, "Ndevelop"] <- 0
#pred[, "Pdevelop"] <- 0
#pred[, "Egalitarian"] <- 0
#pred[, "Communitarian"] <- 0
# Identify columns that start with "K_" and set them to 0
#k_columns <- grep("^K_", colnames(formice1))
#pred[, k_columns] <- 0

# Perform the mice imputation with the updated predictor matrix
#allformice.imp <- mice(allformice, m = 5, maxit = 10, predictorMatrix = pred, seed = 430, print = TRUE)

#summary(allformice.imp)
#allformice.imp$chainMean

#plot(allformice.imp)

```

```{r, echo=FALSE}

# LM with demographic variables and State 
# The pool() function from the mice package creates a list containing the results,
# so we need to access the summary of the model directly.

#pooled_fit4 <- pool(nuclearfit4)

#pooled_summary4 <- summary(pooled_fit4)

#coefficients_df4 <- pooled_summary4
# To ensure the numeric values are rounded to two decimal places, you can use:
#coefficients_df4 <- round(coefficients_df4[, -1], 2)  # Skipping the first column as it's factors (term names)
#coefficients_df4$term <- pooled_summary4$term  # Adding back the unrounded terms

# Print the dataframe
#data.frame(coefficients_df4) %>%
 # dplyr::select(term, estimate, std.error, statistic, df, p.value)
```